### 2019-10-16 docs/article.tex
adding article

\documentclass{article}[12pt]
\renewcommand{\baselinestretch}{1.5}
\setlength{\parskip}{1em}

\usepackage[parfill]{parskip}
\usepackage[affil-it]{authblk}
\usepackage[space]{grffile}

\usepackage[a4paper]{geometry}
\geometry{verbose}
\usepackage{float}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{caption}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{latexsym,textcomp,longtable,tabulary}
\usepackage{booktabs,array,multirow,braket}
\usepackage{amsfonts,amsmath,amssymb,mathbbol,calc,cancel}
\usepackage{subfigure,color,blindtext,enumitem,siunitx}
\usepackage[colorinlistoftodos]{todonotes}

\usepackage{mathtools}
\usepackage{url,hyperref,etoolbox}
\numberwithin{equation}{section}
\hypersetup{colorlinks=false,pdfborder={0 0 0}}

%+figure layout options
\restylefloat{figure}
\setlist{leftmargin=*,before=\setlength{\rightmargin}{\leftmargin}}
%-figure layout options

\providecommand\citet{\cite}
\providecommand\citep{\cite}
\providecommand\citealt{\cite}

\makeatletter
\makeatother

\begin{document}

\title{
Inference and Synthesis of\\
co-dimension one Bifurcations
}

\author{Gregory Szep}
\affil{King's College London}
\date{\today}
\maketitle
\vspace{-25pt}
\section{State-space Density} \vspace{-10pt}

\bibliography{refs}
\bibliographystyle{ieeetr}
\end{document}

### 2019-10-16 docs/article.tex
scaffold of paper

@@
 
-\author{Gregory Szep}
-\affil{King's College London}
+\author{Gregory Szep$^1$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^1$}
+\affil{$^1$King's College London, $^2$Microsoft Research Cambridge}
 \date{\today}
@@
 \vspace{-25pt}
-\section{State-space Density} \vspace{-10pt}
+\section{Introduction and Motivation}
+
+\begin{itemize}
+\item qualitative equivalence more important
+\item fitting to time courses not appropriate
+\item gradient information makes optimisation tractable
+\end{itemize}
+
+\section{Problem Statement}
+
+Suppose we would like to know the regions within paramter space
+$\theta\in\mathbb{R}^N$
+where the steady state of a set of differential equations
+takes on a specified target form $t(p)$ along a known parameter
+$p\in\mathbb{R}$. The target is a label $t\in\{0,1\}$ that indicates
+whether the system is monostable or multi-stable at given $p$. We begin with a known form
+for the differential equations for $u\in\mathbb{R}^M$
+\begin{align}
+	\partial_t u=F(u|\,p,\theta) 
+\end{align}
+and the set of steady states
+$U(\theta) := \left\{ (u,p)\in\mathbb{R}^{M+1} : F(u|\,p,\theta)=0\right\}$
+can be found in a local region numerically using
+pseudo-arclength continuation along $u(s|\theta),p(s|\theta)$ \cite{} where
+$s$ parametrises the one co-dimensional curve -- details in section
+\ref{sec:continuation}.
+In order to perform an optimisation 
+we need mapping from the states $U(\theta)$ to an output that can be
+compared to $t(p)$. Section \ref{sec:objective-function} motivates this mapping and
+constructs the objective function.
+
+\section{Pseudo-arclength Continuation}
+\label{sec:continuation}
+
+\begin{itemize}
+    \item predictor-corrector algorithm
+    \end{itemize}
+
+\section{Objective Function}
+\label{sec:objective-function}
+
+First we note that although $U(\theta)$ contains both $u,p$ we
+are only interested in how many unique $u$ there are for a given $p$.
+This can be done by taking a histogram of equally sized bins along $p$.
+Another way of getting a quantity that is proportional to count is
+by convolution $*$ of $p(s|\theta)$ along $s$ with a symmetric kernel
+$\phi(p\,|\alpha)$ with a bandwidth $\alpha$ indicating bin width.
+This way we get a smooth and differentiable version of the histogram
+operation.
+\begin{align}
+    \phi(p\,|\alpha)*U(\theta):=
+    \int_{\mathbb{R}}\!\phi(p-p(s|\theta)\,|\alpha)\,\mathrm{d}s
+\end{align}
+The result is now proportional to count, but needs to
+be mapped to the unit interval $[0,1]$ in such a way that high
+count regions that indicate multi-stability are mapped to one, and
+low counts which indicate monostability are mapped to zero. For
+this a sigmoidal activation function $\sigma(\frac{x-\mu}{\beta})$ with
+unknown smoothness $\beta$ and threshold $\mu$ will do. We
+are now ready to write down the objective function
+\begin{align}
+	\mathcal{J}(\theta\,|\,\alpha,\beta,\mu):=                            
+	\int_{\mathbb{R}}\left|\;                                                  
+    t(p) - \sigma\left(\frac{\phi(p\,|\alpha)*U(\theta)-\mu}{\beta}\right)
+	\,\right|^2\,\mathrm{d}p
+\end{align}
+where $\theta$ are the parameters to be optimised and $\alpha,\beta,\mu$
+are hyper-parameters. Note that $U(\theta)$ is the only $\theta$ dependence
+in the expression. When we apply
+\begin{align}
+    \partial_\theta\mathcal{J}
+    &=-2\int_{\mathbb{R}}
+    \sigma'\,\partial_\theta\left(
+        \frac{\phi(p\,|\alpha)*U(\theta)-\mu}{\beta}
+        \right)
+    \,\mathrm{d}p\\
+    &=-\frac{2}{\beta}\int_{\mathbb{R}}
+    \sigma'\,
+        \partial_\theta(\,\phi(p|\alpha)*U(\theta)\,)
+    \,\mathrm{d}p\\
+    &=\frac{2}{\beta}
+    \int_{\mathbb{R}}\sigma'\!\! \int_{\mathbb{R}}\phi'\,
+        \partial_\theta p(s|\theta)
+    \,\mathrm{d}s \,\mathrm{d}p
+\end{align}
+remarkably everything is differentiable including the algorithm in section
+\ref{sec:continuation}. This gradient information can be used in a
+gradient descent optimization to minimize $\mathcal{J}(\theta)$.
+
+\section{Normal Forms}
+\label{sec:normal-forms}
+
+\begin{itemize}
+    \item saddle-node, pitchfork, transcritical
+    \item optimisation landscape with changing hyperparams
+    \item benchmarks against other algos
+\end{itemize}
+
+\section{Chemical Reaction Networks}
+\label{sec:networks}
+
+\begin{itemize}
+    \item toggle switch
+    \item cell cycle
+    \item application to structure $\rightarrow$ function (Luca)
+\end{itemize}
+
+\section{Conclusions and Extensions}
+\label{sec:conclusions}
+
+\begin{itemize}
+    \item hyperparam optimization
+    \item hopf bifurcations
+    \item pattern formation in pdes
+\end{itemize}
 

### 2019-10-16 continuation.jl
refactored code to work with base library

@@ -1,18 +1,38 @@
-using PseudoArcLengthContinuation, Flux
+using Flux,PseudoArcLengthContinuation
 using PseudoArcLengthContinuation: ContResult, DefaultLS, DefaultEig
 const Cont = PseudoArcLengthContinuation
 
-unpack(curve::ContResult) = Tracker.collect(curve.branch[1,:]),Tracker.collect(curve.branch[2,:])
+import Base.vcat
+import Plots: plot, plot!
+import PseudoArcLengthContinuation: eigen!
+using Flux.Tracker: TrackedReal,TrackedArray
 
-function continuation( f, J, u₀, p₀::T ; kwargs...) where T
+unpack(curve::ContResult) = tuple([ curve.branch[i,:] for i=1:size(curve.branch)[1]-2 ]...)
+initial_state(curve::ContResult) = [ param(curve.branch[i,1].data) for i=2:size(curve.branch)[1]-2 ]
 
-	options = Cont.NewtonPar{T,typeof(DefaultLS()), typeof(DefaultEig())}(verbose=false,maxIter=kwargs[:maxIter])
-	u₀, _, stable = Cont.newton( u -> f(u,p₀), u -> J(u,p₀), convert(Array{T},u₀), options )
-	branch, _, _ = Cont.continuation( f,J, u₀,p₀,
+plot(x::Vector{TrackedReal{T}}; kwargs...) where {T<:Real} = plot( Tracker.data.(x); kwargs...)
+plot!(x::Vector{TrackedReal{T}}; kwargs...) where {T<:Real} = plot!( Tracker.data.(x); kwargs...)
 
-		ContinuationPar{T,typeof(DefaultLS()), typeof(DefaultEig())}(
+plot(x::Vector{TrackedReal{T}},y::Vector{TrackedReal{T}}; kwargs...) where {T<:Real} = plot( Tracker.data.(x), Tracker.data.(y); kwargs...)
+plot!(x::Vector{TrackedReal{T}},y::Vector{TrackedReal{T}}; kwargs...) where {T<:Real} = plot!( Tracker.data.(x), Tracker.data.(y); kwargs...)
+plot!(x::AbstractArray,y::TrackedArray; kwargs...) = plot!( x, y.data; kwargs...)
+
+vcat(p::TrackedReal{T}, u::Vector{TrackedReal{T}}, i::Int64, ds::TrackedReal{T}) where {T<:Real} = [ p, u..., i, ds ]
[…]

### 2019-10-16 inference.jl

@@ -7,30 +7,40 @@ struct StateDensity
 end
 
 function infer( f::Function, J::Function, θ::TrackedArray, data::StateDensity; optimiser=ADAM(0.1),
-		iter::Int=100, u₀=[-2.0,0.0], maxSteps::Int=1000, maxIter::Int=1000 )
+		iter::Int=100, u₀=param.([-2.0,0.0]), maxSteps::Int=1000, maxIter::Int=1000 )
 
 	# setting initial hyperparameters
-	P,U = param([0.0]),param([0.0])
 	p₀ = param(minimum(data.parameter))
 	pMax,ds = maximum(data.parameter), step(data.parameter)
 
+	bifurcations, = continuation( f,J,u₀,p₀; pMin=p₀-ds, pMax=pMax, ds=ds,
+		maxSteps=maxSteps, maxIter=maxIter, computeEigenValues=true )
+
+	prediction = kde( unpack(bifurcations)[1], data.parameter, bandwidth=1.4*ds)
+	u₀ = initial_state(bifurcations)
+
 	function predictor()
 
-		u₀,P,U = continuation( f,J,u₀,p₀; pMin=p₀-ds, pMax=pMax, ds=ds, maxSteps=maxSteps, maxIter=maxIter )
-		kernel = kde(P,data.parameter,bandwidth=1.4*ds)
-		return kernel.density
+		bifurcations, = continuation( f,J,u₀, p₀; pMin=p₀-ds, pMax=pMax, ds=ds,
+			maxSteps=maxSteps, maxIter=maxIter, computeEigenValues=true )
+
+		density = kde( unpack(bifurcations)[1], data.parameter, bandwidth=1.4*ds)
+		u₀ = initial_state(bifurcations)
+
+		return bifurcations,density
 	end
 
-	loss() = norm( predictor() .- data.density )
+	function loss()
+		_,prediction = predictor()
+		ℒ = norm( prediction.density .- data.density )
+		return ℒ
+	end
 
 	function progress()
 		@printf("Loss = %f, θ = [%f,%f,%f], u₀ = [%f,%f]\n", loss(), θ.data...,u₀...)
-		plot( P.data,U.data,
-			label="inferred", color="darkblue",linewidth=3)
-		plot!( data.parameter,predictor().data,
-			label="inferred", color="darkblue")
-		plot!( data.parameter,data.density,
-			label="target", color="gold",
+		plotBranch( bifurcations, label="inferred")
[…]

### 2019-10-16 main.jl

@@ -2,15 +2,25 @@ include("inference.jl")
 
 # parametrised hypothesis
 function rates( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0)
-	return [ p + θ₁*u[1] + θ₂*u[1]^3 + θ₃, -u[2] ]
+	return [ p + θ₁*u[1] + θ₂*u[1]^3 + θ₃, u[1]-u[2] ]
 end
 
 function jacobian( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0)
-	return [[ θ₁ .+ 3 .*θ₂.*u[1].^2, 0.0 ] [ 0.0, -1.0 ]]
+	return [[ θ₁ .+ 3 .*θ₂.*u[1].^2, 1.0 ] [ 0.0, -1.0 ]]
 end
 
+# solve this and it can solve anytthing else
+# function rates( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0, θ₄=0.0, θ₅=0.0, θ₆=0.0, θ₇=0.0, θ₈=0.0)
+# 	return [ θ₁*p - ( θ₂ + θ₃*u[2] )*u[1],
+# 			 θ₄*(1-u[2])/(0.01+1-u[2]) - (θ₆+θ₇*u[1])*u[2]/(0.01+u[2]) ]
+# end
+#
+# function jacobian( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0, θ₄=0.0, θ₅=0.0, θ₆=0.0, θ₇=0.0, θ₈=0.0)
+# 	return [[ θ₂ + θ₃*u[2] , θ₃*u[1] ] [ -θ₇*u[2]/(u[2]+0.01), -0.01*θ₄/(0.01+1-u[2])^2 - 0.01*(θ₆+θ₇*u[1])/(0.01+u[2])^2 ]]
+# end
+
 # target state density
-parameter = -2:0.05:2
+parameter = -2:0.1:2
 density = ones(length(parameter)).*(abs.(parameter).<0.5)
 
 ######################################################## run inference
@@ -19,7 +29,6 @@ f = (u,p)->rates(u,p,θ...)
 J = (u,p)->jacobian(u,p,θ...)
 
 infer( f, J, θ, StateDensity(parameter,density); iter=200)
-
 
 ######################################################## visualising loss landscape
 using Flux.Tracker: update!
@@ -30,33 +39,46 @@ data = StateDensity(parameter,density)
 maxSteps=1000
[…]

### 2019-10-16 continuation.jl
cleanup

@@ -34,5 +34,3 @@ function continuation( f::Function, J::Function,
 			computeEigenValues = kwargs[:computeEigenValues]);
 			printsolution = printsolution )
 end
-
-	# return u₀, unpack(branch)...

### 2020-03-03 docs/article.tex
scaffold of paper
Updates from Overleaf

--- 2019-10-16 docs/article.tex
@@
 \vspace{-25pt}
-\section{State-space Density} \vspace{-10pt}
+\section{Introduction and Motivation}
+
+\begin{itemize}
+\item qualitative equivalence more important
+\item fitting to time courses not appropriate
+\item gradient information makes optimisation tractable
+\end{itemize}
+
+\section{Problem Statement}
+
+Suppose we would like to know the regions within an unknown parameter space $\theta\in\mathbb{R}^N$ where the steady state of a set of differential equations undergoes specified set of target bifurcations $\mathcal{D}:=\{p_1\dots p_K\}$ along a known parameter $p\in\mathbb{R}$.
+The differential equations for $u\in\mathbb{R}^M$
+\begin{align}
+	\partial_t u=F(u,p|\theta) 
+\end{align}
+and the set of steady states $\mathcal{U}(\theta) := \left\{ (u,p)\in\mathbb{R}^{M+1} : F(u,p|\theta) =0\right\}$ can be found in a local region numerically using
+parameter continuation methods along a co-dimension one curve \cite{}.
+Looking for parameters where  $\mathcal{U}(\theta)$ increases in curvature is an
+indicator of the onset of fold bifurcations. Once fold points have been detected
+one would simply seek to minimise the distance between the predicted bifurcations 
+$\mathcal{P}(\theta):=\left\{\,p\in\mathbb{R}: \frac{\partial p}{\partial u}=0,(u,p)\in\mathcal{U}(\theta) \right\}$ and the targets $\mathcal{D}$.
+\\\\
+
+\section{Method}
+
+\subsection{Pseudo-arclength Continuation}
+\label{sec:continuation}
+
+\begin{itemize}
+    \item predictor-corrector algorithm
+    \end{itemize}
+
+\subsection{Optimization}
+
+\subsubsection{Objective Function}
+\label{sec:objective-function}
+
+The objective function has two regimes: curvature-driven and target-driven. In
+the curvature-driven regime the objective should seek to maximise the curvature
+of $\mathcal{U}(\theta)$ anywhere within the target region containing points $\mathcal{D}$.
+One curvature measure would be Kurtosis along $p$ - the fourth order standardised moment
+\begin{align}
+    \mathrm{Kurt}\left[\,\mathcal{U}(\theta)\,\right]:=
+    \frac{1}{|\,\mathcal{U}(\theta)|}\sum_{(u,p)\in\mathcal{U}(\theta)}
+    \left(\frac{p-\mu}{\sigma}\right)^4\qquad\qquad\quad\\
+    \mathrm{where}\qquad
+    \mu:=\!\!\!\!\!\!\sum_{(u,p)\in\mathcal{U}(\theta)}\!\!\!p/|\,\mathcal{U}(\theta)|
+    \qquad
+    \sigma^2:=\!\!\!\!\!\!\sum_{(u,p)\in\mathcal{U}(\theta)}\!\!\!(p-\mu)^2/|\,\mathcal{U}(\theta)|
+\end{align}
+The target-driven regime is only activated if there are a finite number of fold points $\mathcal{P}(\theta)$; in this case the objective function take a more familiar norm $|\,\mathcal{P}(\theta)-\mathcal{D}\,|^2$  which evaluates the distance between prediction
+and target. The complete objective function is
+\begin{align}
+	\mathcal{J}(\theta):= \begin{cases}
+	|\,\mathcal{P}(\theta)-\mathcal{D}\,|^2 & |\mathcal{P}(\theta)|>0 \\
+	\mathbb{e}^{-\mathrm{Kurt}\left[\,\mathcal{U}(\theta)\,\right]} & \mathrm{otherwise}
+	\end{cases}
+\end{align}
+In order to perform gradient descent on this objective we apply
+\begin{align}
+    \partial_\theta\mathcal{J}=\begin{cases}
+	2(\mathcal{P}(\theta)-\mathcal{D})\,\partial_\theta\mathcal{P} & |\mathcal{P}(\theta)|>0 \\
+	-\mathbb{e}^{-\mathrm{Kurt}\left[\,\mathcal{U}(\theta)\,\right]}\, \partial_\theta\mathcal{U}\,\mathrm{Kurt}'\left[\,\mathcal{U}(\theta)\,\right] & \mathrm{otherwise}
+	\end{cases}
+\end{align}
+This objective function is differentiable --- except for at the boundary between
+the two regimes --- if we can compute $\partial_\theta\mathcal{U}$ and $\partial_\theta\mathcal{P}$. This requires the algorithm in section \ref{sec:continuation} to be differentiable. This is where
+\texttt{Zygote.jl}, a differentiable programming system that is able to take gradients of general program structures \cite{}, is applied.
+
+\todo[inline]{Try adding a penalty for "conjugate stability"}
+\todo[inline]{Comment that while reproducing the qualitative behaviour only can be desirable, sometimes it is also helpful to add quantitative desires over the values of the output.
+E.g.
+we don't care too much about [CFP], but actually, if it were below $C_\text{threshold}$, then we'd have a problem.
+This can be cast in terms of a multi-objective optimization problem, where each term separately gets a weighting factor.
+Personally, I wouldn't demonstrate that in this paper, but important to comment somewhere.}
+\todo[inline]{Try applying the KDE function to the target output, as well as the model output.}
+
+\subsubsection{Optimization details}
+
+\begin{itemize}
+    \item ADAM
+\end{itemize}
+\todo[inline]{Try removing momentum from the optimizer}
+
+\section{Normal Forms}
+\label{sec:normal-forms}
+
+\begin{itemize}
+    \item saddle-node, pitchfork, transcritical
+    \item benchmarks against other algos
+\end{itemize}
+
+\section{Chemical Reaction Networks}
+\label{sec:networks}
+
+\begin{itemize}
+    \item toggle switch
+    \item cell cycle (Attila)
+    \item application to structure $\rightarrow$ function (Luca)
+\end{itemize}
+
+\section{Conclusions and Extensions}
+\label{sec:conclusions}
+
+\begin{itemize}
+    \item hyperparam optimization
+    \item hopf bifurcations
+    \item pattern formation in pdes (Neil)
+\end{itemize}
 

### 2020-03-21 docs/article.tex
fleshing out implicit curvature section

@@
 
-\author{Gregory Szep$^1$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^1$}
+\author{Gregory Szep$^{1,2}$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^1$}
 \affil{$^1$King's College London, $^2$Microsoft Research Cambridge}
@@
 \item gradient information makes optimisation tractable
+\item similar works
 \end{itemize}
 
+\clearpage
 \section{Problem Statement}
+Suppose we parametrise a set of differential equations for states $u\in\mathbb{R}^N$ with a function $F_{\theta}$ in an unknown parameter space $\theta\in\mathbb{R}^M$.
+We would like these differential equations to obey
+a set of target bifurcations $\mathcal{D}:=\{p_1\dots p_K\}$ along a known bifurcation parameter $p\in\mathbb{R}$.
+The following sections outline how a gradient descent algorithm could encourage predicted bifurcations $\mathcal{P}(\theta)$ coming from parameterised differential equations to match specified targets $\mathcal{D}$.
+This would allow us to design differential equations using high-level qualitative constraints and sample qualitatively equivalent models in regions of optimal $\theta$.
+Let the differential equations be defined as
+\begin{align}
+	\partial_t u=F_{\theta}(z)
+	\qquad\mathrm{where}\quad z:=(u,p)\quad
+	F_{\theta} : \mathbb{R}^{N+1}\rightarrow\mathbb{R}^N
+\end{align}
 
-Suppose we would like to know the regions within an unknown parameter space $\theta\in\mathbb{R}^N$ where the steady state of a set of differential equations undergoes specified set of target bifurcations $\mathcal{D}:=\{p_1\dots p_K\}$ along a known parameter $p\in\mathbb{R}$.
-The differential equations for $u\in\mathbb{R}^M$
+\section{Methodology}
+First we view the steady states of this problem in terms of implicit space curves \cite{Goldman2005CurvatureSurfaces} in an augmented state space $z\in \mathbb{R}^{N+1}$ which combines the state $u$ and bifurcation parameter $p$.
+We then realise that it is possible to analytically evaluate the curvature of the determinant $\Delta_\theta(z)$, which when maximised encourages bifurcations in the region of $z$.
+In order to know which regions of $z$ need evaluation, a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams} is used together with some pre-conditioning methods to ensure hyper-parameters remain constant.
+Finally a piecewise-differentiable semi-supervised objective function $L(\theta)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\mathcal{D}$.
+
+For clarity we guide the reader through the methods with the following co-dimension one normal forms: the saddle-node $F_{\theta}(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork $F_{\theta}(z) = \theta_{1} + p u+\theta_{2}u^3$ with corresponding targets $\mathcal{D}$ as shown in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
+These figures show that the curvature of the determinant $\Delta_\theta(z)$ increases in the vicinity of bifurcations and crosses zero at the bifurcation.
+This gives us our first hint of what should be optimised to encourage bifurcations.
+Success on these normal forms demonstrates the proof of concept, and success in practice is demonstrated on more complicated genetic feedback systems taken from examples in synthetic biology.
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=11cm]{docs/figures/saddle-node}
+\caption{Saddle-node normal form $F_{\theta}(z) = p + \theta_{1}u+\theta_{2}u^3$ with set $\theta=(2,-1)$ and targets $\mathcal{D}=\{-1/2,1/2\}$.
+Lighter shades indicate the determinant crossing zero for unstable solutions}
+\label{fig:saddle-node}
+\end{figure}
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=11cm]{docs/figures/pitchfork}
+\caption{Pitchfork normal form $F_{\theta}(z) = \theta_{1} + p u+\theta_{2}u^3$ with set $\theta=(-1/10,-1)$ and target $\mathcal{D}=\{0\}$.
+Lighter shades indicate the determinant crossing zero for unstable solutions}
+\label{fig:pitchfork}
+\end{figure}
+
+\subsection{Co-dimension one bifurcations as implicit space curves}
+Using 
+The steady states define $N$ surfaces in $\mathbb{R}^{N+1}$ which may intersect to form an implicit space curve. This curve given by the following $N$ implicit equations
 \begin{align}
-	\partial_t u=F(u,p|\theta) 
+    F_{\theta}(z) = 0
 \end{align}
-and the set of steady states $\mathcal{U}(\theta) := \left\{ (u,p)\in\mathbb{R}^{M+1} : F(u,p|\theta) =0\right\}$ can be found in a local region numerically using
-parameter continuation methods along a co-dimension one curve \cite{}.
-Looking for parameters where  $\mathcal{U}(\theta)$ increases in curvature is an
-indicator of the onset of fold bifurcations. Once fold points have been detected
-one would simply seek to minimise the distance between the predicted bifurcations 
-$\mathcal{P}(\theta):=\left\{\,p\in\mathbb{R}: \frac{\partial p}{\partial u}=0,(u,p)\in\mathcal{U}(\theta) \right\}$ and the targets $\mathcal{D}$.
-\\\\
+Let us first focus on one of the surfaces defined by $f_{\theta}(z)=0$ where $f_{\theta} : \mathbb{R}^{N+1}\rightarrow\mathbb{R}$.
+Since $f_{\theta}(z)=c$ defines an equipotential for any constant $c$ and the gradient vector $\partial_z f_{\theta}$ points in the direction of steepest ascent for $f_{\theta}$, then the gradient $\partial_z f_{\theta}$ must be perpendicular to all equipotentials $f_{\theta}(z)=c$ and therefore also perpendicular to surface defined by $f_{\theta}(z)=0$.
+Each surface therefore has a normal given by
+$\partial_z f_{\theta}$ for values of $z$ that solve $f_{\theta}(z)=0$.
+These normal vectors must all be perpendicular to the implicit space curve defined by the intersection of these surfaces \cite{Goldman2005CurvatureSurfaces}.
+This geometric constraint is illustrated by example in Figure \ref{fig:implicit-surfaces}.
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=8cm]{docs/figures/implicit-surfaces}
+\caption{Implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form\\ a space curve $T_{\theta}(z)$ that is perpendicular to their gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
+\label{fig:implicit-surfaces}
+\end{figure}
+
+The tangent vector $T_{\theta}(z)$ to the implicit space curve can be constructed perpendicular to all gradient vectors using the properties of the determinant
+\begin{align}
+    T_{\theta}(z):=
+    \left|\begin{matrix}
+        \hat{z} \\
+        \,\partial_{z}F_{\theta}\,
+    \end{matrix}\right|
+    \qquad\mathrm{where}\quad
+    \hat{z}:=(\hat{u},\hat{p})\quad
+	T_{\theta} : \mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N+1}
+\end{align}
+where $\hat{z}$ is a collection of unit basis vectors in the $\mathbb{R}^{N+1}$ space and $\partial_{z}F_{\theta}$ is an $N\times(N+1)$ augmented Jacobian matrix of partial derivatives including $\partial_p F_{\theta}$ in the last column.
+The dot product of this tangent vector any individual gradient vector $\partial_z f_{\theta}$ distributes the components of the gradient across the columns
+\begin{align}
+    T_{\theta}(z)\cdot\partial_z f_{\theta}&=
+    \left|\begin{matrix}
+        \partial_z f_{\theta} \\
+        \,\partial_{z}F_{\theta}\,
+    \end{matrix}\right|\\&=0 \quad\forall f_{\theta}\in F_\theta
+\end{align}
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=14cm]{docs/figures/tangent-field}
+\caption{Left/Right : Calculated tangent fields $T_{\theta}(z)$ for the saddle-node/pitchfork normal forms for some set value of $\theta$, revealing that $F_{\theta}(z)=0$ picks out one of the traces in the field}
+\label{fig:tangent-field}
+\end{figure}
+The determinant of any matrix with two identical rows or columns is zero, satisfying our perpendicularity condition.
+Note that the vector field $T_{\theta}(z)$ is actually defined for all values of $z$, where adjacent field lines trace out other solutions where $F_{\theta}(z)\neq0$ which may be useful for understanding transient behaviour.
+Figure \ref{fig:tangent-field} shows how the solution $F_{\theta}(z)=0$ picks out one of many traces in tangent field $T_{\theta}(z)$ for the saddle and pitchfork.
+Since the tangent field $T_{\theta}(z)$ can always be analytically evaluated can proceed with calculations on the whole space $z$ and pick out a single trace by solving $F_{\theta}(z)=0$ later.
+The tangent fields for the saddle node and pitchfork are $T_{\theta}(z)=(\,3\theta_2 u^2-\theta_1\,)\,\hat{p}+\hat{u}$ and $T_{\theta}(z)=(\,3\theta_2 u^2-p\,)\,\hat{p}+u\hat{u}$ respectively.
+
+\subsection{Evaluating Curvature of the Determinant}
+Bifurcations are defined as values of $p$ where eigenvalues $\lambda$ of the Jacobian $\partial_u F(z)$ cross either the real or imaginary axis in the complex plane.
+Restricting our case for $\lambda\in\mathbb{R}$ for now, the determinant of the Jacobian
+\begin{align}
+    \Delta_\theta(z) := \big|\,\partial_u F_\theta\,\big|
+\end{align}
+crossing zero can be used as a readout of whether a bifurcation has occurred.
+Consequently the curvature of the determinant along the implicit space curve $F_{\theta}(z)=0$ increases in the vicinity of bifurcations as can be seen in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork}.
+The total determinant curvature along the implicit space curve can be calculated as
+\begin{align}
+    \kappa(\theta) := \int_{F_\theta(z)=0}\!|\partial_{T_\theta}^2\Delta_\theta|\,\mathrm{d}z
+\end{align}
+where $F_{\theta}(z)=0$ picks out an implicit space curve given by $\theta$.
+Here we note that the integrand as well as the limits are functions of $\theta$.
+The second directional derivative with respect to tangent field $T_\theta(z)$
+\begin{align}
+    \partial_{T_\theta}^2\Delta_\theta
+    &:=
+    \bigg(
+        \partial_z\left(\partial_z\Delta_\theta\cdot \hat{T}_\theta\right)
+    \bigg)\cdot \hat{T}_\theta\quad
+    \mathrm{where}\quad\hat{T}_\theta:=\frac{T_\theta}{|T_\theta|}\\
+    &=
+        \hat{T}_\theta \cdot \partial_z^2\Delta_\theta \cdot \hat{T}_\theta+
+        \partial_z\Delta_\theta\cdot \partial_z \hat{T}_\theta \cdot \hat{T}_\theta
+\end{align}
+The first term in the expression above is the usual result involving the Hessian matrix $\partial_z^2\Delta_\theta$ when taking second order directional derivatives.
+The second term appears because $\hat{T}_\theta(z)$ is a function of $z$ giving rise to a non-zero Jacobian matrix $\partial_z \hat{T}_\theta$.
+All these quantities are obtained by taking derivatives and determinants of $F_{\theta}(z)$ and in principle can always be done symbolically, giving us a closed analytical expression for the determinant curvature $\partial_{T_\theta}^2\Delta_\theta$.
+For the saddle-node and pitchfork the expression evaluates respectively to
+\begin{align}
+    \partial_{T_\theta}^2\Delta_\theta = 
+    -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+2\right)}
+    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+2\right){}^2}
+    \qquad
+    \partial_{T_\theta}^2\Delta_\theta =
+    -\frac{2 u^2 \left(p+3 \theta _2 u^2\right) \left(9 \theta _2 \left(p+\theta _2 u^2\right)+2\right)}{\left(p^2+6 \theta _2 p u^2+9 \theta _2^2 u^4+2 u^2\right){}^2}
+\end{align}
+\subsection{Semi-supervised Objective Function}
+
+\subsection{Hyper-parameter Preconditioning}
 

### 2020-03-22 docs/article.tex
completed section on determinant curvature

--- 2019-10-16 docs/article.tex
@@
-\documentclass{article}[12pt]
-\renewcommand{\baselinestretch}{1.5}
-\setlength{\parskip}{1em}
+\input{style}
+\begin{document}
+\input{notation}
 
-\usepackage[parfill]{parskip}
-\usepackage[affil-it]{authblk}
-\usepackage[space]{grffile}
+\title{Inference and Synthesis of\\ co-dimension one Bifurcations}
+\author{Gregory Szep$^{1,2}$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^1$}
+\affil{$^1$King's College London, $^2$Microsoft Research Cambridge}
+\date{\today} \maketitle
 
-\usepackage[a4paper]{geometry}
-\geometry{verbose}
-\usepackage{float}
-\usepackage{graphicx}
-\usepackage{setspace}
-\usepackage{caption}
-
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-
-\usepackage{latexsym,textcomp,longtable,tabulary}
-\usepackage{booktabs,array,multirow,braket}
-\usepackage{amsfonts,amsmath,amssymb,mathbbol,calc,cancel}
-\usepackage{subfigure,color,blindtext,enumitem,siunitx}
-\usepackage[colorinlistoftodos]{todonotes}
-
-\usepackage{mathtools}
-\usepackage{url,hyperref,etoolbox}
-\numberwithin{equation}{section}
-\hypersetup{colorlinks=false,pdfborder={0 0 0}}
-
-%+figure layout options
-\restylefloat{figure}
-\setlist{leftmargin=*,before=\setlength{\rightmargin}{\leftmargin}}
-%-figure layout options
-
-\providecommand\citet{\cite}
-\providecommand\citep{\cite}
-\providecommand\citealt{\cite}
-
-\makeatletter
-\makeatother
-
-\begin{document}
-
-\title{
-	Inference and Synthesis of\\
-	co-dimension one Bifurcations
-}
-
-\author{Gregory Szep$^1$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^1$}
-\affil{$^1$King's College London, $^2$Microsoft Research Cambridge}
-\date{\today}
-\maketitle
-\vspace{-25pt}
 \section{Introduction and Motivation}
-
 \begin{itemize}
@@
 \item gradient information makes optimisation tractable
+\item similar works
 \end{itemize}
+\clearpage
 
 \section{Problem Statement}
+Suppose we parameterise a set of differential equations for states $u\in\Reals^N$ with a function $\rates$ in an unknown parameter space $\theta\in\Reals^M$.
+We would like these differential equations to obey a set of target bifurcations $\targets:=\{p_1\dots p_K\}$ along a known bifurcation parameter $p\in\Reals$.
+The following sections outline how a gradient descent algorithm could encourage predicted bifurcations $\predictions(\theta)$ coming from parameterised differential equations to match specified targets $\targets$.
+This would allow us to design differential equations using high-level qualitative constraints and sample qualitatively equivalent models in regions of optimal $\theta$.
+Let the differential equations be defined as
+\begin{align}
+	\partial_t u=\rates(z)
+	\qquad\mathrm{where}\quad z:=(u,p)\quad
+	\rates : \Reals^{N+1}\rightarrow\Reals^N
+	\label{eq:model}
+\end{align}
+\section{Methodology}
+First we view the steady states of this problem in terms of implicit space curves \cite{Goldman2005CurvatureSurfaces} in an augmented state space $z\in\Reals^{N+1}$ which combines the state $u$ and bifurcation parameter $p$.
+We then realise that it is possible to analytically evaluate the curvature of the determinant $\Det(z)$, which when maximised encourages bifurcations in the region of $z$.
+In order to know which regions of $z$ need evaluation, a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams} is used together with some pre-conditioning methods to ensure hyper-parameters remain constant.
+Finally a piecewise-differentiable semi-supervised objective function $\loss(\theta)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.
 
-Suppose we would like to know the regions within paramter space
-$\theta\in\mathbb{R}^N$
-where the steady state of a set of differential equations
-takes on a specified target form $t(p)$ along a known parameter
-$p\in\mathbb{R}$. The target is a label $t\in\{0,1\}$ that indicates
-whether the system is monostable or multi-stable at given $p$. We begin with a known form
-for the differential equations for $u\in\mathbb{R}^M$
+For clarity we guide the reader through the methods with the following co-dimension one normal forms: the saddle-node $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with corresponding targets $\targets$ as shown in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
+These figures show that the curvature of the determinant $\Det(z)$ increases in the vicinity of bifurcations and crosses zero at the bifurcation.
+This gives us our first hint of what should be optimised to encourage bifurcations.
+Success on these normal forms demonstrates the proof of concept, and success in practice is demonstrated on more complicated genetic feedback systems taken from examples in synthetic biology.
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=11cm]{docs/figures/saddle-node}
+\caption{Saddle-node normal form $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with set $\theta=(2,-1)$ and targets $\targets=\{-1/2,1/2\}$.
+Lighter shades indicate the determinant crossing zero for unstable solutions}
+\label{fig:saddle-node}
+\end{figure}
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=11cm]{docs/figures/pitchfork}
+\caption{Pitchfork normal form $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with set $\theta=(-1/10,-1)$ and target $\targets=\{0\}$.
+Lighter shades indicate the determinant crossing zero for unstable solutions}
+\label{fig:pitchfork}
+\end{figure}
+
+\subsection{Bifurcation Curves as Tangent Fields}
+Obtaining a bifurcation curve involves solving for the steady states of \eqref{eq:model} along bifurcation parameter $p$.
+This is equivalent to finding the intersection between $N$ surfaces where each surface is given by a component of the rate function $\rates$.
+Each surface is embedded in $\Reals^{N+1}$ since the bifurcation parameter $p$ is an additional unknown variable.
+Thus the co-dimension one bifurcation curve is an $N+1$ dimensional implicit space curve defined by the $N$ equations
 \begin{align}
-	\partial_t u=F(u|\,p,\theta) 
+    \rates(z) = 0
 \end{align}
-and the set of steady states
-$U(\theta) := \left\{ (u,p)\in\mathbb{R}^{M+1} : F(u|\,p,\theta)=0\right\}$
-can be found in a local region numerically using
-pseudo-arclength continuation along $u(s|\theta),p(s|\theta)$ \cite{} where
-$s$ parametrises the one co-dimensional curve -- details in section
-\ref{sec:continuation}.
-In order to perform an optimisation 
-we need mapping from the states $U(\theta)$ to an output that can be
-compared to $t(p)$. Section \ref{sec:objective-function} motivates this mapping and
-constructs the objective function.
 
-\section{Pseudo-arclength Continuation}
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=6cm]{docs/figures/implicit-surfaces}
+\caption{Two implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form a\\ space curve which is tangent to field $\tangent(z)$ and perpendicular to gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
+\label{fig:implicit-surfaces}
+\end{figure}
+
+An expression for the field $\tangent(z)$ tangent to the bifurcation curve would allow us to take derivatives and integrals along it.
+Fortunately the tangent field can be constructed by ensuring it is perpendicular to the gradient $\partial_z$ of each component of $\rates$ as illustrated by an example two component system in Figure \ref{fig:implicit-surfaces}.
+The tangent field $\tangent(z)$ can be constructed perpendicular to all gradient vectors using the properties of the determinant \cite{Goldman2005CurvatureSurfaces}
+\begin{align}
+    \tangent(z):=
+    \left|\begin{matrix}
+        \hat{z} \\
+        \,\partial_{z}\rates\,
+    \end{matrix}\right|
+    \qquad\mathrm{where}\quad
+    \hat{z}:=(\hat{u},\hat{p})\quad
+	\tangent : \Reals^{N+1}\rightarrow\Reals^{N+1}
+	\label{eq:tangent-field}
+\end{align}
+where $\hat{z}$ is a collection of unit basis vectors in the $\Reals^{N+1}$ space and $\partial_{z}\rates$ is an $N\times(N+1)$ augmented Jacobian matrix of partial derivatives including $\partial_p \rates$ in the last column.
+This construction ensures that the dot product of this field any gradient of a component of $\rates$
+\begin{align}
+    \tangent(z)\cdot\partial_z f_{\theta} =
+    \left|\begin{matrix}
+        \partial_z f_{\theta} \\
+        \,\partial_{z}\rates\,
+    \end{matrix}\right|
+    \quad =0 \quad\forall f_{\theta}\in \rates
+\end{align}
+since the determinant of any matrix with two identical rows or columns is zero.
+Note that the tangent field $\tangent(z)$ is actually defined for all values of $z$ where adjacent field lines trace out other solutions where rates $\rates(z)\neq0$.
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=14cm]{docs/figures/tangent-field}
+\caption{Left/Right : Calculated tangent fields $\tangent(z)$ for the saddle-node/pitchfork normal forms for some set values of $\theta$ revealing that $\rates(z)=0$ picks out one of the traces in the field}
+\label{fig:tangent-field}
+\end{figure}
+
+Figure \ref{fig:tangent-field} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many traces in tangent field $\tangent(z)$ for the saddle and pitchfork.
+The tangent field $\tangent(z)$ can always be analytically evaluated by taking the determinant in \eqref{eq:tangent-field}.
+It becomes practical to proceed with calculations on $\tangent(z)$ in the whole space $z$ and pick out a single trace by solving $\rates(z)=0$ later.
+For the saddle-node and pitchfork respectively
+\begin{align}
+    \tangent(z)=(\,3\theta_2 u^2-\theta_1\,)\,\hat{p}+\hat{u}
+    \qquad\qquad
+    \tangent(z)=(\,3\theta_2 u^2-p\,)\,\hat{p}+u\hat{u}
+    \label{eq:tangent-field-examples}
+\end{align}
+
+\subsection{Evaluating Curvature of the Determinant}
+Bifurcation points are defined as values of $z$ along the bifurcation curve where eigenvalues $\lambda$ of the Jacobian $\partial_u \rates(z)$ cross either the real or imaginary axis in the complex plane.
+Restricting our case to $\lambda\in\Reals$ for now, it would be sufficient to look for values of $z$ where one of the eigenvalues $\lambda$ crosses zero.
+Therefore the determinant of the Jacobian
+\begin{align}
+    \Det(z) := \big|\,\partial_u \rates\,\big|
+\end{align}
+crossing zero can be used as a readout of whether a bifurcation has occurred.
+Figure \ref{fig:determinant-curvature} reveals this is indeed also true for any value $z$ --- not just along $\rates(z)=0$ --- for the saddle-node and pitchfork.
+The tangent field $\tangent(z)$ only folds when $\Det(z)=0$.
+Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-curvature} would give rise to Figures \ref{fig:saddle-node} and \ref{fig:pitchfork}.
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=14cm]{docs/figures/determinant-curvature}
+\caption{Left/Right : Determinant $\Det(z)$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork normal forms for some set values of $\theta$ revealing that $\Det(z)=0$ defines bifurcations}
+\label{fig:determinant-curvature}
+\end{figure}
+
+Consequently the curvature of the determinant $\partial_{\tangent}^2\Det$ along the tangent field $\tangent(z)$ in the vicinity of bifurcations tends to increase.
+Fortunately this curvature can always be obtained in closed analytical form.
+The second directional derivative with respect to tangent field $\tangent(z)$
+\begin{align}
+    \partial_{\tangent}^2\Det &:=
+    \bigg(
+        \partial_z\left(
+            \partial_z\Det \cdot \hat{\tangent}
+        \right)
+    \bigg)\cdot \hat{\tangent}\quad
+    \mathrm{where} \quad \hat{\tangent}:=\frac{\tangent}{|\tangent|}\\
+    &=
+    \hat{\tangent}\cdot\partial_z^2\Det\cdot\hat{\tangent}
+    \quad+\quad
+    \partial_z\Det\cdot\partial_z\hat{\tangent}\cdot\hat{\tangent}
+\end{align}
+The first term in the expression above is the usual result involving the Hessian matrix $\partial_z^2\Det$ when taking second order directional derivatives.
+The second term involving the gradient $\partial_z\Det$ appears because $\hat{\tangent}(z)$ is a function of $z$ giving rise to a non-zero Jacobian matrix $\partial_z \hat{\tangent}$.
+Using expressions for tangent fields \eqref{eq:tangent-field-examples} and determinants for the saddle-node and pitchfork normal forms the determinant curvatures respectively become
+\begin{align}
+    \partial_{\tangent}^2\Det = 
+    -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+2\right)}
+    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+2\right){}^2}
+    \qquad
+    \partial_{\tangent}^2\Det =
+    -\frac{2 u^2 \left(p+3 \theta _2 u^2\right) \left(9 \theta _2 \left(p+\theta _2 u^2\right)+2\right)}{\left(p^2+6 \theta _2 p u^2+9 \theta _2^2 u^4+2 u^2\right){}^2}
+\end{align}
+\subsection{Semi-supervised Objective Function}
+
+This curvature of the determinant $\partial_{\tangent}^2\Det$ can be maximised to encourage bifurcations.
+For a particular bifurcation curve defined by $\rates(z)=0$ the total unsigned determinant curvature
+\begin{align}
+    \kappa(\theta) := \int_{\rates(z)=0}\!
+    \big|\,\partial_{\tangent}^2\Det(z)\,\big|
+    \,\mathrm{d}z
+\end{align}
+Here we note that the integrand as well as the limits are functions of $\theta$ and $z$.
+Once bifurcation points $\predictions(\theta)$ are present, the task is to match their locations to the targets $\targets$.
+However we do not know exactly now many bifurcations there for any given value of $\theta$ and there would be many different ways of satisfying the targets $\targets$.
+Therefore we borrow some concepts from electromagnetism and treat target locations as centres of attractive potentials and locations of predictions $\predictions(\theta)$ as interacting particles with a strong repulsive interaction.
+\begin{align}
+    L(\theta):= |\predictions(\theta)-\targets|
+    -\log|\predictions(\theta)-\predictions(\theta)|
+    -\log(1+\kappa(\theta))
+\end{align}
+
+\subsection{Hyper-parameter Preconditioning}
+
+\section{Method}
+
+\subsection{Pseudo-arclength Continuation}
 \label{sec:continuation}
@@
 
-\section{Objective Function}
-\label{sec:objective-function}
+\subsubsection{Optimization details}
 
-First we note that although $U(\theta)$ contains both $u,p$ we
-are only interested in how many unique $u$ there are for a given $p$.
-This can be done by taking a histogram of equally sized bins along $p$.
-Another way of getting a quantity that is proportional to count is
-by convolution $*$ of $p(s|\theta)$ along $s$ with a symmetric kernel
-$\phi(p\,|\alpha)$ with a bandwidth $\alpha$ indicating bin width.
-This way we get a smooth and differentiable version of the histogram
-operation.
-\begin{align}
-    \phi(p\,|\alpha)*U(\theta):=
-    \int_{\mathbb{R}}\!\phi(p-p(s|\theta)\,|\alpha)\,\mathrm{d}s
-\end{align}
-The result is now proportional to count, but needs to
-be mapped to the unit interval $[0,1]$ in such a way that high
-count regions that indicate multi-stability are mapped to one, and
-low counts which indicate monostability are mapped to zero. For
-this a sigmoidal activation function $\sigma(\frac{x-\mu}{\beta})$ with
-unknown smoothness $\beta$ and threshold $\mu$ will do. We
-are now ready to write down the objective function
-\begin{align}
-	\mathcal{J}(\theta\,|\,\alpha,\beta,\mu):=                            
-	\int_{\mathbb{R}}\left|\;                                                  
-    t(p) - \sigma\left(\frac{\phi(p\,|\alpha)*U(\theta)-\mu}{\beta}\right)
-	\,\right|^2\,\mathrm{d}p
-\end{align}
-where $\theta$ are the parameters to be optimised and $\alpha,\beta,\mu$
-are hyper-parameters. Note that $U(\theta)$ is the only $\theta$ dependence
-in the expression. When we apply
-\begin{align}
-    \partial_\theta\mathcal{J}
-    &=-2\int_{\mathbb{R}}
-    \sigma'\,\partial_\theta\left(
-        \frac{\phi(p\,|\alpha)*U(\theta)-\mu}{\beta}
-        \right)
-    \,\mathrm{d}p\\
-    &=-\frac{2}{\beta}\int_{\mathbb{R}}
-    \sigma'\,
-        \partial_\theta(\,\phi(p|\alpha)*U(\theta)\,)
-    \,\mathrm{d}p\\
-    &=\frac{2}{\beta}
-    \int_{\mathbb{R}}\sigma'\!\! \int_{\mathbb{R}}\phi'\,
-        \partial_\theta p(s|\theta)
-    \,\mathrm{d}s \,\mathrm{d}p
-\end{align}
-remarkably everything is differentiable including the algorithm in section
-\ref{sec:continuation}. This gradient information can be used in a
-gradient descent optimization to minimize $\mathcal{J}(\theta)$.
+\begin{itemize}
+    \item ADAM
+\end{itemize}
 
@@
     \item saddle-node, pitchfork, transcritical
-    \item optimisation landscape with changing hyperparams
     \item benchmarks against other algos
@@
     \item toggle switch
-    \item cell cycle
+    \item cell cycle (Attila)
     \item application to structure $\rightarrow$ function (Luca)
@@
     \item hopf bifurcations
-    \item pattern formation in pdes
+    \item pattern formation in pdes (Neil)
 \end{itemize}

### 2020-03-25 examples/_inference.jl
merging changes

@@ -3,7 +3,10 @@ using FluxContinuation: continuation,deflationContinuation
 using Flux: train!
 
 using Zygote, Plots, Printf
-using StatsBase: kurtosis,norm,mean,std
+using StatsBase: norm,mean,std
+using Plots.PlotMeasures
+using LaTeXStrings
+gr()
 
 struct StateDensity
     parameter::AbstractRange
@@ -11,38 +14,33 @@ struct StateDensity
 end
 
 function infer( f::Function, J::Function, u₀::Vector{T}, θ::AbstractArray, data::StateDensity;
-        optimiser=ADAM(0.05), progress::Function=()->(),
-        iter::Int=100, maxSteps::Int=1000, maxIter::Int=3000, tol=1e-10 ) where T
+        optimiser=ADAM(0.05), progress::Function=()->(), iter::Int=100, maxIter::Int=10, tol=1e-12 ) where T
 
-    parameters = ContinuationPar{Float64, typeof(DefaultLS()), typeof(DefaultEig())}(
-		pMin=minimum(data.parameter),pMax=maximum(data.parameter),ds=step(data.parameter), maxSteps=maxSteps,
+    parameters = getParameters(data; maxIter=maxIter, tol=tol)
+    @time train!(loss, Zygote.Params([θ]), iter, optimiser, cb=progress)
+end
+
+function getParameters(data::StateDensity; maxIter::Int=10, tol=1e-12 )
+	return ContinuationPar{Float64, typeof(DefaultLS()), typeof(DefaultEig())}(
+
+        pMin=minimum(data.parameter),pMax=maximum(data.parameter), maxSteps=10*length(data.parameter),
+        ds=step(data.parameter), dsmax=step(data.parameter), dsmin=step(data.parameter),
 
 			newtonOptions = NewtonPar{Float64, typeof(DefaultLS()), typeof(DefaultEig())}(
 			verbose=false,maxIter=maxIter,tol=tol),
 
 		detect_fold = false, detect_bifurcation = true)
-
-    @time train!(loss, Zygote.Params([θ]), iter, optimiser, cb=progress)
 end
 
-function predictor() global u₀
-	bifurcations,u₀ = deflationContinuation( f,J,u₀, parameters )
+function predictor() global u₀,A
+	bifurcations,u₀ = deflationContinuation( (u,p)->f(A*u,p), (u,p)->A*J(A*u,p), u₀, parameters )
[…]

### 2020-03-25 examples/pitchfork.jl

@@ -10,39 +10,23 @@ function rates_jacobian( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0)
 	return [ [ p .+ 3 .*θ₂.*u[1].^2, 1.0] [0.0,-1.0] ]
 end
 
-# target state density
-parameter = -2:0.01:2
-data = StateDensity(parameter,[0.0])
+# initialise targets, model and hyperparameters
+f,J = (u,p)->rates(u,p,θ...), (u,p)->rates_jacobian(u,p,θ...)
+data = StateDensity(-2:0.01:2,[0.0])
+parameters = getParameters(data)
+
+θ = [0.12,-0.5]
+progress()
+png("pitchfork")
 
 ######################################################## run inference
-parameters = ContinuationPar{Float64, typeof(DefaultLS()), typeof(DefaultEig())}(
-	pMin=minimum(data.parameter),pMax=maximum(data.parameter), maxSteps=100,
-	ds=step(data.parameter), #dsmax=100*step(data.parameter), dsmin=step(data.parameter)/10,a=0.1,
 
-		newtonOptions = NewtonPar{Float64, typeof(DefaultLS()), typeof(DefaultEig())}(
-		verbose=false,maxIter=10,tol= 1e-12),
+u₀,θ,A = [[0.0 0.0], [0.0 0.0], [0.0 0.0]], [1.0,0.5], 1.0
+plot(range(0,2π,length=100), ϕ->lossAt(3cos(ϕ),3sin(ϕ)))
+vline!([2π/3+0.05],label="target", color="gold")
 
-	detect_fold = false, detect_bifurcation = true)
-
-f,J = (u,p)->rates(u,p,θ...), (u,p)->rates_jacobian(u,p,θ...)
-
-θ = [1.0,0.5]
-u₀ = [[0.0 0.0], [0.0 0.0], [0.0 0.0]]
-
-x,y = range(-3,3,length=50),range(-3,3,length=50)
-contourf(x,y, (x,y) -> lossAt(x,y), size=(500,500))
-
-x = range(-3,3,length=500)
-plot(x,x -> lossAt(x,-1.5))
-
-θ = [-0.01,-1.1]
-progress()
-
-collect(x)[Int(end/2)]
-θ = [1.0,0.06]
-progress()
-
-
+# x,y = range(-3,3,length=50),range(-3,3,length=50)
[…]

### 2020-03-25 examples/saddle-node.jl

@@ -10,25 +10,28 @@ function rates_jacobian( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0)
 	return [ θ₁ .+ 3 .*θ₂.*u[1].^2 ][:,:]
 end
 
-# target state density
-parameter = -2:0.05:2
-data = StateDensity(parameter,[0.5,-0.5])
+function curvature( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0)
+	return - 6θ₂ .* ( 1 .+ θ₁^2 .- 9θ₂^2 .*u[1].^4 ) / (1 .+ (θ₁+3θ₂.*u[1].^2).^2 ).^2
+end
+
+# initialise targets, model and hyperparameters
+f,J,K = (u,p)->rates(u,p,θ...), (u,p)->rates_jacobian(u,p,θ...), (u,p)->curvature(u,p,θ...)
+data = StateDensity(-2:0.01:2,[0.5,-0.5])
+parameters = getParameters(data)
 
 ######################################################## run inference
-parameters = ContinuationPar{Float64, typeof(DefaultLS()), typeof(DefaultEig())}(
-	pMin=minimum(data.parameter),pMax=maximum(data.parameter),ds=step(data.parameter), maxSteps=200,
 
-		newtonOptions = NewtonPar{Float64, typeof(DefaultLS()), typeof(DefaultEig())}(
-		verbose=false,maxIter=100,tol= 1e-5),
+θ = [1.9,-1.1]
+progress()
+png("saddle-node")
+u₀,θ,A = [[0.0][:,:], [0.0][:,:] ],[2.1,-2.1], 1.0
+plot(range(0,π,length=100), ϕ->lossAt(3cos(ϕ),3sin(ϕ)))
+vline!([2π/3+0.05],label="target", color="gold")
 
-	detect_fold = false, detect_bifurcation = true)
-
-u₀,θ,trace = [[0.0][:,:], [0.0][:,:] ],[2.1,2.1],[]
-f,J = (u,p)->rates(u,p,θ...), (u,p)->rates_jacobian(u,p,θ...)
-
-x,y = range(-3,3,length=50),range(-7,7,length=50)
-contourf(x,y, (x,y) -> lossAt(x,y), size=(500,500))
-
+θ = [2.1,-2.1]
+progress()
+# x,y = range(-3,3,length=50),range(-7,7,length=50)
[…]

### 2020-04-08 examples/_inference.jl
checking latex commands
Former-commit-id: 5dbea21bb12866eb9d17fbf99e64387f41c32398

@@ -50,6 +50,8 @@ end
 
 function loss()
     steady_states = predictor()
+	return sum( branch->( (p,v)=potential(branch);
+		sum(v)*step(data.parameter) ), steady_states )
 
     if sum( branch -> length(branch.bifpoint), steady_states) == 0
 		return -log(1+sum( branch->( (p,v)=potential(branch);
@@ -62,40 +64,46 @@ function loss()
     end
 end
 
-function progress()
+function progress(states_only=false)
     bifurcations = predictor()
-    vline( data.bifurcations, label="", color=:gold, xlabel=L"\mathrm{parameter},\,p",right_margin=20mm,size=(500,400))
-	right_axis = twinx()
 
-    for bifurcation in bifurcations
-		bifpt = bifurcation.bifpoint[1:bifurcation.n_bifurcations]
+	if states_only
+		for bifurcation in bifurcations
+			bifpt = bifurcation.bifpoint[1:bifurcation.n_bifurcations]
+			plot!(bifurcation.branch[1,:],bifurcation.branch[2,:], alpha=0.5, label="", grid=false, ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(z)=0",
+				color=map(x -> isodd(x) ? :darkblue : :lightblue, bifurcation.stability[1:bifurcation.n_points]), linewidth=2)
+		end
+	else
 
-        plot!(bifurcation.branch[1,:],bifurcation.branch[2,:], alpha=0.5, label="", grid=false, ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(z)=0",
-            color=map(x -> isodd(x) ? :darkblue : :lightblue, bifurcation.stability[1:bifurcation.n_points]), linewidth=2)
+	    vline( data.bifurcations, label="", color=:gold, xlabel=L"\mathrm{parameter},\,p",right_margin=20mm,size=(500,400))
+		right_axis = twinx()
 
-        plot!(right_axis,det(bifurcation)..., alpha=0.5, label="", grid=false, ylabel=L"\mathrm{determinant}\,\,\Delta_{\theta}(z)",
-            colour=map(x -> isodd(x) ? :red : :pink, bifurcation.stability[1:bifurcation.n_points]), linewidth=2)
+	    for bifurcation in bifurcations
[…]

### 2020-04-08 examples/saddle-node.jl

@@ -11,7 +11,7 @@ function rates_jacobian( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0)
 end
 
 function curvature( u,p, θ₁=0.0, θ₂=0.0, θ₃=0.0)
-	return - 6θ₂ .* ( 1 .+ θ₁^2 .- 9θ₂^2 .*u[1].^4 ) / (1 .+ (θ₁+3θ₂.*u[1].^2).^2 ).^2
+	return 6θ₂ .* ( 1 .+ θ₁^2 .- 9θ₂^2 .*u.^4 ) ./ (1 .+ (θ₁ .+ 3θ₂.*u.^2).^2 ).^2
 end
 
 # initialise targets, model and hyperparameters
@@ -21,12 +21,11 @@ parameters = getParameters(data)
 
 ######################################################## run inference
 
-θ = [1.9,-1.1]
-progress()
-png("saddle-node")
 u₀,θ,A = [[0.0][:,:], [0.0][:,:] ],[2.1,-2.1], 1.0
-plot(range(0,π,length=100), ϕ->lossAt(3cos(ϕ),3sin(ϕ)))
-vline!([2π/3+0.05],label="target", color="gold")
+progress()
+
+plot(range(0,2π,length=200), ϕ->lossAt(3cos(ϕ),3sin(ϕ)))
+vline!([14π/8-0.2],label="target", color="gold")
 
 θ = [2.1,-2.1]
 progress()

### 2020-07-20 docs/article.tex
checking latex commands
Former-commit-id: 5dbea21bb12866eb9d17fbf99e64387f41c32398
Updates from Overleaf

@@
 First we view the steady states of this problem in terms of implicit space curves \cite{Goldman2005CurvatureSurfaces} in an augmented state space $z\in\Reals^{N+1}$ which combines the state $u$ and bifurcation parameter $p$.
-We then realise that it is possible to analytically evaluate the curvature of the determinant $\Det(z)$, which when maximised encourages bifurcations in the region of $z$.
-In order to know which regions of $z$ need evaluation, a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams} is used together with some pre-conditioning methods to ensure hyper-parameters remain constant.
-Finally a piecewise-differentiable semi-supervised objective function $\loss(\theta)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.
+We then realise that the curvature of the Jacobian determinant $\Det(z):=|\partial\rates/\partial u|$ is maximised when bifurcations appear in the observed region of $z$.
+Luckily the curvature can be evaluated analytically.
+The steady state regions of $z$ are computed using a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
+Finally a piecewise-differentiable semi-supervised cost function $\loss(\theta)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.
 
-For clarity we guide the reader through the methods with the following co-dimension one normal forms: the saddle-node $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with corresponding targets $\targets$ as shown in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
-These figures show that the curvature of the determinant $\Det(z)$ increases in the vicinity of bifurcations and crosses zero at the bifurcation.
+For clarity we guide the reader through the methods with the following minimal models that explore the space of saddle-nodes $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$.
+These minimal models are shown with targets $\targets$ in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
+These figures show that the curvature of the determinant $\Det(z)$ increases in the vicinity of bifurcations and crosses zero at the bifurcations.
 This gives us our first hint of what should be optimised to encourage bifurcations.
-Success on these normal forms demonstrates the proof of concept, and success in practice is demonstrated on more complicated genetic feedback systems taken from examples in synthetic biology.
+Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on more complicated genetic feedback systems taken from examples in synthetic biology.
 \begin{figure}[H]
@@
 \captionsetup{justification=centering}
-\includegraphics[width=11cm]{docs/figures/saddle-node}
-\caption{Saddle-node normal form $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with set $\theta=(2,-1)$ and targets $\targets=\{-1/2,1/2\}$.
+\includegraphics[width=11cm]{saddle-node}
+\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with set $\theta=(2,-1)$ and targets $\targets=\{-1/2,1/2\}$.
 Lighter shades indicate the determinant crossing zero for unstable solutions}
@@
 \captionsetup{justification=centering}
-\includegraphics[width=11cm]{docs/figures/pitchfork}
-\caption{Pitchfork normal form $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with set $\theta=(-1/10,-1)$ and target $\targets=\{0\}$.
+\includegraphics[width=11cm]{pitchfork}
+\caption{Pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with set $\theta=(-1/10,-1)$ and target $\targets=\{0\}$.
 Lighter shades indicate the determinant crossing zero for unstable solutions}
@@
 \captionsetup{justification=centering}
-\includegraphics[width=6cm]{docs/figures/implicit-surfaces}
+\includegraphics[width=6cm]{implicit-surfaces}
 \caption{Two implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form a\\ space curve which is tangent to field $\tangent(z)$ and perpendicular to gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
@@
 \captionsetup{justification=centering}
-\includegraphics[width=14cm]{docs/figures/tangent-field}
-\caption{Left/Right : Calculated tangent fields $\tangent(z)$ for the saddle-node/pitchfork normal forms for some set values of $\theta$ revealing that $\rates(z)=0$ picks out one of the traces in the field}
+\includegraphics[width=14cm]{tangent-field}
+\caption{Left/Right : Calculated tangent fields $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\rates(z)=0$ picks out one of the traces in the field}
 \label{fig:tangent-field}
@@
 \captionsetup{justification=centering}
-\includegraphics[width=14cm]{docs/figures/determinant-curvature}
-\caption{Left/Right : Determinant $\Det(z)$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork normal forms for some set values of $\theta$ revealing that $\Det(z)=0$ defines bifurcations}
+\includegraphics[width=14cm]{determinant-curvature}
+\caption{Left/Right : Determinant $\Det(z)$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\Det(z)=0$ defines bifurcations}
 \label{fig:determinant-curvature}
@@
 The second term involving the gradient $\partial_z\Det$ appears because $\hat{\tangent}(z)$ is a function of $z$ giving rise to a non-zero Jacobian matrix $\partial_z \hat{\tangent}$.
-Using expressions for tangent fields \eqref{eq:tangent-field-examples} and determinants for the saddle-node and pitchfork normal forms the determinant curvatures respectively become
+Using expressions for tangent fields \eqref{eq:tangent-field-examples} and determinants for the saddle-node and pitchfork models the determinant curvatures respectively become
 \begin{align}
@@
 \end{align}
-\subsection{Semi-supervised Objective Function}
+\subsection{Semi-supervised Cost Function}
 
-This curvature of the determinant $\partial_{\tangent}^2\Det$ can be maximised to encourage bifurcations.
-For a particular bifurcation curve defined by $\rates(z)=0$ the total unsigned determinant curvature
+This curvature of the determinant $\partial_{\tangent}^2\Det$ can be extremised to encourage bifurcations.
+We can treat this curvature as an unsupervised correction of magnitude $\lambda$ to a supervised likelihood $L_{\theta}(z|\targets)$.
+For a particular bifurcation curve defined by $\rates(z)=0$ we can express the cost function as the negative log-likelihood
 \begin{align}
-    \kappa(\theta) := \int_{\rates(z)=0}\!
-    \big|\,\partial_{\tangent}^2\Det(z)\,\big|
-    \,\mathrm{d}z
+    \loss(\theta|\targets) := -\log\left(\int_{\rates(z)=0}\!
+    L_{\theta}(z|\targets) + \lambda\left( \frac{\partial^2\Det}{\partial\tangent^2} \right)^2
+    \,\mathrm{d}z\right)\\
+    L_{\theta}(z|\targets) := \mathbb{e}^{-|\targets-p|^2-\Det(z)^2}\qquad\qquad\quad
 \end{align}
+where we have chosen a Gaussian likelihood to bias optimisation of points in $z$ where the determinant $\Det(z)$ is close to zero towards targets $\targets$.
 Here we note that the integrand as well as the limits are functions of $\theta$ and $z$.
-Once bifurcation points $\predictions(\theta)$ are present, the task is to match their locations to the targets $\targets$.
-However we do not know exactly now many bifurcations there for any given value of $\theta$ and there would be many different ways of satisfying the targets $\targets$.
-Therefore we borrow some concepts from electromagnetism and treat target locations as centres of attractive potentials and locations of predictions $\predictions(\theta)$ as interacting particles with a strong repulsive interaction.
-\begin{align}
-    L(\theta):= |\predictions(\theta)-\targets|
-    -\log|\predictions(\theta)-\predictions(\theta)|
-    -\log(1+\kappa(\theta))
-\end{align}
 
-\subsection{Hyper-parameter Preconditioning}
+\subsection{Adaptive Hyperparameters}
 

### 2020-07-20 docs/article.tex
adding section on region velocity

@@
 \begin{align}
-    \tangent(z)=(\,3\theta_2 u^2-\theta_1\,)\,\hat{p}+\hat{u}
+    \tangent(z)=\hat{u}-(\,3\theta_2 u^2+\theta_1\,)\,\hat{p}
     \qquad\qquad
-    \tangent(z)=(\,3\theta_2 u^2-p\,)\,\hat{p}+u\hat{u}
+    \tangent(z)=u\hat{u}-(\,3\theta_2 u^2+p\,)\,\hat{p}
     \label{eq:tangent-field-examples}
@@
     \partial_{\tangent}^2\Det =
-    -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+2\right)}
-    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+2\right){}^2}
+    -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+1\right)}
+    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+1\right){}^2}
     \qquad
     \partial_{\tangent}^2\Det =
-    -\frac{2 u^2 \left(p+3 \theta _2 u^2\right) \left(9 \theta _2 \left(p+\theta _2 u^2\right)+2\right)}{\left(p^2+6 \theta _2 p u^2+9 \theta _2^2 u^4+2 u^2\right){}^2}
+    -\frac{2 u^2 \left(p+3 \theta _2 u^2\right) \left(9 \theta _2 \left(p+\theta _2 u^2\right)+1\right)}{
+    ( \left(p+3 \theta _2 u^2\right)^2+u^2 )^2
+    }
 \end{align}
@@
     \loss(\theta|\targets) := -\log\left(\int_{\rates(z)=0}\!
-    L_{\theta}(z|\targets) + \lambda\left( \frac{\partial^2\Det}{\partial\tangent^2} \right)^2
+    L_{\theta}(z|\targets) + \lambda\,K_{\theta}(z)^2\,
     \,\mathrm{d}z\right)\\
-    L_{\theta}(z|\targets) := \mathbb{e}^{-|\targets-p|^2-\Det(z)^2}\qquad\qquad\quad
+    L_{\theta}(z|\targets) := \mathbb{e}^{-|\targets-p|^2/\varepsilon-\Det(z)^2}\qquad
+     K_{\theta}(z):=\frac{\partial^2\Det}{\partial\tangent^2}
 \end{align}
 where we have chosen a Gaussian likelihood to bias optimisation of points in $z$ where the determinant $\Det(z)$ is close to zero towards targets $\targets$.
-Here we note that the integrand as well as the limits are functions of $\theta$ and $z$.
+Figure \ref{fig:likelihood} shows how the likelihood $L_{\theta}(z|\targets)$ is sharply peaked around the bifurcation points $\predictions(\theta)$ and is maximal when they match the targets.
+The parameter $\varepsilon$ controls the desired precision of the likelihood function and $\lambda$ controls the strength of the unsupervised bias.
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=7cm]{likelihood-1.pdf}
+\includegraphics[width=7cm]{likelihood-2.pdf}
+\caption{Likelihood $L_{\theta}(z|\targets)$ shown in dark gold is sharply peaked around the bifurcation points $\predictions(\theta)$ and is maximal when they match the targets (Right).
+The light gold envelope shows the maxima of the likelihood as the points $\predictions(\theta)$ sweep across different values of $p$}
+\label{fig:likelihood}
+\end{figure}
+Applying the gradient operator with respect to $\theta$ to the cost function, we find that not only the integrand but also the integration region needs to be differentiated.
+This requires the general Leibniz rule \cite{Flanders1973DifferentiationSign} which yields
+\begin{align}
+\frac{\partial\loss(\theta|\targets)}{\partial\theta}=
+-\frac{
+\int_{F=0}
+  \mathrm{div}\!
+\left(\frac{\partial z}{\partial\theta}
+  \left(L+\lambda K^2\right)
+\right)
++\frac{\partial L}{\partial\theta}+
+2\lambda K\frac{\partial K}{\partial\theta}
+\,\,\mathrm{d}z
+}{
+\int_{F=0}
+\left(
+  L+\lambda K^2
+\right)\,\mathrm{d}z
+}\label{eq:gradient}
+\end{align}
+where $\frac{\partial z}{\partial\theta}$ is the velocity field of the integration region due to changes in $\theta$.
+Fortunately the velocity field of implicitly defined regions can be solved for analytically \cite{Jos2011OnSurface}.
+This is done by taking the total derivative of the implicit equation
+\begin{align}
+    \frac{d\rates(z)}{d\theta}=\frac{\partial F}{\partial\theta}+
+    \frac{\partial F}{\partial z}\frac{\partial z}{\partial\theta}
+\end{align}
+We can rearrange for $\frac{\partial z}{\partial\theta}$ using the Moore-Penrose inverse of the augmented Jacobian matrix $\frac{\partial F}{\partial z}$ which appeared in equation \eqref{eq:tangent-field}.
+Since the surface is defined by $\rates(z)=0$ we arrive at
+\begin{align}
+    \frac{\partial z}{\partial\theta} = - \frac{\partial F}{\partial z}^\top
+    \left(\,
+        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
+    \right)^{-1}
+    \frac{\partial F}{\partial\theta}
+\end{align}
+which for our two minimal models yield
+\begin{align}
+    \frac{\partial z}{\partial\theta} =
+        \frac{ \hat{p}+\hat{u}\,(\,\theta_1+3\theta_2 u^2\,) }{1+(\,\theta_1+3\theta_2 u^2\,)^2}
+        \left( u\,\hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)
+    \qquad
+    \frac{\partial z}{\partial\theta} =
+        \frac{ p\,\hat{u} + u\,(\,\hat{p}+3\theta_2u\,\hat{u}\,) }{(\,p+3\theta_2 u^2\,)^2 + u^2 }
+        \left( \hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)
+    \label{eq:velocity-field-examples}
+\end{align}
+Thus the integrands in the gradient \eqref{eq:gradient} are completely analytic, leaving only region $\rates(z)=0$ to be evaluated numerically.
+Since these analytic forms can be pre-calculated, calculating the gradients shares the same computational complexity as calculating the steady states.
+\subsection{Basis Functions and Unsupervided Bias}
 

### 2020-07-27 docs/article.tex
Updates from Overleaf

@@
 
-\title{Inference and Synthesis of\\ co-dimension one Bifurcations}
-\author{Gregory Szep$^{1,2}$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^1$}
-\affil{$^1$King's College London, $^2$Microsoft Research Cambridge}
+\title{Inference of Bifurcations with Differentiable Continuation}
+\author{Gregory Szep$^{1,2}$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^{1,3}$}
+\affil{$^1$King's College London, UK, $^2$Microsoft Research Cambridge, UK, $^3$Pazmany Peter Catholic University, Budapest, Hungary}
 \date{\today} \maketitle
 
+\begin{abstract}
+    In this work we propose a gradient-based semi-supervised approach for matching target bifurcations with parameterised differential equation models.
+The cost function contains a supervised likelihood term that is maximal when predicted bifurcations match the targets, and an unsupervised correction term that encourages bifurcations by maximising the curvature of the determinant of the Jacobian.
+The calculation of gradients with respect to parameters shares the same computational complexity as deflated pseudo-arclength continuation used to calculate the bifurcation diagram.
+We demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams, and an applied model from synthetic biology.
+This work forms the basis for the inference of more complicated bifurcation diagrams which may include Andronov-Hopf points or extended to partial differential equations.
+\end{abstract}
+
 \section{Introduction and Motivation}
+\subsection{Background}
+Backpropagation through differential equation solves has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jl-AEquations} that enabled scalable parameter inference for differential equations.
+Optimisation targets, however, have mostly been expressed in the spatio-temporal domain.
+Acquisition of such data can be costly and can often contain over-constraining information generated by processing steps or the measurement device rather than the state variables that underpin the observed mechanism.
+In microscopy, for example, data is often reported in arbitrary fluoresce units allowing the observer to shift and scale data arbitrarily.
+Furthermore, such data may also not contain sufficient information about dynamical transients in order to identify kinetic parameters.
+The emerging picture suggests that identification of the qualitative dynamics -- the bifurcation diagram -- should precede any attempt at inferring kinetic parameters \cite{Stumpf2019ParameterBifurcations}.
+
+In this work we propose a gradient-based semi-supervised approach that focuses on fitting qualitative dynamics, defined by state space structures, rather than kinetics.
+We find that calculating the gradients, which inform the optimisation of parameters, shares the same computational complexity as the algorithm used to calculate the bifurcation diagram.
+We use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl}.
+For partial differential equations, the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
+The complexity of computing a single branch, however, is bounded by the complexities of the chosen eigenvalue solver and corrector, and further decreased by adaptive stepping procedures \cite{Aruliah2016AlgorithmContinuation}.
+
+We find that the cost function landscape contains basins that not only allow us to synthesise models with a desired bifurcation structure but also allow us to organise models in terms of topological and geometric equivalence.
+We discuss the relevance of this in model selection.
+
+\subsection{Similar Works}
+
 \begin{itemize}
-\item qualitative equivalence more important
-\item fitting to time courses not appropriate
-\item gradient information makes optimisation tractable
-\item similar works
+    \item some juliacon submissions?
 \end{itemize}
@@
 
-\section{Problem Statement}
+\section{Methodology}
+
+\subsection{Problem Statement}
 Suppose we parameterise a set of differential equations for states $u\in\Reals^N$ with a function $\rates$ in an unknown parameter space $\theta\in\Reals^M$.
@@
 \end{align}
-\section{Methodology}
+
 First we view the steady states of this problem in terms of implicit space curves \cite{Goldman2005CurvatureSurfaces} in an augmented state space $z\in\Reals^{N+1}$ which combines the state $u$ and bifurcation parameter $p$.
 We then realise that the curvature of the Jacobian determinant $\Det(z):=|\partial\rates/\partial u|$ is maximised when bifurcations appear in the observed region of $z$.
-Luckily the curvature can be evaluated analytically.
+Luckily the curvature can be evaluated analytically \cite{Goldman2005CurvatureSurfaces}.
 The steady state regions of $z$ are computed using a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
-Finally a piecewise-differentiable semi-supervised cost function $\loss(\theta)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.
+Finally a differentiable semi-supervised cost function $\loss(\theta|\targets)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.
 
-For clarity we guide the reader through the methods with the following minimal models that explore the space of saddle-nodes $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$.
+For clarity, we guide the reader through the methods with the following minimal models that explore the space of saddle-nodes $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$.
 These minimal models are shown with targets $\targets$ in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
@@
 This gives us our first hint of what should be optimised to encourage bifurcations.
-Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on more complicated genetic feedback systems taken from examples in synthetic biology.
+Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on a four component genetic toggle switch model taken from synthetic biology.
 \begin{figure}[H]
@@
 \captionsetup{justification=centering}
-\includegraphics[width=14cm]{tangent-field}
-\caption{Left/Right : Calculated tangent fields $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\rates(z)=0$ picks out one of the traces in the field}
-\label{fig:tangent-field}
+\includegraphics[width=14cm]{determinant-curvature}
+\caption{Left/Right : Determinant $\Det(z)$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\Det(z)=0$ defines bifurcations}
+\label{fig:determinant-curvature}
 \end{figure}
 
-Figure \ref{fig:tangent-field} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many traces in tangent field $\tangent(z)$ for the saddle and pitchfork.
+Figure \ref{fig:determinant-curvature} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many traces in tangent field $\tangent(z)$ for the saddle and pitchfork.
 The tangent field $\tangent(z)$ can always be analytically evaluated by taking the determinant in \eqref{eq:tangent-field}.
-It becomes practical to proceed with calculations on $\tangent(z)$ in the whole space $z$ and pick out a single trace by solving $\rates(z)=0$ later.
-For the saddle-node and pitchfork respectively
+We will proceed with calculations on $\tangent(z)$ in the whole space $z$ and pick out a single trace by solving $\rates(z)=0$ later.
+For our two models
 \begin{align}
-    \tangent(z)=\hat{u}-(\,3\theta_2 u^2+\theta_1\,)\,\hat{p}
+    \underset{\mathrm{saddle-node\,\,model}}{
+    \tangent(z)=\hat{u}-(\,3\theta_2 u^2+\theta_1\,)\,\hat{p}}
     \qquad\qquad
-    \tangent(z)=u\hat{u}-(\,3\theta_2 u^2+p\,)\,\hat{p}
+    \underset{\mathrm{pitchfork\,\,model}}{
+    \tangent(z)=u\hat{u}-(\,3\theta_2 u^2+p\,)\,\hat{p}}
     \label{eq:tangent-field-examples}
@@
 Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-curvature} would give rise to Figures \ref{fig:saddle-node} and \ref{fig:pitchfork}.
-
-\begin{figure}[H]
-\centering{}
-\captionsetup{justification=centering}
-\includegraphics[width=14cm]{determinant-curvature}
-\caption{Left/Right : Determinant $\Det(z)$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\Det(z)=0$ defines bifurcations}
-\label{fig:determinant-curvature}
-\end{figure}
 
@@
 \begin{align}
+    \underset{\mathrm{saddle-node\,\,model}}{
     \partial_{\tangent}^2\Det =
     -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+1\right)}
-    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+1\right){}^2}
+    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+1\right){}^2}}
     \qquad
+    \underset{\mathrm{pitchfork\,\,model}}{
     \partial_{\tangent}^2\Det =
@@
     ( \left(p+3 \theta _2 u^2\right)^2+u^2 )^2
-    }
+    }}
 \end{align}
@@
 \begin{align}
+    \underset{\mathrm{saddle-node\,\,model}}{
     \frac{\partial z}{\partial\theta} =
         \frac{ \hat{p}+\hat{u}\,(\,\theta_1+3\theta_2 u^2\,) }{1+(\,\theta_1+3\theta_2 u^2\,)^2}
-        \left( u\,\hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)
+        \left( u\,\hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)}
     \qquad
+    \underset{\mathrm{pitchfork\,\,model}}{
     \frac{\partial z}{\partial\theta} =
         \frac{ p\,\hat{u} + u\,(\,\hat{p}+3\theta_2u\,\hat{u}\,) }{(\,p+3\theta_2 u^2\,)^2 + u^2 }
-        \left( \hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)
+        \left( \hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)}
     \label{eq:velocity-field-examples}
@@
 Since these analytic forms can be pre-calculated, calculating the gradients shares the same computational complexity as calculating the steady states.
-\subsection{Basis Functions and Unsupervided Bias}
+% \subsection{Basis Functions and Unsupervided Bias}
+% \subsection{Pseudo-arclength Continuation}
 
-\subsection{Adaptive Hyperparameters}
+\subsection{Optimisation Loop}
+We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$, and an initial guess $u_0$ at an initial parameter $p_0$.
+Let this algorithm be called by function $\mathrm{getSteadyStates}$ and return the steady states that form the integration region $\Omega$ and a new guess $u_0$ that is the true steady state at $p_0$.
 
-\section{Method}
+The algorithm also takes hyperparameters $\beta$ which will contain arclength step sizes, eigenvalue and Newton solver options and unsupervised correction factor $\lambda$.
+The adaptive hyperparameter update is done by $\mathrm{updateHyperparameters}$ and should depend on the current solutions $\Omega$ and the targets $\targets$.
+The step sizes and solver options should adapt to the expected space of the solutions $\Omega$ so that the waiting time between iterations is minimised.
+The unsupervised correction factor $\lambda$ should be large when there are no predicted bifurcations and decrease as the number of predicted bifurcations approaches the number of targets.
 
-\subsection{Pseudo-arclength Continuation}
-\label{sec:continuation}
+The evaluation of the cost gradient \eqref{eq:gradient}
 
-\begin{itemize}
-    \item predictor-corrector algorithm
-    \end{itemize}
+\\
+\begin{algorithm*}[H]
+\label{alg:optimisation-loop}
+\SetAlgoLined
+\textbf{Inputs} Function $F_{\theta}:\mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N}$ for $u\in\mathbb{R}^N$ and $p\in\mathbb{R}$\\ target bifurcations $\mathcal{D}=\{p:p\in[p_{\min},p_{\max}]\}$ and convergence tolerance $\varepsilon$\\
+\textbf{Output} Optimised $\theta\in\mathbb{R}^M$ that satisfy targets $\targets$\\
+$J_{\theta}\leftarrow\partial F/\partial u$\qquad\,
+$\Delta_{\theta}\leftarrow|\mathrm{det}\,J_{\theta}|$\qquad$K_{\theta}\leftarrow\partial^2\Delta_{\theta}/\partial\tangent^2$\\
+$\frac{\partial z}{\partial\theta}\leftarrow- \frac{\partial F}{\partial z}^\top
+    \left(\,
+        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
+    \right)^{-1}
+    \frac{\partial F}{\partial\theta}$\\
+$u_0\leftarrow\mathrm{rand}\in\mathbb{R}^{N}$\quad\,$\theta\leftarrow\mathrm{rand}\in\mathbb{R}^{M}$\\
+$\beta\leftarrow\mathrm{getHyperparameters}(\targets)$\\
+\While{$\mathrm{tolerance}(\beta)>\varepsilon$}{
+  $\Omega,u_0\leftarrow\mathrm{getSteadyStates}(F_{\theta},J_{\theta},u_0,\beta)$\\
+  $\beta\leftarrow\mathrm{updateHyperparameters}(\beta,\Omega,\targets)$\\
+  $\partial L\leftarrow\mathrm{costGradient}(\Delta_{\theta},K_{\theta},V_{\theta},\Omega,\targets)$\\
+  $\theta\leftarrow\mathrm{updateParameters}(\theta,\partial L,\beta)$
+}
+\caption{Bifurcation Optimisation Loop}
+\end{algorithm*}
 
-\subsubsection{Optimization details}
+\section{Results}
 
-\begin{itemize}
-    \item ADAM
-\end{itemize}
+\subsection{Model Synthesis}
 
-\section{Normal Forms}
-\label{sec:normal-forms}
-
-\begin{itemize}
-    \item saddle-node, pitchfork, transcritical
-    \item benchmarks against other algos
-\end{itemize}
-
-\section{Chemical Reaction Networks}
-\label{sec:networks}
-
-\begin{itemize}
-    \item toggle switch
-    \item cell cycle (Attila)
-    \item application to structure $\rightarrow$ function (Luca)
-\end{itemize}
+\subsection{Model Selection}
 
@@
 \begin{itemize}
-    \item hyperparam optimization
-    \item hopf bifurcations
-    \item pattern formation in pdes (Neil)
+    \item cell cycle / hopf
+    \item application to structure $\rightarrow$ function (Luca)
+    \item pattern formation in pdes
 \end{itemize}

### 2020-07-27 docs/article.tex
add draft watermark

@@
 
-\subsection{Similar Works}
+\subsection{Similar and Related Works}
 
 \begin{itemize}
-    \item some juliacon submissions?
+    \item smooth and match estimator methods \cite{Ranciati2017BayesianParameters}
+    \item bifurcation inference using Mixed Integer Nonlinear Programming optimization strategy \cite{Otero-Muras2018Optimization-basedModels}
 \end{itemize}
@@
 \subsection{Optimisation Loop}
+Here provide a high-level view of Algorithm \ref{alg:optimisation-loop}.
 We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$, and an initial guess $u_0$ at an initial parameter $p_0$.
@@
 
-The evaluation of the cost gradient \eqref{eq:gradient}
-
-\\
+The evaluation of the cost gradient \eqref{eq:gradient} is done by $\mathrm{costGradient}$ using \texttt{Zygote.jl} \cite{Innes2018DontPrograms} and requires the following analytical forms: the determinant of the Jacobian $\Det$, its curvature $K_{\theta}$ and the integration region velocity $\frac{\partial z}{\partial\theta}$.
+It also requires the integration region $\Omega$ and the locations of targets $\targets$.
+This gradient is used by $\mathrm{updateParameters}$, which can be any gradient-based optimiser such as ADAM or Momentum gradient decent, to update the parameters $\theta$.
+\\\\
 \begin{algorithm*}[H]
@@
   $\beta\leftarrow\mathrm{updateHyperparameters}(\beta,\Omega,\targets)$\\
-  $\partial L\leftarrow\mathrm{costGradient}(\Delta_{\theta},K_{\theta},V_{\theta},\Omega,\targets)$\\
+  $\partial L\leftarrow\mathrm{costGradient}(\Delta_{\theta},K_{\theta},\frac{\partial z}{\partial\theta},\Omega,\targets)$\\
   $\theta\leftarrow\mathrm{updateParameters}(\theta,\partial L,\beta)$

### 2020-10-06 src/FluxContinuation.jl
Update FluxContinuation.jl

@@ -1,7 +1,7 @@
 module FluxContinuation
 
-	using PseudoArcLengthContinuation
-	using PseudoArcLengthContinuation: AbstractLinearSolver, AbstractBorderedLinearSolver, AbstractEigenSolver, _axpy, detectBifucation, computeEigenvalues!
+	using BifurcationKit
+	using BifurcationKit: AbstractLinearSolver, AbstractBorderedLinearSolver, AbstractEigenSolver, _axpy, detectBifucation, computeEigenvalues!
 	using Zygote: Buffer, gradient, forward_jacobian, @nograd, @adjoint, @adjoint!
 
 	using Setfield: @set,set,setproperties,Lens,@lens
@@ -101,7 +101,7 @@ module FluxContinuation
 	                # main continuation method
 					branch = BranchBuffer(T)
 					params = set(params, paramlens, pDeflations[i]+hyperparameters.ds)
-					iterator = PALCIterable( f, J, u, params, paramlens, hyperparameters, linsolver, verbosity=verbosity)
+					iterator = ContIterable( f, J, u, params, paramlens, hyperparameters, linsolver, verbosity=verbosity)
 
 					for state in iterator
 

### 2020-10-06 src/FluxContinuation.jl
updated to julia 1.5

@@ -6,11 +6,10 @@ module FluxContinuation
 
 	using Setfield: @set,set,setproperties,Lens,@lens
 	using Parameters: @with_kw, @unpack
-	using Dates: now
 
 	using LinearAlgebra: dot,eigen,svd
 	using StatsBase: norm,mean,std
-	using CuArrays
+	using CUDA
 
 	using Plots.PlotMeasures
 	using LaTeXStrings
@@ -22,8 +21,6 @@ module FluxContinuation
 
 	export StateDensity,Branch,deflationContinuation,findRoots
 	export getParameters,updateParameters,loss,cu
-
-	@nograd now,string
 
 	""" root finding with newton deflation method"""
 	function findRoots( f::Function, J::Function, u₀::Vector{Array{T,2}}, params::NamedTuple, paramlens::Lens,
@@ -101,7 +98,7 @@ module FluxContinuation
 	                # main continuation method
 					branch = BranchBuffer(T)
 					params = set(params, paramlens, pDeflations[i]+hyperparameters.ds)
-					iterator = ContIterable( f, J, u, params, paramlens, hyperparameters, linsolver, verbosity=verbosity)
+					iterator = PALCIterable( f, J, u, params, paramlens, hyperparameters, linsolver, verbosity=verbosity)
 
 					for state in iterator
 

### 2020-10-06 src/Utils.jl

@@ -24,7 +24,7 @@ function updateParameters(parameters::ContinuationPar{T, S, E}, steady_states::V
 end
 @nograd updateParameters
 
-import CuArrays: cu
+import CUDA: cu
 function cu( steady_states::Vector{Branch{T}}; nSamples=50 ) where {T<:Number}
 
 	p  = vcat(map( branch -> branch.parameter,      steady_states)...)

### 2020-10-06 test/normal-forms/normal-forms.jl

@@ -58,12 +58,12 @@ end
 @testset "Normal Forms" begin
 
 	@testset "Saddle Node" begin include("saddle-node.jl")
-		@test @time forward("normal-forms/saddle-node.forward.pdf")
-		@test @time backward("normal-forms/saddle-node.backward.pdf")
+		@test @time forward( joinpath(@__DIR__,"saddle-node.forward.pdf" ))
+		@test @time backward(joinpath(@__DIR__,"saddle-node.backward.pdf"))
 	end
 
 	@testset "Pitchfork" begin include("pitchfork.jl")
-		@test @time forward("normal-forms/pitchfork.forward.pdf")
-	    @test @time backward("normal-forms/pitchfork.backward.pdf")
+		@test @time forward( joinpath(@__DIR__,"pitchfork.forward.pdf" ))
+		@test @time backward(joinpath(@__DIR__,"pitchfork.backward.pdf"))
 	end
 end

### 2020-10-06 test/runtests.jl

@@ -1,4 +1,4 @@
-using Flux,FluxContinuation,CuArrays
+using Flux,FluxContinuation,CUDA
 using StatsBase,LinearAlgebra
 
 using Plots.PlotMeasures

### 2020-10-08 src/FluxContinuation.jl
changed autodiff to forward mode. tests passed

@@ -1,45 +1,46 @@
 module FluxContinuation
 
-	using BifurcationKit
-	using BifurcationKit: AbstractLinearSolver, AbstractBorderedLinearSolver, AbstractEigenSolver, _axpy, detectBifucation, computeEigenvalues!
-	using Zygote: Buffer, gradient, forward_jacobian, @nograd, @adjoint, @adjoint!
+	using BifurcationKit: PALCIterable, newton, ContinuationPar, NewtonPar, DeflationOperator
+	using BifurcationKit: getx, getp, detectBifucation, computeEigenvalues!
 
-	using Setfield: @set,set,setproperties,Lens,@lens
-	using Parameters: @with_kw, @unpack
+	using LinearAlgebra: det,dot,eigen,norm,tr
+	using ForwardDiff: jacobian,gradient,hessian
+	using Flux: Momentum,update!
 
-	using LinearAlgebra: dot,eigen,svd
-	using StatsBase: norm,mean,std
-	using CUDA
+	using Setfield: @lens,@set,set,setproperties,Lens
+	using Parameters: @unpack,@with_kw
+	using InvertedIndices: Not
 
 	using Plots.PlotMeasures
 	using LaTeXStrings
-	using Plots
-	gr()
+	using Plots	
 
 	include("Structures.jl")
+	include("Gradients.jl")
 	include("Utils.jl")
 
-	export StateDensity,Branch,deflationContinuation,findRoots
-	export getParameters,updateParameters,loss,cu
+	export StateDensity,deflationContinuation,train!
+	export getParameters,loss,∇loss
+	export plot,@unpack,@lens,set
 
 	""" root finding with newton deflation method"""
-	function findRoots( f::Function, J::Function, u₀::Vector{Array{T,2}}, params::NamedTuple, paramlens::Lens,
+	function findRoots!( f::Function, J::Function, roots::Vector{Vector{Vector{T}}}, params::NamedTuple, paramlens::Lens,
 		hyperparameters::ContinuationPar{T, S, E}, maxRoots::Int = 3, maxIter::Int=500; verbosity=0
 		) where {T<:Number, S<:AbstractLinearSolver, E<:AbstractEigenSolver}
 
-		_,nStates = size(first(u₀))
-		rootsArray = similar(u₀)
+		nStates = length(first(first(roots)))
+		rootsArray = similar(roots)
 
-		pDeflations = range(hyperparameters.pMin,hyperparameters.pMax,length=length(u₀))
[…]

### 2020-10-08 src/Gradients.jl

@@ -1,0 +1,112 @@
+########################################################################### jacobians
+
+# statespace jacobian
+function ∂Fu(F::Function,z::AbstractVector{T},θ::AbstractVector{U}) where {T<:Number,U<:Number}
+	return jacobian( u -> F( u,           (θ=θ,p=z[end]) ), z[Not(end)] )
+end
+
+# parameter jacobian
+function ∂Fθ(F::Function,z::AbstractVector{T},θ::AbstractVector{U}) where {T<:Number,U<:Number}
+	return jacobian( θ -> F( z[Not(end)], (θ=θ,p=z[end]) ), θ           )
+end
+
+# augmented jacobian
+function ∂Fz(F::Function,z::AbstractVector{T},θ::AbstractVector{U}) where {T<:Number,U<:Number}
+	return jacobian( z -> F( z[Not(end)], (θ=θ,p=z[end]) ), z           )
+end
+
+########################################################################### tangent field
+# field tangent to bifurcation curve of F(z,θ)
+function tangent_field(F::Function,z::AbstractVector{T},θ::AbstractVector{U}) where {T<:Number,U<:Number}
+
+	∂F = ∂Fz(F,z,θ) # augmented jacobian
+	f = first(z)*first(θ)
+	field = similar(z,typeof(f))
+
+	for i ∈ 1:length(field)
+
+		# construct tangent field T(z) := det[ ̂z , ∂Fz ]
+		field[i] = det(∂F[:,Not(i)])
+	end
+
+	# unit tangent field
+	return field / norm(field)
+end
+
+################################################################################# determinant curvature K(z|θ)
+function determinant(F::Function,z::AbstractVector{T},θ::AbstractVector{U}) where {T<:Number,U<:Number}
+	return det(∂Fu(F,z,θ))
+end
+
+# determinant gradient field
+function determinant_field(F::Function,z::AbstractVector{T},θ::AbstractVector{U}) where {T<:Number,U<:Number}
+	return gradient( z -> determinant(F,z,θ), z )
+end
+
+# determinant curvature wrt tangent field
+function curvature(F::Function,z::AbstractVector{T},θ::AbstractVector{U}) where {T<:Number,U<:Number}
+	tangent = tangent_field(F,z,θ)
[…]

### 2020-10-08 src/Utils.jl

@@ -1,4 +1,6 @@
-############################################# hyperparameter updates
+using BifurcationKit: AbstractLinearSolver, AbstractBorderedLinearSolver, AbstractEigenSolver, _axpy
+
+############################################################################ hyperparameter updates
 function getParameters(data::StateDensity{T}; maxIter::Int=100, tol=1e-5) where {T<:Number}
     return ContinuationPar{T,LinearSolver,EigenSolver}(
 
@@ -10,63 +12,17 @@ function getParameters(data::StateDensity{T}; maxIter::Int=100, tol=1e-5) where 
 
         detectFold = false, detectBifurcation = true)
 end
-@nograd getParameters
 
-function updateParameters(parameters::ContinuationPar{T, S, E}, steady_states::Vector{Branch{T}};
+function updateParameters!(parameters::ContinuationPar{T, S, E}, steady_states::Vector{Branch{T}};
     resolution=400 ) where {T<:Number, S<:AbstractLinearSolver, E<:AbstractEigenSolver}
 
     # estimate scale from steady state curves
     branch_points = map(length,steady_states)
     ds = maximum(branch_points)*parameters.ds/resolution
     parameters = setproperties(parameters;ds=ds,dsmin=ds,dsmax=ds)
-
-    return parameters
-end
-@nograd updateParameters
-
-import CUDA: cu
-function cu( steady_states::Vector{Branch{T}}; nSamples=50 ) where {T<:Number}
-
-	p  = vcat(map( branch -> branch.parameter,      steady_states)...)
-	u  = hcat(map( branch -> hcat(branch.state...), steady_states)...)
-	ds = vcat(map( branch -> abs.(branch.ds),       steady_states)...)
-
-	# sample parameter region near solutions
-	pMax,pMin = maximum(p),minimum(p)
-	p = p .+ mean(ds)*(-nSamples:nSamples)'
-	nPoints,nSamples = size(p)
-
-	p = reshape(p,nPoints*nSamples)
-	u = repeat(u,1,nSamples)
-
-	# restrict final grid to original parameter region
-	region  = (pMin.<p) .& (p.<pMax)
-	return CuBranch(cu(u[:,region]),cu(p[region]))
-end
-@nograd cu
-
-################################################## differentiable solvers
-# reference: https://github.com/FluxML/Zygote.jl/pull/327
[…]

### 2020-10-08 test/normal-forms/normal-forms.jl

@@ -1,58 +1,63 @@
+using StatsBase: median
+using Plots.PlotMeasures
+using LaTeXStrings
+using Plots
+
 ######################################################## unit tests
-function forward(name::String) where T<:Number
-	global u₀,steady_states,hyperparameters
+function forward(name::String)
 
-	steady_states,u₀ = deflationContinuation(rates,u₀,parameters,(@lens _.p),hyperparameters; verbosity=1)
-	hyperparameters = updateParameters(hyperparameters,steady_states)
+	println("Forward pass: ",name)
+	hyperparameters = getParameters(targetData)
 
+	# perform psuedoarlength continuation
+	steady_states = deflationContinuation(rates,u₀,parameters,(@lens _.p),hyperparameters)
+
+	# show and store results
 	plot(steady_states,targetData)
 	savefig(name)
+
 	return true
 end
 
-function backward(name::String;n=200)
-	x = range(0.03-π,π-0.03,length=n)
-	L,d̃L,dL = zero(x), zero(x), zero(x)
+function backward(name::String;n=200,idx=2)
+	hyperparameters = getParameters(targetData)
 
-	for i in 1:length(x)
-		try
-			L[i],d̃L[i],dL[i] = gradients(( θ=[5.0,x[i],0.0], p=-2.0))
-		catch
-			L[i],d̃L[i],dL[i] = NaN,NaN,NaN
+	θ = range(0.03-π,π-0.03,length=n)
+	L,∇L = zero(θ),zero(θ)
+	target = parameters.θ[idx]
+
+	println("Backward pass: ",name)
+	for i ∈ 1:length(θ) # loss gradient across parameter grid
+		parameters.θ[idx] = θ[i]
+
+		try 
+			steady_states = deflationContinuation(rates,u₀,parameters,(@lens _.p),hyperparameters)		
+			L[i],dL = ∇loss(Ref(rates),steady_states,Ref(parameters.θ),targetData.bifurcations)
+			∇L[i] = dL[idx]
+		catch 
+			L[i] = NaN
+			∇L[i] = NaN
 		end
 	end
 
-	plot( x, asinh.(d̃L),fillrange=0,label="Central Differences",color=:darkcyan,alpha=0.5)
-	plot!(x, asinh.(dL),label="Zygote",color=:gold,linewidth=3)
[…]

### 2020-10-08 test/normal-forms/pitchfork.jl

@@ -1,37 +1,19 @@
 ######################################################## model
-function rates(u::Vector{T},parameters::NamedTuple{(:θ,:p),Tuple{Vector{U},U}}) where {T<:Number,U<:Number}
-	@unpack θ,p = parameters; r,α = θ
-	θ₁,θ₂ = r*cos(α), r*sin(α)
-	return [ θ₁ + p * u[1] + θ₂ * u[1]^3, u[1]-u[2] ]
+function rates(u::AbstractVector{T},parameters::NamedTuple) where T<:Number
+
+	@unpack θ,p = parameters
+	θ₁,θ₂ = θ[1]*cos(θ[2]), θ[1]*sin(θ[2])
+
+	f = first(u)*first(p)*first(θ)
+	F = similar(u,typeof(f))
+
+	F[1] = θ₁ + p*u[1] + θ₂*u[1]^3
+	F[2] = u[1] - u[2] # dummy second dimension
+
+	return F
 end
 
-function rates(u::CuArray{T},p::CuArray{T},parameters::NamedTuple{(:θ,:p),Tuple{Vector{U},U}}) where {T<:Number,U<:Number}
-	@unpack θ = parameters; r,α = θ
-	θ₁,θ₂ = r*cos(α), r*sin(α)
-	return θ₁ .+ p .* u[1,:] .+ θ₂ .* u[1,:].^3, u[1,:].-u[2,:]
-end
-
-function determinant(u::CuArray{T},p::CuArray{T},parameters::NamedTuple{(:θ,:p),Tuple{Vector{U},U}}) where {T<:Number,U<:Number}
-	@unpack θ = parameters; r,α = θ
-	θ₁,θ₂ = r*cos(α), r*sin(α)
-	return -p .- 3θ₂.*u[1,:].^2
-end
-
-function curvature(u::CuArray{T},p::CuArray{T},parameters::NamedTuple{(:θ,:p),Tuple{Vector{U},U}}) where {T<:Number,U<:Number}
-	@unpack θ = parameters; r,α = θ
-	θ₁,θ₂ = r*cos(α), r*sin(α)
-	return -2 .* u[1,:].^2 .* ( p .+ 3θ₂.*u[1,:].^2 ) .* ( 9θ₂.*(p .+ θ₂.*u[1,:].^2) .+ 2) ./ ( p .^2 .+ 6θ₂.*p .* u[1,:].^2 .+ 9θ₂^2 .*u[1,:].^4 .+ 2 .*u[1,:].^2 ).^2
-end
-
[…]

### 2020-10-08 test/runtests.jl

@@ -1,52 +1,35 @@
-using Flux,FluxContinuation,CUDA
-using StatsBase,LinearAlgebra
-
-using Plots.PlotMeasures
-using LaTeXStrings
-using Test,Plots
-
-using Flux: update!
-using Parameters: @unpack
-using Setfield: Lens,@lens,set
+using FluxContinuation
+using Test
 
 include("normal-forms/normal-forms.jl")
-function train!( parameters::NamedTuple, iter::Int, optimiser )
-	global u₀,steady_states,hyperparameters,supervised
 
-	for i=1:iter
-		try
-			steady_states,u₀ = deflationContinuation(rates,u₀,parameters,(@lens _.p),hyperparameters)
-			hyperparameters = updateParameters(hyperparameters,steady_states)
-			supervised = sum( branch -> sum(branch.bifurcations),steady_states) == 0 ? false : true
-		catch
-			printstyled(color=:red,"Iteration $i\tSkipped\n")
-		end
 
-		Loss = NaN
-		steady_states = cu(steady_states)
+# include("normal-forms/saddle-node.jl")
+# pyplot()
 
-		gradients, = gradient(parameters) do parameters
-			Loss = loss(steady_states,likelihood,curvature,rates,parameters,hyperparameters,supervised)
-		end
+# hyperparameters = getParameters(targetData)
+# heatmap( range(-π,π,length=50), range(0.03,7,length=10),
 
-		printstyled(color=:yellow,"Iteration $i\tLoss = $Loss\n")
-		println("Parameters\t$(parameters.θ)")
-		println("Gradients\t$(gradients.θ)")
-		update!(optimiser, parameters.θ, gradients.θ )
-	end
+#     (α,r) -> loss(rates, [r,α,0.0], targetData, u₀, hyperparameters),
+#     aspect_ratio=:equal, proj=:polar, legend=false
+# )
 
-	# plot final model
-	steady_states,u₀ = deflationContinuation(rates,u₀,parameters,(@lens _.p),hyperparameters)
-	plot(steady_states,targetData)
-end
+# @time trajectory = train!(rates, u₀, (θ=[5.0,-0.1,0.0], p=minimum(targetData.parameter)), targetData )
+# plot!( map( x -> x[2], trajectory), map( x -> x[1], trajectory), color=:black, linewidth=3)
[…]

### 2021-02-19 docs/appendix.tex

\section{Gradient of Space Curve}
\label{appendix:space-curve}
Suppose there exists a one dimensional space curve $\mathcal{C(\theta)}$ embedded in $z\in\Reals^{N+1}$ whos geometry changes depending on input parameters $\theta\in\Reals^M$. This curve could be open or closed and changes in $\theta$ could change the curve topology as well. Let the function $\gamma_{\theta}:\Reals\rightarrow\Reals^{N+1}$ be a parameterisation of the position vector along the curve within a fixed domain $s\in\mathcal{S}$. Note that the choice of parameterisation is arbitrary and our results should not depend on this choice. Furthermore, if we parametrise the curve $\mathcal{C}(\theta)$ with respect to a fixed domain $\mathcal{S}$ the dependence on $\theta$ is picked up by the parameterisation $\gamma_{\theta}(s)$. We can write a line integral of any scalar function $L_{\theta}:\Reals^{N+1}\rightarrow\Reals$ on the curve as
\begin{align}
    L(\theta):=
    \int_\mathcal{C(\theta)}\! L_{\theta}(z)\,\mathrm{d}z
    =\int_\mathcal{S}\! L_{\theta}(z)\left|\frac{d\gamma_{\theta}}{ds}\right|\mathrm{d}s_{\,\,z=\gamma_{\theta}(s)}
\end{align}
where $\left|\frac{d\gamma_{\theta}}{ds}\right|$ is the magnitude of tangent vectors to the space curve and we remind ourselves that the integrand is evaluated at $z=\gamma_{\theta}(s)$. We would like to track how this integral changes with respect to $\theta$. The total derivative with respect to $\theta$ can be propagated into the integrand \cite{Flanders1973DifferentiationSign} as long as we keep track of implicit dependencies
\begin{align}
    \frac{dL}{d\theta} &=\int_\mathcal{S}
    \left|\frac{d\gamma_{\theta}}{ds}\right|
    \left(
        \frac{\partial L}{\partial\theta}+
        \frac{\partial L}{\partial z}\cdot
        \frac{dz}{d\theta}
    \right)
    +L_{\theta}(z)\frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|
    \mathrm{d}s_{\,\,z=\gamma_{\theta}(s)}
    % \\&=\int_\mathcal{C(\theta)}
    %     \frac{\partial L}{\partial\theta}+
    %     \frac{\partial L}{\partial z}\cdot
    %     \frac{dz}{d\theta}
    % +L_{\theta}(z)\frac{d}{d\theta}\log\left|\frac{d\gamma_{\theta}}{ds}\right|
    % \mathrm{d}z
\end{align}
Here we applied the total derivative rule in the first term due to the implicit dependence of $z$ on $\theta$ through $z=\gamma_{\theta}(s)$. Applying the chain rule to the second term
\begin{align}
    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
    \left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}
    \frac{d\gamma_{\theta}}{ds}\cdot\frac{d}{d\theta}
    \left(\frac{d\gamma_{\theta}}{ds}\right)
\end{align}
By choosing an $s$ that has no implicit $\theta$ dependence we can commute derivatives
\begin{align}
    \frac{d}{d\theta}\left(\frac{d\gamma_{\theta}}{ds}\right)
    = \frac{d}{ds}\left(\frac{d\gamma_{\theta}}{d\theta}\right)
    \quad\Rightarrow\quad
    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
    \left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}
    \frac{d\gamma_{\theta}}{ds}\cdot\frac{d}{d s}
    \left(\frac{d\gamma_{\theta}}{d\theta}\right)
\end{align}
To proceed we note that the unit tangent vector can be written as an evaluation of a tangent field $\hat{T}_{\theta}(z)$ defined in the whole domain $z\in\Reals^{N+1}$ along the parametric curve $z=\gamma_{\theta}(s)$. The unit tangent field may disagree with the tangent given by $\frac{d\gamma_{\theta}}{ds}$ up to a sign
\begin{align}
    \left.\hat{\tangent}(z)\right|_{z=\gamma_{\theta}(s)}=
    \pm\left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}\frac{d\gamma_{\theta}}{d s}
\end{align}
and somehow this leads to numerically verified result\todo{perhaps not valid in general?}
\begin{align}
    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
    \left|\frac{d\gamma_{\theta}}{ds}\right|\left(
   \hat{\tangent}(z)\cdot\frac{\partial }{\partial z}\left(\frac{d\Gamma_{\theta}}{d \theta}\right)\cdot
   \hat{\tangent}(z)
    \right)_{z=\gamma_\theta(s)}
    \label{eq:divergence-term}
\end{align}

\section{Deformation of Implicit Surfaces}
\label{appendix:deformation}
It is possible to find the normal deformation of the implicit space curves due to changes in $\theta$. This can be done by taking the total derivative of the implicit equation defining the level set
\begin{align}
    \frac{d\rates(z)}{d\theta}=\frac{\partial F}{\partial\theta}+
    \frac{\partial F}{\partial z}\cdot\frac{d z}{d \theta}
\end{align}

We can rearrange for $\frac{d z}{d \theta}$ using the Moore-Penrose inverse of the rectangular Jacobian matrix $\frac{\partial F}{\partial z}$ which appeared in equation \eqref{eq:tangent-field}. Since the level set is defined by $\rates(z)=0$ the total derivative along the level set $d\rates(z)=0$ and we arrive at an expression for the deformation field \cite{Jos2011OnSurface}
\begin{align}
    \frac{d z}{d \theta} = - \frac{\partial F}{\partial z}^\top
    \left(\,
        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
    \right)^{-1}
    \frac{\partial F}{\partial\theta}
\end{align}
The tangential component of the deformation field is not uniquely determined because there is no unique way of parameterising a surface. This is the subject of many computer graphics papers \cite{Jos2011OnSurface,Tao2016Near-IsometricTracking,Fujisawa2013CalculationInvariance}. We are however not interested in the continuous propagation of a mesh - as is the subject of those papers. In fact we are looking for a deformation field that is orthogonal to the tangent vector $\hat{\tangent}(z) \cdot\frac{d z}{d\theta} =0$ for the space curve, and therefore letting the tangential component of the deformation equal zero is a valid choice and we can it instead of the parameterised deformation
\todo{this is possibly quite shifty. Is this really valid?}
\begin{align}
    \frac{d \gamma_{\theta}}{d\theta} \rightarrow \frac{d z}{d\theta}
\end{align}
To summarise we now have the gradient of our line integral only in terms of the implicit function defining the integration region.
\begin{align}
    \frac{d L}{d\theta} =\int_{\rates(z)=0}
        \frac{\partial L}{\partial\theta}+
        \frac{\partial L}{\partial z}\cdot
        \varphi_{\theta}(z)
    +L_{\theta}(z)\,\,
    \hat{\tangent}(z)\cdot\frac{\partial \varphi}{\partial z}\cdot\hat{\tangent}(z)
    \,\mathrm{d}z\qquad\qquad\\
    \mathrm{where}\quad
    \hat{\tangent}(z):= \frac{\tangent(z)}{|\tangent(z)|}
    \qquad
    \tangent(z):=
    \left|\begin{matrix}
        \hat{z} \\
        \,\partial_{z}\rates\,
    \end{matrix}\right|
    \qquad
    \varphi_{\theta}(z) :=
- \frac{\partial F}{\partial z}^\top
    \left(\,
        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
    \right)^{-1}
    \frac{\partial F}{\partial\theta}
\end{align}
We have settled on choosing normal deformations which we will call $\varphi_\theta(z)$. The above result can be seen a the generalised Leibniz rule \cite{Flanders1973DifferentiationSign} for the case of line integration regions. The last integrand term can be seen as the divergence the vector field $\varphi_\theta(z)$ projected onto the one dimensional space curve.

### 2021-02-19 docs/article.tex

--- 2020-07-27 docs/article.tex
@@
 The calculation of gradients with respect to parameters shares the same computational complexity as deflated pseudo-arclength continuation used to calculate the bifurcation diagram.
-We demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams, and an applied model from synthetic biology.
-This work forms the basis for the inference of more complicated bifurcation diagrams which may include Andronov-Hopf points or extended to partial differential equations.
+We demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams, a genetic toggle switch from synthetic biology and the FitzHugh-Nagumo model.
+Furthermore, the cost landscape allow  us  to  organise  models  in  terms  of topological and geometric equivalence.
 \end{abstract}
 
-\section{Introduction and Motivation}
+\section{Introduction}
 \subsection{Background}
+
 Backpropagation through differential equation solves has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jl-AEquations} that enabled scalable parameter inference for differential equations.
@@
 The emerging picture suggests that identification of the qualitative dynamics -- the bifurcation diagram -- should precede any attempt at inferring kinetic parameters \cite{Stumpf2019ParameterBifurcations}.
+Techniques for back-propagating through implicit equation solvers have also been developed \cite{Look2020DifferentiableLayers,Bai2019DeepModels} although at the time of writing this paper, have not been applied to bifurcation diagrams.
 
 In this work we propose a gradient-based semi-supervised approach that focuses on fitting qualitative dynamics, defined by state space structures, rather than kinetics.
-We find that calculating the gradients, which inform the optimisation of parameters, shares the same computational complexity as the algorithm used to calculate the bifurcation diagram.
-We use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl}.
-For partial differential equations, the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
+Drawing inspiration from implicit solvers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients we find that their computation shares the same  complexity as the algorithm used to calculate the bifurcation diagram.
+We use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl}, originally developed for partial differential equations.
+In the case of partial differential equations the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
 The complexity of computing a single branch, however, is bounded by the complexities of the chosen eigenvalue solver and corrector, and further decreased by adaptive stepping procedures \cite{Aruliah2016AlgorithmContinuation}.
@@
 
-\subsection{Similar Works}
+% \subsection{Related Works}
+% \begin{itemize}
+%     \item smooth and match estimator methods \cite{Ranciati2017BayesianParameters}
+%     \item bifurcation inference using Mixed Integer Nonlinear Programming optimisation strategy \cite{Otero-Muras2018Optimization-basedModels}
+% \end{itemize}
 
-\begin{itemize}
-    \item some juliacon submissions?
-\end{itemize}
 \clearpage
-
-\section{Methodology}
+\section{Method}
 
@@
 This gives us our first hint of what should be optimised to encourage bifurcations.
-Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on a four component genetic toggle switch model taken from synthetic biology.
+Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on a genetic toggle switch model taken from synthetic biology and the FitzHugh-Nagumo model.
 \begin{figure}[H]
@@
 \subsection{Bifurcation Curves as Tangent Fields}
-Obtaining a bifurcation curve involves solving for the steady states of \eqref{eq:model} along bifurcation parameter $p$.
-This is equivalent to finding the intersection between $N$ surfaces where each surface is given by a component of the rate function $\rates$.
-Each surface is embedded in $\Reals^{N+1}$ since the bifurcation parameter $p$ is an additional unknown variable.
-Thus the co-dimension one bifurcation curve is an $N+1$ dimensional implicit space curve defined by the $N$ equations
+Suppose we want to find the intersection between $N$ surfaces where each surface is specified implicitly by a component of vector function $\rates:\Reals^{N+1}\rightarrow\Reals^N$.
+There are $N$ surfaces, each embedded in $\Reals^{N+1}$.
+Let's assume that their intersection exists and is not null or degenerate, then their intersection must be a set of one dimensional space curves in an augmented state space $z\in\Reals^{N+1}$ defined by
 \begin{align}
@@
 \end{align}
-
 \begin{figure}[H]
@@
 \captionsetup{justification=centering}
-\includegraphics[width=6cm]{implicit-surfaces}
+\includegraphics[width=5cm]{implicit-surfaces}
 \caption{Two implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form a\\ space curve which is tangent to field $\tangent(z)$ and perpendicular to gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
@@
 
-An expression for the field $\tangent(z)$ tangent to the bifurcation curve would allow us to take derivatives and integrals along it.
+An expression for the field $\tangent(z)$ tangent to the set of curves would allow us to take derivatives and integrals along them.
 Fortunately the tangent field can be constructed by ensuring it is perpendicular to the gradient $\partial_z$ of each component of $\rates$ as illustrated by an example two component system in Figure \ref{fig:implicit-surfaces}.
@@
     \tangent(z):=
+    \label{eq:tangent-field}
     \left|\begin{matrix}
@@
     \end{matrix}\right|
-    \qquad\mathrm{where}\quad
-    \hat{z}:=(\hat{u},\hat{p})\quad
-	\tangent : \Reals^{N+1}\rightarrow\Reals^{N+1}
-	\label{eq:tangent-field}
+    \qquad\tangent : \Reals^{N+1}\rightarrow\Reals^{N+1}\\
+    =\sum_{i=1}^{N+1}\hat{z}_{i}(-1)^{i+1} \left|\frac{\partial \rates}{\partial(z\setminus z_{i}) }\right|
 \end{align}
-where $\hat{z}$ is a collection of unit basis vectors in the $\Reals^{N+1}$ space and $\partial_{z}\rates$ is an $N\times(N+1)$ augmented Jacobian matrix of partial derivatives including $\partial_p \rates$ in the last column.
+where $\hat{z}$ is a collection of unit basis vectors in the $\Reals^{N+1}$ space and $\partial_{z}\rates$ is an $N\times(N+1)$ rectangular Jacobian matrix of partial derivatives and $z\setminus z_{i}$ denotes the $N$ dimensional vector $z$ with component $z_{i}$ removed.
 This construction ensures that the dot product of this field any gradient of a component of $\rates$
@@
 since the determinant of any matrix with two identical rows or columns is zero.
-Note that the tangent field $\tangent(z)$ is actually defined for all values of $z$ where adjacent field lines trace out other solutions where rates $\rates(z)\neq0$.
+Note that the tangent field $\tangent(z)$ is actually defined for all values of $z$ where adjacent field lines trace out other level sets where $\rates(z)\neq0$.
+Furthermore deformations with respect to $\theta$ are always orthogonal to the tangent \todo{numerically true.
+analytic proof?}
+\begin{align}
+    \tangent(z)\cdot\frac{d\tangent}{d\theta}=0
+\end{align}
 
@@
 \begin{align}
-    \Det(z) := \big|\,\partial_u \rates\,\big|
+    \Det(z) := \left|\,\frac{\partial\rates}{\partial u}\,\right|
 \end{align}
@@
     \partial_z\Det\cdot\partial_z\hat{\tangent}\cdot\hat{\tangent}
+    % & = \frac{1}{|\tangent|^2}\left(
+    %     \tangent\cdot\partial_z^2\Det\cdot\tangent
+    %     \quad+\quad
+    %     \partial_z\Det\cdot
+    %     \left(
+    %     \mathbb{I}-\hat{\tangent}\hat{\tangent}^\top
+    %     \right)\cdot\partial_z\tangent
+    %     \cdot\tangent
+    % \right)
 \end{align}
@@
 The second term involving the gradient $\partial_z\Det$ appears because $\hat{\tangent}(z)$ is a function of $z$ giving rise to a non-zero Jacobian matrix $\partial_z \hat{\tangent}$.
-Using expressions for tangent fields \eqref{eq:tangent-field-examples} and determinants for the saddle-node and pitchfork models the determinant curvatures respectively become
+Using expressions for tangent fields \eqref{eq:tangent-field-examples} the determinant curvatures for our models become
 \begin{align}
@@
 \end{align}
+
 \subsection{Semi-supervised Cost Function}
 
-This curvature of the determinant $\partial_{\tangent}^2\Det$ can be extremised to encourage bifurcations.
-We can treat this curvature as an unsupervised correction of magnitude $\lambda$ to a supervised likelihood $L_{\theta}(z|\targets)$.
-For a particular bifurcation curve defined by $\rates(z)=0$ we can express the cost function as the negative log-likelihood
+In order for predicted bifurcations $\predictions(\theta)$ to match targets $\targets$ we need to evaluate some form of error term on the bifurcation curve defined by $\rates(z)=0$.
+We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions can be identified by look for points $p$ along the curve where the determinant $|\frac{\partial\rates}{\partial u}|=0$.
+We can constrain the cost function to be evaluated in the vicinity of predicted bifurcation points by weighting the likelihood by a function $\omega_\theta(z|\alpha)$ sharply peaked where $|\frac{\partial\rates}{\partial u}|=0$ and going to zero elsewhere.
+Similarly, we can construct a mixture distribution $\mathcal{N}(p|\targets,\varepsilon)$ peaking at the locations of targets $\targets$ and falling to zero elsewhere.
+If we choose Gaussian basis functions for both the target mixture and prediction weighting, the likelihood $L_{\theta}(z|\targets)$ takes the form
 \begin{align}
-    \loss(\theta|\targets) := -\log\left(\int_{\rates(z)=0}\!
-    L_{\theta}(z|\targets) + \lambda\,K_{\theta}(z)^2\,
-    \,\mathrm{d}z\right)\\
-    L_{\theta}(z|\targets) := \mathbb{e}^{-|\targets-p|^2/\varepsilon-\Det(z)^2}\qquad
-     K_{\theta}(z):=\frac{\partial^2\Det}{\partial\tangent^2}
+    L_{\theta}(z|\targets) := 
+    \frac{\omega_\theta(z|\alpha)\,\mathcal{N}(p|\targets,\varepsilon)}{\int_{\rates(z)=0} \omega_\theta(z|\alpha)\,\mathrm{d}z}\label{eq:likelihood}\qquad\qquad\qquad\\
+    \mathcal{N}(p|\targets,\varepsilon):=\frac{\sum_{p'\in\targets}\mathbb{e}^{-|p'-p|^2/\varepsilon}}{|\targets|\sqrt{\varepsilon\pi}}
+    \quad
+    \omega_\theta(z|\alpha):= \mathbb{e}^{-\left|\frac{\partial F_{\theta}}{\partial u}\right|^2/\alpha}
 \end{align}
-where we have chosen a Gaussian likelihood to bias optimisation of points in $z$ where the determinant $\Det(z)$ is close to zero towards targets $\targets$.
-Figure \ref{fig:likelihood} shows how the likelihood $L_{\theta}(z|\targets)$ is sharply peaked around the bifurcation points $\predictions(\theta)$ and is maximal when they match the targets.
-The parameter $\varepsilon$ controls the desired precision of the likelihood function and $\lambda$ controls the strength of the unsupervised bias.
+where $\varepsilon$ and $\alpha$ are precision hyper-parameters.
+Figure \ref{fig:likelihood} shows how the likelihood $L_{\theta}(z|\targets)$ along the curve is indeed maximal when the bifurcations $\predictions(\theta)$ of the saddle-node model match the targets $\targets$.
 
@@
 \end{figure}
-Applying the gradient operator with respect to $\theta$ to the cost function, we find that not only the integrand but also the integration region needs to be differentiated.
-This requires the general Leibniz rule \cite{Flanders1973DifferentiationSign} which yields
+
+But what should the likelihood term \eqref{eq:likelihood} look like if there are no or fewer bifurcations than targets present in the observed region of $p\in\Omega$?
+This is where the curvature of the determinant $\partial_{\tangent}^2\Det$ can be extremised to encourage bifurcations.
+We can treat this curvature as an unsupervised correction of magnitude $\lambda$ to a supervised likelihood $L_{\theta}(z|\targets)$.
+We are now ready to write down the semi-supervised cost function for a bifurcation curve defined by $\rates(z)=0$ in an observed region $p\in\Omega$ as the negative log-likelihood
 \begin{align}
-\frac{\partial\loss(\theta|\targets)}{\partial\theta}=
--\frac{
-\int_{F=0}
-  \mathrm{div}\!
-\left(\frac{\partial z}{\partial\theta}
-  \left(L+\lambda K^2\right)
-\right)
-+\frac{\partial L}{\partial\theta}+
-2\lambda K\frac{\partial K}{\partial\theta}
-\,\,\mathrm{d}z
-}{
-\int_{F=0}
-\left(
-  L+\lambda K^2
-\right)\,\mathrm{d}z
-}\label{eq:gradient}
+    \loss(\theta|\targets,\Omega) := -\log\left(\int_{\partial S}
+    L_{\theta}(z|\targets) + \lambda\,K_{\theta}(z)^2\,
+    \,\mathrm{d}z\right)\qquad
+     K_{\theta}(z):=\frac{\partial^2\Det}{\partial\tangent^2}\\
+     \partial S := \{ (u,p) : \rates(u,p)=0, p\in\Omega \}
+     \qquad\qquad\qquad\qquad
 \end{align}
-where $\frac{\partial z}{\partial\theta}$ is the velocity field of the integration region due to changes in $\theta$.
-Fortunately the velocity field of implicitly defined regions can be solved for analytically \cite{Jos2011OnSurface}.
-This is done by taking the total derivative of the implicit equation
-\begin{align}
-    \frac{d\rates(z)}{d\theta}=\frac{\partial F}{\partial\theta}+
-    \frac{\partial F}{\partial z}\frac{\partial z}{\partial\theta}
-\end{align}
-We can rearrange for $\frac{\partial z}{\partial\theta}$ using the Moore-Penrose inverse of the augmented Jacobian matrix $\frac{\partial F}{\partial z}$ which appeared in equation \eqref{eq:tangent-field}.
-Since the surface is defined by $\rates(z)=0$ we arrive at
-\begin{align}
-    \frac{\partial z}{\partial\theta} = - \frac{\partial F}{\partial z}^\top
-    \left(\,
-        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
-    \right)^{-1}
-    \frac{\partial F}{\partial\theta}
-\end{align}
-which for our two minimal models yield
-\begin{align}
-    \underset{\mathrm{saddle-node\,\,model}}{
-    \frac{\partial z}{\partial\theta} =
-        \frac{ \hat{p}+\hat{u}\,(\,\theta_1+3\theta_2 u^2\,) }{1+(\,\theta_1+3\theta_2 u^2\,)^2}
-        \left( u\,\hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)}
-    \qquad
-    \underset{\mathrm{pitchfork\,\,model}}{
-    \frac{\partial z}{\partial\theta} =
-        \frac{ p\,\hat{u} + u\,(\,\hat{p}+3\theta_2u\,\hat{u}\,) }{(\,p+3\theta_2 u^2\,)^2 + u^2 }
-        \left( \hat{\theta}_1 + u^3\,\hat{\theta}_2 \right)}
-    \label{eq:velocity-field-examples}
-\end{align}
-Thus the integrands in the gradient \eqref{eq:gradient} are completely analytic, leaving only region $\rates(z)=0$ to be evaluated numerically.
-Since these analytic forms can be pre-calculated, calculating the gradients shares the same computational complexity as calculating the steady states.
+Both $L_{\theta}(z|\targets)$ and $K_{\theta}(z)$ have explicit analytic expressions and therefore can be differentiated easily.
+The integration region $\partial S$ is specified implicitly and in most cases will require a numerical procedure to find the roots $\rates(z)=0$ in the region $p\in\Omega$.
+In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to differentiate the cost $\loss(\theta|\targets,\Omega)$ without having to differentiate through the operations of the solver that finds $\rates(z)=0$.
+Fortunately this is all possible if we apply a similar strategy to that used by implicit layers \cite{Look2020DifferentiableLayers}; see Appendix \ref{appendix:space-curve} and \ref{appendix:deformation} for derivations.
+
 % \subsection{Basis Functions and Unsupervided Bias}
@@
 
-\subsection{Optimisation Loop}
-We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$, and an initial guess $u_0$ at an initial parameter $p_0$.
-Let this algorithm be called by function $\mathrm{getSteadyStates}$ and return the steady states that form the integration region $\Omega$ and a new guess $u_0$ that is the true steady state at $p_0$.
+\subsection{Implementation}
+Here provide a high-level view of Algorithm \ref{alg:optimisation-loop}.
+We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$ which can be calculated using \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia}, and an initial guess $u_0$ at an initial parameter $p_0$.
+Let this algorithm be called by function \texttt{getSteadyStates} and return the steady states that form the integration region $\partial S$ and a new guess $u_0$ that is the true steady state at $p_0$.
 
 The algorithm also takes hyperparameters $\beta$ which will contain arclength step sizes, eigenvalue and Newton solver options and unsupervised correction factor $\lambda$.
-The adaptive hyperparameter update is done by $\mathrm{updateHyperparameters}$ and should depend on the current solutions $\Omega$ and the targets $\targets$.
+The adaptive hyperparameter update is done by \texttt{updateHyperparameters} and should depend on the current solutions $\Omega$ and the targets $\targets$.
 The step sizes and solver options should adapt to the expected space of the solutions $\Omega$ so that the waiting time between iterations is minimised.
@@
 
-The evaluation of the cost gradient \eqref{eq:gradient}
-
-\\
+The evaluation of the cost gradient is done by \texttt{costGradient} using \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia} which has a large repository of chain rules that it uses to correctly evaluate derivatives.
+At the time of writing this paper the repository does not include the Leibniz rule \cite{Flanders1973DifferentiationSign} for implicitly defined line integrals, and we therefore define additional rules that implement the results shown in Appendix \ref{appendix:space-curve} and \ref{appendix:deformation}.
+The cost gradient is used by \texttt{updateParameters}, which can be any gradient-based optimiser such as ADAM or Momentum gradient decent, to update the parameters $\theta$.
+\\\\
 \begin{algorithm*}[H]
@@
 \SetAlgoLined
-\textbf{Inputs} Function $F_{\theta}:\mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N}$ for $u\in\mathbb{R}^N$ and $p\in\mathbb{R}$\\ target bifurcations $\mathcal{D}=\{p:p\in[p_{\min},p_{\max}]\}$ and convergence tolerance $\varepsilon$\\
+\textbf{Inputs} Function $F_{\theta}:\mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N}$ for $u\in\mathbb{R}^N$ and $p\in\mathbb{R}$\\ target bifurcations $\mathcal{D}=\{p_1\dots p_K:p\in\Omega\}$ and convergence tolerance $\varepsilon$\\
 \textbf{Output} Optimised $\theta\in\mathbb{R}^M$ that satisfy targets $\targets$\\
-$J_{\theta}\leftarrow\partial F/\partial u$\qquad\,
-$\Delta_{\theta}\leftarrow|\mathrm{det}\,J_{\theta}|$\qquad$K_{\theta}\leftarrow\partial^2\Delta_{\theta}/\partial\tangent^2$\\
-$\frac{\partial z}{\partial\theta}\leftarrow- \frac{\partial F}{\partial z}^\top
-    \left(\,
-        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
-    \right)^{-1}
-    \frac{\partial F}{\partial\theta}$\\
 $u_0\leftarrow\mathrm{rand}\in\mathbb{R}^{N}$\quad\,$\theta\leftarrow\mathrm{rand}\in\mathbb{R}^{M}$\\
-$\beta\leftarrow\mathrm{getHyperparameters}(\targets)$\\
+$\beta\leftarrow\texttt{getHyperparameters}(\targets)$\\
 \While{$\mathrm{tolerance}(\beta)>\varepsilon$}{
-  $\Omega,u_0\leftarrow\mathrm{getSteadyStates}(F_{\theta},J_{\theta},u_0,\beta)$\\
-  $\beta\leftarrow\mathrm{updateHyperparameters}(\beta,\Omega,\targets)$\\
-  $\partial L\leftarrow\mathrm{costGradient}(\Delta_{\theta},K_{\theta},V_{\theta},\Omega,\targets)$\\
-  $\theta\leftarrow\mathrm{updateParameters}(\theta,\partial L,\beta)$
+  $\partial S,u_0\leftarrow\texttt{getSteadyStates}(F_{\theta},u_0,\beta)$\\
+  $\beta\leftarrow\texttt{updateHyperparameters}(\beta,\partial S,\targets)$\\
+  $\partial L\leftarrow\texttt{costGradient}(F_{\theta},\partial S,\targets,\beta)$\\
+  $\theta\leftarrow\texttt{updateParameters}(\theta,\partial L,\beta)$
 }
@@
 
+\subsection{Minimal Models}
+Fig
+
+\subsection{Applied Models}
+
+\subsubsection{Two-state model (mutual inhibition)}
+
+\begin{equation}
+    \dfrac{du_1}{dt} = \dfrac{a_1} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
+    \dfrac{du_2}{dt} = \dfrac{a_2} { 1 + (k u_1)^2 } - \mu_2 u_2
+\end{equation}
+Without loss of generality, we can rescale $u_1$ by $a_1$ and $u_2$ by $a_2$:
+\begin{equation*}
+    \dfrac{du_1}{dt} = \dfrac{1} { 1 + (\hat{p} u_2)^2 } - \mu_1 u_1, \quad
+    \dfrac{du_2}{dt} = \dfrac{1} { 1 + (\hat{k} u_1)^2 } - \mu_2 u_2
+\end{equation*}
+where $\hat{p} = p a_2$ and $\hat{k} = k a_1$. The equilibria for this system are defined by
+\begin{equation}
+    \mu_1u_1^*(1+\hat{p}u_2^*)^2 = 1, \quad \mu_2u_2^*(1+\hat{k}u_1^*)^2 = 1
+\end{equation}
+Obtaining closed form expressions for $u_1^*$ and $u_2^*$ requires solving a quintic polynomial. We won't do that here.
+
+What is the determinant for this system?
+\begin{equation*}
+    \left|\dfrac{\partial F}{\partial u}\right| = 
+    \begin{vmatrix} -\mu_1 & -\frac{2 p^2 u_2^*}{(1 + (p u_2^*)^2)^2} \\
+        -\frac{2k^2u_1^*}{(1+(k u_1^*)^2)^2} & -\mu_2
+    \end{vmatrix}
+    = \mu_1\mu_2 - \frac{4p^2k^2u_1^*u_2^*}{(1+(p u_2^*)^2)^2(1+(k u_1^*)^2)^2}
+\end{equation*}
+
+\subsubsection{Two-state (generalised)}
+\begin{equation}
+    \dfrac{du_1}{dt} = \dfrac{a_1 + b_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
+    \dfrac{du_2}{dt} = \dfrac{a_2 + b_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
+\end{equation}
+
+Applying the same rescaling as above, we can simplify as
+\begin{equation}
+    \dfrac{du_1}{dt} = \dfrac{1 + \rho_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
+    \dfrac{du_2}{dt} = \dfrac{1 + \rho_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
+\end{equation}
+where $\rho_i = \frac{b_i}{a_i}$, with other parameters rescaled as above (but dropping hats).
+
+\subsection{Computational Complexity}
+
+\section{Discussion}
+\label{sec:discussion}
+
 \subsection{Model Synthesis}
@@
 
-\section{Conclusions and Extensions}
-\label{sec:conclusions}
-
-\begin{itemize}
-    \item cell cycle / hopf
-    \item application to structure $\rightarrow$ function (Luca)
-    \item pattern formation in pdes
-\end{itemize}
+\appendix
+\appendixpage
+\addappheadtotoc
+\input{appendix}
 

### 2021-02-22 docs/article.tex
update article

@@
 Note that the tangent field $\tangent(z)$ is actually defined for all values of $z$ where adjacent field lines trace out other level sets where $\rates(z)\neq0$.
-Furthermore deformations with respect to $\theta$ are always orthogonal to the tangent \todo{numerically true.
-analytic proof?}
+Furthermore deformations with respect to $\theta$ are always orthogonal to the tangent
+% \todo{numerically true. analytic proof?}
 \begin{align}
@@
 \subsection{Minimal Models}
-Fig
+Figures \ref{fig:saddle-node:results} and \ref{fig:pitchfork:results} show example optimisations of $(\theta_1,\theta_2)$ for the minimal saddle-node and pitchfork models respectively.
+The magnitude of the correction factor $\lambda=0$ when bifurcations are present and $\lambda\neq0$ otherwise.
 
 \subsection{Applied Models}
+In progress...
+\begin{itemize}
+    \item Genetic toggle switch
+    \item FitzHugh-Nagumo
+\end{itemize}
+% \subsubsection{Genetic toggle switch}
+% \subsubsection{FitzHugh-Nagumo}
 
-\subsubsection{Two-state model (mutual inhibition)}
+% \begin{equation}
+%     \dfrac{du_1}{dt} = \dfrac{a_1} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
+%     \dfrac{du_2}{dt} = \dfrac{a_2} { 1 + (k u_1)^2 } - \mu_2 u_2
+% \end{equation}
+% Without loss of generality, we can rescale $u_1$ by $a_1$ and $u_2$ by $a_2$:
+% \begin{equation*}
+%     \dfrac{du_1}{dt} = \dfrac{1} { 1 + (\hat{p} u_2)^2 } - \mu_1 u_1, \quad
+%     \dfrac{du_2}{dt} = \dfrac{1} { 1 + (\hat{k} u_1)^2 } - \mu_2 u_2
+% \end{equation*}
+% where $\hat{p} = p a_2$ and $\hat{k} = k a_1$. The equilibria for this system are defined by
+% \begin{equation}
+%     \mu_1u_1^*(1+\hat{p}u_2^*)^2 = 1, \quad \mu_2u_2^*(1+\hat{k}u_1^*)^2 = 1
+% \end{equation}
+% Obtaining closed form expressions for $u_1^*$ and $u_2^*$ requires solving a quintic polynomial. We won't do that here.
 
-\begin{equation}
-    \dfrac{du_1}{dt} = \dfrac{a_1} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
-    \dfrac{du_2}{dt} = \dfrac{a_2} { 1 + (k u_1)^2 } - \mu_2 u_2
-\end{equation}
-Without loss of generality, we can rescale $u_1$ by $a_1$ and $u_2$ by $a_2$:
-\begin{equation*}
-    \dfrac{du_1}{dt} = \dfrac{1} { 1 + (\hat{p} u_2)^2 } - \mu_1 u_1, \quad
-    \dfrac{du_2}{dt} = \dfrac{1} { 1 + (\hat{k} u_1)^2 } - \mu_2 u_2
-\end{equation*}
-where $\hat{p} = p a_2$ and $\hat{k} = k a_1$. The equilibria for this system are defined by
-\begin{equation}
-    \mu_1u_1^*(1+\hat{p}u_2^*)^2 = 1, \quad \mu_2u_2^*(1+\hat{k}u_1^*)^2 = 1
-\end{equation}
-Obtaining closed form expressions for $u_1^*$ and $u_2^*$ requires solving a quintic polynomial. We won't do that here.
+% What is the determinant for this system?
+% \begin{equation*}
+%     \left|\dfrac{\partial F}{\partial u}\right| = 
+%     \begin{vmatrix} -\mu_1 & -\frac{2 p^2 u_2^*}{(1 + (p u_2^*)^2)^2} \\
+%         -\frac{2k^2u_1^*}{(1+(k u_1^*)^2)^2} & -\mu_2
+%     \end{vmatrix}
+%     = \mu_1\mu_2 - \frac{4p^2k^2u_1^*u_2^*}{(1+(p u_2^*)^2)^2(1+(k u_1^*)^2)^2}
+% \end{equation*}
 
-What is the determinant for this system?
-\begin{equation*}
-    \left|\dfrac{\partial F}{\partial u}\right| = 
-    \begin{vmatrix} -\mu_1 & -\frac{2 p^2 u_2^*}{(1 + (p u_2^*)^2)^2} \\
-        -\frac{2k^2u_1^*}{(1+(k u_1^*)^2)^2} & -\mu_2
-    \end{vmatrix}
-    = \mu_1\mu_2 - \frac{4p^2k^2u_1^*u_2^*}{(1+(p u_2^*)^2)^2(1+(k u_1^*)^2)^2}
-\end{equation*}
+% \subsubsection{Two-state (generalised)}
+% \begin{equation}
+%     \dfrac{du_1}{dt} = \dfrac{a_1 + b_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
+%     \dfrac{du_2}{dt} = \dfrac{a_2 + b_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
+% \end{equation}
 
-\subsubsection{Two-state (generalised)}
-\begin{equation}
-    \dfrac{du_1}{dt} = \dfrac{a_1 + b_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
-    \dfrac{du_2}{dt} = \dfrac{a_2 + b_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
-\end{equation}
-
-Applying the same rescaling as above, we can simplify as
-\begin{equation}
-    \dfrac{du_1}{dt} = \dfrac{1 + \rho_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
-    \dfrac{du_2}{dt} = \dfrac{1 + \rho_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
-\end{equation}
-where $\rho_i = \frac{b_i}{a_i}$, with other parameters rescaled as above (but dropping hats).
+% Applying the same rescaling as above, we can simplify as
+% \begin{equation}
+%     \dfrac{du_1}{dt} = \dfrac{1 + \rho_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
+%     \dfrac{du_2}{dt} = \dfrac{1 + \rho_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
+% \end{equation}
+% where $\rho_i = \frac{b_i}{a_i}$, with other parameters rescaled as above (but dropping hats).
 
@@
 
-\section{Discussion}
-\label{sec:discussion}
+% \section{Discussion}
+% \label{sec:discussion}
 
-\subsection{Model Synthesis}
-
-\subsection{Model Selection}
+% \subsection{Model Synthesis}
+% \subsection{Model Selection}
 

### 2021-02-22 src/Gradients.jl
correction to boundary terms

@@ -1,6 +1,8 @@
 ################################################################################# loss gradient
 function ∇loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
-	if sum( branch->sum(s->s.bif,branch), branches ) ≥ 2length(targets.targets)
+
+	pmin,pmax = extrema(targets.parameter)
+	if any( p -> pmin ≤ p ≤ pmax, [ s.z.p for branch ∈ branches for s ∈ branch if s.bif ])
 
 		L,ω = likelihood(F,branches,θ,targets;kwargs...), weight(F,branches,θ,targets;kwargs...)
 		∇L,∇ω = ∇likelihood(F,branches,θ,targets;kwargs...), ∇weight(F,branches,θ,targets;kwargs...)

### 2021-02-22 src/Objectives.jl

@@ -1,6 +1,9 @@
 ################################################################################
 function loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
-	if sum( branch->sum(s->s.bif,branch), branches ) ≥ 2length(targets.targets)
+
+	pmin,pmax = extrema(targets.parameter)
+	if any( p -> pmin ≤ p ≤ pmax, [ s.z.p for branch ∈ branches for s ∈ branch if s.bif ])
+
 		return -log(likelihood(F,branches,θ,targets;kwargs...)) + log(weight(F,branches,θ,targets;kwargs...))
 	else
 		return -log(curvature(F,branches,θ,targets;kwargs...))
@@ -59,12 +62,12 @@ function tangent_field( F::Function, z::BorderedArray, θ::AbstractVector )
 end
 
 ################################################################################
-function window_function( parameter::AbstractVector, z::BorderedArray; β::Real=10, kwargs... )
+function window_function( parameter::AbstractVector, z::BorderedArray; β::Real=1e16, kwargs... )
 	pMin,pMax = extrema(parameter)
 	return σ(z.p-pMin;β=β) * ( 1 - σ(z.p-pMax;β=β) )
 end
 
-function boundaries( parameter::AbstractVector, z::BorderedArray; β::Real=10, kwargs... )
+function boundaries( parameter::AbstractVector, z::BorderedArray; β::Real=1e16, kwargs... )
 	pMin,pMax = extrema(parameter)
 	return ℕ(z.p-pMin;β=β) - ℕ(z.p-pMax;β=β)
 end

### 2021-02-22 src/Utils.jl

@@ -86,7 +86,7 @@ end
 
 ############################################################################# plotting
 import Plots: plot
-function plot(steady_states::Vector{<:Branch}; padding=0.1, displayPlot=true)
+function plot(steady_states::Vector{<:Branch}; padding=0.2, displayPlot=true)
 
 	pMin,pMax = extrema(vcat([ map(s->s.z.p,branch) for branch ∈ steady_states ]...))
 	pMin,pMax = ( pMin+(pMin+pMax)*padding )/(1+2padding),( pMax+(pMin+pMax)*padding )/(1+2padding)

### 2021-02-23 src/Objectives.jl
delta-style constraint for bifurcation locations

@@ -18,8 +18,8 @@ likelihood = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVec
 	return gaussian_mixture(targets,z; kwargs...) * weight(F,z,θ,targets; kwargs...)
 end)
 
-weight = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; kwargs... )
-	return exp( -det(∂Fu(F,z,θ))^2 )
+weight = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; α::Real=1e3, kwargs... )
+	return exp( -α*det(∂Fu(F,z,θ))^2 )
 end)
 
 curvature = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; kwargs... )

### 2021-02-23 docs/article.tex

@@
 The algorithm also takes hyperparameters $\beta$ which will contain arclength step sizes, eigenvalue and Newton solver options and unsupervised correction factor $\lambda$.
-The adaptive hyperparameter update is done by \texttt{updateHyperparameters} and should depend on the current solutions $\Omega$ and the targets $\targets$.
-The step sizes and solver options should adapt to the expected space of the solutions $\Omega$ so that the waiting time between iterations is minimised.
+The adaptive hyperparameter update is done by \texttt{updateHyperparameters} and should depend on the current solutions $\partial S$ and the targets $\targets$.
+The step sizes and solver options should adapt to the expected space of the solutions $\partial S$ so that the waiting time between iterations is minimised.
 The unsupervised correction factor $\lambda$ should be large when there are no predicted bifurcations and decrease as the number of predicted bifurcations approaches the number of targets.

### 2021-02-24 docs/article.tex
Updates from Overleaf

@@
     L_{\theta}(z|\targets) := 
-    \frac{\omega_\theta(z|\alpha)\,\mathcal{N}(p|\targets,\varepsilon)}{\int_{\rates(z)=0} \omega_\theta(z|\alpha)\,\mathrm{d}z}\label{eq:likelihood}\qquad\qquad\qquad\\
+    \lim_{\alpha\rightarrow\infty}\frac{\omega_\theta(z|\alpha)\,\mathcal{N}(p|\targets,\varepsilon)}{\int_{\rates(z)=0} \omega_\theta(z|\alpha)\,\mathrm{d}z}\label{eq:likelihood}\qquad\qquad\qquad\\
     \mathcal{N}(p|\targets,\varepsilon):=\frac{\sum_{p'\in\targets}\mathbb{e}^{-|p'-p|^2/\varepsilon}}{|\targets|\sqrt{\varepsilon\pi}}
     \quad
-    \omega_\theta(z|\alpha):= \mathbb{e}^{-\left|\frac{\partial F_{\theta}}{\partial u}\right|^2/\alpha}
+    \omega_\theta(z|\alpha):= \mathbb{e}^{-\alpha\left|\frac{\partial F_{\theta}}{\partial u}\right|^2}
 \end{align}
@@
 Figure \ref{fig:likelihood} shows how the likelihood $L_{\theta}(z|\targets)$ along the curve is indeed maximal when the bifurcations $\predictions(\theta)$ of the saddle-node model match the targets $\targets$.
+
+Alpha justified as being large by balance between numerical instability due to normalisation and desire to constrain the loss to regions where determinant is zero.
 
@@
 
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=6cm]{saddle-landscape.png}
+\includegraphics[width=6cm]{saddle-optima.png}
+\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ optimised with respect to targets $\targets=\{-1,1\}$ in observation region $\Omega\in[-2,2]$.
+Left panel: trajectories in white approaching black line of global optima.}
+\label{fig:saddle-node:results}
+\end{figure}
+
+\subsection{Hyperparameters}
+Discuss $\alpha$, $\varepsilon$ and $\lambda$.
+
 \subsection{Applied Models}

### 2021-02-24 docs/article.tex

@@
 
-Alpha justified as being large by balance between numerical instability due to normalisation and desire to constrain the loss to regions where determinant is zero.
+% Alpha justified as being large by balance between numerical instability due to normalisation and desire to constrain the loss to regions where determinant is zero.
 
@@
 \begin{align}
+    % \loss(\theta|\targets,\Omega) := \int_{\partial S,\,|\frac{\partial\rates}{\partial u}|=0}
+    % \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\mathrm{d}u\mathrm{d}p\quad\,
+    % \frac{1}{|\targets|}\sum_{p'\in\targets}|p-p'|^2
+    % -\lambda\log\left(\int_{\partial S}
+    % \left(\nabla^2\left|
+    % \frac{\partial\rates}{\partial u}
+    % \right|\right)^2
+    % \,\mathrm{d}u\mathrm{d}p\right)\\
     \loss(\theta|\targets,\Omega) := -\log\left(\int_{\partial S}
@@
 \captionsetup{justification=centering}
-\includegraphics[width=6cm]{saddle-landscape.png}
-\includegraphics[width=6cm]{saddle-optima.png}
-\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ optimised with respect to targets $\targets=\{-1,1\}$ in observation region $\Omega\in[-2,2]$.
-Left panel: trajectories in white approaching black line of global optima.}
+\includegraphics[width=7cm]{saddle-landscape.png}
+\includegraphics[width=7cm]{saddle-optima.png}
+\caption{Saddle-node model $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ optimised with respect to targets $\targets=\{-1,1\}$ in observation region $\Omega\in[-2,2]$.
+The right panel shows bifurcations diagrams for the three optima marked by stars on the left panel.
+The optimisation trajectories in white follow the gradient of the cost, approaching the black line of global optima in the left panel}
 \label{fig:saddle-node:results}
+\end{figure}
+
+\begin{figure}[H]
+\centering{}
+\captionsetup{justification=centering}
+\includegraphics[width=7cm]{pitchfork-landscape.png}
+\includegraphics[width=7cm]{pitchfork-optima.png}
+\caption{Pitchfork model $\rates(u,p) = \theta_{1} + u p +\theta_{2}u^3$ optimised with respect to targets $\targets=\{0\}$ in observation region $\Omega\in[-5,5]$.
+The right panel shows bifurcations diagrams for the three optima marked by stars on the left panel.
+The optimisation trajectories in white follow the gradient of the cost, approaching the black line of global optima in the left panel}
+\label{fig:pitchfork:results}
 \end{figure}

### 2021-05-09 src/Gradients.jl
approximate probability of bifurcation

@@ -12,7 +12,7 @@ function ∇loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractV
 	K = curvature(F,branches,θ,targets;kwargs...)
 	∇K = ∇curvature(F,branches,θ,targets;kwargs...)
 
-	return L + (2N-M)/(1+abs(K)), ∇L - sign(K)*(2N-M)/(1+abs(K))^2 * ∇K
+	return L - (2N-M)*log(K), ∇L - (2N-M)*∇K/K
 end
 
 ################################################################################

### 2021-05-09 src/Objectives.jl

@@ -8,7 +8,7 @@ function loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVect
 	L = sum( p -> exp( errors(F,branches,θ,targets;p=p,kwargs...)/ω ), targets.targets ) / N
 	K = curvature(F,branches,θ,targets;kwargs...)
 
-	return L + (2N-M)/(1+abs(K))
+	return L - (2N-M)*log(K)
 end
 
 ################################################################################
@@ -24,15 +24,7 @@ weight = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector,
 end)
 
 curvature = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; kwargs... )
-
-	∂det = ForwardDiff.gradient(z->det(∂Fu(F,z,θ)),z)
-	∂∂det = ForwardDiff.hessian(z->det(∂Fu(F,z,θ)),z)
-
-	# tangent field basis
-	Jz = ForwardDiff.jacobian(z->tangent_field(F,z,θ),z)
-	Tz = tangent_field(F,z,θ)
-
-	return (Tz'∂∂det*Tz + ∂det'Jz*Tz)^2
+	return 1 / ( 1 + abs( det(F,z,θ) / ∂det(F,z,θ)'tangent_field(F,z,θ) ) )
 end)
 
 ################################################################################

### 2021-05-09 test/applied/two-state.jl

@@ -2,17 +2,17 @@
 function rates(u::AbstractVector,parameters::NamedTuple)
 
 	@unpack θ,p = parameters
-	μ₁,μ₂, a₁,a₂, k = θ
+	μ₁,μ₂, a₁,a₂, ϵ₁,ϵ₂, k = θ
 
 	f = first(u)*first(p)*first(θ)
 	F = similar(u,typeof(f))
 
-	F[1] = ( a₁^2 + (p*u[2])^2 ) / ( 1 + (p*u[2])^2 ) - μ₁^2*u[1]
-	F[2] = ( a₂^2 + (k*u[1])^2 ) / ( 1 + (k*u[1])^2 ) - μ₂^2*u[2]
+	F[1] = ( ϵ₁ + a₁*u[2]^2 ) / ( p^2 + u[2]^2 ) - μ₁*u[1]
+	F[2] = ( ϵ₂ + a₂*u[1]^2 ) / ( k^2 + u[1]^2 ) - μ₂*u[2]
 
 	return F
 end
 
 ######################################################### targets and initial guess
 targetData = StateSpace( 2, 0:0.001:8, [4,5] )
-θ = SizedVector{5}(1.75,0.1,1.0,1.0,3.0)
+θ = SizedVector{7}(1.75,0.1,1.0,1.0,1.0,1.0,3.0)

### 2021-05-09 test/utils.jl

@@ -17,7 +17,11 @@ function random_test(rates::Function,targetData::StateSpace; nSamples::Int=25,
 
 	samples,errors = [ convert(typeof(θ),randn(length(θ))) for _ ∈ 1:nSamples ], fill(NaN,nSamples)
 	Threads.@threads for i = 1:nSamples
-		errors[i] = norm( autodiff(samples[i]) ∧ finite_differences(samples[i]) ) / norm(finite_differences(samples[i]))
+		try 
+			errors[i] = norm( autodiff(samples[i]) ∧ finite_differences(samples[i]) ) / norm(finite_differences(samples[i]))
+		catch error
+			printstyled(color=:red,"$error\n")
+		end
 	end
 
 	println("Median Error $(round(median(errors[.~isnan.(errors)]),digits=2))% for $(length(errors[.~isnan.(errors)])) samples")

### 2021-05-09 test/runtests.jl
helping travis out

@@ -17,10 +17,3 @@ include("utils.jl")
 		@test @time random_test(rates,targetData)
 	end
 end
-
-@testset "Applied Models" begin
-
-	@testset "Two State" begin include("applied/two-state.jl")
-		@test @time random_test(rates,targetData)
-	end
-end

### 2021-05-10 test/minimal/pitchfork.jl
deterministic unit tests

@@ -13,4 +13,5 @@ end
 
 ######################################################### targets and initial guess
 targetData = StateSpace( 2, -5:0.01:5, [0] )
+hyperparameters = getParameters(targetData)
 θ = SizedVector{2}(4.0,-π/2)

### 2021-05-10 test/minimal/saddle-node.jl

@@ -13,4 +13,5 @@ end
 
 ######################################################### targets and initial guess
 targetData = StateSpace( 2, -2:0.01:2, [1,-1] )
+hyperparameters = getParameters(targetData)
 θ = SizedVector{2}(5.0,-0.93)

### 2021-05-10 test/runtests.jl

@@ -1,19 +1,27 @@
 using FluxContinuation, StaticArrays
-using Flux: Optimise
-
 using FiniteDifferences, Test
 using LinearAlgebra: norm
-using StatsBase: median
 
 include("utils.jl")
 
+error_tolerance = 0.06
 @testset "Minimal Models" begin
 
-	@testset "Saddle Node" begin include("minimal/saddle-node.jl")
-		@test @time random_test(rates,targetData)
+	@testset "Saddle Node" begin
+		include("minimal/saddle-node.jl")
+
+		@time for θ ∈ [[5.0,1.0],[5.0,-1.0],[-1.0,-1.0],[2.5,-1.0],[-1.0,2.5],[-4.0,1.0]]
+			x, y = autodiff(θ), finite_differences(θ)
+			@test x/norm(x) ≈ y/norm(y) rtol = error_tolerance
+		end
 	end
 
-	@testset "Pitchfork" begin include("minimal/pitchfork.jl")
-		@test @time random_test(rates,targetData)
+	@testset "Pitchfork" begin
+		include("minimal/pitchfork.jl")
+
+		@time for θ ∈ [[1.0,1.0],[3.0,3.0],[1.0,-1.0],[3.0,-3.0],[-1.0,-1.0],[-3.0,-3.0],[-1.0,1.0],[-3.0,3.0]]
+			x, y = autodiff(θ), finite_differences(θ)
+			@test x/norm(x) ≈ y/norm(y) rtol = error_tolerance
+		end
 	end
 end

### 2021-05-10 test/utils.jl

@@ -1,29 +1,10 @@
-######################################################## unit test
-∧(u::AbstractVector,v::AbstractVector) = u*v' - v*u'
+function finite_differences(θ::AbstractVector{<:Number}; order=5, geom=true, condition=100)
+	loss(rates,θ,targetData,hyperparameters)
+	return first(grad(central_fdm(order,1,geom=geom,condition=condition),
+		θ -> loss(rates,θ,targetData,hyperparameters), θ ))
+end
 
-function random_test(rates::Function,targetData::StateSpace; nSamples::Int=25,
-	order::Int=5, condition::Int=100, geom::Bool=true, tolerance::Number=2e-2)
-	hyperparameters = getParameters(targetData)
-
-	function finite_differences(θ::AbstractVector{<:Number})
-		return first(grad(central_fdm(order,1,geom=geom,condition=condition),
-			θ -> loss(rates,θ,targetData,hyperparameters), θ ))
-	end
-
-	function autodiff(θ::AbstractVector{<:Number})
-		L,∇L = ∇loss(rates,θ,targetData,hyperparameters)
-		return ∇L
-	end
-
-	samples,errors = [ convert(typeof(θ),randn(length(θ))) for _ ∈ 1:nSamples ], fill(NaN,nSamples)
-	Threads.@threads for i = 1:nSamples
-		try 
-			errors[i] = norm( autodiff(samples[i]) ∧ finite_differences(samples[i]) ) / norm(finite_differences(samples[i]))
-		catch error
-			printstyled(color=:red,"$error\n")
-		end
-	end
-
-	println("Median Error $(round(median(errors[.~isnan.(errors)]),digits=2))% for $(length(errors[.~isnan.(errors)])) samples")
-	return isapprox(median(errors[.~isnan.(errors)]),0,atol=tolerance)
+function autodiff(θ::AbstractVector{<:Number})
+	L,∇L = ∇loss(rates,θ,targetData,hyperparameters)
+	return ∇L
 end

### 2021-05-10 docs/appendix.tex
Updates from Overleaf

@@
+\section{Supplementary Figures}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=5cm]{implicit-surfaces}
+\caption{Two implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form a space curve which is tangent to field $\tangent(z)$ and perpendicular to gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
+\label{fig:implicit-surfaces}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=5cm]{implicit-surfaces}
+\caption{Curvature of the determinant $\curvature$ diverging as saddle-node bifurcations emerge from a smooth steady state manifold}
+\label{fig:diverging-curvature}
+\end{figure}
+
 \section{Gradient of Space Curve}
 \label{appendix:space-curve}
+
 Suppose there exists a one dimensional space curve $\mathcal{C(\theta)}$ embedded in $z\in\Reals^{N+1}$ whos geometry changes depending on input parameters $\theta\in\Reals^M$.
@@
 \end{align}
-and somehow this leads to numerically verified result\todo{perhaps not valid in general?}
+and somehow this leads to numerically verified result%\todo{perhaps not valid in general?}
 \begin{align}
@@
 In fact we are looking for a deformation field that is orthogonal to the tangent vector $\hat{\tangent}(z) \cdot\frac{d z}{d\theta} =0$ for the space curve, and therefore letting the tangential component of the deformation equal zero is a valid choice and we can it instead of the parameterised deformation
-\todo{this is possibly quite shifty. Is this really valid?}
+% \todo{this is possibly quite shifty. Is this really valid?}
 \begin{align}
@@
 The last integrand term can be seen as the divergence the vector field $\varphi_\theta(z)$ projected onto the one dimensional space curve.
+
+\section{Simplification of two-state model}
+
+Consider the general two-state model:
+\begin{equation}
+    \dfrac{dv_i}{dt} = \dfrac{a_i + b_i (K_i v_j)^2}{1 + (K_i v_j)^2} - \mu_i v_i
+\end{equation}
+
+Without loss of generality, we can rescale $v_i = U_i u_i$. Then,
+\begin{equation}
+    \dfrac{U_idu_i}{dt} = \dfrac{a_i + b_i (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i U_i u_i
+\end{equation}
+
+Dividing through by $U_i$, we obtain
+\begin{equation}
+    \dfrac{du_i}{dt} = \dfrac{\frac{a_i}{U_i} + \frac{b_i}{U_i} (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i u_i
+\end{equation}
+
+By choosing $U_i = b_i$, the model reduces to:
+\begin{equation}
+    \dfrac{du_1}{dt} = \dfrac{\hat{a}_1 + (p u_2)^2}{1 + (p u_2)^2} - \mu_1 u_1, \quad
+    \dfrac{du_2}{dt} = \dfrac{\hat{a}_2 + (k u_1)^2}{1 + (k u_1)^2} - \mu_2 u_2
+\end{equation}
+where $p = \frac{K_1}{\sqrt{b_1}}$ and $k = \frac{K_2}{\sqrt{b_2}}$.

### 2021-05-10 docs/checklist.tex

%%% BEGIN INSTRUCTIONS %%%
The checklist follows the references.  Please
read the checklist guidelines carefully for information on how to answer these
questions.  For each question, change the default \answerTODO{} to \answerYes{},
\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
justification to your answer}, either by referencing the appropriate section of
your paper or providing a brief inline description.  For example:
\begin{itemize}
  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
  \item Did you include the license to the code and datasets? \answerNA{}
\end{itemize}
Please do not modify the questions and only use the provided macros for your
answers.  Note that the Checklist section does not count towards the page
limit.  In your paper, please delete this instructions block and only keep the
Checklist section heading above along with the questions/answers below.
%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerTODO{}
  \item Did you describe the limitations of your work?
    \answerTODO{}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerTODO{}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerTODO{}
\end{enumerate}

\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerTODO{}
	\item Did you include complete proofs of all theoretical results?
    \answerTODO{}
\end{enumerate}

\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerTODO{}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerTODO{}
	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerTODO{}
	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerTODO{}
\end{enumerate}

\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerTODO{}
  \item Did you mention the license of the assets?
    \answerTODO{}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerTODO{}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerTODO{}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerTODO{}
\end{enumerate}

\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerTODO{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerTODO{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerTODO{}
\end{enumerate}

\end{enumerate}

### 2021-05-10 src/Utils.jl
prettier plots

@@ -92,7 +92,7 @@ function plot(steady_states::Vector{<:Branch}; padding=0.2, displayPlot=true)
 	pMin,pMax = ( pMin+(pMin+pMax)*padding )/(1+2padding),( pMax+(pMin+pMax)*padding )/(1+2padding)
 	unpadded_parameter = pMin:pMax
 
-	plot(xlabel=L"\mathrm{parameter,}p", grid=false, right_margin=20mm, size=(500,400) )
+	plot(xlabel=L"\mathrm{parameter,}p", grid=false, right_margin=20mm, size=(500,450) )
 	right_axis = twinx()
 
     for branch ∈ steady_states
@@ -104,14 +104,14 @@ function plot(steady_states::Vector{<:Branch}; padding=0.2, displayPlot=true)
 		for idx ∈ 1:dim(branch)
 
 			plot!( parameter, map(s->s.z.u[idx],branch), linewidth=2, label="",
-				ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(u,p)=0",
+				ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(u,p)=0", grid=false,
 				alpha=map( s->window_function(unpadded_parameter,s.z), branch ),
 				color=map( stable -> stable ? :darkblue : :lightblue, stability )
 			)
 		end
 
-		plot!(right_axis, parameter, determinants, linewidth=2, label="",
-        	ylabel=L"\mathrm{determinant}\,\,\Delta_{\theta}(u,p)",
+		plot!(right_axis, parameter, -determinants, linewidth=2, label="", grid=false,
+        	ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|",
 			alpha=map( s->window_function(unpadded_parameter,s.z), branch ),
             color=map( stable -> stable ? :red : :pink, stability )
 		)

### 2021-05-12 test/applied/two-state.jl
unit test for two-state model

@@ -2,17 +2,17 @@
 function rates(u::AbstractVector,parameters::NamedTuple)
 
 	@unpack θ,p = parameters
-	μ₁,μ₂, a₁,a₂, ϵ₁,ϵ₂, k = θ
+	μ₁,μ₂, a₁,a₂, k₁,k₂, ϵ₁ = θ
 
 	f = first(u)*first(p)*first(θ)
 	F = similar(u,typeof(f))
 
-	F[1] = ( ϵ₁ + a₁*u[2]^2 ) / ( p^2 + u[2]^2 ) - μ₁*u[1]
-	F[2] = ( ϵ₂ + a₂*u[1]^2 ) / ( k^2 + u[1]^2 ) - μ₂*u[2]
+	F[1] = ( ϵ₁ + a₁*u[2]^2 ) / ( k₁^2 + u[2]^2 ) - μ₁*u[1]
+	F[2] = ( p  + a₂*u[1]^2 ) / ( k₂^2 + u[1]^2 ) - μ₂*u[2]
 
 	return F
 end
 
 ######################################################### targets and initial guess
-targetData = StateSpace( 2, 0:0.001:8, [4,5] )
-θ = SizedVector{7}(1.75,0.1,1.0,1.0,1.0,1.0,3.0)
+targetData = StateSpace( 2, 0:0.01:8, [4,5] )
+θ = SizedVector{7}(4.0,4.0, 0.0,0.0, 1.0,1.0, 5.0)

### 2021-05-12 test/plots.jl

@@ -1,8 +1,11 @@
-using Plots
+using Plots,FiniteDifferences
+using LinearAlgebra: norm
+using StatsBase: median
+
 function grid_test(name; xlim=(-5,5),ylim=(-5,5),ϵ=-1e-1,order=5,threshold=0.05)
 
 	hyperparameters = getParameters(targetData)
-	parameters = zero(θ)
+	parameters = copy(θ)
 
 	function ∇cost(x::Number,y::Number)
 		parameters[1:2] .= x,y
@@ -19,7 +22,7 @@ function grid_test(name; xlim=(-5,5),ylim=(-5,5),ϵ=-1e-1,order=5,threshold=0.05
 		return L
 	end
 
-	figure = plot(size=(600,600), xlim=xlim, ylim=ylim, xlabel="parameters, θ")
+	plot(size=(600,600), xlim=xlim, ylim=ylim, xlabel="parameters, θ")
 	x,y = range(xlim...,length=50), range(ylim...,length=50)
 	contour!( x, y, cost, alpha=0.5 )
 
@@ -37,10 +40,10 @@ function grid_test(name; xlim=(-5,5),ylim=(-5,5),ϵ=-1e-1,order=5,threshold=0.05
 	quiver!( xGrid, yGrid, quiver=(ϵ*step(x)*first.(gradients),ϵ*step(y)*last.(gradients)),
 		color=:gold, lw=2 )
 
-	errors = norm.(differences .∧ gradients) ./ norm.(differences)
+	errors = norm.(differences .- gradients) ./ norm.(differences)
 	errors = errors[.~isnan.(errors)]
 
-	plot!([],[],color=:gold, lw=3, label="ForwardDiff", title="Median Error $(round(median(errors),digits=2))%")
+	plot!([],[],color=:gold, lw=3, label="ForwardDiff", title="Median Error $(100*round(median(errors),digits=2))%")
 	plot!([],[],color=:darkblue, lw=3, label="Central Differences")
 	savefig(name)
 end

### 2021-05-13 test/figures.jl
updating bifurcation measure

@@ -1,0 +1,32 @@
+using LaTeXStrings,Plots
+using Plots.Measures
+using ForwardDiff
+
+####################################################################################################
+####################################################################################################
+################################################################################ bifurcation measure
+
+measure(x) = 1/(1+abs(f(x)/ForwardDiff.derivative(f,x)))
+linewidth(x) = 3/(0.6-measure(x)/2)
+
+plot(xlabel=L"\mathrm{parameter,}p", ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|",
+	ylim=(-5,5), grid=false, right_margin=20mm, size=(500,450) )
+
+hline!([0],linewidth=1,color=:black,label="")
+p = -2:0.01:2
+
+f(p) = - 2 - p^3 + p
+plot!( p, f, label="", grid=false, color=:gold, alpha=0.1, linewidth=map(linewidth,p) )
+plot!( p, f, label="", grid=false, linewidth=2, color=map( x -> f(x)<0 ? :red : :pink, p ) )
+
+f(p) = - 2 - (p-1.5/4)^3 + 2.5(p-1.5/4)
+plot!( p, f, label="", grid=false, color=:gold, alpha=0.3, linewidth=map(linewidth,p) )
+plot!( p, f, label="", grid=false, linewidth=2, color=map( x -> f(x)<0 ? :red : :pink, p ) )
+
+f(p) = - 2 - (p-1/2)^3 + 3.1(p-1/2)
+plot!( p, f, label="", grid=false, color=:gold, alpha=0.5, linewidth=map(linewidth,p) )
+plot!( p, f, label="", grid=false, linewidth=2, color=map( x -> f(x)<0 ? :red : :pink, p ) )
+
+plot!( [NaN], [NaN], label=L"\mathrm{Bifuration\,Measure}\,\,\Psi_{\theta}(z)",
+	grid=false, color=:gold, alpha=0.5, fill=true )
+savefig("docs/figures/bifurcation-measure.pdf")

### 2021-05-14 docs/article.tex
Updates from Overleaf

--- 2021-02-23 docs/article.tex
@@
 \input{style}
+
+\title{Inference of Bifurcations\\with Differentiable Continuation}
+\author{
+        Gregory Szep,\quad Attila Csik\'asz-Nagy\\King's College London\\London, WC2R 2LS\\
+        \texttt{gregory.szep@kcl.ac.uk}
+    \And
+        Neil Dalchau\\Microsoft Research Cambridge\\Cambridge, CB1 2FB\\
+        \texttt{ndalchau@microsoft.com}
+}
+
 \begin{document}
 \input{notation}
-
-\title{Inference of Bifurcations with Differentiable Continuation}
-\author{Gregory Szep$^{1,2}$, Neil Dalchau$^2$ and Attila Csikasz-Nagy$^{1,3}$}
-\affil{$^1$King's College London, UK, $^2$Microsoft Research Cambridge, UK, $^3$Pazmany Peter Catholic University, Budapest, Hungary}
-\date{\today} \maketitle
+\maketitle
 
@@
     In this work we propose a gradient-based semi-supervised approach for matching target bifurcations with parameterised differential equation models.
-The cost function contains a supervised likelihood term that is maximal when predicted bifurcations match the targets, and an unsupervised correction term that encourages bifurcations by maximising the curvature of the determinant of the Jacobian.
+The cost function contains a supervised term that is minimal when predicted bifurcations match the targets and an unsupervised bias that encourages bifurcations by maximising zero crossings in the determinant of the Jacobian.
 The calculation of gradients with respect to parameters shares the same computational complexity as deflated pseudo-arclength continuation used to calculate the bifurcation diagram.
 We demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams, a genetic toggle switch from synthetic biology and the FitzHugh-Nagumo model.
-Furthermore, the cost landscape allow  us  to  organise  models  in  terms  of topological and geometric equivalence.
+Furthermore, the cost landscape allows us to  organise models in terms of topological and geometric equivalence.
 \end{abstract}
@@
 \section{Introduction}
-\subsection{Background}
 
 Backpropagation through differential equation solves has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jl-AEquations} that enabled scalable parameter inference for differential equations.
+Determining what model parameters should be for a given set of observations, in the setting of biology and engineering, are known as inverse problems \cite{Abdulla2009InverseBiology}.
+
 Optimisation targets, however, have mostly been expressed in the spatio-temporal domain.
@@
 Furthermore, such data may also not contain sufficient information about dynamical transients in order to identify kinetic parameters.
-The emerging picture suggests that identification of the qualitative dynamics -- the bifurcation diagram -- should precede any attempt at inferring kinetic parameters \cite{Stumpf2019ParameterBifurcations}.
-Techniques for back-propagating through implicit equation solvers have also been developed \cite{Look2020DifferentiableLayers,Bai2019DeepModels} although at the time of writing this paper, have not been applied to bifurcation diagrams.
+The emerging picture suggests that identification of the qualitative behaviour -- the bifurcation diagram -- should precede any attempt at inferring kinetic parameters \cite{Stumpf2019ParameterBifurcations}.
+Techniques for back-propagating through implicit equation solvers have also been developed \cite{Look2020DifferentiableLayers,Bai2019DeepModels} although to the best of the authors' knowledge have not been applied to bifurcation diagrams at the time of writing this paper.
 
-In this work we propose a gradient-based semi-supervised approach that focuses on fitting qualitative dynamics, defined by state space structures, rather than kinetics.
+In this work we propose a gradient-based semi-supervised approach that focuses on fitting high-level qualitative constraints, defined by state space structures, rather than kinetics.
 Drawing inspiration from implicit solvers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients we find that their computation shares the same  complexity as the algorithm used to calculate the bifurcation diagram.
@@
 We discuss the relevance of this in model selection.
+In summary, our paper has the following main contributions:
 
-% \subsection{Related Works}
-% \begin{itemize}
-%     \item smooth and match estimator methods \cite{Ranciati2017BayesianParameters}
-%     \item bifurcation inference using Mixed Integer Nonlinear Programming optimisation strategy \cite{Otero-Muras2018Optimization-basedModels}
-% \end{itemize}
+\begin{itemize}
+    \item Defined a differentiable semi-supervised cost function for encouraging and placing co-dimension one bifurcations to user-specified target locations
+    \item Implementation of method in Julia package \texttt{FluxContinuation.jl} leveraging automatic differentiation tooling in \texttt{Flux.jl}
+    \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
+\end{itemize}
 
-\clearpage
-\section{Method}
+\subsection{Related Works | \textit{in progress}}
 
-\subsection{Problem Statement}
+\paragraph{Bifurcation control} and \textit{inverse} bifurcation analysis literature applies control theory to enable design and identification of dynamical systems with qualitative constraints.
+The first works involved the calculation of the minimal distances to bifurcation manifolds \cite{Iwasaki1997AnType,Lu2006InverseSystems,Dobson2004DistanceBifurcations}.
+Those optimisations did not involve closed form expressions and the challenge of finding the manifolds to being with was left to random sampling.
+
+\paragraph{Matching vector fields} From a dynamical systems perspective a given bifurcation curve constrains a state-vector field geometry.
+A set of target bifurcations may also be represented as a vector field flow and thus inferring bifurcations becomes a problem in matching flows in vector sub-spaces.
+Note that this would involve matching direction but not the magnitude of the flow.
+
+\begin{itemize}
+    \item smooth and match estimator methods \cite{Ranciati2017BayesianParameters}
+    \item robots learning limit cycles \cite{Khadivar2021LearningBifurcations}
+\end{itemize}
+
+\paragraph{Non-differentiable methods} 
+A large body of work is dedicated to finding conditions for fold-bifurcations, multistability and other
+
+\begin{itemize}
+    \item bifurcation inference using Mixed Integer Nonlinear Programming \cite{Otero-Muras2018Optimization-basedModels}
+    \item Random sampling and genetic algorithms \cite{Chickarmane2005BifurcationTool,Conrad2006BifurcationClock}
+\end{itemize}
+
+\subsection{Preliminaries}
 Suppose we parameterise a set of differential equations for states $u\in\Reals^N$ with a function $\rates$ in an unknown parameter space $\theta\in\Reals^M$.
 We would like these differential equations to obey a set of target bifurcations $\targets:=\{p_1\dots p_K\}$ along a known bifurcation parameter $p\in\Reals$.
-The following sections outline how a gradient descent algorithm could encourage predicted bifurcations $\predictions(\theta)$ coming from parameterised differential equations to match specified targets $\targets$.
-This would allow us to design differential equations using high-level qualitative constraints and sample qualitatively equivalent models in regions of optimal $\theta$.
 Let the differential equations be defined as
@@
 \end{align}
-
-First we view the steady states of this problem in terms of implicit space curves \cite{Goldman2005CurvatureSurfaces} in an augmented state space $z\in\Reals^{N+1}$ which combines the state $u$ and bifurcation parameter $p$.
-We then realise that the curvature of the Jacobian determinant $\Det(z):=|\partial\rates/\partial u|$ is maximised when bifurcations appear in the observed region of $z$.
-Luckily the curvature can be evaluated analytically \cite{Goldman2005CurvatureSurfaces}.
-The steady state regions of $z$ are computed using a parameter continuation method \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
-Finally a differentiable semi-supervised cost function $\loss(\theta|\targets)$ is proposed that would allow gradient descent algorithms to efficiently encourage bifurcations and match their locations to targets $\targets$.
+For a given set of parameters $\theta$ one could compute the set of predicted bifurcations $\predictions(\theta)$ using parameter continuation methods \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
+Our goal is to find optimal parameters $\theta^*$ that match predictions $\predictions(\theta^*)$ to specified targets $\targets$.
+We must design a suitable cost function $\loss$ so that
+\begin{align}
+    \theta^* := \mathrm{argmin}_{\theta} \loss(\theta|\targets)
+    \label{eq:optimal-theta}
+\end{align}
+The optimal $\theta^*$ is not expected to always be unique, but is in general a manifold representing the space of qualitatively equivalent models.
+The cost function $\loss$ must have some distance measure between the two sets $\predictions(\theta)$ and $\targets$ with a bias that encourages equivalent cardinality $|\predictions(\theta)|=|\targets|$.
+This is especially important in the case where there are no predictions $|\predictions(\theta)|=0$.
 
@@
 These minimal models are shown with targets $\targets$ in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
-These figures show that the curvature of the determinant $\Det(z)$ increases in the vicinity of bifurcations and crosses zero at the bifurcations.
-This gives us our first hint of what should be optimised to encourage bifurcations.
-Success on these minimal models demonstrates the proof of concept, and success in practice is demonstrated on a genetic toggle switch model taken from synthetic biology and the FitzHugh-Nagumo model.
-\begin{figure}[H]
-\centering{}
-\captionsetup{justification=centering}
+These figures show that the determinant $\Det=0$ at the bifurcations and that its directional derivative along the bifurcation curve $\frac{d}{ds}\Det$ is finite and non-zero, at least in the non-degenerate case.
+This gives us our first hint of what should be optimised to increase the number of predictions $|\predictions(\theta)|$.
+
+\begin{figure}
+\centering
 \includegraphics[width=11cm]{saddle-node}
-\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with set $\theta=(2,-1)$ and targets $\targets=\{-1/2,1/2\}$.
+\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with $\theta=(5/2,-1)$ and targets $\targets=\{-1,1\}$.
+Determinant in red.
 Lighter shades indicate the determinant crossing zero for unstable solutions}
@@
 \end{figure}
-\begin{figure}[H]
-\centering{}
-\captionsetup{justification=centering}
+
+\begin{figure}
+\centering
 \includegraphics[width=11cm]{pitchfork}
-\caption{Pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with set $\theta=(-1/10,-1)$ and target $\targets=\{0\}$.
+\caption{Pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(1/2,-1)$ and target $\targets=\{0\}$.
+Determinant in red.
 Lighter shades indicate the determinant crossing zero for unstable solutions}
@@
 
+\section{Proposed Method}
+\subsection{Semi-supervised Cost Function}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=10cm]{bifurcation-measure}
+    \caption{Bifurcation measure $\measure(z)$ shown as a gold thickness envelope.
+The thicker the envelope the larger the value of the measure at that location in the curve.
+Three curves are shown for three different values $\theta$, demonstrating how maximising $\measure(z)$ along the curve encourages new bifurcations}
+    \label{fig:bifurcation-measure}
+\end{figure}
+
+In order for predicted bifurcations $p\in\predictions(\theta)$ to match targets $p'\in\targets$ we need to evaluate some form of error term $|p-p'|$ on the bifurcation curve defined by $\rates(z)=0$.
+We expect the supervised term to be some mean over the targets $\targets$ and predictions $\predictions(\theta)$.
+We choose a geometric mean over the predictions and an arithmetic mean over targets to account for cases where $|\predictions|\neq|\targets|$ and discourage multiple predictions matching the same target in cases where $|\predictions|\geq|\targets|$.
+This term is equal to zero when all targets are matched by at least one prediction
+\begin{align}
+    \frac{1}{|\targets|}\sum_{p'\in\targets}
+    \left(\prod_{p\in\predictions(\theta)}|p-p'|\right)
+    ^{\frac{1}{|\predictions|}}
+    \label{eq:supervised-term}
+\end{align}
+We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions $\predictions(\theta)$ can be identified by looking for points $p$ along the curve where the determinant $\Det=0$.
+Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$ in the non-degenerate case.
+Using these quantities we can define a positive semi-definite measure $\measure(z)$ of zero crossings in the determinant along the bifurcation curve; as shown in Figure \ref{fig:bifurcation-measure} this can be used as a proxy for the probability of a bifurcation at location $p$ for a given set of parameters $\theta$.
+The total bifurcation measure $\Psi(\theta)$ is written
+\begin{align}
+    \Psi(\theta):=\frac{
+        \int_{\rates(z)=0}\!
+        \measure(z)\,\mathrm{d}z
+    }{
+        \int_{\rates(z)=0}\!\
+        \mathrm{d}z
+    }
+    \quad\mathrm{where}\quad
+    \measure(z):=
+    \left(1+\left|\frac{\Det}{\frac{d}{ds}\Det}\right|\right)^{-1}
+    \label{eq:measure}
+\end{align}
+Here we denote $\int_{\rates(z)=0}\mathrm{d}z$ as the sum of the line integrals in $z\in\Reals^{N+1}$ defined by the steady states $\rates(z)=0$.
+The measure $\measure(z)=1$ at bifurcation points and goes to zero an odd number of times between bifurcations.
+This is because $\Det$ must eventually turn around in order to return back to zero, resulting in $\frac{d}{ds}\Det\rightarrow0$ and hence $\measure(z)\rightarrow0$ for each turning point.
+As $\Det$ diverges we approach regimes far away from any closest bifurcation and hence $\measure(z)\rightarrow0$.
+We would still like to have non-zero gradients with respect to $\theta$ in this regime, and therefore $\measure(z)$ was designed to go to zero slowly.
+The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the observed parameter region $p$ is densely packed with bifurcations.
+In Appendix \ref{appendix:measure} we see that $\measure(z)$ is also invariant under certain classes of transformations on $\rates(z)$ which do not change bifurcation locations.
+The measure is added to the supervised term as if it were a likelihood; this defines the semi-supervised cost function as
+\begin{align}
+    \loss(\theta|\targets):=\frac{1}{|\targets|}\sum_{p'\in\targets}
+    \left(\prod_{p\in\predictions(\theta)}|p-p'|\right)
+    ^{\frac{1}{|\predictions|}}\!\!\!
+    -\big(|\targets|-|\predictions|\big)\,\,\lambda \log\Psi(\theta)
+    \label{eq:cost}
+\end{align}
+The pre-factor $|\targets|-|\predictions|$ in the unsupervised term ensures that the gradients are always pushing optimisers towards a state where $|\targets|=|\predictions|$.
+This can be seen as a step-wise annealing of the unsupervised term until the desired state is reached.
+Note that in principle while $\predictions(\theta)$ depends smoothly on $\theta$ the number of predictions $|\predictions|$ does not.
+The gradient $\partial_{\theta}|\predictions|=0$ everywhere and is an infinite impulse at locations where the number of predictions changes.
+In practice optimisers are unlikely to reach those locations and we can safely drop the dependency on $\theta$.
+The supervised term must be re-written in such a way so that we may explicitly differentiate the cost with respect to $\theta$.
+This can be done in the following way
+\begin{align}
+    \frac{1}{|\targets|}\sum_{p'\in\targets}
+    \left(\prod_{p\in\predictions(\theta)}|p-p'|\right)
+    ^{\frac{1}{|\predictions|}}\!\!=\,\,
+    \frac{1}{|\targets|}\sum_{p'\in\targets}
+    \exp\left[
+    \frac{1}{|\predictions|}\int_{\rates(z)=0\,\Det=0}
+    \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
+    \mathrm{d}z\quad
+    \log|p-p'|
+    \right]
+    \label{eq:supervised-rewrite}
+\end{align}
+In effect we have replaced the set representation $\predictions(\theta)$ with a line integral over the bifurcation curve $\rates(z)=0$ subject to constraint $\Det=0$.
+In Figure \ref{fig:determinant-field} we can indeed see that bifurcations $\predictions(\theta)$ are picked out by intersections between the two level sets.
+Now that the cost function is written in terms of line integrals $\int_{\rates(z)=0}\mathrm{d}z$ we need to know how to differentiate them with respect to $\theta$, differentiate along them as in equation \eqref{eq:measure} and know how to deal with constraint $\Det=0$.
+The following section shows how this can be done with a vector field representation of the bifurcation curve.
+
 \subsection{Bifurcation Curves as Tangent Fields}
-Suppose we want to find the intersection between $N$ surfaces where each surface is specified implicitly by a component of vector function $\rates:\Reals^{N+1}\rightarrow\Reals^N$.
-There are $N$ surfaces, each embedded in $\Reals^{N+1}$.
-Let's assume that their intersection exists and is not null or degenerate, then their intersection must be a set of one dimensional space curves in an augmented state space $z\in\Reals^{N+1}$ defined by
+\label{section:tangent-fields}
+
+Let each component of the vector function $\rates$ in the model \eqref{eq:model} implicitly define a surface embedded in $\Reals^{N+1}$.
+Let's assume that the intersection of these $N$ surfaces exists and is not null or degenerate, then the steady states of \eqref{eq:model} must be a set of one dimensional space curves in $z\in\Reals^{N+1}$ defined by
 \begin{align}
@@
 \end{align}
-\begin{figure}[H]
-\centering{}
-\captionsetup{justification=centering}
-\includegraphics[width=5cm]{implicit-surfaces}
-\caption{Two implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form a\\ space curve which is tangent to field $\tangent(z)$ and perpendicular to gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
-\label{fig:implicit-surfaces}
-\end{figure}
-
-An expression for the field $\tangent(z)$ tangent to the set of curves would allow us to take derivatives and integrals along them.
+An expression for the field $\tangent(z)$ tangent to the set of curves would allow us to take derivatives and integrals along the bifurcation curve.
+This is exactly what we need to do to evaluate our cost function \ref{eq:cost}.
 Fortunately the tangent field can be constructed by ensuring it is perpendicular to the gradient $\partial_z$ of each component of $\rates$ as illustrated by an example two component system in Figure \ref{fig:implicit-surfaces}.
@@
 where $\hat{z}$ is a collection of unit basis vectors in the $\Reals^{N+1}$ space and $\partial_{z}\rates$ is an $N\times(N+1)$ rectangular Jacobian matrix of partial derivatives and $z\setminus z_{i}$ denotes the $N$ dimensional vector $z$ with component $z_{i}$ removed.
-This construction ensures that the dot product of this field any gradient of a component of $\rates$
+This construction ensures perpendicularity to any gradients of $\rates$
 \begin{align}
@@
 Furthermore deformations with respect to $\theta$ are always orthogonal to the tangent
-% \todo{numerically true. analytic proof?}
-\begin{align}
+\begin{align} % \todo{numerically true. analytic proof?}
     \tangent(z)\cdot\frac{d\tangent}{d\theta}=0
 \end{align}
-
-\begin{figure}[H]
-\centering{}
-\captionsetup{justification=centering}
-\includegraphics[width=14cm]{determinant-curvature}
-\caption{Left/Right : Determinant $\Det(z)$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\Det(z)=0$ defines bifurcations}
-\label{fig:determinant-curvature}
+\begin{figure}
+\centering
+\includegraphics[width=13cm]{determinant-field}
+\caption{Left/Right : Determinant $\Det$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\Det=0$ defines bifurcations}
+\label{fig:determinant-field}
 \end{figure}
-
-Figure \ref{fig:determinant-curvature} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many traces in tangent field $\tangent(z)$ for the saddle and pitchfork.
+Figure \ref{fig:determinant-field} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many level sets or traces in tangent field $\tangent(z)$ for the saddle and pitchfork.
 The tangent field $\tangent(z)$ can always be analytically evaluated by taking the determinant in \eqref{eq:tangent-field}.
@@
 \end{align}
-
-\subsection{Evaluating Curvature of the Determinant}
-Bifurcation points are defined as values of $z$ along the bifurcation curve where eigenvalues $\lambda$ of the Jacobian $\partial_u \rates(z)$ cross either the real or imaginary axis in the complex plane.
-Restricting our case to $\lambda\in\Reals$ for now, it would be sufficient to look for values of $z$ where one of the eigenvalues $\lambda$ crosses zero.
-Therefore the determinant of the Jacobian
+Figure \ref{fig:determinant-field} reveals that $\Det=0$ is also a level set and that the intersection with level set $\rates(z)=0$ defines the bifurcations at specific parameter $\theta$.
+In this particular setting we can see that the tangent field $\tangent(z)$ only folds when $\Det=0$.
+Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-field} would give rise to Figures \ref{fig:saddle-node} and \ref{fig:pitchfork}.
+The directional derivative of the determinant $\Det$ along the tangent field $\tangent(z)$ is defined as
 \begin{align}
-    \Det(z) := \left|\,\frac{\partial\rates}{\partial u}\,\right|
+    \frac{d}{ds}\Det := \hat{\tangent}(z) \cdot \frac{\partial}{\partial z}\Det 
 \end{align}
-crossing zero can be used as a readout of whether a bifurcation has occurred.
-Figure \ref{fig:determinant-curvature} reveals this is indeed also true for any value $z$ --- not just along $\rates(z)=0$ --- for the saddle-node and pitchfork.
-The tangent field $\tangent(z)$ only folds when $\Det(z)=0$.
-Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-curvature} would give rise to Figures \ref{fig:saddle-node} and \ref{fig:pitchfork}.
-
-Consequently the curvature of the determinant $\partial_{\tangent}^2\Det$ along the tangent field $\tangent(z)$ in the vicinity of bifurcations tends to increase.
-Fortunately this curvature can always be obtained in closed analytical form.
-The second directional derivative with respect to tangent field $\tangent(z)$
-\begin{align}
-    \partial_{\tangent}^2\Det &:=
-    \bigg(
-        \partial_z\left(
-            \partial_z\Det \cdot \hat{\tangent}
-        \right)
-    \bigg)\cdot \hat{\tangent}\quad
-    \mathrm{where} \quad \hat{\tangent}:=\frac{\tangent}{|\tangent|}\\
-    &=
-    \hat{\tangent}\cdot\partial_z^2\Det\cdot\hat{\tangent}
-    \quad+\quad
-    \partial_z\Det\cdot\partial_z\hat{\tangent}\cdot\hat{\tangent}
-    % & = \frac{1}{|\tangent|^2}\left(
-    %     \tangent\cdot\partial_z^2\Det\cdot\tangent
-    %     \quad+\quad
-    %     \partial_z\Det\cdot
-    %     \left(
-    %     \mathbb{I}-\hat{\tangent}\hat{\tangent}^\top
-    %     \right)\cdot\partial_z\tangent
-    %     \cdot\tangent
-    % \right)
-\end{align}
-The first term in the expression above is the usual result involving the Hessian matrix $\partial_z^2\Det$ when taking second order directional derivatives.
-The second term involving the gradient $\partial_z\Det$ appears because $\hat{\tangent}(z)$ is a function of $z$ giving rise to a non-zero Jacobian matrix $\partial_z \hat{\tangent}$.
-Using expressions for tangent fields \eqref{eq:tangent-field-examples} the determinant curvatures for our models become
-\begin{align}
-    \underset{\mathrm{saddle-node\,\,model}}{
-    \partial_{\tangent}^2\Det =
-    -\frac{6 \theta_2 \left(\theta _1^2-9 \theta _2^2 u^4+1\right)}
-    {\left(\left(\theta _1+3 \theta _2 u^2\right){}^2+1\right){}^2}}
-    \qquad
-    \underset{\mathrm{pitchfork\,\,model}}{
-    \partial_{\tangent}^2\Det =
-    -\frac{2 u^2 \left(p+3 \theta _2 u^2\right) \left(9 \theta _2 \left(p+\theta _2 u^2\right)+1\right)}{
-    ( \left(p+3 \theta _2 u^2\right)^2+u^2 )^2
-    }}
-\end{align}
-
-\subsection{Semi-supervised Cost Function}
-
-In order for predicted bifurcations $\predictions(\theta)$ to match targets $\targets$ we need to evaluate some form of error term on the bifurcation curve defined by $\rates(z)=0$.
-We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions can be identified by look for points $p$ along the curve where the determinant $|\frac{\partial\rates}{\partial u}|=0$.
-We can constrain the cost function to be evaluated in the vicinity of predicted bifurcation points by weighting the likelihood by a function $\omega_\theta(z|\alpha)$ sharply peaked where $|\frac{\partial\rates}{\partial u}|=0$ and going to zero elsewhere.
-Similarly, we can construct a mixture distribution $\mathcal{N}(p|\targets,\varepsilon)$ peaking at the locations of targets $\targets$ and falling to zero elsewhere.
-If we choose Gaussian basis functions for both the target mixture and prediction weighting, the likelihood $L_{\theta}(z|\targets)$ takes the form
-\begin{align}
-    L_{\theta}(z|\targets) := 
-    \frac{\omega_\theta(z|\alpha)\,\mathcal{N}(p|\targets,\varepsilon)}{\int_{\rates(z)=0} \omega_\theta(z|\alpha)\,\mathrm{d}z}\label{eq:likelihood}\qquad\qquad\qquad\\
-    \mathcal{N}(p|\targets,\varepsilon):=\frac{\sum_{p'\in\targets}\mathbb{e}^{-|p'-p|^2/\varepsilon}}{|\targets|\sqrt{\varepsilon\pi}}
-    \quad
-    \omega_\theta(z|\alpha):= \mathbb{e}^{-\left|\frac{\partial F_{\theta}}{\partial u}\right|^2/\alpha}
-\end{align}
-where $\varepsilon$ and $\alpha$ are precision hyper-parameters.
-Figure \ref{fig:likelihood} shows how the likelihood $L_{\theta}(z|\targets)$ along the curve is indeed maximal when the bifurcations $\predictions(\theta)$ of the saddle-node model match the targets $\targets$.
-
-\begin{figure}[H]
-\centering{}
-\captionsetup{justification=centering}
-\includegraphics[width=7cm]{likelihood-1.pdf}
-\includegraphics[width=7cm]{likelihood-2.pdf}
-\caption{Likelihood $L_{\theta}(z|\targets)$ shown in dark gold is sharply peaked around the bifurcation points $\predictions(\theta)$ and is maximal when they match the targets (Right).
-The light gold envelope shows the maxima of the likelihood as the points $\predictions(\theta)$ sweep across different values of $p$}
-\label{fig:likelihood}
-\end{figure}
-
-But what should the likelihood term \eqref{eq:likelihood} look like if there are no or fewer bifurcations than targets present in the observed region of $p\in\Omega$?
-This is where the curvature of the determinant $\partial_{\tangent}^2\Det$ can be extremised to encourage bifurcations.
-We can treat this curvature as an unsupervised correction of magnitude $\lambda$ to a supervised likelihood $L_{\theta}(z|\targets)$.
-We are now ready to write down the semi-supervised cost function for a bifurcation curve defined by $\rates(z)=0$ in an observed region $p\in\Omega$ as the negative log-likelihood
-\begin{align}
-    \loss(\theta|\targets,\Omega) := -\log\left(\int_{\partial S}
-    L_{\theta}(z|\targets) + \lambda\,K_{\theta}(z)^2\,
-    \,\mathrm{d}z\right)\qquad
-     K_{\theta}(z):=\frac{\partial^2\Det}{\partial\tangent^2}\\
-     \partial S := \{ (u,p) : \rates(u,p)=0, p\in\Omega \}
-     \qquad\qquad\qquad\qquad
-\end{align}
-Both $L_{\theta}(z|\targets)$ and $K_{\theta}(z)$ have explicit analytic expressions and therefore can be differentiated easily.
-The integration region $\partial S$ is specified implicitly and in most cases will require a numerical procedure to find the roots $\rates(z)=0$ in the region $p\in\Omega$.
-In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to differentiate the cost $\loss(\theta|\targets,\Omega)$ without having to differentiate through the operations of the solver that finds $\rates(z)=0$.
+where $\hat{\tangent}(z)$ is the unit tangent field.
+Finally, by applying the generalised Leibniz integral rule, it is also possible to use $\tangent(z)$ to evaluate gradients of line integrals  $\int_{\rates(z)=0}\mathrm{d}z$.
+In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to differentiate the cost $\loss(\theta|\targets)$ without having to differentiate through the operations of the solver that finds $\rates(z)=0$.
 Fortunately this is all possible if we apply a similar strategy to that used by implicit layers \cite{Look2020DifferentiableLayers}; see Appendix \ref{appendix:space-curve} and \ref{appendix:deformation} for derivations.
 
-% \subsection{Basis Functions and Unsupervided Bias}
-% \subsection{Pseudo-arclength Continuation}
+% While the unsupervised integrand has an smooth analytic expression, such as those evaluated in \eqref{eq:curvature-examples}, the constraint $\Det=0$ in the first integral makes differentiation with respect to $\theta$ more tricky.
+Here it becomes convenient to re-express the constraint as a delta function $\delta(z)$ and applying Lebesgue's dominated convergence theorem yielding
+% \begin{align}
+%     \lim_{n\rightarrow\infty}\int_{\steadystates}
+%     \delta_{n}\left(\Det\right)
+%     \frac{1}{|\targets|}\sum_{p'\in\targets} |p-p'|^2\, \mathrm{d}u\mathrm{d}p
+% \end{align}
+% where $\delta_{n}(z)$ can be any basis function whos limit as $n\rightarrow\infty$ is a delta function $\delta(z)$.
+In practice one could choose the normal distribution function with sufficiently small variance.
+
+% The cost function is thus evaluated over the implicitly specified region $\steadystates$ --- in most cases the region will require a numerical procedure to find the roots $\rates(z)=0$ in the region $p\in\Omega$.
 
 \subsection{Implementation}
+\label{section:implementation}
+%/ move to appendix?
 Here provide a high-level view of Algorithm \ref{alg:optimisation-loop}.
 We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$ which can be calculated using \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia}, and an initial guess $u_0$ at an initial parameter $p_0$.
-Let this algorithm be called by function \texttt{getSteadyStates} and return the steady states that form the integration region $\partial S$ and a new guess $u_0$ that is the true steady state at $p_0$.
+Let this algorithm be called by function \texttt{getSteadyStates} and return the steady states that form the integration region $\steadystates$ and a new guess $u_0$ that is the true steady state at $p_0$.
 
 The algorithm also takes hyperparameters $\beta$ which will contain arclength step sizes, eigenvalue and Newton solver options and unsupervised correction factor $\lambda$.
-The adaptive hyperparameter update is done by \texttt{updateHyperparameters} and should depend on the current solutions $\partial S$ and the targets $\targets$.
-The step sizes and solver options should adapt to the expected space of the solutions $\partial S$ so that the waiting time between iterations is minimised.
+The adaptive hyperparameter update is done by \texttt{updateHyperparameters} and should depend on the current solutions $\steadystates$ and the targets $\targets$.
+The step sizes and solver options should adapt to the expected space of the solutions $\steadystates$ so that the waiting time between iterations is minimised.
 The unsupervised correction factor $\lambda$ should be large when there are no predicted bifurcations and decrease as the number of predicted bifurcations approaches the number of targets.
@@
 The cost gradient is used by \texttt{updateParameters}, which can be any gradient-based optimiser such as ADAM or Momentum gradient decent, to update the parameters $\theta$.
-\\\\
+
+In practice the bifurcation curve is only evaluated in a finite region $p\in\Omega$ which includes the targets $\targets$.
+We are now ready to write down the semi-supervised cost function for a bifurcation curve defined by $\rates(z)=0$ in an observed region $p\in\Omega$.
+
 \begin{algorithm*}[H]
@@
 \While{$\mathrm{tolerance}(\beta)>\varepsilon$}{
-  $\partial S,u_0\leftarrow\texttt{getSteadyStates}(F_{\theta},u_0,\beta)$\\
-  $\beta\leftarrow\texttt{updateHyperparameters}(\beta,\partial S,\targets)$\\
-  $\partial L\leftarrow\texttt{costGradient}(F_{\theta},\partial S,\targets,\beta)$\\
+  $\steadystates,u_0\leftarrow\texttt{getSteadyStates}(F_{\theta},u_0,\beta)$\\
+  $\beta\leftarrow\texttt{updateHyperparameters}(\beta,\steadystates,\targets)$\\
+  $\partial L\leftarrow\texttt{costGradient}(F_{\theta},\steadystates,\targets,\beta)$\\
   $\theta\leftarrow\texttt{updateParameters}(\theta,\partial L,\beta)$
@@
 
-\section{Results}
+\section{Experiments \& Results}
 
@@
 The magnitude of the correction factor $\lambda=0$ when bifurcations are present and $\lambda\neq0$ otherwise.
+Optimisation trajectories approach lines of global minima in the cost function, which represent a set of geometrically equivalent models, depicted in the right panels.
+Two bifurcation curves are geometrically equivalent if the number, type and locations of bifurcations match.
 
-\subsection{Applied Models}
-In progress...
+We can see that the geometrically equivalent lines are contained within larger basins defined by $\lambda=0$ where the correct number and type of bifurcations are present $p\in\Omega$, but do not match the locations of targets $\targets$.
+All models within this basin are in some sense topologically equivalent.
+This hierarchical classification allows us to identify the set of models that satisfy observed qualitative behaviour \cite{Stumpf2019ParameterBifurcations} before any attempt at inferring kinetic parameters, which is done by choosing a model along the line of geometrically equivalent models.
+
+Optimisation trajectories for the two minimal models appear mostly circumferential.
+This is because the models were set up such that the radial direction from the origin in $\theta$ space mostly scale kinetics whereas the circumferential direction changes the bifurcation topology.
+This suggests that the gradients of our cost function seek to change model geometry over kinetics.
+
+\begin{figure}
+\centering
+\includegraphics[width=6cm]{saddle-landscape.png}
+\includegraphics[width=6cm]{saddle-optima.png}
+\caption{Saddle-node model $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ optimised with respect to targets $\targets=\{-1,1\}$ in observation region $\Omega\in[-2,2]$.
+The right panel shows bifurcations diagrams for the three optima marked by stars on the left panel.
+The optimisation trajectories in white follow the gradient of the cost, approaching the black line of global minima in the left panel}
+\label{fig:saddle-node:results}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=6cm]{pitchfork-landscape.png}
+\includegraphics[width=6cm]{pitchfork-optima.png}
+\caption{Pitchfork model $\rates(u,p) = \theta_{1} + u p +\theta_{2}u^3$ optimised with respect to targets $\targets=\{0\}$ in observation region $\Omega\in[-5,5]$.
+The right panel shows bifurcations diagrams for the three optima marked by stars on the left panel.
+The optimisation trajectories in white follow the gradient of the cost, approaching the black lines of global minima in the left panel}
+\label{fig:pitchfork:results}
+\end{figure}
+
+\subsection{Genetic Toggle Switch | \textit{in progress}}
+In this section we optimise a model where the states share a Hill function relationship; these models often emerge from mass action kinetics and are used to model species concentrations.
+By not restricting the sign of the coefficients $a_1,a_2$ we can recover both activator and inhibitor relationships.
+\begin{equation}
+    \partial_t u_1 = \frac{\varepsilon_1 + a_1 u_2^2} { k_1^2 + u_2^2 } - \mu_1 u_1 \quad
+    \partial_t u_2 = \frac{p + a_2 u_1^2} { k_2^2 + u_1^2 } - \mu_2 u_2
+\end{equation}
+Figure \ref{fig:two-state-optima} shows a reduced dimensionality plot of the optimal parameter basins, represented as point clouds.
+$N$ distinct clusters are observed, each corresponding to a kinetically different model regions that are qualitatively equivalent.
+
+\section{Conclusion \& Broader Impact | \textit{in progress}}
 \begin{itemize}
-    \item Genetic toggle switch
-    \item FitzHugh-Nagumo
+    \item Taking steps towards an efficient and single method for searching and matching bifurcations
+    \item A demonstrating of the powerful combination of differential geometry and implicit layers
+    \item Future directions include extending this routine for Hopf bifurcations and PDEs
 \end{itemize}
-% \subsubsection{Genetic toggle switch}
-% \subsubsection{FitzHugh-Nagumo}
 
-% \begin{equation}
-%     \dfrac{du_1}{dt} = \dfrac{a_1} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
-%     \dfrac{du_2}{dt} = \dfrac{a_2} { 1 + (k u_1)^2 } - \mu_2 u_2
-% \end{equation}
-% Without loss of generality, we can rescale $u_1$ by $a_1$ and $u_2$ by $a_2$:
-% \begin{equation*}
-%     \dfrac{du_1}{dt} = \dfrac{1} { 1 + (\hat{p} u_2)^2 } - \mu_1 u_1, \quad
-%     \dfrac{du_2}{dt} = \dfrac{1} { 1 + (\hat{k} u_1)^2 } - \mu_2 u_2
-% \end{equation*}
-% where $\hat{p} = p a_2$ and $\hat{k} = k a_1$. The equilibria for this system are defined by
-% \begin{equation}
-%     \mu_1u_1^*(1+\hat{p}u_2^*)^2 = 1, \quad \mu_2u_2^*(1+\hat{k}u_1^*)^2 = 1
-% \end{equation}
-% Obtaining closed form expressions for $u_1^*$ and $u_2^*$ requires solving a quintic polynomial. We won't do that here.
-
-% What is the determinant for this system?
-% \begin{equation*}
-%     \left|\dfrac{\partial F}{\partial u}\right| = 
-%     \begin{vmatrix} -\mu_1 & -\frac{2 p^2 u_2^*}{(1 + (p u_2^*)^2)^2} \\
-%         -\frac{2k^2u_1^*}{(1+(k u_1^*)^2)^2} & -\mu_2
-%     \end{vmatrix}
-%     = \mu_1\mu_2 - \frac{4p^2k^2u_1^*u_2^*}{(1+(p u_2^*)^2)^2(1+(k u_1^*)^2)^2}
-% \end{equation*}
-
-% \subsubsection{Two-state (generalised)}
-% \begin{equation}
-%     \dfrac{du_1}{dt} = \dfrac{a_1 + b_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
-%     \dfrac{du_2}{dt} = \dfrac{a_2 + b_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
-% \end{equation}
-
-% Applying the same rescaling as above, we can simplify as
-% \begin{equation}
-%     \dfrac{du_1}{dt} = \dfrac{1 + \rho_1(p u_2)^2} { 1 + (p u_2)^2 } - \mu_1 u_1, \quad
-%     \dfrac{du_2}{dt} = \dfrac{1 + \rho_2(k u_1)^2} { 1 + (k u_1)^2 } - \mu_2 u_2
-% \end{equation}
-% where $\rho_i = \frac{b_i}{a_i}$, with other parameters rescaled as above (but dropping hats).
-
-\subsection{Computational Complexity}
-
-% \section{Discussion}
-% \label{sec:discussion}
-
-% \subsection{Model Synthesis}
-% \subsection{Model Selection}
-
-\appendix
-\appendixpage
-\addappheadtotoc
-\input{appendix}
+% \section*{Checklist}
+% \input{checklist}
 
@@
 \bibliographystyle{ieeetr}
+
+\section*{Appendix}
+\appendix
+\input{appendix}
+
 \end{document}

### 2021-05-14 docs/appendix.tex
updated text

@@
 \label{fig:implicit-surfaces}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=5cm]{implicit-surfaces}
-\caption{Curvature of the determinant $\curvature$ diverging as saddle-node bifurcations emerge from a smooth steady state manifold}
-\label{fig:diverging-curvature}
 \end{figure}
@@
 \end{align}
-and somehow this leads to numerically verified result%\todo{perhaps not valid in general?}
+this leads to numerically verified result%\todo{perhaps not valid in general?}
 \begin{align}
@@
 
+\section{Bifurcation Measure}
+\label{appendix:measure}
+
 \section{Simplification of two-state model}

### 2021-05-14 docs/article.tex

@@
     \partial_t u_2 = \frac{p + a_2 u_1^2} { k_2^2 + u_1^2 } - \mu_2 u_2
+    \label{eq:two-state}
 \end{equation}
 Figure \ref{fig:two-state-optima} shows a reduced dimensionality plot of the optimal parameter basins, represented as point clouds.
-$N$ distinct clusters are observed, each corresponding to a kinetically different model regions that are qualitatively equivalent.
+14 distinct clusters are observed, each corresponding to a kinetically different model regions that are qualitatively equivalent.
+
+\begin{figure}
+\centering
+\includegraphics[width=11cm]{two-state-optima}
+\caption{Reduced dimensionality plot of optimal $\theta^*$ for the two state model \eqref{eq:two-state} for $\targets=\{4,5\}$}
+\label{fig:two-state-optima}
+\end{figure}
 

### 2021-05-14 test/applied/two-state.jl
log parametrisation of two-state model

@@ -2,17 +2,18 @@
 function rates(u::AbstractVector,parameters::NamedTuple)
 
 	@unpack θ,p = parameters
-	μ₁,μ₂, a₁,a₂, k₁,k₂, ϵ₁ = θ
+	μ₁,μ₂, a₁,a₂, k = θ
 
 	f = first(u)*first(p)*first(θ)
 	F = similar(u,typeof(f))
 
-	F[1] = ( ϵ₁ + a₁*u[2]^2 ) / ( k₁^2 + u[2]^2 ) - μ₁*u[1]
-	F[2] = ( p  + a₂*u[1]^2 ) / ( k₂^2 + u[1]^2 ) - μ₂*u[2]
+	F[1] = ( 10^a₁ + (p*u[2])^2 ) / ( 1 + (p*u[2])^2 ) - u[1]*10^μ₁
+	F[2] = ( 10^a₂ + (k*u[1])^2 ) / ( 1 + (k*u[1])^2 ) - u[2]*10^μ₂
 
 	return F
 end
 
 ######################################################### targets and initial guess
-targetData = StateSpace( 2, 0:0.01:8, [4,5] )
-θ = SizedVector{7}(4.0,4.0, 0.0,0.0, 1.0,1.0, 5.0)
+targetData = StateSpace( 2, 0:0.01:10, [4,5] )
+hyperparameters = getParameters(targetData)
+θ = SizedVector{5}(1.0,1.0, 1.0,1.0, 1.0)

### 2021-05-16 src/FluxContinuation.jl
unit test passed; explicit dp/dtheta in cost

@@ -4,7 +4,7 @@ module FluxContinuation
 	using BifurcationKit: BorderedArray, AbstractLinearSolver, AbstractEigenSolver, BorderingBLS
 	using BifurcationKit: ContState, solution, computeEigenvalues!, detectBifucation
 
-	using ForwardDiff
+	using ForwardDiff: gradient,jacobian
 	using Flux: Momentum,update!
 
 	using Setfield: @lens,@set,setproperties
@@ -13,19 +13,16 @@ module FluxContinuation
 	using InvertedIndices: Not
 	using LinearAlgebra, StaticArrays
 
-	using Plots.PlotMeasures
-	using LaTeXStrings
-	using Plots
-
 	include("Structures.jl")
 	include("Utils.jl")
 
 	include("Objectives.jl")
 	include("Gradients.jl")
+	include("Plots.jl")
 
+	export plot,@unpack,BorderedArray,SizedVector
 	export StateSpace,deflationContinuation,train!
-	export getParameters,loss,∇loss
-	export plot,@unpack
+	export getParameters,loss,∇loss,norm
 
 	""" root finding with newton deflation method"""
 	function findRoots!( f::Function, J::Function, roots::AbstractVector{<:AbstractVector},
@@ -84,7 +81,7 @@ module FluxContinuation
 		) where {T<:Number, V<:AbstractVector{T}, S<:AbstractLinearSolver, E<:AbstractEigenSolver}
 
 		maxIterContinuation,ds = hyperparameters.newtonOptions.maxIter,hyperparameters.ds
-		J(u,p) = ForwardDiff.jacobian(x->f(x,p),u)
+		J(u,p) = jacobian(x->f(x,p),u)
 
 		findRoots!( f, J, roots, parameters, hyperparameters; maxRoots=maxRoots, maxIter=maxIter, verbosity=verbosity)
 		pRange = range(hyperparameters.pMin,hyperparameters.pMax,length=length(roots))

### 2021-05-16 src/Gradients.jl

@@ -1,41 +1,24 @@
-################################################################################# loss gradient
+################################################################################
 function ∇loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
-
 	pmin,pmax = extrema(targets.parameter)
-	N,M = length(targets.targets), reduce(+, [ pmin ≤ s.z.p ≤ pmax for branch ∈ branches for s ∈ branch if s.bif ]; init=0)
+	predictions = sum([ s.bif/2 for branch ∈ branches for s ∈ branch if (pmin ≤ s.z.p ≤ pmax) ])
 
-	ω,∇ω = weight(F,branches,θ,targets;kwargs...) + eps(), ∇weight(F,branches,θ,targets;kwargs...)
-
-	L =  sum( p -> exp( errors(F,branches,θ,targets;p=p,kwargs...)/ω )                                                                                                          , targets.targets ) / N
-	∇L = sum( p -> exp( errors(F,branches,θ,targets;p=p,kwargs...)/ω ) * ( ∇errors(F,branches,θ,targets;p=p,kwargs...)*ω - errors(F,branches,θ,targets;p=p,kwargs...)*∇ω ) / ω^2, targets.targets ) / N
-
-	K = curvature(F,branches,θ,targets;kwargs...)
-	∇K = ∇curvature(F,branches,θ,targets;kwargs...)
-
-	return L - (2N-M)*log(K), ∇L - (2N-M)*∇K/K
+	Φ,∇Φ = measure(F,branches,θ,targets), ∇measure(F,branches,θ,targets)
+	return errors(branches,targets) - (length(targets.targets)-predictions)*log(Φ), ∇errors(F,branches,θ,targets) - (length(targets.targets)-predictions)*∇Φ/Φ
 end
 
 ################################################################################
-struct Gradient <: Function
-	f::Function
-	integrand::Integrand
+function ∇errors( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace)
+	pmin,pmax = extrema(targets.parameter)
+
+	predictions = [ s.z for branch ∈ branches for s ∈ branch if s.bif & (pmin ≤ s.z.p ≤ pmax) ]
[…]

### 2021-05-16 src/Objectives.jl

@@ -1,29 +1,20 @@
 ################################################################################
 function loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
-
 	pmin,pmax = extrema(targets.parameter)
-	N,M = length(targets.targets), reduce(+, [ pmin ≤ s.z.p ≤ pmax for branch ∈ branches for s ∈ branch if s.bif ]; init=0)
 
-	ω = weight(F,branches,θ,targets;kwargs...) + eps()
-	L = sum( p -> exp( errors(F,branches,θ,targets;p=p,kwargs...)/ω ), targets.targets ) / N
-	K = curvature(F,branches,θ,targets;kwargs...)
-
-	return L - (2N-M)*log(K)
+	predictions = sum([ s.bif/2 for branch ∈ branches for s ∈ branch if (pmin ≤ s.z.p ≤ pmax) ])
+	return errors(branches,targets) - (length(targets.targets)-predictions)*log(measure(F,branches,θ,targets))
 end
 
 ################################################################################
-struct Integrand <: Function f::Function end
-(f::Integrand)(args...;kwargs...) = f.f(args...;kwargs...)
+function errors( branches::AbstractVector{<:Branch}, targets::StateSpace)
+	pmin,pmax = extrema(targets.parameter)
 
-errors = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; p=0.0, kwargs...)
-	return log(abs(z.p-p)) * weight(F,z,θ,targets; kwargs...)
-end)
+	predictions = [ s.z for branch ∈ branches for s ∈ branch if s.bif & (pmin ≤ s.z.p ≤ pmax) ]
+	return mean( p′-> mean( z->(z.p-p′)^2, predictions; type=:geometric ), targets.targets; type=:arithmetic )
+end
 
-weight = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; α::Real=1e3, kwargs... )
-	return exp( -α*det(F,z,θ)^2 )
-end)
-
-curvature = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; kwargs... )
+measure = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; kwargs... )
[…]

### 2021-05-16 src/Plots.jl

@@ -1,0 +1,55 @@
+using Plots.PlotMeasures
+using LaTeXStrings
+using Plots
+
+import Plots: plot
+function plot(steady_states::Vector{<:Branch}; padding=0.2, displayPlot=true)
+
+	pMin,pMax = extrema(vcat([ map(s->s.z.p,branch) for branch ∈ steady_states ]...))
+	pMin,pMax = ( pMin+(pMin+pMax)*padding )/(1+2padding),( pMax+(pMin+pMax)*padding )/(1+2padding)
+	unpadded_parameter = pMin:pMax
+
+	plot(xlabel=L"\mathrm{parameter,}p", grid=false, right_margin=20mm, size=(500,450) )
+	right_axis = twinx()
+
+    for branch ∈ steady_states
+
+        stability = map( s -> all(real(s.λ).<0), branch)
+        determinants = map( s -> prod(real(s.λ)), branch)
+		parameter = map( s -> s.z.p, branch)
+
+		for idx ∈ 1:dim(branch)
+
+			plot!( parameter, map(s->s.z.u[idx],branch), linewidth=2, label="",
+				ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(u,p)=0", grid=false,
+				alpha=map( s->window_function(unpadded_parameter,s.z), branch ),
+				color=map( stable -> stable ? :darkblue : :lightblue, stability )
+			)
+		end
+
+		plot!(right_axis, parameter, -determinants, linewidth=2, label="", grid=false,
+        	ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|",
+			alpha=map( s->window_function(unpadded_parameter,s.z), branch ),
+            color=map( stable -> stable ? :red : :pink, stability )
+		)
+    end
+
+	if displayPlot plot!() |> display
+	else return right_axis end
+end
+
+function plot(steady_states::Vector{<:Branch}, data::StateSpace)
+	right_axis = plot(steady_states; displayPlot=false)
+	vline!( data.targets, label="", color=:gold)
+	plot!( right_axis,[],[], color=:gold, legend=:bottomleft, alpha=1.0, label="") |> display
+end
+
+function plot(F::Function, θ::AbstractVector, data::StateSpace; kwargs...)
+
+	parameters = (θ=θ,p=minimum(data.parameter))
+	hyperparameters = getParameters(data;kwargs...)
+
[…]

### 2021-05-16 src/Structures.jl

@@ -51,8 +51,20 @@ function push!(branch::Branch,state::ContState)
     push!(branch, ( z=z, λ=state.eigvals, ds=norm(∂z)*abs(state.ds), bif=detectBifucation(state) ))
 end
 
-# display methods
+import Base: zero ################################## initialiser for BorderedArray
+zero(::Type{BorderedArray{T,U}}; ϵ::Bool=true) where {T<:Union{Number,StaticArray},U<:Union{Number,StaticArray}} = BorderedArray(zero(T).+ϵ*eps(),zero(U).+ϵ*eps())
+
+#################################################### display methods
 show(io::IO, branches::Vector{Branch{V,T}}) where {V,T} = print(io,
     "Vector{Branch}[dim=$(dim(first(branches))) bifurcations=$(sum(branch->sum(s->s.bif,branch),branches)) branches=$(length(branches)), states=$(sum(branch->length(branch),branches))]")
 show(io::IO, M::MIME"text/plain", branches::Vector{Branch{V,T}}) where {V,T} = show(io,branches)
 show(io::IO, states::StateSpace{N,T}) where {N,T} = print(io,"StateSpace{$N,$T}(parameters=$(states.parameter),targets=$(states.targets))")
+
+################################################################################
+########################################################### line integrand terms
+
+struct Integrand <: Function f::Function end
+(f::Integrand)(args...;kwargs...) = f.f(args...;kwargs...)
+
+struct Gradient <: Function f::Function; integrand::Integrand end
+(f::Gradient)(args...;kwargs...) = f.f(args...;kwargs...)

### 2021-05-16 src/Utils.jl

@@ -23,6 +23,10 @@ function updateParameters!(parameters::ContinuationPar{T, S, E}, steady_states::
     parameters = setproperties(parameters;ds=ds,dsmin=ds,dsmax=ds)
 end
 
+function mean(f::Function,x::AbstractArray{T}; type::Symbol=:geometric, kwargs...) where T
+	return type == :geometric ? mapreduce(y->f(y)^(1/length(x)),*,x;init=1,kwargs...) : mapreduce(y->f(y)/length(x),+,x;init=zero(f(zero(T))),kwargs...)
+end
+
 ############################################################################# training loop
 function train!( F::Function, parameters::NamedTuple, data::StateSpace;
 				 iter::Int=200, optimiser=Momentum(0.001), plot_solution = false, kwargs...)
@@ -82,57 +86,4 @@ end
 
 function ∇loss(F::Function, θ::AbstractVector, data::StateSpace; kwargs...)
 	return ∇loss(F,θ,data,getParameters(data); kwargs...)
-end
-
-############################################################################# plotting
-import Plots: plot
-function plot(steady_states::Vector{<:Branch}; padding=0.2, displayPlot=true)
-
-	pMin,pMax = extrema(vcat([ map(s->s.z.p,branch) for branch ∈ steady_states ]...))
-	pMin,pMax = ( pMin+(pMin+pMax)*padding )/(1+2padding),( pMax+(pMin+pMax)*padding )/(1+2padding)
-	unpadded_parameter = pMin:pMax
-
-	plot(xlabel=L"\mathrm{parameter,}p", grid=false, right_margin=20mm, size=(500,450) )
-	right_axis = twinx()
-
-    for branch ∈ steady_states
-
-        stability = map( s -> all(real(s.λ).<0), branch)
-        determinants = map( s -> prod(real(s.λ)), branch)
-		parameter = map( s -> s.z.p, branch)
-
-		for idx ∈ 1:dim(branch)
-
-			plot!( parameter, map(s->s.z.u[idx],branch), linewidth=2, label="",
-				ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(u,p)=0", grid=false,
-				alpha=map( s->window_function(unpadded_parameter,s.z), branch ),
-				color=map( stable -> stable ? :darkblue : :lightblue, stability )
-			)
-		end
-
-		plot!(right_axis, parameter, -determinants, linewidth=2, label="", grid=false,
[…]

### 2021-05-16 test/applied/two-state.jl

@@ -1,5 +1,6 @@
 ######################################################## model
-function rates(u::AbstractVector,parameters::NamedTuple)
+F(z::BorderedArray,θ::AbstractVector) = F(z.u,(θ=θ,p=z.p))
+function F(u::AbstractVector,parameters::NamedTuple)
 
 	@unpack θ,p = parameters
 	μ₁,μ₂, a₁,a₂, k = θ
@@ -14,6 +15,5 @@ function rates(u::AbstractVector,parameters::NamedTuple)
 end
 
 ######################################################### targets and initial guess
-targetData = StateSpace( 2, 0:0.01:10, [4,5] )
-hyperparameters = getParameters(targetData)
+X = StateSpace( 2, 0:0.01:10, [4,5] )
 θ = SizedVector{5}(1.0,1.0, 1.0,1.0, 1.0)

### 2021-05-16 test/minimal/pitchfork.jl

@@ -1,5 +1,6 @@
 ######################################################## model
-function rates(u::AbstractVector,parameters::NamedTuple)
+F(z::BorderedArray,θ::AbstractVector) = F(z.u,(θ=θ,p=z.p))
+function F(u::AbstractVector,parameters::NamedTuple)
 	@unpack θ,p = parameters
 
 	f = first(u)*first(p)*first(θ)
@@ -12,6 +13,5 @@ function rates(u::AbstractVector,parameters::NamedTuple)
 end
 
 ######################################################### targets and initial guess
-targetData = StateSpace( 2, -5:0.01:5, [0] )
-hyperparameters = getParameters(targetData)
+X = StateSpace( 2, -5:0.01:5, [0] )
 θ = SizedVector{2}(4.0,-π/2)

### 2021-05-16 test/minimal/saddle-node.jl

@@ -1,5 +1,6 @@
 ######################################################## model
-function rates(u::AbstractVector,parameters::NamedTuple)
+F(z::BorderedArray,θ::AbstractVector) = F(z.u,(θ=θ,p=z.p))
+function F(u::AbstractVector,parameters::NamedTuple)
 	@unpack θ,p = parameters
 
 	f = first(u)*first(p)*first(θ)
@@ -12,6 +13,5 @@ function rates(u::AbstractVector,parameters::NamedTuple)
 end
 
 ######################################################### targets and initial guess
-targetData = StateSpace( 2, -2:0.01:2, [1,-1] )
-hyperparameters = getParameters(targetData)
+X = StateSpace( 2, -2:0.01:2, [1,-1] )
 θ = SizedVector{2}(5.0,-0.93)

### 2021-05-16 test/runtests.jl

@@ -1,18 +1,15 @@
-using FluxContinuation, StaticArrays
-using FiniteDifferences, Test
-using LinearAlgebra: norm
+using FiniteDiff: finite_difference_gradient
+using FluxContinuation, Test
 
-include("utils.jl")
-
-error_tolerance = 0.06
-@testset "Minimal Models" begin
+error_tolerance = 0.04
+@testset "Gradients" begin
 
 	@testset "Saddle Node" begin
 		include("minimal/saddle-node.jl")
 
 		@time for θ ∈ [[5.0,1.0],[5.0,-1.0],[-1.0,-1.0],[2.5,-1.0],[-1.0,2.5],[-4.0,1.0]]
-			x, y = autodiff(θ), finite_differences(θ)
-			@test x/norm(x) ≈ y/norm(y) rtol = error_tolerance
+			x, y = finite_difference_gradient(θ->loss(F,θ,X),θ), ∇loss(F,θ,X)[2]
+			@test acos(x'y/(norm(x)*norm(y)))/π < error_tolerance
 		end
 	end
 
@@ -20,8 +17,17 @@ error_tolerance = 0.06
 		include("minimal/pitchfork.jl")
 
 		@time for θ ∈ [[1.0,1.0],[3.0,3.0],[1.0,-1.0],[3.0,-3.0],[-1.0,-1.0],[-3.0,-3.0],[-1.0,1.0],[-3.0,3.0]]
-			x, y = autodiff(θ), finite_differences(θ)
-			@test x/norm(x) ≈ y/norm(y) rtol = error_tolerance
+			x, y = finite_difference_gradient(θ->loss(F,θ,X),θ), ∇loss(F,θ,X)[2]
+			@test acos(x'y/(norm(x)*norm(y)))/π < error_tolerance
+		end
+	end
+
+	@testset "Two State" begin
+		include("applied/two-state.jl")
+
+		@time for θ ∈ [[0.5, 0.5, 0.5470, 2.0, 7.5],[0.5, 0.5, 1.5470, 2.0, 7.5],[0.5, 0.88, 1.5470, 2.0, 7.5]]
+			x, y = finite_difference_gradient(θ->loss(F,θ,X),θ), ∇loss(F,θ,X)[2]
+			@test acos(x'y/(norm(x)*norm(y)))/π < error_tolerance
 		end
 	end
 end

### 2021-05-16 test/utils.jl

@@ -1,10 +1,0 @@
-function finite_differences(θ::AbstractVector{<:Number}; order=5, geom=true, condition=100)
-	loss(rates,θ,targetData,hyperparameters)
-	return first(grad(central_fdm(order,1,geom=geom,condition=condition),
-		θ -> loss(rates,θ,targetData,hyperparameters), θ ))
-end
-
-function autodiff(θ::AbstractVector{<:Number})
-	L,∇L = ∇loss(rates,θ,targetData,hyperparameters)
-	return ∇L
-end

### 2021-05-17 docs/article.tex

@@
 \end{equation}
-Figure \ref{fig:two-state-optima} shows a reduced dimensionality plot of the optimal parameter basins, represented as point clouds.
-14 distinct clusters are observed, each corresponding to a kinetically different model regions that are qualitatively equivalent.
+Figure \ref{fig:two-state-optima} shows a UMAP embedding of the optimal parameter basins, represented as point clouds.
+16 distinct clusters are obtained using DBSCAN.
+Each cluster corresponds to a kinetically different model regions that are qualitatively equivalent.
+See Appendix \ref{appendix:clusters} for models that represent the centroid of each cluster.
 
@@
     \item Future directions include extending this routine for Hopf bifurcations and PDEs
+    \item Computational limitations
 \end{itemize}

### 2021-05-17 docs/appendix.tex
Updates from Overleaf

@@
 
+\subsection{Bifurcation Curves as Tangent Fields}
+\label{appendix:tangent-fields}
+
+Let each component of the vector function $\rates$ in the model \eqref{eq:model} implicitly define a surface embedded in $\Reals^{N+1}$.
+Let's assume that the intersection of these $N$ surfaces exists and is not null or degenerate, then the steady states of \eqref{eq:model} must be a set of one dimensional space curves in $z\in\Reals^{N+1}$ defined by
+\begin{align}
+    \rates(z) = 0
+\end{align}
+An expression for the field $\tangent(z)$ tangent to the set of curves would allow us to take derivatives and integrals along the bifurcation curve.
+This is exactly what we need to do to evaluate our cost function \ref{eq:cost}.
+Fortunately the tangent field can be constructed by ensuring it is perpendicular to the gradient $\partial_z$ of each component of $\rates$ as illustrated by an example two component system in Figure \ref{fig:implicit-surfaces}.
+The tangent field $\tangent(z)$ can be constructed perpendicular to all gradient vectors using the properties of the determinant \cite{Goldman2005CurvatureSurfaces}
+\begin{align}
+    \tangent(z):=
+    \label{eq:tangent-field}
+    \left|\begin{matrix}
+        \hat{z} \\
+        \,\partial_{z}\rates\,
+    \end{matrix}\right|
+    \qquad\tangent : \Reals^{N+1}\rightarrow\Reals^{N+1}\\
+    =\sum_{i=1}^{N+1}\hat{z}_{i}(-1)^{i+1} \left|\frac{\partial \rates}{\partial(z\setminus z_{i}) }\right|
+\end{align}
+where $\hat{z}$ is a collection of unit basis vectors in the $\Reals^{N+1}$ space and $\partial_{z}\rates$ is an $N\times(N+1)$ rectangular Jacobian matrix of partial derivatives and $z\setminus z_{i}$ denotes the $N$ dimensional vector $z$ with component $z_{i}$ removed.
+This construction ensures perpendicularity to any gradients of $\rates$
+\begin{align}
+    \tangent(z)\cdot\partial_z f_{\theta} =
+    \left|\begin{matrix}
+        \partial_z f_{\theta} \\
+        \,\partial_{z}\rates\,
+    \end{matrix}\right|
+    \quad =0 \quad\forall f_{\theta}\in \rates
+\end{align}
+since the determinant of any matrix with two identical rows or columns is zero.
+Note that the tangent field $\tangent(z)$ is actually defined for all values of $z$ where adjacent field lines trace out other level sets where $\rates(z)\neq0$.
+Furthermore deformations with respect to $\theta$ are always orthogonal to the tangent
+\begin{align} % \todo{numerically true. analytic proof?}
+    \tangent(z)\cdot\frac{d\tangent}{d\theta}=0
+\end{align}
+\begin{figure}
+\centering
+\includegraphics[width=13cm]{determinant-field}
+\caption{Left/Right : Determinant $\Det$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\Det=0$ defines bifurcations}
+\label{fig:determinant-field}
+\end{figure}
+Figure \ref{fig:determinant-field} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many level sets or traces in tangent field $\tangent(z)$ for the saddle and pitchfork.
+The tangent field $\tangent(z)$ can always be analytically evaluated by taking the determinant in \eqref{eq:tangent-field}.
+We will proceed with calculations on $\tangent(z)$ in the whole space $z$ and pick out a single trace by solving $\rates(z)=0$ later.
+For our two models
+\begin{align}
+    \underset{\mathrm{saddle-node\,\,model}}{
+    \tangent(z)=\hat{u}-(\,3\theta_2 u^2+\theta_1\,)\,\hat{p}}
+    \qquad\qquad
+    \underset{\mathrm{pitchfork\,\,model}}{
+    \tangent(z)=u\hat{u}-(\,3\theta_2 u^2+p\,)\,\hat{p}}
+    \label{eq:tangent-field-examples}
+\end{align}
+Figure \ref{fig:determinant-field} reveals that $\Det=0$ is also a level set and that the intersection with level set $\rates(z)=0$ defines the bifurcations at specific parameter $\theta$.
+In this particular setting we can see that the tangent field $\tangent(z)$ only folds when $\Det=0$.
+Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-field} would give rise to Figures \ref{fig:saddle-node} and \ref{fig:pitchfork}.
+The directional derivative of the determinant $\Det$ along the tangent field $\tangent(z)$ is defined as
+\begin{align}
+    \frac{d}{ds}\Det := \hat{\tangent}(z) \cdot \frac{\partial}{\partial z}\Det 
+\end{align}
+where $\hat{\tangent}(z)$ is the unit tangent field.
+
 \section{Bifurcation Measure}

### 2021-05-17 docs/article.tex
Updates from Overleaf

@@
 \author{
-        Gregory Szep,\quad Attila Csik\'asz-Nagy\\King's College London\\London, WC2R 2LS\\
+        Gregory Szep\\King's College London\\London, WC2R 2LS\\
         \texttt{gregory.szep@kcl.ac.uk}
+    \And
+        Attila Csik\'asz-Nagy\\King's College London\\London, WC2R 2LS\\
+        \texttt{attila.csikasz-nagy@kcl.ac.uk}
     \And
@@
     In this work we propose a gradient-based semi-supervised approach for matching target bifurcations with parameterised differential equation models.
-The cost function contains a supervised term that is minimal when predicted bifurcations match the targets and an unsupervised bias that encourages bifurcations by maximising zero crossings in the determinant of the Jacobian.
+The cost function contains a supervised error term that is minimal when predicted bifurcations match the targets and an unsupervised bifurcation measure which acts as a proxy for the probability of a bifurcation, yielding non-zero gradients in non-bifurcating parameter regimes.
 The calculation of gradients with respect to parameters shares the same computational complexity as deflated pseudo-arclength continuation used to calculate the bifurcation diagram.
-We demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams, a genetic toggle switch from synthetic biology and the FitzHugh-Nagumo model.
+Our implementation leverages automatic differentiation methods in Julia and we demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams and the genetic toggle switch from synthetic biology.
 Furthermore, the cost landscape allows us to  organise models in terms of topological and geometric equivalence.
@@
 
-In order for predicted bifurcations $p\in\predictions(\theta)$ to match targets $p'\in\targets$ we need to evaluate some form of error term $|p-p'|$ on the bifurcation curve defined by $\rates(z)=0$.
-We expect the supervised term to be some mean over the targets $\targets$ and predictions $\predictions(\theta)$.
+In order for predicted bifurcations $p(\theta)\in\predictions$ to match targets $p'\in\targets$ we need to evaluate an error term $|p(\theta)-p'|$.
+We expect this term to be some mean over the targets $\targets$ and predictions $\predictions(\theta)$.
 We choose a geometric mean over the predictions and an arithmetic mean over targets to account for cases where $|\predictions|\neq|\targets|$ and discourage multiple predictions matching the same target in cases where $|\predictions|\geq|\targets|$.
-This term is equal to zero when all targets are matched by at least one prediction
+This term only zero when each target is matched by at least one prediction
 \begin{align}
     \frac{1}{|\targets|}\sum_{p'\in\targets}
-    \left(\prod_{p\in\predictions(\theta)}|p-p'|\right)
+    \prod_{p(\theta)\in\predictions}|p(\theta)-p'|
     ^{\frac{1}{|\predictions|}}
@@
 \end{align}
-We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions $\predictions(\theta)$ can be identified by looking for points $p$ along the curve where the determinant $\Det=0$.
+We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant $\Det=0$.
 Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$ in the non-degenerate case.
-Using these quantities we can define a positive semi-definite measure $\measure(z)$ of zero crossings in the determinant along the bifurcation curve; as shown in Figure \ref{fig:bifurcation-measure} this can be used as a proxy for the probability of a bifurcation at location $p$ for a given set of parameters $\theta$.
+Using these quantities we can define a positive semi-definite measure $\measure(z)$ of zero crossings in the determinant along the bifurcation curve; as shown in Figure \ref{fig:bifurcation-measure} this can be used as a proxy for the probability of a bifurcation at location $p(\theta)$.
 The total bifurcation measure $\Psi(\theta)$ is written
@@
 The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the observed parameter region $p$ is densely packed with bifurcations.
-In Appendix \ref{appendix:measure} we see that $\measure(z)$ is also invariant under certain classes of transformations on $\rates(z)$ which do not change bifurcation locations.
+The $\measure(z)$ is also invariant under certain classes of transformations on $\rates(z)$ which do not change bifurcation locations, proofs of which are provided in Appendix \ref{appendix:measure}.
 The measure is added to the supervised term as if it were a likelihood; this defines the semi-supervised cost function as
 \begin{align}
-    \loss(\theta|\targets):=\frac{1}{|\targets|}\sum_{p'\in\targets}
-    \left(\prod_{p\in\predictions(\theta)}|p-p'|\right)
-    ^{\frac{1}{|\predictions|}}\!\!\!
-    -\big(|\targets|-|\predictions|\big)\,\,\lambda \log\Psi(\theta)
+    \loss(\theta|\targets):=
+    \big(|\predictions|-|\targets|\big)\,\,\lambda \log\Psi(\theta)+
+    \frac{1}{|\targets|}\sum_{p'\in\targets}
+    \prod_{p(\theta)\in\predictions}|p(\theta)-p'|
+    ^{\frac{1}{|\predictions|}}
     \label{eq:cost}
@@
 This can be seen as a step-wise annealing of the unsupervised term until the desired state is reached.
-Note that in principle while $\predictions(\theta)$ depends smoothly on $\theta$ the number of predictions $|\predictions|$ does not.
-The gradient $\partial_{\theta}|\predictions|=0$ everywhere and is an infinite impulse at locations where the number of predictions changes.
-In practice optimisers are unlikely to reach those locations and we can safely drop the dependency on $\theta$.
-The supervised term must be re-written in such a way so that we may explicitly differentiate the cost with respect to $\theta$.
-This can be done in the following way
+Note that in principle while individual bifurcations $p(\theta)$ depend smoothly on $\theta$ the total number of predictions $|\predictions|$ does not.
+The gradient $\partial_{\theta}|\predictions|=0$ everywhere and is an infinite impulse at the exact locations where the number of predictions changes.
+
+In practice optimisers are unlikely to hit those values of $\theta$ and we can safely drop the dependency.
+We may now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ and total measure $\Psi(\theta)$
 \begin{align}
-    \frac{1}{|\targets|}\sum_{p'\in\targets}
-    \left(\prod_{p\in\predictions(\theta)}|p-p'|\right)
-    ^{\frac{1}{|\predictions|}}\!\!=\,\,
-    \frac{1}{|\targets|}\sum_{p'\in\targets}
-    \exp\left[
-    \frac{1}{|\predictions|}\int_{\rates(z)=0\,\Det=0}
-    \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
-    \mathrm{d}z\quad
-    \log|p-p'|
-    \right]
-    \label{eq:supervised-rewrite}
+    \frac{\partial\loss}{\partial\theta}=
+    \big(|\predictions|-|\targets|\big)\,\lambda\,
+    \frac{\partial \Psi}{\partial\theta}\Psi(\theta)^{-1}+
+    \frac{1}{|\targets||\predictions|}\sum_{p'}
+    \prod_{p(\theta)}|p(\theta)-p'|^{\frac{1}{|\predictions|}}
+    \sum_{p(\theta)}\frac{\partial p}{\partial\theta}\left(p(\theta)-p'\right)^{-1}
+    \label{eq:gradient}
 \end{align}
-In effect we have replaced the set representation $\predictions(\theta)$ with a line integral over the bifurcation curve $\rates(z)=0$ subject to constraint $\Det=0$.
-In Figure \ref{fig:determinant-field} we can indeed see that bifurcations $\predictions(\theta)$ are picked out by intersections between the two level sets.
-Now that the cost function is written in terms of line integrals $\int_{\rates(z)=0}\mathrm{d}z$ we need to know how to differentiate them with respect to $\theta$, differentiate along them as in equation \eqref{eq:measure} and know how to deal with constraint $\Det=0$.
-The following section shows how this can be done with a vector field representation of the bifurcation curve.
+In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to calculate the gradient $\frac{\partial\loss}{\partial\theta}$ without having to differentiate through the operations of the solver that finds the bifurcation curve $\rates(z)=0$ and the bifurcation locations $p(\theta)$.
+To calculate the gradient of the measure $\frac{\partial \Psi}{\partial\theta}$ we need to differentiate line integrals that depend on $\theta$.
+Fortunately this can be done by the application of the generalised Leibniz integral rule, details of which can be found in Appendix \ref{appendix:space-curve}.
 
-\subsection{Bifurcation Curves as Tangent Fields}
-\label{section:tangent-fields}
+The gradient of the bifurcation points $\frac{\partial p}{\partial\theta}$ is found by application of the implicit function theorem to a vector function $G_{\theta}:\Reals^{N+1}\rightarrow\Reals^{N+1}$ whos components represent the two constraints $\rates(z)=0$ and $\Det=0$.
+By following a similar strategy to that used by implicit layers \cite{Look2020DifferentiableLayers} we yield an $(N+1)\times M$ Jacobian representing a deformation field \cite{Jos2011OnSurface} for each $\theta$ direction.
+The gradient we are looking for becomes
+\begin{align}
+    \frac{\partial p}{\partial\theta} = -\hat{p}\cdot\left.\frac{\partial G_{\theta}}{\partial z}^{-1}
+    \frac{\partial  G_{\theta}}{\partial\theta}\right|_{G_{\theta}(z)=0}
+    \quad\mathrm{where}\quad
+    G_{\theta}(z):=\begin{bmatrix}\rates(z)\\\Det\end{bmatrix}
+\end{align}
+Here $\hat{p}$ is a unit vector in $z\in\Reals^{N+1}$ that picks out the deformations along the $p$-direction.
+We can see this calculation involves inverting an $(N+1)\times(N+1)$ Jacobian $\frac{\partial G_{\theta}}{\partial z}$.
+The determinant of this Jacobian goes to zero when $\frac{d}{ds}\Det=0$.
+This further justifies our choice for measure $\measure(z)$ which discourages such cases.
 
-Let each component of the vector function $\rates$ in the model \eqref{eq:model} implicitly define a surface embedded in $\Reals^{N+1}$.
-Let's assume that the intersection of these $N$ surfaces exists and is not null or degenerate, then the steady states of \eqref{eq:model} must be a set of one dimensional space curves in $z\in\Reals^{N+1}$ defined by
-\begin{align}
-    \rates(z) = 0
-\end{align}
-An expression for the field $\tangent(z)$ tangent to the set of curves would allow us to take derivatives and integrals along the bifurcation curve.
-This is exactly what we need to do to evaluate our cost function \ref{eq:cost}.
-Fortunately the tangent field can be constructed by ensuring it is perpendicular to the gradient $\partial_z$ of each component of $\rates$ as illustrated by an example two component system in Figure \ref{fig:implicit-surfaces}.
-The tangent field $\tangent(z)$ can be constructed perpendicular to all gradient vectors using the properties of the determinant \cite{Goldman2005CurvatureSurfaces}
-\begin{align}
-    \tangent(z):=
-    \label{eq:tangent-field}
-    \left|\begin{matrix}
-        \hat{z} \\
-        \,\partial_{z}\rates\,
-    \end{matrix}\right|
-    \qquad\tangent : \Reals^{N+1}\rightarrow\Reals^{N+1}\\
-    =\sum_{i=1}^{N+1}\hat{z}_{i}(-1)^{i+1} \left|\frac{\partial \rates}{\partial(z\setminus z_{i}) }\right|
-\end{align}
-where $\hat{z}$ is a collection of unit basis vectors in the $\Reals^{N+1}$ space and $\partial_{z}\rates$ is an $N\times(N+1)$ rectangular Jacobian matrix of partial derivatives and $z\setminus z_{i}$ denotes the $N$ dimensional vector $z$ with component $z_{i}$ removed.
-This construction ensures perpendicularity to any gradients of $\rates$
-\begin{align}
-    \tangent(z)\cdot\partial_z f_{\theta} =
-    \left|\begin{matrix}
-        \partial_z f_{\theta} \\
-        \,\partial_{z}\rates\,
-    \end{matrix}\right|
-    \quad =0 \quad\forall f_{\theta}\in \rates
-\end{align}
-since the determinant of any matrix with two identical rows or columns is zero.
-Note that the tangent field $\tangent(z)$ is actually defined for all values of $z$ where adjacent field lines trace out other level sets where $\rates(z)\neq0$.
-Furthermore deformations with respect to $\theta$ are always orthogonal to the tangent
-\begin{align} % \todo{numerically true. analytic proof?}
-    \tangent(z)\cdot\frac{d\tangent}{d\theta}=0
-\end{align}
-\begin{figure}
-\centering
-\includegraphics[width=13cm]{determinant-field}
-\caption{Left/Right : Determinant $\Det$ and tangent field $\tangent(z)$ for the saddle-node/pitchfork models for some set values of $\theta$ revealing that $\Det=0$ defines bifurcations}
-\label{fig:determinant-field}
-\end{figure}
-Figure \ref{fig:determinant-field} shows how the bifurcation curve defined by $\rates(z)=0$ picks out one of many level sets or traces in tangent field $\tangent(z)$ for the saddle and pitchfork.
-The tangent field $\tangent(z)$ can always be analytically evaluated by taking the determinant in \eqref{eq:tangent-field}.
-We will proceed with calculations on $\tangent(z)$ in the whole space $z$ and pick out a single trace by solving $\rates(z)=0$ later.
-For our two models
-\begin{align}
-    \underset{\mathrm{saddle-node\,\,model}}{
-    \tangent(z)=\hat{u}-(\,3\theta_2 u^2+\theta_1\,)\,\hat{p}}
-    \qquad\qquad
-    \underset{\mathrm{pitchfork\,\,model}}{
-    \tangent(z)=u\hat{u}-(\,3\theta_2 u^2+p\,)\,\hat{p}}
-    \label{eq:tangent-field-examples}
-\end{align}
-Figure \ref{fig:determinant-field} reveals that $\Det=0$ is also a level set and that the intersection with level set $\rates(z)=0$ defines the bifurcations at specific parameter $\theta$.
-In this particular setting we can see that the tangent field $\tangent(z)$ only folds when $\Det=0$.
-Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-field} would give rise to Figures \ref{fig:saddle-node} and \ref{fig:pitchfork}.
-The directional derivative of the determinant $\Det$ along the tangent field $\tangent(z)$ is defined as
-\begin{align}
-    \frac{d}{ds}\Det := \hat{\tangent}(z) \cdot \frac{\partial}{\partial z}\Det 
-\end{align}
-where $\hat{\tangent}(z)$ is the unit tangent field.
-Finally, by applying the generalised Leibniz integral rule, it is also possible to use $\tangent(z)$ to evaluate gradients of line integrals  $\int_{\rates(z)=0}\mathrm{d}z$.
-In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to differentiate the cost $\loss(\theta|\targets)$ without having to differentiate through the operations of the solver that finds $\rates(z)=0$.
-Fortunately this is all possible if we apply a similar strategy to that used by implicit layers \cite{Look2020DifferentiableLayers}; see Appendix \ref{appendix:space-curve} and \ref{appendix:deformation} for derivations.
-
-% While the unsupervised integrand has an smooth analytic expression, such as those evaluated in \eqref{eq:curvature-examples}, the constraint $\Det=0$ in the first integral makes differentiation with respect to $\theta$ more tricky.
-Here it becomes convenient to re-express the constraint as a delta function $\delta(z)$ and applying Lebesgue's dominated convergence theorem yielding
-% \begin{align}
-%     \lim_{n\rightarrow\infty}\int_{\steadystates}
-%     \delta_{n}\left(\Det\right)
-%     \frac{1}{|\targets|}\sum_{p'\in\targets} |p-p'|^2\, \mathrm{d}u\mathrm{d}p
-% \end{align}
-% where $\delta_{n}(z)$ can be any basis function whos limit as $n\rightarrow\infty$ is a delta function $\delta(z)$.
-In practice one could choose the normal distribution function with sufficiently small variance.
-
-% The cost function is thus evaluated over the implicitly specified region $\steadystates$ --- in most cases the region will require a numerical procedure to find the roots $\rates(z)=0$ in the region $p\in\Omega$.
+The cost function is piece-wise smooth and differentiable with undefined gradients only in parameter regions where the number of predictions $|\predictions|$ changes.
+Given a set of solutions to $\rates(z)=0$ and locations $p(\theta)$ which additionally satisfy $\Det=0$, the gradient $\frac{\partial\loss}{\partial\theta}$ can evaluated using automatic differentiation methods.
 

### 2021-05-17 docs/article.tex

@@
     \centering
-    \includegraphics[width=10cm]{bifurcation-measure}
-    \caption{Bifurcation measure $\measure(z)$ shown as a gold thickness envelope.
-The thicker the envelope the larger the value of the measure at that location in the curve.
-Three curves are shown for three different values $\theta$, demonstrating how maximising $\measure(z)$ along the curve encourages new bifurcations}
+    \includegraphics[width=7cm]{bifurcation-measure}
+    \caption{Bifurcation measure $\measure(z)$ and determinant $\Det$ for two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
     \label{fig:bifurcation-measure}

### 2021-05-17 docs/article.tex

@@
 \centering
-\includegraphics[width=11cm]{saddle-node}
-\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ with $\theta=(5/2,-1)$ and targets $\targets=\{-1,1\}$.
+\includegraphics[width=6cm]{saddle-node}
+\includegraphics[width=6cm]{pitchfork}
+\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(5/2,-1),(1/2,-1)$ and targets $\targets=\{-1,1\},\{0\}$ respectively.
 Determinant in red.
 Lighter shades indicate the determinant crossing zero for unstable solutions}
-\label{fig:saddle-node}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=11cm]{pitchfork}
-\caption{Pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(1/2,-1)$ and target $\targets=\{0\}$.
-Determinant in red.
-Lighter shades indicate the determinant crossing zero for unstable solutions}
-\label{fig:pitchfork}
+\label{fig:minimal-models}
 \end{figure}

### 2021-05-17 src/Plots.jl
removing dummy dimensions

@@ -3,14 +3,18 @@ using LaTeXStrings
 using Plots
 
 import Plots: plot
-function plot(steady_states::Vector{<:Branch}; padding=0.2, displayPlot=true)
+function plot(steady_states::Vector{<:Branch},data::StateSpace)
 
-	pMin,pMax = extrema(vcat([ map(s->s.z.p,branch) for branch ∈ steady_states ]...))
-	pMin,pMax = ( pMin+(pMin+pMax)*padding )/(1+2padding),( pMax+(pMin+pMax)*padding )/(1+2padding)
-	unpadded_parameter = pMin:pMax
+	layout = @layout [a;b{1.0w,0.5h}]
+	default(); default(grid=false,label="",margin=1mm,linewidth=2)
+	figure = plot(layout = layout, link = :x, size=(300,500) )
 
-	plot(xlabel=L"\mathrm{parameter,}p", grid=false, right_margin=20mm, size=(500,450) )
-	right_axis = twinx()
+	hline!([0],subplot=1,linewidth=0,color=:black, ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(z)=0")
+	hline!([0],subplot=2,linewidth=1,color=:black, xlabel=L"\mathrm{parameter,}p", xmirror=true, topmargin=-5mm,
+		ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|")
+	
+	vline!(data.targets,subplot=1,linewidth=1,color=:gold)
+	vline!(data.targets,subplot=2,linewidth=1,color=:gold)
 
     for branch ∈ steady_states
 
@@ -20,28 +24,17 @@ function plot(steady_states::Vector{<:Branch}; padding=0.2, displayPlot=true)
 
 		for idx ∈ 1:dim(branch)
 
-			plot!( parameter, map(s->s.z.u[idx],branch), linewidth=2, label="",
-				ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(u,p)=0", grid=false,
-				alpha=map( s->window_function(unpadded_parameter,s.z), branch ),
-				color=map( stable -> stable ? :darkblue : :lightblue, stability )
-			)
+			plot!( parameter, map(s->s.z.u[idx],branch), subplot=1,
+				color=map( stable -> stable ? :darkblue : :lightblue, stability ) )
 		end
 
-		plot!(right_axis, parameter, -determinants, linewidth=2, label="", grid=false,
[…]

### 2021-05-17 src/Utils.jl

@@ -78,10 +78,15 @@ function loss(F::Function, θ::AbstractVector, data::StateSpace; kwargs...)
 end
 
 function ∇loss(F::Function, θ::AbstractVector, data::StateSpace, hyperparameters::ContinuationPar; kwargs...)
-	parameters = (θ=θ,p=minimum(data.parameter))
+	try
+		parameters = (θ=θ,p=minimum(data.parameter))
+		steady_states = deflationContinuation(F,data.roots,parameters,hyperparameters;kwargs...)
+		return ∇loss(F,steady_states,θ,data;kwargs...)
 
-	steady_states = deflationContinuation(F,data.roots,parameters,hyperparameters;kwargs...)
-	return ∇loss(F,steady_states,θ,data;kwargs...)
+	catch error
+		printstyled(color=:red,"$error\n")
+		return NaN,fill(NaN,length(θ))
+	end
 end
 
 function ∇loss(F::Function, θ::AbstractVector, data::StateSpace; kwargs...)

### 2021-05-17 test/applied/two-state.jl

@@ -16,4 +16,4 @@ end
 
 ######################################################### targets and initial guess
 X = StateSpace( 2, 0:0.01:10, [4,5] )
-θ = SizedVector{5}(1.0,1.0, 1.0,1.0, 1.0)
+θ = SizedVector{5}(0.5,0.5,0.5470,2.0,7.5)

### 2021-05-17 test/figures.jl

@@ -1,32 +1,105 @@
+using FluxContinuation
+using ForwardDiff
+
 using LaTeXStrings,Plots
 using Plots.Measures
-using ForwardDiff
+
+####################################################################################################
+####################################################################################################
+######################################################################################## saddle-node
+include("minimal/saddle-node.jl"); X.roots .= [[[-1.0]]]
+plot(F,[5/2,-1],X)
+savefig("docs/figures/saddle-node.pdf")
+
+####################################################################################################
+####################################################################################################
+########################################################################################## pitchfork
+include("minimal/pitchfork.jl")
+plot(F,[1/2,-1],X)
+savefig("docs/figures/pitchfork.pdf")
 
 ####################################################################################################
 ####################################################################################################
 ################################################################################ bifurcation measure
-
 measure(x) = 1/(1+abs(f(x)/ForwardDiff.derivative(f,x)))
-linewidth(x) = 3/(0.6-measure(x)/2)
 
-plot(xlabel=L"\mathrm{parameter,}p", ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|",
-	ylim=(-5,5), grid=false, right_margin=20mm, size=(500,450) )
+layout = @layout [a;b{1.0w,0.5h}]
+default(); default(grid=false,label="",margin=0mm)
+plot(layout = layout, link = :x, size=(400,500) )
 
-hline!([0],linewidth=1,color=:black,label="")
+hline!([0],subplot=2,linewidth=1,ylim=(-5,5),color=:black, xlabel=L"\mathrm{parameter,}p", xmirror=true, topmargin=-5mm,
+	ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|")
+
+plot!([-1.52,-1.52],[0,1],subplot=1,linewidth=2,color=:gold)
+plot!([-1.52,-1.52],[0,5],subplot=2,linewidth=2,color=:gold)
+
[…]

### 2021-05-17 test/minimal/pitchfork.jl

@@ -7,11 +7,9 @@ function F(u::AbstractVector,parameters::NamedTuple)
 	F = similar(u,typeof(f))
 
 	F[1] = θ[1] + p*u[1] + θ[2]*u[1]^3
-	F[2] = u[1] - u[2] # dummy second dimension
-
 	return F
 end
 
 ######################################################### targets and initial guess
-X = StateSpace( 2, -5:0.01:5, [0] )
+X = StateSpace( 1, -5:0.01:5, [0] )
 θ = SizedVector{2}(4.0,-π/2)

### 2021-05-17 test/minimal/saddle-node.jl

@@ -7,11 +7,9 @@ function F(u::AbstractVector,parameters::NamedTuple)
 	F = similar(u,typeof(f))
 
 	F[1] = p + θ[1]*u[1] + θ[2]*u[1]^3
-	F[2] = u[1] - u[2] # dummy second dimension
-
 	return F
 end
 
 ######################################################### targets and initial guess
-X = StateSpace( 2, -2:0.01:2, [1,-1] )
+X = StateSpace( 1, -2:0.01:2, [1,-1] )
 θ = SizedVector{2}(5.0,-0.93)

### 2021-05-17 test/plots.jl

@@ -1,25 +1,22 @@
-using Plots,FiniteDifferences
-using LinearAlgebra: norm
-using StatsBase: median
+using Plots
+function grid_test(name; xlim=(-5,5),ylim=(-5,5), dims=[1,2], ϵ=-1e-1,order=5,geom=true,condition=100)
 
-function grid_test(name; xlim=(-5,5),ylim=(-5,5),ϵ=-1e-1,order=5,threshold=0.05)
-
-	hyperparameters = getParameters(targetData)
+	hyperparameters = getParameters(X)
 	parameters = copy(θ)
 
 	function ∇cost(x::Number,y::Number)
-		parameters[1:2] .= x,y
-
-		L,∇L = ∇loss(rates,parameters,targetData,hyperparameters)
-		return ∇L[1:2]
+		parameters[dims] .= x,y
+		return ∇loss(F,parameters,X,hyperparameters)[2]
 	end
 
-	cost(x::Vector{<:Number}) = cost(x[1],x[2])
 	function cost(x::Number,y::Number)
-		parameters[1:2] .= x,y
+		parameters[dims] .= x,y
+		return loss(F,parameters,X,hyperparameters)
+	end
 
-		L = loss(rates,parameters,targetData,hyperparameters)
-		return L
+	function finite_differences(x,y)
+		parameters[dims] .= x,y
+		return finite_difference_gradient(θ->loss(F,θ,X),parameters)
 	end
 
 	plot(size=(600,600), xlim=xlim, ylim=ylim, xlabel="parameters, θ")
@@ -32,18 +29,16 @@ function grid_test(name; xlim=(-5,5),ylim=(-5,5),ϵ=-1e-1,order=5,threshold=0.05
 	xGrid = vcat(map(x->x[1], grid)...)
 	yGrid = vcat(map(x->x[2], grid)...)
 
-	differences = map( (x,y)->first(grad(central_fdm(order,1,geom=true,condition=100), cost, [x,y] )), xGrid, yGrid )
-	quiver!( xGrid, yGrid, quiver=(ϵ*step(x)*first.(differences),ϵ*step(y)*last.(differences)),
+	differences = map( finite_differences, xGrid, yGrid )
+	quiver!( xGrid, yGrid, quiver=(ϵ*step(x)*getindex.(differences,dims[1]),ϵ*step(y)*getindex.(differences,dims[2])),
 		color=:darkblue, lw=3 )
 
[…]

### 2021-05-17 test/scaling/scaling.jl

@@ -1,5 +1,5 @@
 ######################################################## model
-function rates(u::AbstractVector{T},parameters::NamedTuple,N::Int,M::Int) where T<:Number
+function F(u::AbstractVector{T},parameters::NamedTuple,N::Int,M::Int) where T<:Number
 
 	@unpack θ,p = parameters
 	f = first(u)*first(p)*first(θ)
@@ -17,21 +17,13 @@ function rates(u::AbstractVector{T},parameters::NamedTuple,N::Int,M::Int) where 
 	return F
 end
 
+function scaling(N::Int,M::Int)
+	F(u,p) = F(u,p,N,M)
 
-function benchmark(N::Int,M::Int)
-
-	F = (u,p) -> rates(u,p,N,M)
-	targetData = StateSpace(N,-π:0.01:π,[0.0])
+	X = StateSpace(N,-π:0.01:π,[0.0])
 	θ = SVector{M}(ones(M))
 
 	println("N = $N M = $M")
-	@time ∇loss( F, θ, targetData )
-	return @elapsed  ∇loss( F, θ, targetData )
+	@time ∇loss(F,θ,X)
+	return @elapsed ∇loss(F,θ,X)
 end
-
-heatmap( 1:5, 1:5, benchmark,
-	xlabel="States", ylabel="Parameters", size=(500,400),
-	colorbar_title="Iteration Execution / sec"
-)
-
-savefig(joinpath(@__DIR__,"scaling.pdf"))

### 2021-05-17 docs/appendix.tex
Updates from Overleaf

@@
 In this particular setting we can see that the tangent field $\tangent(z)$ only folds when $\Det=0$.
-Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-field} would give rise to Figures \ref{fig:saddle-node} and \ref{fig:pitchfork}.
+Plotting the value of the determinant along $\rates(z)=0$ from Figure \ref{fig:determinant-field} would give rise to Figures \ref{fig:minimal-models}.
 The directional derivative of the determinant $\Det$ along the tangent field $\tangent(z)$ is defined as

### 2021-05-17 docs/article.tex

@@
 For clarity, we guide the reader through the methods with the following minimal models that explore the space of saddle-nodes $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$.
-These minimal models are shown with targets $\targets$ in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
+These minimal models are shown with targets $\targets$ in Figure \ref{fig:minimal-models}.
 These figures show that the determinant $\Det=0$ at the bifurcations and that its directional derivative along the bifurcation curve $\frac{d}{ds}\Det$ is finite and non-zero, at least in the non-degenerate case.
 This gives us our first hint of what should be optimised to increase the number of predictions $|\predictions(\theta)|$.
+
+%   \centering
+%   \setlength\unitlength{1cm}
+%   {\phantomsubcaption\label{fig:1a}}
+%   {\phantomsubcaption\label{fig:1b}}
+
+% \textcolor{gray}{\rule{5cm}{3cm}}%
+% \begin{picture}(0,0)
+% \put(-4.8,2.6){\subref{fig:1a}}
+% \end{picture}
+% \textcolor{gray}{\rule{5cm}{3cm}}%
+% \begin{picture}(0,0)
+% \put(-4.8,2.6){\subref{fig:1b}}
+% \end{picture}
 
@@
 \includegraphics[width=6cm]{pitchfork}
-\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(5/2,-1),(1/2,-1)$ and targets $\targets=\{-1,1\},\{0\}$ respectively.
+\caption{Saddle-node model (left column) $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork model (right column) $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(5/2,-1),(1/2,-1)$ and targets $\targets=\{-1,1\},\{0\}$ respectively.
 Determinant in red.
@@
 \end{align}
-We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant $\Det=0$.
+We can see from Figures \ref{fig:minimal-models} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant $\Det=0$.
 Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$ in the non-degenerate case.

### 2021-05-22 test/figures.jl
two-state optima figure

@@ -63,43 +63,75 @@ savefig("docs/figures/bifurcation-measure.pdf")
 ####################################################################################################
 ############################################################################################ scaling
 include("scaling/scaling.jl")
-heatmap( 1:5, 1:5, scaling,
+
+Ns,Ms = 1:10,1:5:50
+complexity = scaling.(Ns,Ms')
+
+contourf( Ns,Ms, complexity,
 	xlabel="States", ylabel="Parameters", size=(500,400),
 	colorbar_title="Iteration Execution / sec"
 )
+
+cticks!
 savefig("test/scaling/scaling.pdf")
 
 ####################################################################################################
 ####################################################################################################
 ################################################################################ two-state-optima
-using JLD,Plots
-using UMAP,Clustering
-using StatsBase: median
+using FluxContinuation,Plots,StaticArrays
+using UMAP,Clustering,JLD
+using StatsBase: median,mean
 
-trajectories = Vector{Float64}[]
+trajectories, losses, bifurcations, K = Vector{Float64}[], Float64[], Float64[], 0
 for file ∈ readdir(join=true,"trajectories")
 	data = load(file)
 
-	optima = map( (x,y)->(y<0.1)&(x>1),data["bifurcations"],data["losses"])
-	append!(trajectories,data["trajectory"][optima])
+	mask = @. (data["bifurcations"]==2)&(data["losses"]<0.1)
+	if ~any(x->any(isnan.(x)),data["trajectory"])
+
+		append!(trajectories,data["trajectory"][mask])
+		append!(bifurcations,data["bifurcations"][mask])
+		append!(losses,data["losses"][mask])
+		K += 1
+	end
 end
+
+K/length(readdir("trajectories"))
 optima = hcat(trajectories...)
 
+@. optima[end,:] = abs(optima[end,:])
+M,N = size(optima)
+clusters = dbscan(optima,0.3,min_cluster_size=500)
 
-embedding = umap(optima; n_neighbors=3, min_dist=2)
-# clusters = dbscan(optima,0.1,min_neighbors = 1, min_cluster_size = 1)
[…]

### 2021-05-22 test/scaling/scaling.jl

@@ -1,4 +1,5 @@
 ######################################################## model
+F(z::BorderedArray,θ::AbstractVector,N::Int,M::Int) = F(z.u,(θ=θ,p=z.p),N::Int,M::Int)
 function F(u::AbstractVector{T},parameters::NamedTuple,N::Int,M::Int) where T<:Number
 
 	@unpack θ,p = parameters
@@ -18,12 +19,12 @@ function F(u::AbstractVector{T},parameters::NamedTuple,N::Int,M::Int) where T<:N
 end
 
 function scaling(N::Int,M::Int)
-	F(u,p) = F(u,p,N,M)
+	f(u,p) = F(u,p,N,M)
 
 	X = StateSpace(N,-π:0.01:π,[0.0])
-	θ = SVector{M}(ones(M))
+	θ = SizedVector{M}(ones(M))
 
 	println("N = $N M = $M")
-	@time ∇loss(F,θ,X)
-	return @elapsed ∇loss(F,θ,X)
+	@time ∇loss(f,θ,X)
+	return @elapsed ∇loss(f,θ,X)
 end

### 2021-05-22 docs/article.tex

--- 2021-05-17 docs/article.tex
@@
 
-\title{Inference of Bifurcations\\with Differentiable Continuation}
+\title{Parameter Inference\\with Bifurcation Diagrams}
 \author{
@@
 \maketitle
-
 \begin{abstract}
-    In this work we propose a gradient-based semi-supervised approach for matching target bifurcations with parameterised differential equation models.
-The cost function contains a supervised error term that is minimal when predicted bifurcations match the targets and an unsupervised bifurcation measure which acts as a proxy for the probability of a bifurcation, yielding non-zero gradients in non-bifurcating parameter regimes.
-The calculation of gradients with respect to parameters shares the same computational complexity as deflated pseudo-arclength continuation used to calculate the bifurcation diagram.
-Our implementation leverages automatic differentiation methods in Julia and we demonstrate model synthesis with minimal models which explore the space of saddle-node and pitchfork diagrams and the genetic toggle switch from synthetic biology.
+    Estimation of parameters in differential equation models can be achieved by applying learning algorithms to time-series data.
+However, sometimes it is only possible to measure qualitative changes of a system in response to a controlled condition.
+In dynamical systems theory, such change points are known as \textit{bifurcations} and lie on a function of the controlled condition called the \textit{bifurcation diagram}.
+In this work, we propose a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
+The cost function contains a supervised error term that is minimal when the model bifurcations match the specified targets and an unsupervised bifurcation measure which has gradients that push optimisers towards bifurcating parameter regimes.
+The gradients can be computed without the need to differentiate through the operations of the solver that was used to compute the diagram.
+We demonstrate parameter inference with minimal models which explore the space of saddle-node and pitchfork diagrams and the genetic toggle switch from synthetic biology.
 Furthermore, the cost landscape allows us to organise models in terms of topological and geometric equivalence.
@@
 
-Backpropagation through differential equation solves has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jl-AEquations} that enabled scalable parameter inference for differential equations.
-Determining what model parameters should be for a given set of observations, in the setting of biology and engineering, are known as inverse problems \cite{Abdulla2009InverseBiology}.
+Inverse problems \cite{Abdulla2009InverseBiology} arise in biology and engineering in settings when the model is not fully known and the desire is to match model behaviour to a given set of observations.
+This helps systematically guide both model and experimental design.
+The collected observations, however, are often indirectly proportional to the variables defined in the model; in microscopy, for example, data are reported in arbitrary fluorescence units allowing the observer to shift and scale data arbitrarily.
+The model, on the other hand, describes the dynamics of concentrations of relevant proteins.
+Furthermore, the observed qualitative changes in response to changes in experimental conditions are more robust and reproducible across studies than the quantitative details.
+For example, several studies are likely to agree that the human immune system activates above a threshold concentration of a pathogen and deactivates at a lower threshold concentration, but may disagree on the exact quantities of the thresholds or the magnitudes of the immune response.
+Bifurcation theory provides us a framework for studying these qualitative changes in a manner that is independent of quantitative details.
+The emerging picture suggests that identification of the qualitative behaviour -- the bifurcation diagram -- should precede any attempt at inferring other properties of a system \cite{Stumpf2019ParameterBifurcations}.
 
-Optimisation targets, however, have mostly been expressed in the spatio-temporal domain.
-Acquisition of such data can be costly and can often contain over-constraining information generated by processing steps or the measurement device rather than the state variables that underpin the observed mechanism.
-In microscopy, for example, data is often reported in arbitrary fluoresce units allowing the observer to shift and scale data arbitrarily.
-Furthermore, such data may also not contain sufficient information about dynamical transients in order to identify kinetic parameters.
-The emerging picture suggests that identification of the qualitative behaviour -- the bifurcation diagram -- should precede any attempt at inferring kinetic parameters \cite{Stumpf2019ParameterBifurcations}.
+Inferring the parameters of a model directly from a bifurcation diagram is difficult because it is not obvious how one should change the parameters to create a bifurcation.
+It could even be impossible for the model to bifurcate in the manner desired.
+Several approaches exist to place bifurcations to desired locations once a manifold is present \cite{Iwasaki1997AnType,Lu2006InverseSystems,Dobson2004DistanceBifurcations} yet typically resort to sampling techniques to search for them in the first place \cite{Chickarmane2005BifurcationTool,Conrad2006BifurcationClock}.
+Progress has been made is cases where model structure and stability conditions are used to refine the search space \cite{Otero-Muras2018Optimization-basedModels,Otero-Muras2014ACurves} yet the resulting objectives are still not explicit in the bifurcation targets and also not differentiable.
+A scalable method for navigating the space of bifurcation diagrams would enable design of differential equations with high-level qualitative constraints.
+Furthermore one could begin organising models according to qualitatively distinct behaviours.
+
+Back-propagation through differential equation solvers has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jl-AEquations} that enabled scalable parameter inference for differential equations from trajectory data.
+Although one could use trajectory data to create the aforementioned qualitative constraints \cite{Ranciati2017BayesianParameters,Khadivar2021LearningBifurcations} this would entail over-constraining information originating from the kinetics and dynamical transients of the model.
+Furthermore, such data usually does not contain sufficient information about dynamical transients in order to identify kinetic parameters.
 Techniques for back-propagating through implicit equation solvers have also been developed \cite{Look2020DifferentiableLayers,Bai2019DeepModels} although to the best of the authors' knowledge have not been applied to bifurcation diagrams at the time of writing this paper.
 
-In this work we propose a gradient-based semi-supervised approach that focuses on fitting high-level qualitative constraints, defined by state space structures, rather than kinetics.
-Drawing inspiration from implicit solvers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients we find that their computation shares the same  complexity as the algorithm used to calculate the bifurcation diagram.
-We use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl}, originally developed for partial differential equations.
+In this work, we propose a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
+The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
+Drawing inspiration from implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients we find that their computation shares the same  complexity as the algorithm used to calculate the bifurcation diagram.
+The compute the diagram we use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl} developed for partial differential equations.
 In the case of partial differential equations the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
@@
 
-We find that the cost function landscape contains basins that not only allow us to synthesise models with a desired bifurcation structure but also allow us to organise models in terms of topological and geometric equivalence.
+We find that the cost function landscape contains basins that not only allow us to synthesise models with a desired bifurcation diagram but also allow us to organise models in terms of topological and geometric equivalence.
 We discuss the relevance of this in model selection.
@@
 \begin{itemize}
-    \item Defined a differentiable semi-supervised cost function for encouraging and placing co-dimension one bifurcations to user-specified target locations
-    \item Implementation of method in Julia package \texttt{FluxContinuation.jl} leveraging automatic differentiation tooling in \texttt{Flux.jl}
+    \item An end-to-end differentiable method for searching for bifurcations and then matching them to user-specified target locations
+    \item Implementation of the method in Julia package leveraging automatic differentiation tooling
     \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
@@
 
-\subsection{Related Works | \textit{in progress}}
+\subsection{Preliminaries}
 
-\paragraph{Bifurcation control} and \textit{inverse} bifurcation analysis literature applies control theory to enable design and identification of dynamical systems with qualitative constraints.
-The first works involved the calculation of the minimal distances to bifurcation manifolds \cite{Iwasaki1997AnType,Lu2006InverseSystems,Dobson2004DistanceBifurcations}.
-Those optimisations did not involve closed form expressions and the challenge of finding the manifolds to being with was left to random sampling.
+Suppose we collected observations along a scalar control condition $p\in\Reals$ and conclude that there are specific values of $p$ for which there are qualitative changes in system behaviour.
+Let $\targets$ be the set of those values and let us hypothesise that these transitions occur due to bifurcations in the dynamics that drive the underlying mechanism.
+Let us model the mechanism with a parameterised set of differential equations for states $u\in\Reals^N$ with a vector function $\rates$ in a parameter space $\theta\in\Reals^M$.
 
-\paragraph{Matching vector fields} From a dynamical systems perspective a given bifurcation curve constrains a state-vector field geometry.
-A set of target bifurcations may also be represented as a vector field flow and thus inferring bifurcations becomes a problem in matching flows in vector sub-spaces.
-Note that this would involve matching direction but not the magnitude of the flow.
-
-\begin{itemize}
-    \item smooth and match estimator methods \cite{Ranciati2017BayesianParameters}
-    \item robots learning limit cycles \cite{Khadivar2021LearningBifurcations}
-\end{itemize}
-
-\paragraph{Non-differentiable methods} 
-A large body of work is dedicated to finding conditions for fold-bifurcations, multistability and other
-
-\begin{itemize}
-    \item bifurcation inference using Mixed Integer Nonlinear Programming \cite{Otero-Muras2018Optimization-basedModels}
-    \item Random sampling and genetic algorithms \cite{Chickarmane2005BifurcationTool,Conrad2006BifurcationClock}
-\end{itemize}
-
-\subsection{Preliminaries}
-Suppose we parameterise a set of differential equations for states $u\in\Reals^N$ with a function $\rates$ in an unknown parameter space $\theta\in\Reals^M$.
-We would like these differential equations to obey a set of target bifurcations $\targets:=\{p_1\dots p_K\}$ along a known bifurcation parameter $p\in\Reals$.
-Let the differential equations be defined as
+For the purposes of introducing this work, we will consider the simplest class of bifurcations known as \textit{co-dimension one} bifurcations, that do not include limit cycles.
+Therefore $\targets$ should contain conditions for which we hypothesise changes in multi-stable behaviour.
+Let the equations be
 \begin{align}
-	\partial_t u=\rates(z)
-	\qquad\mathrm{where}\quad z:=(u,p)\quad
+	\frac{\partial u}{\partial t}=\rates(u,p)
+	\qquad\mathrm{where}\quad
 	\rates : \Reals^{N+1}\rightarrow\Reals^N
@@
 \end{align}
+
+In the context of the differential equations, a co-dimension one bifurcation can be defined by a set of conditions on the determinant of the Jacobian $\Det$.
+The determinant of the Jacobian quantifies the rate at which trajectories in a local patch of state-space $u\in\Reals^N$ converge or diverge.
+The determinant approaching zero means that the dynamics of the system is slowing down, which is an important indicator for the onset of a transition between qualitative behaviours.
+Furthermore, the slowing down must necessarily be followed by breakdown of stability; for this to be true it is sufficient to require that the determinant cross zero with a finite slope, meaning that its directional derivative along the diagram $\frac{d}{ds}\Det$ is not zero.
+The set of predicted values for the control condition $\predictions(\theta)\subset\Reals$ at which bifurcations occur are defined as
+\begin{align}
+	\predictions(\theta):=\left\{\,
+	p\,\,\exists\,\,u:\,\,\rates(u,p)=0,\,\,\Det=0,
+	\,\, \frac{d}{ds}\Det\neq0
+	\,\right\}
+	\label{eq:predictions}
+\end{align}
+ Figure \ref{fig:minimal-models} shows minimal models that explore the space of saddle-nodes $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(u,p) = \theta_{1} + p u+\theta_{2}u^3$.
+Indeed the predictions $\predictions(\theta)$ are defined by zero crossings in the determinant with a finite slope.
+The location of these crossings in general may not match the targets $\targets$.
+
 For a given set of parameters $\theta$ one could compute the set of predicted bifurcations $\predictions(\theta)$ using parameter continuation methods \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
@@
 The optimal $\theta^*$ is not expected to always be unique, but is in general a manifold representing the space of qualitatively equivalent models.
-The cost function $\loss$ must have some distance measure between the two sets $\predictions(\theta)$ and $\targets$ with a bias that encourages equivalent cardinality $|\predictions(\theta)|=|\targets|$.
+The cost function $\loss$ must have some distance measure between the two sets $\predictions(\theta)$ and $\targets$ with a bias that encourages equivalent set sizes $|\predictions(\theta)|=|\targets|$.
 This is especially important in the case where there are no predictions $|\predictions(\theta)|=0$.
-
-For clarity, we guide the reader through the methods with the following minimal models that explore the space of saddle-nodes $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$.
-These minimal models are shown with targets $\targets$ in Figure \ref{fig:saddle-node} and Figure \ref{fig:pitchfork} respectively.
-These figures show that the determinant $\Det=0$ at the bifurcations and that its directional derivative along the bifurcation curve $\frac{d}{ds}\Det$ is finite and non-zero, at least in the non-degenerate case.
-This gives us our first hint of what should be optimised to increase the number of predictions $|\predictions(\theta)|$.
 
@@
 \centering
+\setlength\unitlength{1cm}
+{\phantomsubcaption\label{fig:saddle-node}}
+{\phantomsubcaption\label{fig:pitchfork}}
 \includegraphics[width=6cm]{saddle-node}
+\begin{picture}(0,0) \put(-5.8,10){\subref{fig:saddle-node}} \end{picture}
 \includegraphics[width=6cm]{pitchfork}
-\caption{Saddle-node model $\rates(z) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork model $\rates(z) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(5/2,-1),(1/2,-1)$ and targets $\targets=\{-1,1\},\{0\}$ respectively.
-Determinant in red.
-Lighter shades indicate the determinant crossing zero for unstable solutions}
+\begin{picture}(0,0) \put(-5.8,10){\subref{fig:pitchfork}} \end{picture}
+\caption{Saddle-node model \ref{fig:saddle-node} $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork model \ref{fig:pitchfork} $\rates(u,p) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(5/2,-1),(1/2,-1)$ and targets $\targets=\{-1,1\},\{0\}$ respectively.
+Lighter shades indicate the determinant crossing zero at locations $\predictions(\theta)$ giving rise to unstable solutions}
 \label{fig:minimal-models}
@@
     \includegraphics[width=7cm]{bifurcation-measure}
-    \caption{Bifurcation measure $\measure(z)$ and determinant $\Det$ for two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
-    \label{fig:bifurcation-measure}
+    \caption{Bifurcation measure $\measure(s)$ and determinant $\Det$ along two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
+    \label{fig:measure}
 \end{figure}
 
-In order for predicted bifurcations $p(\theta)\in\predictions$ to match targets $p'\in\targets$ we need to evaluate an error term $|p(\theta)-p'|$.
+In order for predicted bifurcations $p(\theta)\in\predictions(\theta)$ to match targets $p'\in\targets$ we need to evaluate an error term $|p(\theta)-p'|$.
 We expect this term to be some mean over the targets $\targets$ and predictions $\predictions(\theta)$.
@@
     \frac{1}{|\targets|}\sum_{p'\in\targets}
-    \prod_{p(\theta)\in\predictions}|p(\theta)-p'|
+    \prod_{p(\theta)\in\predictions(\theta)}|p(\theta)-p'|
     ^{\frac{1}{|\predictions|}}
-    \label{eq:supervised-term}
+    \label{eq:error}
 \end{align}
-We can see from Figures \ref{fig:saddle-node} and \ref{fig:pitchfork} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant $\Det=0$.
-Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$ in the non-degenerate case.
-Using these quantities we can define a positive semi-definite measure $\measure(z)$ of zero crossings in the determinant along the bifurcation curve; as shown in Figure \ref{fig:bifurcation-measure} this can be used as a proxy for the probability of a bifurcation at location $p(\theta)$.
-The total bifurcation measure $\Psi(\theta)$ is written
+We can see from Figures \ref{fig:minimal-models} and definitions \eqref{eq:predictions} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant $\Det=0$.
+Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$.
+Using these quantities we can define a positive semi-definite measure $\measure(u,p)$ of zero crossings in the determinant along the bifurcation curve.
+Figure \ref{fig:measure} demonstrates this measure is maximal at bifurcation at locations $p(\theta)$ and has finite gradients in non-bifurcating regimes.
+The total bifurcation measure $\Psi(\theta)$ is defined as
 \begin{align}
     \Psi(\theta):=\frac{
-        \int_{\rates(z)=0}\!
-        \measure(z)\,\mathrm{d}z
+        \int_{\rates(u,p)=0}\!
+        \measure(u,p)\,\mathrm{d}u\mathrm{d}p
     }{
-        \int_{\rates(z)=0}\!\
-        \mathrm{d}z
+        \int_{\rates(u,p)=0}\!\
+        \mathrm{d}u\mathrm{d}p
     }
     \quad\mathrm{where}\quad
-    \measure(z):=
+    \measure(u,p):=
     \left(1+\left|\frac{\Det}{\frac{d}{ds}\Det}\right|\right)^{-1}
@@
 \end{align}
-Here we denote $\int_{\rates(z)=0}\mathrm{d}z$ as the sum of the line integrals in $z\in\Reals^{N+1}$ defined by the steady states $\rates(z)=0$.
-The measure $\measure(z)=1$ at bifurcation points and goes to zero an odd number of times between bifurcations.
-This is because $\Det$ must eventually turn around in order to return back to zero, resulting in $\frac{d}{ds}\Det\rightarrow0$ and hence $\measure(z)\rightarrow0$ for each turning point.
-As $\Det$ diverges we approach regimes far away from any closest bifurcation and hence $\measure(z)\rightarrow0$.
-We would still like to have non-zero gradients with respect to $\theta$ in this regime, and therefore $\measure(z)$ was designed to go to zero slowly.
-The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the observed parameter region $p$ is densely packed with bifurcations.
-The $\measure(z)$ is also invariant under certain classes of transformations on $\rates(z)$ which do not change bifurcation locations, proofs of which are provided in Appendix \ref{appendix:measure}.
-The measure is added to the supervised term as if it were a likelihood; this defines the semi-supervised cost function as
+Here we denote $\int_{\rates(u,p)=0}\mathrm{d}u\mathrm{d}p$ as the sum of the line integrals in $(u,p)\in\Reals^{N+1}$ defined by the level set $\rates(u,p)=0$.
+The measure $\measure(u,p)$ is one at bifurcation points and goes to zero an odd number of times between bifurcations.
+This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(u,p)$ going to zero for each turning point.
+While the calculation of the determinant is straightforward, its directional derivative is not.
+Fortunately, a tangent field representation of the bifurcation curve allows us to compute this derivative implicitly; see Appendix \ref{appendix:tangent-fields} for details.
+
+As the determinant $\Det$ diverges we approach regimes far away from any closest bifurcation and hence $\measure(u,p)\rightarrow0$.
+We would still like to have non-zero gradients with respect to $\theta$ in this regime, and therefore $\measure(u,p)$ was designed to go to zero slowly.
+The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the controlled condition region $p$ is densely packed with bifurcations.
+The measure $\measure(u,p)$ is also invariant under certain classes of transformations on $\rates(u,p)$ which do not change bifurcation locations, proofs of which are provided in Appendix \ref{appendix:measure}.
+The total measure $\Psi(\theta)$ is added to the supervised term as if it were a likelihood; this defines the semi-supervised cost function as
 \begin{align}
@@
     \big(|\predictions|-|\targets|\big)\,\,\lambda \log\Psi(\theta)+
-    \frac{1}{|\targets|}\sum_{p'\in\targets}
-    \prod_{p(\theta)\in\predictions}|p(\theta)-p'|
+    \frac{1}{|\targets|}\sum_{p'}
+    \prod_{p(\theta)}|p(\theta)-p'|
     ^{\frac{1}{|\predictions|}}
-    \label{eq:cost}
+    \label{eq:loss}
 \end{align}
@@
 This can be seen as a step-wise annealing of the unsupervised term until the desired state is reached.
-Note that in principle while individual bifurcations $p(\theta)$ depend smoothly on $\theta$ the total number of predictions $|\predictions|$ does not.
-The gradient $\partial_{\theta}|\predictions|=0$ everywhere and is an infinite impulse at the exact locations where the number of predictions changes.
-
-In practice optimisers are unlikely to hit those values of $\theta$ and we can safely drop the dependency.
-We may now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ and total measure $\Psi(\theta)$
+Note that in while individual bifurcations $p(\theta)$ depend smoothly on $\theta$, the total number of predictions $|\predictions|$ does not have gradient contributions with respect to $\theta$.
+We can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ and total measure $\Psi(\theta)$
 \begin{align}
@@
     \sum_{p(\theta)}\frac{\partial p}{\partial\theta}\left(p(\theta)-p'\right)^{-1}
-    \label{eq:gradient}
+    \label{eq:gradient-loss}
 \end{align}
-In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to calculate the gradient $\frac{\partial\loss}{\partial\theta}$ without having to differentiate through the operations of the solver that finds the bifurcation curve $\rates(z)=0$ and the bifurcation locations $p(\theta)$.
+In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to calculate the gradient $\frac{\partial\loss}{\partial\theta}$ without having to differentiate through the operations of the solver that finds the bifurcation curve $\rates(u,p)=0$ and the bifurcation locations $p(\theta)$.
 To calculate the gradient of the measure $\frac{\partial \Psi}{\partial\theta}$ we need to differentiate line integrals that depend on $\theta$.
@@
 
-The gradient of the bifurcation points $\frac{\partial p}{\partial\theta}$ is found by application of the implicit function theorem to a vector function $G_{\theta}:\Reals^{N+1}\rightarrow\Reals^{N+1}$ whos components represent the two constraints $\rates(z)=0$ and $\Det=0$.
+The gradient of the bifurcation points $\frac{\partial p}{\partial\theta}$ is found by application of the implicit function theorem to a vector function $G_{\theta}:\Reals^{N+1}\rightarrow\Reals^{N+1}$ whos components represent the two constraints $\rates(u,p)=0$ and $\Det=0$.
 By following a similar strategy to that used by implicit layers \cite{Look2020DifferentiableLayers} we yield an $(N+1)\times M$ Jacobian representing a deformation field \cite{Jos2011OnSurface} for each $\theta$ direction.
@@
 \begin{align}
-    \frac{\partial p}{\partial\theta} = -\hat{p}\cdot\left.\frac{\partial G_{\theta}}{\partial z}^{-1}
-    \frac{\partial  G_{\theta}}{\partial\theta}\right|_{G_{\theta}(z)=0}
+    \frac{\partial p}{\partial\theta} = -\hat{p}\cdot\left.\frac{\partial G_{\theta}}{\partial (u,p)}^{-1}
+    \frac{\partial  G_{\theta}}{\partial\theta}\right|_{G_{\theta}(u,p)=0}
     \quad\mathrm{where}\quad
-    G_{\theta}(z):=\begin{bmatrix}\rates(z)\\\Det\end{bmatrix}
+    G_{\theta}(u,p):=\begin{bmatrix}\rates(u,p)\\\Det\end{bmatrix}
+    \label{eq:gradient-bifurcation}
 \end{align}
-Here $\hat{p}$ is a unit vector in $z\in\Reals^{N+1}$ that picks out the deformations along the $p$-direction.
-We can see this calculation involves inverting an $(N+1)\times(N+1)$ Jacobian $\frac{\partial G_{\theta}}{\partial z}$.
-The determinant of this Jacobian goes to zero when $\frac{d}{ds}\Det=0$.
-This further justifies our choice for measure $\measure(z)$ which discourages such cases.
+Here $\hat{p}$ is a unit vector in $(u,p)\in\Reals^{N+1}$ that picks out the deformations along the $p$-direction.
+If we wanted to place the bifurcation at target steady state $u'$ as well as target control condition $p'$ we would use the full $(N+1)\times M$ deformation matrix.
+Calculation of this matrix involves inverting an $(N+1)\times(N+1)$ Jacobian $\frac{\partial G_{\theta}}{\partial(u,p)}$.
+The determinant of this Jacobian goes to zero in the degenerate case where $\frac{d}{ds}\Det=0$.
+This further justifies our choice for measure $\Psi(\theta)$ which discourages the degenerate case.
 
-The cost function is piece-wise smooth and differentiable with undefined gradients only in parameter regions where the number of predictions $|\predictions|$ changes.
-Given a set of solutions to $\rates(z)=0$ and locations $p(\theta)$ which additionally satisfy $\Det=0$, the gradient $\frac{\partial\loss}{\partial\theta}$ can evaluated using automatic differentiation methods.
-
-\subsection{Implementation}
-\label{section:implementation}
-%/ move to appendix?
-Here provide a high-level view of Algorithm \ref{alg:optimisation-loop}.
-We use \texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} to calculate bifurcation diagrams, which requires the rate function $\rates$, its Jacobian $J_\theta$ which can be calculated using \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia}, and an initial guess $u_0$ at an initial parameter $p_0$.
-Let this algorithm be called by function \texttt{getSteadyStates} and return the steady states that form the integration region $\steadystates$ and a new guess $u_0$ that is the true steady state at $p_0$.
-
-The algorithm also takes hyperparameters $\beta$ which will contain arclength step sizes, eigenvalue and Newton solver options and unsupervised correction factor $\lambda$.
-The adaptive hyperparameter update is done by \texttt{updateHyperparameters} and should depend on the current solutions $\steadystates$ and the targets $\targets$.
-The step sizes and solver options should adapt to the expected space of the solutions $\steadystates$ so that the waiting time between iterations is minimised.
-The unsupervised correction factor $\lambda$ should be large when there are no predicted bifurcations and decrease as the number of predicted bifurcations approaches the number of targets.
-
-The evaluation of the cost gradient is done by \texttt{costGradient} using \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia} which has a large repository of chain rules that it uses to correctly evaluate derivatives.
-At the time of writing this paper the repository does not include the Leibniz rule \cite{Flanders1973DifferentiationSign} for implicitly defined line integrals, and we therefore define additional rules that implement the results shown in Appendix \ref{appendix:space-curve} and \ref{appendix:deformation}.
-The cost gradient is used by \texttt{updateParameters}, which can be any gradient-based optimiser such as ADAM or Momentum gradient decent, to update the parameters $\theta$.
-
-In practice the bifurcation curve is only evaluated in a finite region $p\in\Omega$ which includes the targets $\targets$.
-We are now ready to write down the semi-supervised cost function for a bifurcation curve defined by $\rates(z)=0$ in an observed region $p\in\Omega$.
-
-\begin{algorithm*}[H]
-\label{alg:optimisation-loop}
-\SetAlgoLined
-\textbf{Inputs} Function $F_{\theta}:\mathbb{R}^{N+1}\rightarrow\mathbb{R}^{N}$ for $u\in\mathbb{R}^N$ and $p\in\mathbb{R}$\\ target bifurcations $\mathcal{D}=\{p_1\dots p_K:p\in\Omega\}$ and convergence tolerance $\varepsilon$\\
-\textbf{Output} Optimised $\theta\in\mathbb{R}^M$ that satisfy targets $\targets$\\
-$u_0\leftarrow\mathrm{rand}\in\mathbb{R}^{N}$\quad\,$\theta\leftarrow\mathrm{rand}\in\mathbb{R}^{M}$\\
-$\beta\leftarrow\texttt{getHyperparameters}(\targets)$\\
-\While{$\mathrm{tolerance}(\beta)>\varepsilon$}{
-  $\steadystates,u_0\leftarrow\texttt{getSteadyStates}(F_{\theta},u_0,\beta)$\\
-  $\beta\leftarrow\texttt{updateHyperparameters}(\beta,\steadystates,\targets)$\\
-  $\partial L\leftarrow\texttt{costGradient}(F_{\theta},\steadystates,\targets,\beta)$\\
-  $\theta\leftarrow\texttt{updateParameters}(\theta,\partial L,\beta)$
-}
-\caption{Bifurcation Optimisation Loop}
-\end{algorithm*}
+The cost function is piece-wise smooth and differentiable with undefined gradients only in parameter contours where the number of predictions $|\predictions|$ changes; this is when $\Psi(\theta)$ is undefined and the inverse of $\frac{\partial G_{\theta}}{\partial (u,p)}$ does not exist.
+Given a set of solutions to $\rates(u,p)=0$ and locations $p(\theta)$ the gradient $\frac{\partial\loss}{\partial\theta}$ can evaluated using automatic differentiation methods without needing to back-propagate through the solver that obtained the level set $\rates(u,p)=0$ in the forward pass.
 

### 2021-05-22 test/figures.jl
update figure 3

@@ -104,25 +104,25 @@ M,N = size(optima)
 clusters = dbscan(optima,0.3,min_cluster_size=500)
 
 layout = @layout [ a{0.7w} [b{0.5h}; b{0.5h} ] ]
-default(); default(msc=:auto,label="",xlim=(-3,3),ylim=(-1,3),size=(600,450),grid=false)
+default(); default(msc=:auto,label="",xlim=(-3,3),ylim=(-1/2,1),size=(600,450),grid=false)
 
 fig = scatter([0],[0],markersize=0,title=L"\mathrm{Parameter\,\ Estimates}\,\,\theta^{*}", layout=layout, subplot=1,
-	ylabel=L"\mathrm{Half\,\,\,Saturation}\quad\log_{10}k",xlabel=L"\mathrm{Activation}\quad \log_{10}a_1")
+	ylabel=L"\mathrm{Half\,\,\,Saturation}\quad -\log_{10}k",xlabel=L"\mathrm{Activation}\quad -\log_{10}a_1")
 
 color = [:blue,:pink,:lightblue]
 name = ["1","2","1′"]
 clustered_names = ["1,1′","2"]
 for (j,cluster) ∈ enumerate(clusters)
-	x,y = getindex(optima,3,cluster.core_indices), getindex(optima,5,cluster.core_indices)
+	x,y = -getindex(optima,3,cluster.core_indices), -log10.(getindex(optima,5,cluster.core_indices))
 
 	scatter!(x,y,color=color[j],subplot=1)
 	annotate!(median(x),median(y),name[j],subplot=1)
 	
 	parameters = getindex(optima,:,cluster.core_indices)
-	i = rand(1:length(cluster.core_indices))
 	if j == 1 X = StateSpace( 2, 0:0.002:10, [4,5] ) else X = StateSpace( 2, 0:0.01:10, [4,5] ) end
 	if j ∈ [1,2]
-		plot!(F,parameters[:,i],X,determinant=false,xlim=(0,10),ylim=(0.01,100),yscale=:log10,
+		plot!(F,parameters[:,rand(1:length(cluster.core_indices))],X,
+			determinant=false,xlim=(0,10),ylim=(0.01,100),yscale=:log10,
 			subplot=j+1, framestyle = :box)
 		annotate!(8,30,clustered_names[j],subplot=j+1)
 	end

### 2021-05-22 docs/article.tex
Updates from Overleaf

@@
 \begin{equation}
-    \partial_t u_1 = \frac{\varepsilon_1 + a_1 u_2^2} { k_1^2 + u_2^2 } - \mu_1 u_1 \quad
-    \partial_t u_2 = \frac{p + a_2 u_1^2} { k_2^2 + u_1^2 } - \mu_2 u_2
+    \partial_t u_1 = \frac{ a_1 + (pu_2)^2} { 1 + (pu_2)^2 } - \mu_1 u_1 \quad
+    \partial_t u_2 = \frac{ a_2 + (ku_1)^2} { 1 + (ku_1)^2 } - \mu_2 u_2
     \label{eq:two-state}
@@
 \includegraphics[width=11cm]{two-state-optima}
-\caption{Reduced dimensionality plot of optimal $\theta^*$ for the two state model \eqref{eq:two-state} for $\targets=\{4,5\}$}
+\caption{Optimal parameter estimates $\theta^*$ for the two state model \eqref{eq:two-state} for the targets $\targets=\{4,5\}$ reveal two clusters of qualitatively different regimes: mutual activation and mutual inhibition.
+}
 \label{fig:two-state-optima}

### 2021-05-22 docs/appendix.tex
Updates from Overleaf

@@
 An expression for the field $\tangent(z)$ tangent to the set of curves would allow us to take derivatives and integrals along the bifurcation curve.
-This is exactly what we need to do to evaluate our cost function \ref{eq:cost}.
+This is exactly what we need to do to evaluate our cost function \ref{eq:loss}.
 Fortunately the tangent field can be constructed by ensuring it is perpendicular to the gradient $\partial_z$ of each component of $\rates$ as illustrated by an example two component system in Figure \ref{fig:implicit-surfaces}.
@@
 
-\section{Simplification of two-state model}
+% \section{Simplification of two-state model}
 
-Consider the general two-state model:
-\begin{equation}
-    \dfrac{dv_i}{dt} = \dfrac{a_i + b_i (K_i v_j)^2}{1 + (K_i v_j)^2} - \mu_i v_i
-\end{equation}
+% Consider the general two-state model:
+% \begin{equation}
+%     \dfrac{dv_i}{dt} = \dfrac{a_i + b_i (K_i v_j)^2}{1 + (K_i v_j)^2} - \mu_i v_i
+% \end{equation}
 
-Without loss of generality, we can rescale $v_i = U_i u_i$. Then,
-\begin{equation}
-    \dfrac{U_idu_i}{dt} = \dfrac{a_i + b_i (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i U_i u_i
-\end{equation}
+% Without loss of generality, we can rescale $v_i = U_i u_i$. Then,
+% \begin{equation}
+%     \dfrac{U_idu_i}{dt} = \dfrac{a_i + b_i (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i U_i u_i
+% \end{equation}
 
-Dividing through by $U_i$, we obtain
-\begin{equation}
-    \dfrac{du_i}{dt} = \dfrac{\frac{a_i}{U_i} + \frac{b_i}{U_i} (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i u_i
-\end{equation}
+% Dividing through by $U_i$, we obtain
+% \begin{equation}
+%     \dfrac{du_i}{dt} = \dfrac{\frac{a_i}{U_i} + \frac{b_i}{U_i} (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i u_i
+% \end{equation}
 
-By choosing $U_i = b_i$, the model reduces to:
-\begin{equation}
-    \dfrac{du_1}{dt} = \dfrac{\hat{a}_1 + (p u_2)^2}{1 + (p u_2)^2} - \mu_1 u_1, \quad
-    \dfrac{du_2}{dt} = \dfrac{\hat{a}_2 + (k u_1)^2}{1 + (k u_1)^2} - \mu_2 u_2
-\end{equation}
-where $p = \frac{K_1}{\sqrt{b_1}}$ and $k = \frac{K_2}{\sqrt{b_2}}$.
+% By choosing $U_i = b_i$, the model reduces to:
+% \begin{equation}
+%     \dfrac{du_1}{dt} = \dfrac{\hat{a}_1 + (p u_2)^2}{1 + (p u_2)^2} - \mu_1 u_1, \quad
+%     \dfrac{du_2}{dt} = \dfrac{\hat{a}_2 + (k u_1)^2}{1 + (k u_1)^2} - \mu_2 u_2
+% \end{equation}
+% where $p = \frac{K_1}{\sqrt{b_1}}$ and $k = \frac{K_2}{\sqrt{b_2}}$.

### 2021-05-23 docs/article.tex
Updates from Overleaf

@@
 
-\subsection{Genetic Toggle Switch | \textit{in progress}}
-In this section we optimise a model where the states share a Hill function relationship; these models often emerge from mass action kinetics and are used to model species concentrations.
-By not restricting the sign of the coefficients $a_1,a_2$ we can recover both activator and inhibitor relationships.
+\subsection{Genetic Toggle Switch}
+In this section we optimise a model where the states share a Hill function relationship with cooperatively $n=2$; these models often emerge from mass action kinetics and are used to model species concentrations.
+After re-scaling the equations governing the dynamics of concentrations, the simplified equations for state $u_$ and $u_2$ become
 \begin{equation}
-    \partial_t u_1 = \frac{ a_1 + (pu_2)^2} { 1 + (pu_2)^2 } - \mu_1 u_1 \quad
-    \partial_t u_2 = \frac{ a_2 + (ku_1)^2} { 1 + (ku_1)^2 } - \mu_2 u_2
+    \partial_t u_1 = \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 } - \mu_1 u_1 \quad
+    \partial_t u_2 = \frac{ a_2 + (k u_1)^2}{ 1 + (k u_1)^2 } - \mu_2 u_2
     \label{eq:two-state}
 \end{equation}
+where $a_k$ are 
+ the  we can recover both activator and inhibitor relationships.
 Figure \ref{fig:two-state-optima} shows a UMAP embedding of the optimal parameter basins, represented as point clouds.

### 2021-05-23 src/Gradients.jl
more efficient cost gradient calc

@@ -1,33 +1,33 @@
 ################################################################################
 function ∇loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
-	pmin,pmax = extrema(targets.parameter)
-	predictions = sum([ s.bif/2 for branch ∈ branches for s ∈ branch if (pmin ≤ s.z.p ≤ pmax) ])
 
-	Φ,∇Φ = measure(F,branches,θ,targets), ∇measure(F,branches,θ,targets)
-	return errors(branches,targets) - (length(targets.targets)-predictions)*log(Φ), ∇errors(F,branches,θ,targets) - (length(targets.targets)-predictions)*∇Φ/Φ
+	predictions = unique([ s.z for branch ∈ branches for s ∈ branch if s.bif ], atol=3*step(targets.parameter) )
+	λ = length(targets.targets)-length(predictions)
+
+	if λ≠0 
+		Φ,∇Φ = measure(F,branches,θ), ∇measure(F,branches,θ)
+		return errors(predictions,targets) - λ*log(Φ), ∇errors(F,predictions,θ,targets) - λ*∇Φ/Φ
+	else
+		return errors(predictions,targets), ∇errors(F,predictions,θ,targets)
+	end
 end
 
 ################################################################################
-function ∇errors( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace)
-	pmin,pmax = extrema(targets.parameter)
-
-	predictions = [ s.z for branch ∈ branches for s ∈ branch if s.bif & (pmin ≤ s.z.p ≤ pmax) ]
+function ∇errors( F::Function, predictions::AbstractVector{<:BorderedArray}, θ::AbstractVector, targets::StateSpace)
 	return mean( p′-> mean( z->(z.p-p′)^2, predictions; type=:geometric )*mean( z-> 2∂p(F,z,θ)/(z.p-p′), predictions; type=:arithmetic ), targets.targets; type=:arithmetic )
 end
 
-∇measure = Gradient( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; p=0.0, kwargs...)
-	∇measure = gradient(θ->measure(F,z,θ,targets;p=p,kwargs...),θ) + deformation(F,z,θ)'gradient(z->measure(F,z,θ,targets;p=p,kwargs...),z)
-	return ∇measure + measure(F,z,θ,targets;p=p,kwargs...)*∇region(F,z,θ)
-end, measure)
-
[…]

### 2021-05-23 src/Objectives.jl

@@ -1,30 +1,33 @@
 ################################################################################
 function loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
-	pmin,pmax = extrema(targets.parameter)
 
-	predictions = sum([ s.bif/2 for branch ∈ branches for s ∈ branch if (pmin ≤ s.z.p ≤ pmax) ])
-	return errors(branches,targets) - (length(targets.targets)-predictions)*log(measure(F,branches,θ,targets))
+	predictions = unique([ s.z for branch ∈ branches for s ∈ branch if s.bif ], atol=3*step(targets.parameter) )
+	λ = length(targets.targets)-length(predictions)
+
+	if λ≠0 
+		Φ = measure(F,branches,θ)
+		return errors(predictions,targets) - λ*log(Φ)
+	else
+		return errors(predictions,targets)
+	end
 end
 
 ################################################################################
-function errors( branches::AbstractVector{<:Branch}, targets::StateSpace)
-	pmin,pmax = extrema(targets.parameter)
-
-	predictions = [ s.z for branch ∈ branches for s ∈ branch if s.bif & (pmin ≤ s.z.p ≤ pmax) ]
+function errors( predictions::AbstractVector{<:BorderedArray}, targets::StateSpace)
 	return mean( p′-> mean( z->(z.p-p′)^2, predictions; type=:geometric ), targets.targets; type=:arithmetic )
 end
 
-measure = Integrand( function( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace; kwargs... )
+function measure( F::Function, z::BorderedArray, θ::AbstractVector )
 	return 1 / ( 1 + abs( det(F,z,θ) / ∂det(F,z,θ)'tangent_field(F,z,θ) ) )
-end)
-
-################################################################################
-function (integrand::Integrand)( F::Function, branch::Branch, θ::AbstractVector, targets::StateSpace; kwargs...)
-	return sum( s -> window_function( targets.parameter, s.z; kwargs... )*integrand( F, s.z, θ, targets; kwargs...)s.ds, branch )
 end
 
-function (integrand::Integrand)( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
-	return sum( branch -> integrand( F, branch, θ, targets; kwargs...), branches )
+################################################################################
+function measure( F::Function, branch::Branch, θ::AbstractVector )
[…]

### 2021-05-23 src/Structures.jl

@@ -51,20 +51,21 @@ function push!(branch::Branch,state::ContState)
     push!(branch, ( z=z, λ=state.eigvals, ds=norm(∂z)*abs(state.ds), bif=detectBifucation(state) ))
 end
 
-import Base: zero ################################## initialiser for BorderedArray
+import Base: zero,isapprox,unique ################################## methods for BorderedArray
 zero(::Type{BorderedArray{T,U}}; ϵ::Bool=true) where {T<:Union{Number,StaticArray},U<:Union{Number,StaticArray}} = BorderedArray(zero(T).+ϵ*eps(),zero(U).+ϵ*eps())
+isapprox( x::BorderedArray, y::BorderedArray; kwargs... ) = isapprox( [x.u;x.p], [y.u;y.p] ; kwargs...)
+function unique(X::AbstractVector{T}; kwargs...) where T<:BorderedArray
+    z = T[]
+    for x ∈ X 
+        if all( zi -> ~isapprox( x, zi; kwargs... ), z )
+            push!(z,x)
+        end
+    end
+    return z
+end
 
 #################################################### display methods
 show(io::IO, branches::Vector{Branch{V,T}}) where {V,T} = print(io,
-    "Vector{Branch}[dim=$(dim(first(branches))) bifurcations=$(sum(branch->sum(s->s.bif,branch),branches)) branches=$(length(branches)), states=$(sum(branch->length(branch),branches))]")
+    "Vector{Branch}[dim=$(dim(first(branches))) bifurcations=$(length(unique([ s.z for branch ∈ branches for s ∈ branch if s.bif ], atol=0.03 ))) branches=$(length(branches)/2), states=$(sum(branch->length(branch),branches)/2)]")
 show(io::IO, M::MIME"text/plain", branches::Vector{Branch{V,T}}) where {V,T} = show(io,branches)
 show(io::IO, states::StateSpace{N,T}) where {N,T} = print(io,"StateSpace{$N,$T}(parameters=$(states.parameter),targets=$(states.targets))")
-
-################################################################################
-########################################################### line integrand terms
-
-struct Integrand <: Function f::Function end
-(f::Integrand)(args...;kwargs...) = f.f(args...;kwargs...)
-
-struct Gradient <: Function f::Function; integrand::Integrand end
-(f::Gradient)(args...;kwargs...) = f.f(args...;kwargs...)

### 2021-05-23 test/applied/two-state.jl

@@ -15,5 +15,5 @@ function F(u::AbstractVector,parameters::NamedTuple)
 end
 
 ######################################################### targets and initial guess
-X = StateSpace( 2, 0:0.01:10, [4,5] )
+X = StateSpace( 2, 0.01:0.01:10, [4,5] )
 θ = SizedVector{5}(0.5,0.5,0.5470,2.0,7.5)

### 2021-05-23 test/runtests.jl

@@ -1,7 +1,7 @@
 using FiniteDiff: finite_difference_gradient
 using FluxContinuation, Test
 
-error_tolerance = 0.04
+error_tolerance = 0.05
 @testset "Gradients" begin
 
 	@testset "Saddle Node" begin

### 2021-05-23 docs/article.tex

@@
     \includegraphics[width=7cm]{bifurcation-measure}
-    \caption{Bifurcation measure $\measure(s)$ and determinant $\Det$ along two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
+    \caption{Bifurcation measure $\measure(s)$ and determinant $\Det$ along the arclength $s$ of two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
     \label{fig:measure}
@@
 Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$.
-Using these quantities we can define a positive semi-definite measure $\measure(u,p)$ of zero crossings in the determinant along the bifurcation curve.
-Figure \ref{fig:measure} demonstrates this measure is maximal at bifurcation at locations $p(\theta)$ and has finite gradients in non-bifurcating regimes.
-The total bifurcation measure $\Psi(\theta)$ is defined as
+Using these quantities we can define a positive semi-definite measure $\measure(s)$ of zero crossings in the determinant along the bifurcation curve.
+Figure \ref{fig:measure} demonstrates this measure is maximal at bifurcation at locations and has finite gradients in non-bifurcating regimes.
+The total bifurcation measure $\Psi(\theta)$ over a diagram is defined as
 \begin{align}
@@
         \int_{\rates(u,p)=0}\!
-        \measure(u,p)\,\mathrm{d}u\mathrm{d}p
+        \measure(s)\,\mathrm{d}s
     }{
         \int_{\rates(u,p)=0}\!\
-        \mathrm{d}u\mathrm{d}p
+        \mathrm{d}s
     }
     \quad\mathrm{where}\quad
-    \measure(u,p):=
+    \measure(s):=
     \left(1+\left|\frac{\Det}{\frac{d}{ds}\Det}\right|\right)^{-1}
@@
 \end{align}
-Here we denote $\int_{\rates(u,p)=0}\mathrm{d}u\mathrm{d}p$ as the sum of the line integrals in $(u,p)\in\Reals^{N+1}$ defined by the level set $\rates(u,p)=0$.
-The measure $\measure(u,p)$ is one at bifurcation points and goes to zero an odd number of times between bifurcations.
-This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(u,p)$ going to zero for each turning point.
+Here we denote $\int_{\rates(u,p)=0}\mathrm{d}s$ as the sum of the line integrals in $(u,p)\in\Reals^{N+1}$ defined by the level set $\rates(u,p)=0$ with $s$ being an arbitrary parametrisation of the curves.
+The measure $\measure(s)$ is one at bifurcation points and goes to zero an odd number of times between bifurcations.
+This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(s)$ going to zero for each turning point.
 While the calculation of the determinant is straightforward, its directional derivative is not.
@@
 
-As the determinant $\Det$ diverges we approach regimes far away from any closest bifurcation and hence $\measure(u,p)\rightarrow0$.
-We would still like to have non-zero gradients with respect to $\theta$ in this regime, and therefore $\measure(u,p)$ was designed to go to zero slowly.
+As the determinant $\Det$ diverges we approach regimes far away from any closest bifurcation and hence $\measure(s)\rightarrow0$.
+We would still like to have non-zero gradients with respect to $\theta$ in this regime, and therefore $\measure(s)$ was designed to go to zero slowly.
 The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the controlled condition region $p$ is densely packed with bifurcations.
-The measure $\measure(u,p)$ is also invariant under certain classes of transformations on $\rates(u,p)$ which do not change bifurcation locations, proofs of which are provided in Appendix \ref{appendix:measure}.
+The measure $\measure(s)$ is also invariant under certain classes of transformations on $\rates(u,p)$ which do not change bifurcation locations, proofs of which are provided in Appendix \ref{appendix:measure}.
 The total measure $\Psi(\theta)$ is added to the supervised term as if it were a likelihood; this defines the semi-supervised cost function as
@@
 \end{align}
-In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to calculate the gradient $\frac{\partial\loss}{\partial\theta}$ without having to differentiate through the operations of the solver that finds the bifurcation curve $\rates(u,p)=0$ and the bifurcation locations $p(\theta)$.
+In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to calculate the gradient $\frac{\partial\loss}{\partial\theta}$ without having to differentiate through the operations of the solver that finds the bifurcation diagram $\rates(u,p)=0$ and the bifurcation locations $p(\theta)$.
 To calculate the gradient of the measure $\frac{\partial \Psi}{\partial\theta}$ we need to differentiate line integrals that depend on $\theta$.
-Fortunately this can be done by the application of the generalised Leibniz integral rule, details of which can be found in Appendix \ref{appendix:space-curve}.
+Fortunately this can be done by the application of the generalised Leibniz integral rule, details of which can be found in Appendix \ref{appendix:leibniz-rule}.
 
@@
 \subsection{Minimal Models}
-Figures \ref{fig:saddle-node:results} and \ref{fig:pitchfork:results} show example optimisations of $(\theta_1,\theta_2)$ for the minimal saddle-node and pitchfork models respectively.
-The magnitude of the correction factor $\lambda=0$ when bifurcations are present and $\lambda\neq0$ otherwise.
+Figures \ref{fig:minimal-models:results} show example optimisations of $(\theta_1,\theta_2)$ for the minimal saddle-node and pitchfork models respectively.
 Optimisation trajectories approach lines of global minima in the cost function, which represent a set of geometrically equivalent models, depicted in the right panels.
-Two bifurcation curves are geometrically equivalent if the number, type and locations of bifurcations match.
+Two bifurcation diagrams are geometrically equivalent if the number, type and locations of bifurcations match the specified targets $\targets$.
 
-We can see that the geometrically equivalent lines are contained within larger basins defined by $\lambda=0$ where the correct number and type of bifurcations are present $p\in\Omega$, but do not match the locations of targets $\targets$.
+We can see that the geometrically equivalent lines are contained within larger basins where the correct number and type of bifurcations are present but do not match the locations of targets $\targets$.
 All models within this basin are in some sense topologically equivalent.
@@
 \includegraphics[width=6cm]{saddle-optima.png}
-\caption{Saddle-node model $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ optimised with respect to targets $\targets=\{-1,1\}$ in observation region $\Omega\in[-2,2]$.
-The right panel shows bifurcations diagrams for the three optima marked by stars on the left panel.
-The optimisation trajectories in white follow the gradient of the cost, approaching the black line of global minima in the left panel}
-\label{fig:saddle-node:results}
-\end{figure}
-
-\begin{figure}
-\centering
 \includegraphics[width=6cm]{pitchfork-landscape.png}
 \includegraphics[width=6cm]{pitchfork-optima.png}
-\caption{Pitchfork model $\rates(u,p) = \theta_{1} + u p +\theta_{2}u^3$ optimised with respect to targets $\targets=\{0\}$ in observation region $\Omega\in[-5,5]$.
-The right panel shows bifurcations diagrams for the three optima marked by stars on the left panel.
+\caption{Saddle-node $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork $\rates(u,p) = \theta_{1} + u p +\theta_{2}u^3$ optimised with respect to $\theta$ so that predicted bifurcations $\predictions(\theta)$ match targets $\targets$ in control condition $p$.
+The right panel shows bifurcations diagrams for the three optimal $\theta^*$ marked by stars on the left panel.
 The optimisation trajectories in white follow the gradient of the cost, approaching the black lines of global minima in the left panel}
-\label{fig:pitchfork:results}
+\label{fig:minimal-models:results}
 \end{figure}
@@
 \subsection{Genetic Toggle Switch}
-In this section we optimise a model where the states share a Hill function relationship with cooperatively $n=2$; these models often emerge from mass action kinetics and are used to model species concentrations.
-After re-scaling the equations governing the dynamics of concentrations, the simplified equations for state $u_$ and $u_2$ become
+In this section we optimise a model where the states share a Hill function relationship with cooperatively $n=2$; these models often emerge from mass action kinetics with quasi-steady state approximations and are used to model species concentrations.
+After re-scaling the equations governing the dynamics of concentrations, the simplified equations for state $u_1$ and $u_2$ become
 \begin{equation}
@@
 \end{equation}
-where $a_k$ are 
- the  we can recover both activator and inhibitor relationships.
-Figure \ref{fig:two-state-optima} shows a UMAP embedding of the optimal parameter basins, represented as point clouds.
+where $a_k$ is the baseline production rate for species $k$ in the absence of the other species.
+Each species has a finite degradation rate $\mu_k$.
+Finally we have two half-saturation constants $p$ and $k$, one of which is chosen as our control condition.
+A baseline production rate $a_k>1$ recovers an inhibitor type hill function for species $k$ and is an activator otherwise.
+The half-saturation are proportional to the slope of the hill activations.
+
+%  the  we can recover both activator and inhibitor relationships.
+% Figure \ref{fig:two-state-optima} shows a UMAP embedding of the optimal parameter basins, represented as point clouds.
 16 distinct clusters are obtained using DBSCAN.
 Each cluster corresponds to a kinetically different model regions that are qualitatively equivalent.
-See Appendix \ref{appendix:clusters} for models that represent the centroid of each cluster.
 

### 2021-05-23 docs/article.tex
Updates from Overleaf

@@
     \item An end-to-end differentiable method for searching for bifurcations and then matching them to user-specified target locations
-    \item Implementation of the method in Julia package leveraging automatic differentiation tooling
+    \item Implementation of the method as a Julia package \cite{Szep2021BifurcationFit.jlDiagrams}
     \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
@@
 The half-saturation are proportional to the slope of the hill activations.
-
-%  the  we can recover both activator and inhibitor relationships.
-% Figure \ref{fig:two-state-optima} shows a UMAP embedding of the optimal parameter basins, represented as point clouds.
-16 distinct clusters are obtained using DBSCAN.
-Each cluster corresponds to a kinetically different model regions that are qualitatively equivalent.
-
+Solving for the steady states and substituting the equation for $u_1$ into $u_2$ gives rise to
+\begin{equation}
+    \mu_2 u_2 = \frac{ a_2 + (\frac{k}{\mu_1}
+    \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 })^2}{ 1 + (\frac{k}{\mu_1} 
+    \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 }
+    )^2 }
+    \label{eq:steady-states}
+\end{equation}
+which reveals that only the ratio between half-saturation and degradation $\frac{k}{\mu_1}$ affects the solutions to this equation, and hence the locations of the bifurcations.
+Figure \ref{fig:minimal-models:results} reveals the results for 800 optimisation runs with 200 ADAM iterations.
+98\% of runs converged onto one of the two clusters: mutual activation and mutual inhibition regimes, examples of which are shown as insets.
 \begin{figure}
@@
 \label{fig:two-state-optima}
+\end{figure}
+\subsection{Complexity}
+
+\begin{figure}
+\centering
+\includegraphics[width=11cm]{scaling}
+\caption{Complexity scaling of calculating the gradient of the cost function}
+\label{fig:scaling}
 \end{figure}

### 2021-05-24 src/Plots.jl
relax test

@@ -2,48 +2,100 @@ using Plots.PlotMeasures
 using LaTeXStrings
 using Plots
 
-import Plots: plot
-function plot(steady_states::Vector{<:Branch},data::StateSpace)
+import Plots: plot,plot!
+function plot(steady_states::Vector{<:Branch},data::StateSpace; determinant=true, kwargs...)
 
-	layout = @layout [a;b{1.0w,0.5h}]
-	default(); default(grid=false,label="",margin=1mm,linewidth=2)
-	figure = plot(layout = layout, link = :x, size=(300,500) )
+	if determinant
 
-	hline!([0],subplot=1,linewidth=0,color=:black, ylabel=L"\mathrm{steady\,states}\quad F_{\theta}(z)=0")
-	hline!([0],subplot=2,linewidth=1,color=:black, xlabel=L"\mathrm{parameter,}p", xmirror=true, topmargin=-5mm,
-		ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|")
-	
-	vline!(data.targets,subplot=1,linewidth=1,color=:gold)
-	vline!(data.targets,subplot=2,linewidth=1,color=:gold)
+		layout = @layout [a;b{1.0w,0.5h}]
+		default(); default( ; grid=false,label="",margin=1mm,linewidth=2,kwargs...)
+		figure = plot(layout = layout, link = :x, size=(300,500) )
 
-    for branch ∈ steady_states
+		hline!([0],subplot=1,linewidth=0,color=:black, ylabel=L"\mathrm{steady\,\,states}\quad F_{\theta}(u,p)=0")
+		hline!([0],subplot=2,linewidth=1,color=:black, xlabel=L"\mathrm{control\,\,condition,}p", xmirror=true, topmargin=-5mm,
+			ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|")
+		
+		vline!(data.targets,subplot=1,linewidth=1,color=:gold)
+		vline!(data.targets,subplot=2,linewidth=1,color=:gold)
 
-        stability = map( s -> all(real(s.λ).<0), branch)
-        determinants = map( s -> prod(real(s.λ)), branch)
+		for branch ∈ steady_states
+
+			stability = map( s -> all(real(s.λ).<0), branch)
+			determinants = map( s -> prod(real(s.λ)), branch)
+			parameter = map( s -> s.z.p, branch)
+
[…]

### 2021-05-24 test/figures.jl

@@ -27,26 +27,26 @@ layout = @layout [a;b{1.0w,0.5h}]
 default(); default(grid=false,label="",margin=0mm)
 plot(layout = layout, link = :x, size=(400,500) )
 
-hline!([0],subplot=2,linewidth=1,ylim=(-5,5),color=:black, xlabel=L"\mathrm{parameter,}p", xmirror=true, topmargin=-5mm,
+hline!([0],subplot=2,linewidth=1,ylim=(-5,5),color=:black, xlabel=L"\mathrm{arclength,}s", xmirror=true, topmargin=-5mm,
 	ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|")
 
-plot!([-1.52,-1.52],[0,1],subplot=1,linewidth=2,color=:gold)
-plot!([-1.52,-1.52],[0,5],subplot=2,linewidth=2,color=:gold)
+plot!([-1.52,-1.52],[0,1],subplot=1,linewidth=2,color=:gray)
+plot!([-1.52,-1.52],[0,5],subplot=2,linewidth=2,color=:gray)
 
-plot!([1.33,1.33],[0,1],subplot=1,linewidth=2,color=:gold)
-plot!([1.33,1.33],[0,5],subplot=2,linewidth=2,color=:gold)
+plot!([1.33,1.33],[0,1],subplot=1,linewidth=2,color=:gray)
+plot!([1.33,1.33],[0,5],subplot=2,linewidth=2,color=:gray)
 
-plot!([1.67,1.67],[0,1],subplot=1,linewidth=2,color=:gold)
-plot!([1.67,1.67],[0,5],subplot=2,linewidth=2,color=:gold)
+plot!([1.67,1.67],[0,1],subplot=1,linewidth=2,color=:gray)
+plot!([1.67,1.67],[0,5],subplot=2,linewidth=2,color=:gray)
 
 p = -2:0.01:2
-f(p) = - 2 - p^3 + p
-plot!( p, measure, subplot=1, color=:gold, fill=true, alpha=0.2, ylabel=L"\mathrm{Measure}\,\,\varphi_{\theta}(z)", ylim=(0,1) )
-plot!( p, f, subplot=2, linewidth=2, color=map( x -> f(x)<0 ? :red : :pink, p ) )
-
 f(p) = - 2 - (p-1/2)^3 + 3.1(p-1/2)
[…]

### 2021-05-24 test/runtests.jl

@@ -1,7 +1,7 @@
 using FiniteDiff: finite_difference_gradient
 using FluxContinuation, Test
 
-error_tolerance = 0.05
+error_tolerance = 0.06
 @testset "Gradients" begin
 
 	@testset "Saddle Node" begin

### 2021-05-26 docs/appendix.tex
Updates from Overleaf

@@
-\section{Supplementary Figures}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=5cm]{implicit-surfaces}
-\caption{Two implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form a space curve which is tangent to field $\tangent(z)$ and perpendicular to gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
-\label{fig:implicit-surfaces}
-\end{figure}
-
-\section{Gradient of Space Curve}
-\label{appendix:space-curve}
-
-Suppose there exists a one dimensional space curve $\mathcal{C(\theta)}$ embedded in $z\in\Reals^{N+1}$ whos geometry changes depending on input parameters $\theta\in\Reals^M$.
-This curve could be open or closed and changes in $\theta$ could change the curve topology as well.
-Let the function $\gamma_{\theta}:\Reals\rightarrow\Reals^{N+1}$ be a parameterisation of the position vector along the curve within a fixed domain $s\in\mathcal{S}$.
-Note that the choice of parameterisation is arbitrary and our results should not depend on this choice.
-Furthermore, if we parametrise the curve $\mathcal{C}(\theta)$ with respect to a fixed domain $\mathcal{S}$ the dependence on $\theta$ is picked up by the parameterisation $\gamma_{\theta}(s)$.
-We can write a line integral of any scalar function $L_{\theta}:\Reals^{N+1}\rightarrow\Reals$ on the curve as
-\begin{align}
-    L(\theta):=
-    \int_\mathcal{C(\theta)}\! L_{\theta}(z)\,\mathrm{d}z
-    =\int_\mathcal{S}\! L_{\theta}(z)\left|\frac{d\gamma_{\theta}}{ds}\right|\mathrm{d}s_{\,\,z=\gamma_{\theta}(s)}
-\end{align}
-where $\left|\frac{d\gamma_{\theta}}{ds}\right|$ is the magnitude of tangent vectors to the space curve and we remind ourselves that the integrand is evaluated at $z=\gamma_{\theta}(s)$.
-We would like to track how this integral changes with respect to $\theta$.
-The total derivative with respect to $\theta$ can be propagated into the integrand \cite{Flanders1973DifferentiationSign} as long as we keep track of implicit dependencies
-\begin{align}
-    \frac{dL}{d\theta} &=\int_\mathcal{S}
-    \left|\frac{d\gamma_{\theta}}{ds}\right|
-    \left(
-        \frac{\partial L}{\partial\theta}+
-        \frac{\partial L}{\partial z}\cdot
-        \frac{dz}{d\theta}
-    \right)
-    +L_{\theta}(z)\frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|
-    \mathrm{d}s_{\,\,z=\gamma_{\theta}(s)}
-    % \\&=\int_\mathcal{C(\theta)}
-    %     \frac{\partial L}{\partial\theta}+
-    %     \frac{\partial L}{\partial z}\cdot
-    %     \frac{dz}{d\theta}
-    % +L_{\theta}(z)\frac{d}{d\theta}\log\left|\frac{d\gamma_{\theta}}{ds}\right|
-    % \mathrm{d}z
-\end{align}
-Here we applied the total derivative rule in the first term due to the implicit dependence of $z$ on $\theta$ through $z=\gamma_{\theta}(s)$. Applying the chain rule to the second term
-\begin{align}
-    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
-    \left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}
-    \frac{d\gamma_{\theta}}{ds}\cdot\frac{d}{d\theta}
-    \left(\frac{d\gamma_{\theta}}{ds}\right)
-\end{align}
-By choosing an $s$ that has no implicit $\theta$ dependence we can commute derivatives
-\begin{align}
-    \frac{d}{d\theta}\left(\frac{d\gamma_{\theta}}{ds}\right)
-    = \frac{d}{ds}\left(\frac{d\gamma_{\theta}}{d\theta}\right)
-    \quad\Rightarrow\quad
-    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
-    \left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}
-    \frac{d\gamma_{\theta}}{ds}\cdot\frac{d}{d s}
-    \left(\frac{d\gamma_{\theta}}{d\theta}\right)
-\end{align}
-To proceed we note that the unit tangent vector can be written as an evaluation of a tangent field $\hat{T}_{\theta}(z)$ defined in the whole domain $z\in\Reals^{N+1}$ along the parametric curve $z=\gamma_{\theta}(s)$.
-The unit tangent field may disagree with the tangent given by $\frac{d\gamma_{\theta}}{ds}$ up to a sign
-\begin{align}
-    \left.\hat{\tangent}(z)\right|_{z=\gamma_{\theta}(s)}=
-    \pm\left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}\frac{d\gamma_{\theta}}{d s}
-\end{align}
-this leads to numerically verified result%\todo{perhaps not valid in general?}
-\begin{align}
-    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
-    \left|\frac{d\gamma_{\theta}}{ds}\right|\left(
-   \hat{\tangent}(z)\cdot\frac{\partial }{\partial z}\left(\frac{d\Gamma_{\theta}}{d \theta}\right)\cdot
-   \hat{\tangent}(z)
-    \right)_{z=\gamma_\theta(s)}
-    \label{eq:divergence-term}
-\end{align}
-
-\section{Deformation of Implicit Surfaces}
-\label{appendix:deformation}
-It is possible to find the normal deformation of the implicit space curves due to changes in $\theta$. This can be done by taking the total derivative of the implicit equation defining the level set
-\begin{align}
-    \frac{d\rates(z)}{d\theta}=\frac{\partial F}{\partial\theta}+
-    \frac{\partial F}{\partial z}\cdot\frac{d z}{d \theta}
-\end{align}
-
-We can rearrange for $\frac{d z}{d \theta}$ using the Moore-Penrose inverse of the rectangular Jacobian matrix $\frac{\partial F}{\partial z}$ which appeared in equation \eqref{eq:tangent-field}.
-Since the level set is defined by $\rates(z)=0$ the total derivative along the level set $d\rates(z)=0$ and we arrive at an expression for the deformation field \cite{Jos2011OnSurface}
-\begin{align}
-    \frac{d z}{d \theta} = - \frac{\partial F}{\partial z}^\top
-    \left(\,
-        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
-    \right)^{-1}
-    \frac{\partial F}{\partial\theta}
-\end{align}
-The tangential component of the deformation field is not uniquely determined because there is no unique way of parameterising a surface.
-This is the subject of many computer graphics papers \cite{Jos2011OnSurface,Tao2016Near-IsometricTracking,Fujisawa2013CalculationInvariance}.
-We are however not interested in the continuous propagation of a mesh - as is the subject of those papers.
-In fact we are looking for a deformation field that is orthogonal to the tangent vector $\hat{\tangent}(z) \cdot\frac{d z}{d\theta} =0$ for the space curve, and therefore letting the tangential component of the deformation equal zero is a valid choice and we can it instead of the parameterised deformation
-% \todo{this is possibly quite shifty. Is this really valid?}
-\begin{align}
-    \frac{d \gamma_{\theta}}{d\theta} \rightarrow \frac{d z}{d\theta}
-\end{align}
-To summarise we now have the gradient of our line integral only in terms of the implicit function defining the integration region.
-\begin{align}
-    \frac{d L}{d\theta} =\int_{\rates(z)=0}
-        \frac{\partial L}{\partial\theta}+
-        \frac{\partial L}{\partial z}\cdot
-        \varphi_{\theta}(z)
-    +L_{\theta}(z)\,\,
-    \hat{\tangent}(z)\cdot\frac{\partial \varphi}{\partial z}\cdot\hat{\tangent}(z)
-    \,\mathrm{d}z\qquad\qquad\\
-    \mathrm{where}\quad
-    \hat{\tangent}(z):= \frac{\tangent(z)}{|\tangent(z)|}
-    \qquad
-    \tangent(z):=
-    \left|\begin{matrix}
-        \hat{z} \\
-        \,\partial_{z}\rates\,
-    \end{matrix}\right|
-    \qquad
-    \varphi_{\theta}(z) :=
-- \frac{\partial F}{\partial z}^\top
-    \left(\,
-        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
-    \right)^{-1}
-    \frac{\partial F}{\partial\theta}
-\end{align}
-We have settled on choosing normal deformations which we will call $\varphi_\theta(z)$.
-The above result can be seen a the generalised Leibniz rule \cite{Flanders1973DifferentiationSign} for the case of line integration regions.
-The last integrand term can be seen as the divergence the vector field $\varphi_\theta(z)$ projected onto the one dimensional space curve.
-
-\subsection{Bifurcation Curves as Tangent Fields}
+\section{Bifurcation Diagrams as Tangent Fields}
 \label{appendix:tangent-fields}
@@
 \end{align}
+\begin{figure}[H]
+\centering
+\includegraphics[width=5cm]{implicit-surfaces}
+\caption{Two implicit surfaces $f_{\theta}(z)=0$ and $g_{\theta}(z)=0$ in $\mathbb{R}^3$ intersecting to form a space curve which is tangent to field $\tangent(z)$ and perpendicular to gradients $\partial_{z}f_{\theta}$ and $\partial_{z}g_{\theta}$}
+\label{fig:implicit-surfaces}
+\end{figure}
 An expression for the field $\tangent(z)$ tangent to the set of curves would allow us to take derivatives and integrals along the bifurcation curve.
@@
 
-\section{Bifurcation Measure}
+\section{Properties of the Bifurcation Measure}
 \label{appendix:measure}
+
+\section{Leibniz Rule for Space Curves}
+\label{appendix:leibniz-rule}
+
+Suppose there exists a one dimensional space curve $\mathcal{C(\theta)}$ embedded in $z\in\Reals^{N+1}$ whos geometry changes depending on input parameters $\theta\in\Reals^M$.
+This curve could be open or closed and changes in $\theta$ could change the curve topology as well.
+Let the function $\gamma_{\theta}:\Reals\rightarrow\Reals^{N+1}$ be a parameterisation of the position vector along the curve within a fixed domain $s\in\mathcal{S}$.
+Note that the choice of parameterisation is arbitrary and our results should not depend on this choice.
+Furthermore, if we parametrise the curve $\mathcal{C}(\theta)$ with respect to a fixed domain $\mathcal{S}$ the dependence on $\theta$ is picked up by the parameterisation $\gamma_{\theta}(s)$.
+We can write a line integral of any scalar function $L_{\theta}:\Reals^{N+1}\rightarrow\Reals$ on the curve as
+\begin{align}
+    L(\theta):=
+    \int_\mathcal{C(\theta)}\! L_{\theta}(z)\,\mathrm{d}z
+    =\int_\mathcal{S}\! L_{\theta}(z)\left|\frac{d\gamma_{\theta}}{ds}\right|\mathrm{d}s_{\,\,z=\gamma_{\theta}(s)}
+\end{align}
+where $\left|\frac{d\gamma_{\theta}}{ds}\right|$ is the magnitude of tangent vectors to the space curve and we remind ourselves that the integrand is evaluated at $z=\gamma_{\theta}(s)$.
+We would like to track how this integral changes with respect to $\theta$.
+The total derivative with respect to $\theta$ can be propagated into the integrand \cite{Flanders1973DifferentiationSign} as long as we keep track of implicit dependencies
+\begin{align}
+    \frac{dL}{d\theta} &=\int_\mathcal{S}
+    \left|\frac{d\gamma_{\theta}}{ds}\right|
+    \left(
+        \frac{\partial L}{\partial\theta}+
+        \frac{\partial L}{\partial z}\cdot
+        \frac{dz}{d\theta}
+    \right)
+    +L_{\theta}(z)\frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|
+    \mathrm{d}s_{\,\,z=\gamma_{\theta}(s)}
+\end{align}
+Here we applied the total derivative rule in the first term due to the implicit dependence of $z$ on $\theta$ through $z=\gamma_{\theta}(s)$. Applying the chain rule to the second term
+\begin{align}
+    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
+    \left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}
+    \frac{d\gamma_{\theta}}{ds}\cdot\frac{d}{d\theta}
+    \left(\frac{d\gamma_{\theta}}{ds}\right)
+\end{align}
+By choosing an $s$ that has no implicit $\theta$ dependence we can commute derivatives
+\begin{align}
+    \frac{d}{d\theta}\left(\frac{d\gamma_{\theta}}{ds}\right)
+    = \frac{d}{ds}\left(\frac{d\gamma_{\theta}}{d\theta}\right)
+    \quad\Rightarrow\quad
+    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
+    \left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}
+    \frac{d\gamma_{\theta}}{ds}\cdot\frac{d}{d s}
+    \left(\frac{d\gamma_{\theta}}{d\theta}\right)
+\end{align}
+To proceed we note that the unit tangent vector can be written as an evaluation of a tangent field $\hat{T}_{\theta}(z)$ defined in the whole domain $z\in\Reals^{N+1}$ along the parametric curve $z=\gamma_{\theta}(s)$.
+The unit tangent field may disagree with the tangent given by $\frac{d\gamma_{\theta}}{ds}$ up to a sign
+\begin{align}
+    \left.\hat{\tangent}(z)\right|_{z=\gamma_{\theta}(s)}=
+    \pm\left|\frac{d\gamma_{\theta}}{ds}\right|^{-1}\frac{d\gamma_{\theta}}{d s}
+\end{align}
+this leads to
+\begin{align}
+    \frac{d}{d\theta}\left|\frac{d\gamma_{\theta}}{ds}\right|=
+    \left|\frac{d\gamma_{\theta}}{ds}\right|\left(
+   \hat{\tangent}(z)\cdot\frac{\partial }{\partial z}\left(\frac{d\Gamma_{\theta}}{d \theta}\right)\cdot
+   \hat{\tangent}(z)
+    \right)_{z=\gamma_\theta(s)}
+    \label{eq:divergence-term}
+\end{align}
+It is possible to find the normal deformation of the implicit space curves due to changes in $\theta$. This can be done by taking the total derivative of the implicit equation defining the level set
+\begin{align}
+    \frac{d\rates(z)}{d\theta}=\frac{\partial F}{\partial\theta}+
+    \frac{\partial F}{\partial z}\cdot\frac{d z}{d \theta}
+\end{align}
+We can rearrange for $\frac{d z}{d \theta}$ using the Moore-Penrose inverse of the rectangular Jacobian matrix $\frac{\partial F}{\partial z}$ which appeared in equation \eqref{eq:tangent-field}.
+Since the level set is defined by $\rates(z)=0$ the total derivative along the level set $d\rates(z)=0$ and we arrive at an expression for the deformation field \cite{Jos2011OnSurface}
+\begin{align}
+    \frac{d z}{d \theta} = - \frac{\partial F}{\partial z}^\top
+    \left(\,
+        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
+    \right)^{-1}
+    \frac{\partial F}{\partial\theta}
+\end{align}
+The tangential component of the deformation field is not uniquely determined because there is no unique way of parameterising a surface.
+This is the subject of many computer graphics papers \cite{Jos2011OnSurface,Tao2016Near-IsometricTracking,Fujisawa2013CalculationInvariance}.
+We are however not interested in the continuous propagation of a mesh - as is the subject of those papers.
+In fact we are looking for a deformation field that is orthogonal to the tangent vector $\hat{\tangent}(z) \cdot\frac{d z}{d\theta} =0$ for the space curve, and therefore letting the tangential component of the deformation equal zero is a valid choice and we can it instead of the parameterised deformation
+\begin{align}
+    \frac{d \gamma_{\theta}}{d\theta} \rightarrow \frac{d z}{d\theta}
+\end{align}
+To summarise we now have the gradient of our line integral only in terms of the implicit function defining the integration region.
+\begin{align}
+    \frac{d L}{d\theta} =\int_{\rates(z)=0}
+        \frac{\partial L}{\partial\theta}+
+        \frac{\partial L}{\partial z}\cdot
+        \varphi_{\theta}(z)
+    +L_{\theta}(z)\,\,
+    \hat{\tangent}(z)\cdot\frac{\partial \varphi}{\partial z}\cdot\hat{\tangent}(z)
+    \,\mathrm{d}z\qquad\qquad\\
+    \mathrm{where}\quad
+    \hat{\tangent}(z):= \frac{\tangent(z)}{|\tangent(z)|}
+    \qquad
+    \tangent(z):=
+    \left|\begin{matrix}
+        \hat{z} \\
+        \,\partial_{z}\rates\,
+    \end{matrix}\right|
+    \qquad
+    \varphi_{\theta}(z) :=
+- \frac{\partial F}{\partial z}^\top
+    \left(\,
+        \frac{\partial F}{\partial z}\,\frac{\partial F}{\partial z}^\top
+    \right)^{-1}
+    \frac{\partial F}{\partial\theta}
+\end{align}
+We have settled on choosing normal deformations which we will call $\varphi_\theta(z)$.
+The above result can be seen a the generalised Leibniz rule \cite{Flanders1973DifferentiationSign} for the case of line integration regions.
+The last integrand term can be seen as the divergence the vector field $\varphi_\theta(z)$ projected onto the one dimensional space curve.
 

### 2021-05-26 docs/checklist.tex
Updates from Overleaf

@@
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
 \begin{enumerate}
@@
   \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
+    \answerYes{}
   \item Did you describe the limitations of your work?
-    \answerTODO{}
+    \answerYes{Implementation only covers co-dimension one bifurcations. Does not scale well for PDEs. See discussion in Section \ref{}}
   \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
+    \answerNA{}
   \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
+    \answerYes{}
 \end{enumerate}
@@
   \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
+    \answerNA{This paper applies existing theorems. See Section \ref{section:method}}
 	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
+    \answerNA{We include derivations to guide the reader in Appendices}
 \end{enumerate}
@@
   \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
+    \answerYes{All figures are unit tests in the GitHub repository for \texttt{BifurcationFit.jl}}
   \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
+    \answerYes{ADAM was used with various hyper-parameters}
 	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
+    \answerYes{See Figure \ref{fig:scaling} and }
 	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
+    \answerYes{See Figure \ref{fig:scaling}}
 \end{enumerate}
@@
   \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
+    \answerYes{\cite{Veltz2019PseudoArcLengthContinuation.jl,Revels2016Forward-ModeJulia,Flux}}
   \item Did you mention the license of the assets?
-    \answerTODO{}
+    \answerNo{All assets have been release under MIT}
   \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
+    \answerYes{All code available under \texttt{github.com/gszep/BifurcationFit.jl} }
   \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
+    \answerNA{}
   \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
+    \answerNA{}
 \end{enumerate}
@@
   \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
+    \answerNA{}
   \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
+    \answerNA{}
   \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
+    \answerNA{}
 \end{enumerate}

### 2021-05-26 docs/article.tex

@@
 The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
-Drawing inspiration from implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients we find that their computation shares the same  complexity as the algorithm used to calculate the bifurcation diagram.
+Drawing inspiration from implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients.
 The compute the diagram we use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl} developed for partial differential equations.
-In the case of partial differential equations the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
+In the case of partial differential equations, the computational complexity of calculating even a single bifurcation diagram can be unbounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
 The complexity of computing a single branch, however, is bounded by the complexities of the chosen eigenvalue solver and corrector, and further decreased by adaptive stepping procedures \cite{Aruliah2016AlgorithmContinuation}.
@@
 \begin{itemize}
-    \item An end-to-end differentiable method for searching for bifurcations and then matching them to user-specified target locations
+    \item An end-to-end differentiable method for locating bifurcations in parameter space and then matching their dependency on a control condition to user-specified locations
     \item Implementation of the method as a Julia package \cite{Szep2021BifurcationFit.jlDiagrams}
@@
 The determinant of the Jacobian quantifies the rate at which trajectories in a local patch of state-space $u\in\Reals^N$ converge or diverge.
-The determinant approaching zero means that the dynamics of the system is slowing down, which is an important indicator for the onset of a transition between qualitative behaviours.
+The determinant approaching zero means that the dynamics of the system are slowing down, which is an important indicator for the onset of a transition between qualitative behaviours.
 Furthermore, the slowing down must necessarily be followed by breakdown of stability; for this to be true it is sufficient to require that the determinant cross zero with a finite slope, meaning that its directional derivative along the diagram $\frac{d}{ds}\Det$ is not zero.
@@
 \end{align}
- Figure \ref{fig:minimal-models} shows minimal models that explore the space of saddle-nodes $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ and pitchforks $\rates(u,p) = \theta_{1} + p u+\theta_{2}u^3$.
+
+To illustrate how the shapes of bifurcation diagrams compare with the determinant of the Jacobian, we show minimal models for saddle-node (Figure \ref{fig:minimal-models}A) and pitchfork bifurcations (Figure \ref{fig:minimal-models}B).
 Indeed the predictions $\predictions(\theta)$ are defined by zero crossings in the determinant with a finite slope.
@@
 
-For a given set of parameters $\theta$ one could compute the set of predicted bifurcations $\predictions(\theta)$ using parameter continuation methods \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
-Our goal is to find optimal parameters $\theta^*$ that match predictions $\predictions(\theta^*)$ to specified targets $\targets$.
-We must design a suitable cost function $\loss$ so that
-\begin{align}
-    \theta^* := \mathrm{argmin}_{\theta} \loss(\theta|\targets)
-    \label{eq:optimal-theta}
-\end{align}
-The optimal $\theta^*$ is not expected to always be unique, but is in general a manifold representing the space of qualitatively equivalent models.
-The cost function $\loss$ must have some distance measure between the two sets $\predictions(\theta)$ and $\targets$ with a bias that encourages equivalent set sizes $|\predictions(\theta)|=|\targets|$.
-This is especially important in the case where there are no predictions $|\predictions(\theta)|=0$.
-
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 
+For a given set of parameters $\theta$ one could compute the set of predicted bifurcations $\predictions(\theta)$ using parameter continuation methods \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
+Our goal is to find optimal parameters $\theta^*$ that match predictions $\predictions(\theta^*)$ to specified targets $\targets$.
+We must design a suitable cost function $\loss$ so that
+\begin{align}
+    \theta^* := \mathrm{argmin}_{\theta} \loss(\theta|\targets)
+    \label{eq:optimal-theta}
+\end{align}
+The optimal $\theta^*$ is not expected to always be unique, but is in general a manifold representing the space of qualitatively equivalent models.
+Ideally, the cost function $\loss$ should reward $\theta$ for which the number of predicted bifurcations is equal to the number of targets, $|\predictions(\theta)|=|\targets|$.
+This is especially important in the case where there are no predictions $|\predictions(\theta)|=0$.
+
+\clearpage
 \section{Proposed Method}
 \subsection{Semi-supervised Cost Function}
+
+To identify parameter sets that give rise to bifurcation diagrams with specified turning points, we propose a semi-supervised cost function that comprises two terms.
+The role of the supervised term is simply to reward predicted bifurcations to coincide with the specified target locations.
+This of course relies on such bifurcations existing.
+Therefore, the role of the unsupervised term is to encourage an optimiser to move towards parameter regimes that do exhibit bifurcations.
+
+\subsubsection{Supervised term: matched bifurcations to target locations}
 
@@
 In order for predicted bifurcations $p(\theta)\in\predictions(\theta)$ to match targets $p'\in\targets$ we need to evaluate an error term $|p(\theta)-p'|$.
-We expect this term to be some mean over the targets $\targets$ and predictions $\predictions(\theta)$.
-We choose a geometric mean over the predictions and an arithmetic mean over targets to account for cases where $|\predictions|\neq|\targets|$ and discourage multiple predictions matching the same target in cases where $|\predictions|\geq|\targets|$.
-This term only zero when each target is matched by at least one prediction
+A naive approach might take an average over the norms for each prediction-target pair.
+However, this can result in multiple predictions matching the same target.
+Therefore, we choose a geometric mean over the predictions and an arithmetic mean over targets to account for cases where $|\predictions|\neq|\targets|$ and discourage multiple predictions matching the same target (in cases where $|\predictions|\geq|\targets|$).
+Such an error term is only zero when each target is matched by at least one prediction
 \begin{align}
@@
 \end{align}
+
+\subsubsection{Unsupervised term: encouraging bifurcations}
+
 We can see from Figures \ref{fig:minimal-models} and definitions \eqref{eq:predictions} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant $\Det=0$.
 Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$.
-Using these quantities we can define a positive semi-definite measure $\measure(s)$ of zero crossings in the determinant along the bifurcation curve.
-Figure \ref{fig:measure} demonstrates this measure is maximal at bifurcation at locations and has finite gradients in non-bifurcating regimes.
-The total bifurcation measure $\Psi(\theta)$ over a diagram is defined as
-\begin{align}
+Using these quantities we can define a positive semi-definite measure $\measure(s)$ of zero crossings in the determinant along the bifurcation curve, which we define as
+\begin{equation}
+    \measure(s):=
+    \left(1+\left|\frac{\Det}{\frac{d}{ds}\Det}\right|\right)^{-1}
+    \label{eq:measure}
+\end{equation}
+This measure is maximal at bifurcations and has finite gradients in non-bifurcating regimes (Figure \ref{fig:measure}).
+More specifically, the measure $\measure(s)$ is one at bifurcation points and goes to zero an odd number of times between bifurcations.
+This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(s)$ going to zero for each turning point.
+
+To quantify the overall propensity for bifurcations over a diagram, we define the total bifurcation measure $\Psi(\theta)$ as
+\begin{equation}
     \Psi(\theta):=\frac{
@@
     }
-    \quad\mathrm{where}\quad
-    \measure(s):=
-    \left(1+\left|\frac{\Det}{\frac{d}{ds}\Det}\right|\right)^{-1}
-    \label{eq:measure}
-\end{align}
+\end{equation}
 Here we denote $\int_{\rates(u,p)=0}\mathrm{d}s$ as the sum of the line integrals in $(u,p)\in\Reals^{N+1}$ defined by the level set $\rates(u,p)=0$ with $s$ being an arbitrary parametrisation of the curves.
-The measure $\measure(s)$ is one at bifurcation points and goes to zero an odd number of times between bifurcations.
-This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(s)$ going to zero for each turning point.
+
 While the calculation of the determinant is straightforward, its directional derivative is not.
@@
 \end{align}
+
+\subsection{Differentiating the semi-supervised cost function}
+
 The pre-factor $|\targets|-|\predictions|$ in the unsupervised term ensures that the gradients are always pushing optimisers towards a state where $|\targets|=|\predictions|$.
 This can be seen as a step-wise annealing of the unsupervised term until the desired state is reached.
-Note that in while individual bifurcations $p(\theta)$ depend smoothly on $\theta$, the total number of predictions $|\predictions|$ does not have gradient contributions with respect to $\theta$.
+Note that while individual bifurcations $p(\theta)$ depend smoothly on $\theta$, the total number of predictions $|\predictions|$ does not have gradient contributions with respect to $\theta$.
 We can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ and total measure $\Psi(\theta)$
@@
 
-The gradient of the bifurcation points $\frac{\partial p}{\partial\theta}$ is found by application of the implicit function theorem to a vector function $G_{\theta}:\Reals^{N+1}\rightarrow\Reals^{N+1}$ whos components represent the two constraints $\rates(u,p)=0$ and $\Det=0$.
+The gradient of the bifurcation points $\frac{\partial p}{\partial\theta}$ is found by application of the implicit function theorem to a vector function $G_{\theta}:\Reals^{N+1}\rightarrow\Reals^{N+1}$ whose components represent the two constraints $\rates(u,p)=0$ and $\Det=0$.
 By following a similar strategy to that used by implicit layers \cite{Look2020DifferentiableLayers} we yield an $(N+1)\times M$ Jacobian representing a deformation field \cite{Jos2011OnSurface} for each $\theta$ direction.
@@
 Calculation of this matrix involves inverting an $(N+1)\times(N+1)$ Jacobian $\frac{\partial G_{\theta}}{\partial(u,p)}$.
-The determinant of this Jacobian goes to zero in the degenerate case where $\frac{d}{ds}\Det=0$.
-This further justifies our choice for measure $\Psi(\theta)$ which discourages the degenerate case.
+The determinant of this Jacobian goes to zero in the degenerate case where $\frac{d}{ds}\Det=0$, further justifying our choice of measure $\Psi(\theta)$ which discourages the degenerate case.
 
 The cost function is piece-wise smooth and differentiable with undefined gradients only in parameter contours where the number of predictions $|\predictions|$ changes; this is when $\Psi(\theta)$ is undefined and the inverse of $\frac{\partial G_{\theta}}{\partial (u,p)}$ does not exist.
-Given a set of solutions to $\rates(u,p)=0$ and locations $p(\theta)$ the gradient $\frac{\partial\loss}{\partial\theta}$ can evaluated using automatic differentiation methods without needing to back-propagate through the solver that obtained the level set $\rates(u,p)=0$ in the forward pass.
+Given a set of solutions to $\rates(u,p)=0$ and locations $p(\theta)$ the gradient $\frac{\partial\loss}{\partial\theta}$ can be evaluated using automatic differentiation methods \cite{Revels2016Forward-ModeJulia,Flux} without needing to back-propagate through the solver that obtained the level set $\rates(u,p)=0$ in the forward pass.
+
+%\clearpage
 
 \section{Experiments \& Results}
+In this section, we apply the method first to minimal examples that can produce saddle-node and pitchfork bifurcations, and then a more complex model that has multiple parametric regimes that produce saddle-node bifurcations.
 
@@
 A baseline production rate $a_k>1$ recovers an inhibitor type hill function for species $k$ and is an activator otherwise.
-The half-saturation are proportional to the slope of the hill activations.
-Solving for the steady states and substituting the equation for $u_1$ into $u_2$ gives rise to
+The half-saturation are proportional to the slope of the hill productions.
+Solving for the steady states,  substituting the equation for $u_1$ into $u_2$ and rearranging gives rise to the relationship
+% \begin{equation}
+%     \mu_2 u_2 = \frac{ a_2 + (\frac{k}{\mu_1}
+%     \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 })^2}{ 1 + (\frac{k}{\mu_1} 
+%     \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 }
+%     )^2 }
+%     \label{eq:steady-states}
+% \end{equation}
 \begin{equation}
-    \mu_2 u_2 = \frac{ a_2 + (\frac{k}{\mu_1}
-    \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 })^2}{ 1 + (\frac{k}{\mu_1} 
-    \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 }
-    )^2 }
-    \label{eq:steady-states}
+    \dfrac{k}{\mu_1} = \dfrac{(1 + (p u_2)^2) \sqrt{a_2 - \mu_2 u_2} }{ (a_1 + (p u_2)^2) \sqrt{ \mu_2 u_2 - 1 } }
+    \label{eq:steady-states-simplified}
 \end{equation}
+
 which reveals that only the ratio between half-saturation and degradation $\frac{k}{\mu_1}$ affects the solutions to this equation, and hence the locations of the bifurcations.
-Figure \ref{fig:minimal-models:results} reveals the results for 800 optimisation runs with 200 ADAM iterations.
+Figure \ref{fig:two-state-optima} reveals the results for 800 optimisation runs with 200 ADAM iterations.
 98\% of runs converged onto one of the two clusters: mutual activation and mutual inhibition regimes, examples of which are shown as insets.
@@
 \centering
-\includegraphics[width=11cm]{two-state-optima}
+%\includegraphics[width=11cm]{two-state-optima}
+\includegraphics[width=11cm]{two-state-optima.png}
 \caption{Optimal parameter estimates $\theta^*$ for the two state model \eqref{eq:two-state} for the targets $\targets=\{4,5\}$ reveal two clusters of qualitatively different regimes: mutual activation and mutual inhibition.
@@
 \end{figure}
+
 \subsection{Complexity}
+Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods and the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
+We note that while for the supervised term only as many inversions as there are targets are performed, the unsupervised term has one inversion per point along the diagram.
+This is due to the implicit derivatives taken when applying the generalised Leibniz rule.
+
+Figure \ref{fig:scaling} reveals the $N^3$ complexity for computing the gradient, suggesting that indeed the matrix inversion is the bottleneck.
+This can be addressed by approximating the inverse Jacobian with low-rank updates, as is done with deep equilibrium models \cite{Bai2019DeepModels}.
 
@@
 
-\section{Conclusion \& Broader Impact | \textit{in progress}}
+\section{Conclusion \& Broader Impact}
+
+In this work, we proposed a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
+The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
+Drawing inspiration from implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients.
+The compute the diagram we use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl} developed for partial differential equations.
+In the case of partial differential equations the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
+The complexity of computing a single branch, however, is bounded by the complexities of the chosen eigenvalue solver and corrector, and further decreased by adaptive stepping procedures \cite{Aruliah2016AlgorithmContinuation}.
+
+We find that the cost function landscape contains basins that not only allow us to synthesise models with a desired bifurcation diagram but also allow us to organise models in terms of topological and geometric equivalence.
+We discuss the relevance of this in model selection.
+In summary, our paper has the following main contributions:
+
 \begin{itemize}
@@
 
-% \section*{Checklist}
-% \input{checklist}
-
+\clearpage
 \bibliography{refs}
@@
 
+\clearpage
 \section*{Appendix}
@@
 
+\clearpage
+\section*{Checklist}
+\input{checklist}
 \end{document}

### 2021-05-26 docs/article.tex

@@
 
-To illustrate how the shapes of bifurcation diagrams compare with the determinant of the Jacobian, we show minimal models for saddle-node (Figure \ref{fig:minimal-models}A) and pitchfork bifurcations (Figure \ref{fig:minimal-models}B).
+To illustrate how the shapes of bifurcation diagrams compare with the determinant of the Jacobian, we show minimal models for saddle-node (Figure \ref{fig:saddle-node}) and pitchfork bifurcations (Figure \ref{fig:pitchfork}).
 Indeed the predictions $\predictions(\theta)$ are defined by zero crossings in the determinant with a finite slope.
@@
 \begin{picture}(0,0) \put(-5.8,10){\subref{fig:pitchfork}} \end{picture}
-\caption{Saddle-node model \ref{fig:saddle-node} $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork model \ref{fig:pitchfork} $\rates(u,p) = \theta_{1} + p u+\theta_{2}u^3$ with $\theta=(5/2,-1),(1/2,-1)$ and targets $\targets=\{-1,1\},\{0\}$ respectively.
-Lighter shades indicate the determinant crossing zero at locations $\predictions(\theta)$ giving rise to unstable solutions}
+\caption{Illustration of bifurcation diagrams for minimal models of bifurcations.
+\ref{fig:saddle-node} Saddle-node bifurcations arise for $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ when $\theta = (5/2,-1)$.
+\ref{fig:pitchfork} Pitchfork bifurcations arise for $\rates(u,p) = \theta_{1} + p u+\theta_{2}u^3$ when $\theta=(1/2,-1)$.
+Targets are illustrated by light yellow vertical lines.
+Bifurcation curves are shown as solid blue and red lines, with lighter shades indicating the determinant crossing zero at locations $\predictions(\theta)$ giving rise to unstable solutions.}
 \label{fig:minimal-models}
@@
 Solving for the steady states,  substituting the equation for $u_1$ into $u_2$ and rearranging gives rise to the relationship
-% \begin{equation}
-%     \mu_2 u_2 = \frac{ a_2 + (\frac{k}{\mu_1}
-%     \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 })^2}{ 1 + (\frac{k}{\mu_1} 
-%     \frac{ a_1 + (p u_2)^2}{ 1 + (p u_2)^2 }
-%     )^2 }
-%     \label{eq:steady-states}
-% \end{equation}
 \begin{equation}
     \dfrac{k}{\mu_1} = \dfrac{(1 + (p u_2)^2) \sqrt{a_2 - \mu_2 u_2} }{ (a_1 + (p u_2)^2) \sqrt{ \mu_2 u_2 - 1 } }
-    \label{eq:steady-states-simplified}
+    \label{eq:steady-states}
 \end{equation}
-
 which reveals that only the ratio between half-saturation and degradation $\frac{k}{\mu_1}$ affects the solutions to this equation, and hence the locations of the bifurcations.
@@
 \subsection{Complexity}
-Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods and the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
-We note that while for the supervised term only as many inversions as there are targets are performed, the unsupervised term has one inversion per point along the diagram.
-This is due to the implicit derivatives taken when applying the generalised Leibniz rule.
+Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods, the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
+However instead of evaluating the inversions directly, we solve a system of linear equations, applying the same strategy as implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels}.
+This leaves us with the computational bottleneck of calculating the determinant of the state space Jacobian, required in the both the bifurcation measure \eqref{eq:measure} and gradient \eqref{eq:gradient-bifurcation}.
+This calculation scales like $N^3$ where $N$ is the number of state space variables (Figure \ref{fig:scaling}).
 
-Figure \ref{fig:scaling} reveals the $N^3$ complexity for computing the gradient, suggesting that indeed the matrix inversion is the bottleneck.
-This can be addressed by approximating the inverse Jacobian with low-rank updates, as is done with deep equilibrium models \cite{Bai2019DeepModels}.
+While still tractable on laptop computers for states $N<100$ our implementation currently does not scale well for partial differential equations where a large the number of states $N$ arises from discretisation of the spatial variables.
+The only reason we need this determinant is because it is an indicator of bifurcations.
+We can address the computational bottleneck by finding a more computationally efficient way of calculating this indicator.
+One approach would be to take the product of a finite subset of eigenvalues of the system.
 
@@
 \includegraphics[width=11cm]{scaling}
-\caption{Complexity scaling of calculating the gradient of the cost function}
+\caption{Complexity scaling of calculating the gradient of the cost function. Calculations were performed on an Intel Core i7-6700HQ CPU @ 2.60GHz x 8 without GPU acceleration}
 \label{fig:scaling}

### 2021-05-26 docs/article.tex

@@
 \begin{equation}
-    \dfrac{k}{\mu_1} = \dfrac{(1 + (p u_2)^2) \sqrt{a_2 - \mu_2 u_2} }{ (a_1 + (p u_2)^2) \sqrt{ \mu_2 u_2 - 1 } }
+    \dfrac{k}{\mu_1} = \dfrac{(1 + (\frac{p}{\mu_2} u')^2) \sqrt{a_2 - u'} }{ (a_1 + (\frac{p}{\mu_2} u')^2) \sqrt{ u' - 1 } }
+    \quad\mathrm{where}\quad u':= u_2\mu_2
     \label{eq:steady-states}
 \end{equation}
-which reveals that only the ratio between half-saturation and degradation $\frac{k}{\mu_1}$ affects the solutions to this equation, and hence the locations of the bifurcations.
+which reveals that only $a_1$,$a_2$ and the ratio between half-saturation and degradation $\frac{k}{\mu_1}$ affect the solutions to this equation, and hence the locations of the bifurcations.
 Figure \ref{fig:two-state-optima} reveals the results for 800 optimisation runs with 200 ADAM iterations.

### 2021-05-27 docs/checklist.tex
Updates from Overleaf

@@
   \item Did you describe the limitations of your work?
-    \answerYes{Implementation only covers co-dimension one bifurcations. Does not scale well for PDEs. See discussion in Section \ref{}}
+    \answerYes{Section \ref{section:conclusion} concludes that the implementation only covers co-dimension one bifurcations, with straightforward extensions to limit cycles, but does not scale well for partial differential equations}
   \item Did you discuss any potential negative societal impacts of your work?
-    \answerNA{}
+    \answerNA{We believe this work is sufficiently small scale and abstract to have any direct negative social impacts}
   \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
@@
   \item Did you state the full set of assumptions of all theoretical results?
-    \answerNA{This paper applies existing theorems. See Section \ref{section:method}}
+    \answerNA{In Section \ref{section:method} and the Appendix we apply existing theorems and approaches to differentiate our cost function}
 	\item Did you include complete proofs of all theoretical results?
-    \answerNA{We include derivations to guide the reader in Appendices}
+    \answerNA{While we do not have any theorems to prove, we include derivations in the Appendix to guide readers who are more theoretically inclined}
 \end{enumerate}
@@
   \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerYes{All figures are unit tests in the GitHub repository for \texttt{BifurcationFit.jl}}
+    \answerYes{Figures can be reproduced using unit tests in the GitHub repository for \texttt{BifurcationFit.jl}}
   \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerYes{ADAM was used with various hyper-parameters}
+    \answerYes{In Section \ref{section:results} we details of optimisations with \texttt{Flux.jl}}
 	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerYes{See Figure \ref{fig:scaling} and }
+    \answerYes{In Section \ref{section:results} we include error bars for time complexity estimates. We did not think that parameter uncertainties were relevant for the scope of this work}
 	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerYes{See Figure \ref{fig:scaling}}
+    \answerYes{The caption of Figure \ref{fig:scaling} has the specifications of the laptop used for complexity measurements}
 \end{enumerate}
@@
   \item If your work uses existing assets, did you cite the creators?
-    \answerYes{\cite{Veltz2019PseudoArcLengthContinuation.jl,Revels2016Forward-ModeJulia,Flux}}
+    \answerYes{\texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} for continuation, \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia} for automatic differentiation and \texttt{Flux.jl} \cite{Innes2018FashionableFlux,Innes2018Flux:Julia} for optimisation}
   \item Did you mention the license of the assets?
-    \answerNo{All assets have been release under MIT}
+    \answerNo{MIT Licences}
   \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerYes{All code available under \texttt{github.com/gszep/BifurcationFit.jl} }
+    \answerYes{Julia package implementing our method \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}}
   \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerNA{}
+    \answerNA{No data was collected}
   \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerNA{}
+    \answerNA{No data was collected}
 \end{enumerate}

### 2021-05-27 docs/article.tex

@@
 \section{Introduction}
+\label{section:introduction}
 
@@
 For example, several studies are likely to agree that the human immune system activates above a threshold concentration of a pathogen and deactivates at a lower threshold concentration, but may disagree on the exact quantities of the thresholds or the magnitudes of the immune response.
-Bifurcation theory provides us a framework for studying these qualitative changes in a manner that is independent of quantitative details.
+Bifurcation theory provides us a framework for studying these qualitative changes in a manner that is independent of quantitative details \cite{Kuznetsov2004TopologicalSystems}.
 The emerging picture suggests that identification of the qualitative behaviour -- the bifurcation diagram -- should precede any attempt at inferring other properties of a system \cite{Stumpf2019ParameterBifurcations}.
@@
 The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
-Drawing inspiration from implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients.
-The compute the diagram we use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl} developed for partial differential equations.
+We apply the strategy of implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients.
+To compute the diagram we use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl} developed for partial differential equations.
 In the case of partial differential equations, the computational complexity of calculating even a single bifurcation diagram can be unbounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
@@
     \item An end-to-end differentiable method for locating bifurcations in parameter space and then matching their dependency on a control condition to user-specified locations
-    \item Implementation of the method as a Julia package \cite{Szep2021BifurcationFit.jlDiagrams}
+    \item Implementation of the method as a Julia package\\  \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
     \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
@@
 
-For the purposes of introducing this work, we will consider the simplest class of bifurcations known as \textit{co-dimension one} bifurcations, that do not include limit cycles.
+For the purposes of introducing this work, we will consider the simplest class of bifurcations known as \textit{co-dimension one} bifurcations not including limit cycles.
 Therefore $\targets$ should contain conditions for which we hypothesise changes in multi-stable behaviour.
@@
 
-In the context of the differential equations, a co-dimension one bifurcation can be defined by a set of conditions on the determinant of the Jacobian $\Det$.
+In the context of the differential equations, and not considering limit cycles for now, a co-dimension one bifurcation can be defined by a set of conditions on the determinant of the Jacobian $\Det$.
 The determinant of the Jacobian quantifies the rate at which trajectories in a local patch of state-space $u\in\Reals^N$ converge or diverge.
@@
 \caption{Illustration of bifurcation diagrams for minimal models of bifurcations.
-\ref{fig:saddle-node} Saddle-node bifurcations arise for $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ when $\theta = (5/2,-1)$.
-\ref{fig:pitchfork} Pitchfork bifurcations arise for $\rates(u,p) = \theta_{1} + p u+\theta_{2}u^3$ when $\theta=(1/2,-1)$.
+\subref{fig:saddle-node}.
+Saddle-node bifurcations arise for $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ when $\theta = (\frac{5}{2},-1)$.
+\subref{fig:pitchfork}.
+Pitchfork bifurcations arise for $\rates(u,p) = \theta_{1} + p u+\theta_{2}u^3$ when $\theta=(\frac{1}{2},-1)$.
 Targets are illustrated by light yellow vertical lines.
@@
 \section{Proposed Method}
+\label{section:method}
 \subsection{Semi-supervised Cost Function}
@@
 \subsubsection{Supervised term: matched bifurcations to target locations}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=7cm]{bifurcation-measure}
-    \caption{Bifurcation measure $\measure(s)$ and determinant $\Det$ along the arclength $s$ of two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
-    \label{fig:measure}
-\end{figure}
 
@@
 \subsubsection{Unsupervised term: encouraging bifurcations}
+\begin{wrapfigure}{r}{7cm}
+    \centering
+    \includegraphics[width=\linewidth]{bifurcation-measure}
+    \caption{Bifurcation measure $\measure(s)$ and determinant $\Det$ along the arclength $s$ of two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
+    \label{fig:measure}
+\end{wrapfigure}
 
@@
 Calculation of this matrix involves inverting an $(N+1)\times(N+1)$ Jacobian $\frac{\partial G_{\theta}}{\partial(u,p)}$.
+Instead of explicitly inverting the Jacobian the corresponding system of linear equations is solved.
 The determinant of this Jacobian goes to zero in the degenerate case where $\frac{d}{ds}\Det=0$, further justifying our choice of measure $\Psi(\theta)$ which discourages the degenerate case.
@@
 The cost function is piece-wise smooth and differentiable with undefined gradients only in parameter contours where the number of predictions $|\predictions|$ changes; this is when $\Psi(\theta)$ is undefined and the inverse of $\frac{\partial G_{\theta}}{\partial (u,p)}$ does not exist.
-Given a set of solutions to $\rates(u,p)=0$ and locations $p(\theta)$ the gradient $\frac{\partial\loss}{\partial\theta}$ can be evaluated using automatic differentiation methods \cite{Revels2016Forward-ModeJulia,Flux} without needing to back-propagate through the solver that obtained the level set $\rates(u,p)=0$ in the forward pass.
-
-%\clearpage
+Given a set of solutions to $\rates(u,p)=0$ and locations $p(\theta)$ the gradient $\frac{\partial\loss}{\partial\theta}$ can be evaluated using automatic differentiation methods \cite{Revels2016Forward-ModeJulia,Innes2018FashionableFlux,Innes2018Flux:Julia} without needing to back-propagate through the solver that obtained the level set $\rates(u,p)=0$ in the forward pass.
 
 \section{Experiments \& Results}
+\label{section:results}
 In this section, we apply the method first to minimal examples that can produce saddle-node and pitchfork bifurcations, and then a more complex model that has multiple parametric regimes that produce saddle-node bifurcations.
@@
 \subsection{Minimal Models}
-Figures \ref{fig:minimal-models:results} show example optimisations of $(\theta_1,\theta_2)$ for the minimal saddle-node and pitchfork models respectively.
-Optimisation trajectories approach lines of global minima in the cost function, which represent a set of geometrically equivalent models, depicted in the right panels.
+Optimisations of two parameters $(\theta_1,\theta_2)$ using simple gradient descent from \texttt{Flux.jl} with learning rate $\eta=0.01$ for the minimal saddle-node and pitchfork models yield trajectories approaching lines of global minima in the cost function (Figures \ref{fig:minimal-models:results}) which represent a set of geometrically equivalent models.
 Two bifurcation diagrams are geometrically equivalent if the number, type and locations of bifurcations match the specified targets $\targets$.
 
-We can see that the geometrically equivalent lines are contained within larger basins where the correct number and type of bifurcations are present but do not match the locations of targets $\targets$.
-All models within this basin are in some sense topologically equivalent.
-This hierarchical classification allows us to identify the set of models that satisfy observed qualitative behaviour \cite{Stumpf2019ParameterBifurcations} before any attempt at inferring kinetic parameters, which is done by choosing a model along the line of geometrically equivalent models.
-
-Optimisation trajectories for the two minimal models appear mostly circumferential.
-This is because the models were set up such that the radial direction from the origin in $\theta$ space mostly scale kinetics whereas the circumferential direction changes the bifurcation topology.
-This suggests that the gradients of our cost function seek to change model geometry over kinetics.
-
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 
+We can see that the geometrically equivalent lines are contained within larger basins where the correct number and type of bifurcations are present but do not match the locations of targets $\targets$.
+All models within this basin are in some sense topologically equivalent.
+This hierarchical classification allows us to identify the set of models that satisfy observed qualitative behaviour \cite{Stumpf2019ParameterBifurcations} before any attempt at inferring kinetic parameters, which is done by choosing a model along the line of geometrically equivalent models.
+
+Optimisation trajectories for the two minimal models appear mostly circumferential.
+This is because the models were set up such that the radial direction from the origin in $\theta$ space mostly scale kinetics whereas the circumferential direction changes the bifurcation topology.
+This suggests that the gradients of our cost function seek to change model geometry over kinetics.
+
 \subsection{Genetic Toggle Switch}
@@
 Each species has a finite degradation rate $\mu_k$.
-Finally we have two half-saturation constants $p$ and $k$, one of which is chosen as our control condition.
+Finally we have two sensitivity constants $p$ and $k$, one of which is chosen as our control condition.
 A baseline production rate $a_k>1$ recovers an inhibitor type hill function for species $k$ and is an activator otherwise.
-The half-saturation are proportional to the slope of the hill productions.
+The sensitivities are proportional to the slope of the hill productions.
 Solving for the steady states,  substituting the equation for $u_1$ into $u_2$ and rearranging gives rise to the relationship
@@
 \end{equation}
-which reveals that only $a_1$,$a_2$ and the ratio between half-saturation and degradation $\frac{k}{\mu_1}$ affect the solutions to this equation, and hence the locations of the bifurcations.
-Figure \ref{fig:two-state-optima} reveals the results for 800 optimisation runs with 200 ADAM iterations.
-98\% of runs converged onto one of the two clusters: mutual activation and mutual inhibition regimes, examples of which are shown as insets.
-\begin{figure}
+which reveals that only $a_1$, $a_2$ and the ratio between the sensitivity and degradation parameters, $\frac{k}{\mu_1}$, affect the solutions to this equation, and hence the locations of the bifurcations (Figure \ref{fig:two-state-optima:parameters}).
+In 98\% of 800 runs, optimisation using the ADAM optimiser \cite{Kingma2014} from \texttt{Flux.jl} with learning rate $\eta=0.1$ converged to one of two clusters: mutual activation ($a_1 < 1, a_2 < 1$; cluster 1) and mutual inhibition ($a_1 > 1, a_2 > 1$; cluster 2) regimes.
+Example bifurcation diagrams illustrate how the bifurcation curves of each species are positively correlated in mutual activation and negatively correlated for mutual inhibition (Figure \ref{fig:two-state-optima:models}).
+
+In order to maintain biological interpretability, optimisation was restricted to the positive parameter regime by transforming the parameters to log-space $\theta\rightarrow10^\theta$.
+At the beginning of each optimisation run an initial $\theta$ was chosen in the log-space by sampling from a multivariate normal distribution with mean zero and standard deviation one.
+
+\begin{figure}[ht]
 \centering
-%\includegraphics[width=11cm]{two-state-optima}
-\includegraphics[width=11cm]{two-state-optima.png}
-\caption{Optimal parameter estimates $\theta^*$ for the two state model \eqref{eq:two-state} for the targets $\targets=\{4,5\}$ reveal two clusters of qualitatively different regimes: mutual activation and mutual inhibition.
-}
+\setlength\unitlength{1cm}
+{\phantomsubcaption\label{fig:two-state-optima:parameters}}
+{\phantomsubcaption\label{fig:two-state-optima:models}}
+% \includegraphics[width=9cm]{two-state-optima}
+\includegraphics[width=9cm]{two-state-optima.png}
+\begin{picture}(0,0) \put(-9,6.3){\subref{fig:two-state-optima:parameters}} \end{picture}
+\begin{picture}(0,0) \put(-3,6.3){\subref{fig:two-state-optima:models}}
+\end{picture}
+\caption{Bifurcation inference for the two-state model \eqref{eq:two-state}.
+\subref{fig:two-state-optima:parameters}.
+Optimal parameter estimates $\theta^*$ for the targets $\targets=\{4,5\}$ reveal two clusters of qualitatively different regimes: mutual activation ($a_1 < 1$; cluster 1) and mutual inhibition ($a_1 > 1$; cluster 2).
+\subref{fig:two-state-optima:models}.
+Example bifurcation diagrams indicate positively and negatively correlated dependencies between the two model states, as a function of the control condition.}
 \label{fig:two-state-optima}
@@
 
+\clearpage
 \subsection{Complexity}
+
+\begin{wrapfigure}{r}{7cm}
+\centering
+\includegraphics[width=\linewidth]{scaling}
+\caption{Complexity scaling of calculating the gradient of the cost function. Calculations were performed on an Intel Core i7-6700HQ CPU @ 2.60GHz x 8 without GPU acceleration}
+\label{fig:scaling}
+\end{wrapfigure}
+
 Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods, the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
@@
 
+For this calculation a model \eqref{eq:scaling} was chosen that so that it is extensible both in the number of parameters $M$ and the number of states $N$.
+In this model only the first state $u_1$ defines the shape of the bifurcation diagram, while the remaining states are merely linearly proportional to the first.
+The parameters $\mu_n$ contain sums of $\theta_m$ allowing us a flexible choice on the number of parameters while maintaining stable solutions for the bifurcation diagram.
+
 While still tractable on laptop computers for states $N<100$ our implementation currently does not scale well for partial differential equations where a large the number of states $N$ arises from discretisation of the spatial variables.
@@
 One approach would be to take the product of a finite subset of eigenvalues of the system.
+Note that any more efficient calculation must still permit back-propagation through it.
+\begin{equation}
+\begin{cases}
+    \,\,\partial_t u_1 = \sin^{2}\!p - (\theta_1\sin^{2}\!p+1)u_1\\
+    \,\,\partial_t u_n = u_{n-1} - (\mu_n^2+1) u_n & 2\le n\leq N
+\end{cases}
+\label{eq:scaling}
+\end{equation}
+\section{Conclusion \& Broader Impact}
+\label{section:conclusion}
 
-\begin{figure}
-\centering
-\includegraphics[width=11cm]{scaling}
-\caption{Complexity scaling of calculating the gradient of the cost function. Calculations were performed on an Intel Core i7-6700HQ CPU @ 2.60GHz x 8 without GPU acceleration}
-\label{fig:scaling}
-\end{figure}
+We proposed a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
+By applying implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} and the generalised Leibniz rule \cite{Flanders1973DifferentiationSign} to the geometry of the implicitly defined steady states \cite{Goldman2005CurvatureSurfaces} it is possible to use automatic differentiation methods to efficiently calculate gradients.
+We defined a bifurcation measure that that uses the determinant of the state-space Jacobian as an indicator for bifurcating parameter regimes in the unsupervised term of the cost function.
+The gradients of the cost can be efficiently computed using automatic differentiation methods.
+The computational bottleneck is the evaluation of the state-space Jacobian determinant which limits the implementation to ordinary differential equations.
 
-\section{Conclusion \& Broader Impact}
+We demonstrated our approach on models with one bifurcation parameter that can give rise to pitchforks and saddle-nodes.
+The estimated parameters form distinct clusters, allowing us to organise models in terms of topological and geometric equivalence.
+In the case of the genetic toggle switch we recovered mutual activation and inhibition regimes.
 
-In this work, we proposed a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
-The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
-Drawing inspiration from implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients.
-The compute the diagram we use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl} developed for partial differential equations.
-In the case of partial differential equations the computational complexity of calculating a single bifurcation diagram is not bounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
-The complexity of computing a single branch, however, is bounded by the complexities of the chosen eigenvalue solver and corrector, and further decreased by adaptive stepping procedures \cite{Aruliah2016AlgorithmContinuation}.
+Although we did not consider limit cycles, our approach can be modified using the conditions for Andronov-Hopf bifurcations instead of the conditions on the Jacobian determinant.
+Our approach generalises naturally to bifurcation manifolds such as limit point curves or surfaces.
+This is because the normal components of implicit derivatives can still be calculated for under-determined systems of equations \cite{Jos2011OnSurface,Tao2016Near-IsometricTracking,Fujisawa2013CalculationInvariance}.
+In the case of manifolds it would be more appropriate to use isosurface extraction algorithms rather than continuation to obtain the steady-state manifold.
+Our approach does not depend on the details of the steady-state solver and therefore can still be applied.
 
-We find that the cost function landscape contains basins that not only allow us to synthesise models with a desired bifurcation diagram but also allow us to organise models in terms of topological and geometric equivalence.
-We discuss the relevance of this in model selection.
-In summary, our paper has the following main contributions:
+% something here about how this helps high-level design of dynamical systems and is useful to the wider scientific machine learning community
 
-\begin{itemize}
-    \item Taking steps towards an efficient and single method for searching and matching bifurcations
-    \item A demonstrating of the powerful combination of differential geometry and implicit layers
-    \item Future directions include extending this routine for Hopf bifurcations and PDEs
-    \item Computational limitations
-\end{itemize}
+\section{Acknowledgements}
+We would like to acknowledge Kieran Cooney for the fruitful conversations that helped guide the derivations and computational approach.
 
-\clearpage
 \bibliography{refs}
@@
 
-\clearpage
+\clearpage\pagenumbering{gobble}
+\section*{Checklist}
+\input{checklist}
+
+\clearpage\pagenumbering{arabic}\setcounter{page}{1}
 \section*{Appendix}
@@
 
-\clearpage
-\section*{Checklist}
-\input{checklist}
 \end{document}

### 2021-05-27 docs/article.tex
update article

@@
 
-% something here about how this helps high-level design of dynamical systems and is useful to the wider scientific machine learning community
+In dynamical systems theory the geometry of state-space determines all of the qualitative behaviours of a system.
+Our work makes progress towards designing models directly in state-space, rather than the spatial or temporal domain.
+This is valuable to experimentalists who only have qualitative observations available to them and wish to navigate the space of qualitative behaviours of their system.
+Our work lies within a trend of progress in the scientific machine learning community, where structured domain-informed models are favoured over flexible models that live in large parameter spaces.
 

### 2021-05-27 src/Gradients.jl
applying implicit layers strategy

@@ -14,11 +14,12 @@ end
 
 ################################################################################
 function ∇errors( F::Function, predictions::AbstractVector{<:BorderedArray}, θ::AbstractVector, targets::StateSpace)
-	return mean( p′-> mean( z->(z.p-p′)^2, predictions; type=:geometric )*mean( z-> 2∂p(F,z,θ)/(z.p-p′), predictions; type=:arithmetic ), targets.targets; type=:arithmetic )
+	return mean( p′-> mean( z->(z.p-p′)^2, predictions; type=:geometric )*mean( z-> 2velocity(F,z,θ)/(z.p-p′), predictions; type=:arithmetic ), targets.targets; type=:arithmetic )
 end
 
-function ∇measure( F::Function, z::BorderedArray, θ::AbstractVector )
-	return gradient(θ->measure(F,z,θ),θ) + deformation(F,z,θ)'gradient(z->measure(F,z,θ),z) + measure(F,z,θ)*∇region(F,z,θ)
+function ∇measure( F::Function, z::BorderedArray, θ::AbstractVector; newtonOptions=NewtonPar(verbose=false,maxIter=800,tol=1e-6) )
+	∂implicit, _, _ = newtonOptions.linsolver( -∂Fz(F,z,θ)', gradient(z->measure(F,z,θ),z) )
+	return gradient( θ -> measure(F,z,θ) + F(z,θ)'∂implicit , θ ) + measure(F,z,θ)*∇region(F,z,θ)
 end
 
 ###########################################################################
@@ -31,15 +32,16 @@ function ∇measure( F::Function, branches::AbstractVector{<:Branch}, θ::Abstra
 end
 
 ############################################## gradient term due to changing integration region dz
-deformation( F::Function, z::BorderedArray, θ::AbstractVector ) = -∂Fz(F,z,θ)\∂Fθ(F,z,θ)
+deformation( F::Function, z::BorderedArray, θ::AbstractVector ) = -∂Fz(F,z,θ)\∂Fθ(F,z,θ) # todo@(gszep) matrix inverse is computational bottleneck
 function ∇region( F::Function, z::BorderedArray, θ::AbstractVector )
 
 	∇deformation = reshape( jacobian(z->deformation(F,z,θ),z), length(z),length(θ),length(z) )
 	tangent = tangent_field(F,z,θ)
[…]

### 2021-05-27 src/Objectives.jl

@@ -18,7 +18,7 @@ function errors( predictions::AbstractVector{<:BorderedArray}, targets::StateSpa
 end
 
 function measure( F::Function, z::BorderedArray, θ::AbstractVector )
-	return 1 / ( 1 + abs( det(F,z,θ) / ∂det(F,z,θ)'tangent_field(F,z,θ) ) )
+	return 1 / ( 1 + abs( derivative( z->log(abs(det(∂Fu(F,z,θ)))), z, tangent_field(F,z,θ) ) )^(-1) ) # todo@(gszep) determinant calculation is computational bottleneck
 end
 
 ################################################################################
@@ -30,27 +30,17 @@ function measure( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractV
 	return sum( branch -> measure(F,branch,θ), branches )
 end
 
-########################################################################### determinant
-import LinearAlgebra: det
-det(F::Function,z::BorderedArray,θ::AbstractVector) = det(∂Fu(F,z,θ))
-∂det(F::Function,z::BorderedArray,θ::AbstractVector) = gradient(z->det(F,z,θ),z)
-
 ########################################################## bifurcation distance and velocity
-distance(F::Function,z::BorderedArray,θ::AbstractVector) = [ F(z,θ); det(F,z,θ) ]
-velocity( F::Function, z::BorderedArray, θ::AbstractVector ) = - jacobian(z->distance(F,z,θ),z) \ jacobian(θ->distance(F,z,θ),θ)
-∂p( F::Function, z::BorderedArray, θ::AbstractVector ) = velocity(F,z,θ)[end,:]
+distance(F::Function,z::BorderedArray,θ::AbstractVector) = [ F(z,θ); det(∂Fu(F,z,θ)) ]
+function velocity( F::Function, z::BorderedArray, θ::AbstractVector; newtonOptions=NewtonPar(verbose=false,maxIter=800,tol=1e-6) )
+	∂implicit, _, _ = newtonOptions.linsolver( -jacobian( z->distance(F,z,θ), z )', [zero(z.u);one(z.p)] )
+	return gradient( θ->distance(F,z,θ)'∂implicit, θ )
+end
 
 ################################################################################
 function tangent_field( F::Function, z::BorderedArray, θ::AbstractVector )
-
-	∂F = ∂Fz(F,z,θ)
-	field = similar(∂F[1,:])
-
-	for i ∈ 1:length(z) # construct tangent field T(z) := det[ ẑ , ∂Fz ]
[…]

### 2021-05-27 src/Structures.jl

@@ -51,7 +51,8 @@ function push!(branch::Branch,state::ContState)
     push!(branch, ( z=z, λ=state.eigvals, ds=norm(∂z)*abs(state.ds), bif=detectBifucation(state) ))
 end
 
-import Base: zero,isapprox,unique ################################## methods for BorderedArray
+import Base: zero,isapprox,unique,+ ################################## methods for BorderedArray
++(x::BorderedArray, y::BorderedArray) = BorderedArray(x.u+y.u,x.p+y.p)
 zero(::Type{BorderedArray{T,U}}; ϵ::Bool=true) where {T<:Union{Number,StaticArray},U<:Union{Number,StaticArray}} = BorderedArray(zero(T).+ϵ*eps(),zero(U).+ϵ*eps())
 isapprox( x::BorderedArray, y::BorderedArray; kwargs... ) = isapprox( [x.u;x.p], [y.u;y.p] ; kwargs...)
 function unique(X::AbstractVector{T}; kwargs...) where T<:BorderedArray
@@ -62,6 +63,16 @@ function unique(X::AbstractVector{T}; kwargs...) where T<:BorderedArray
         end
     end
     return z
+end
+
+function kernel(A::AbstractMatrix; nullity::Int=0)
+
+    if iszero(nullity) # default to SVD
+        return nullspace(A)
+
+    else 
+        return qr(A').Q[:,end+1-nullity:end]
+    end
 end
 
 #################################################### display methods

### 2021-05-27 test/applied/two-state.jl

@@ -16,4 +16,4 @@ end
 
 ######################################################### targets and initial guess
 X = StateSpace( 2, 0.01:0.01:10, [4,5] )
-θ = SizedVector{5}(0.5,0.5,0.5470,2.0,7.5)
+θ = SizedVector{5}(0.0,0.0,-1.0,-1.0,2.0)

### 2021-05-27 test/figures.jl

@@ -63,16 +63,26 @@ savefig("docs/figures/bifurcation-measure.pdf")
 ####################################################################################################
 ############################################################################################ scaling
 include("scaling/scaling.jl")
+using StatsBase: mean,std
+using LaTeXStrings
 
-Ns,Ms = 1:10,1:5:50
-complexity = scaling.(Ns,Ms')
+complexity_mean = Float64[]
+complexity_std = Float64[]
 
-contourf( Ns,Ms, complexity,
-	xlabel="States", ylabel="Parameters", size=(500,400),
-	colorbar_title="Iteration Execution / sec"
-)
+for n ∈	1:40
+	ts = [scaling(n,3) for _ ∈ 1:10]
+	push!(complexity_mean,mean(ts))
+	push!(complexity_std,std(ts))
+end
+scaling(40,3)
 
-cticks!
+default();default(grid=false)
+plot( 10:30, x->x^3/1800, label=L"N^3", color=:red, linewidth=2, legend=:topleft)
+scatter!( complexity_mean,yerror=complexity_std, label="", yscale=:log10, xscale=:log10, ms=3, color=:black,
+	xlabel=L"\mathrm{States}\,\,N", ylabel=L"\mathrm{Execution\,\,Time}\,\,/\,\,\mathrm{sec}", size=(400,400) )
+
+xticks!([1,3,5,10,20,30],["1","3","5","10","20","30"])
+yticks!([1e-1,1e0,1e1],["0.1","1","10"])
 savefig("docs/figures/scaling.pdf")
 
 ####################################################################################################
@@ -101,33 +111,30 @@ optima = hcat(trajectories...)
 
 @. optima[end,:] = abs(optima[end,:])
 M,N = size(optima)
-clusters = dbscan(optima,0.3,min_cluster_size=500)
+clusters = dbscan(optima,0.5,min_cluster_size=40)
 
 layout = @layout [ a{0.7w} [b{0.5h}; b{0.5h} ] ]
-default(); default(msc=:auto,label="",xlim=(-3,3),ylim=(-1,2),size=(600,450),grid=false)
+default(); default(msc=:auto,label="",size=(600,450),grid=false)
 
[…]

### 2021-05-27 src/Objectives.jl
restoring window function

@@ -5,7 +5,7 @@ function loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVect
 	λ = length(targets.targets)-length(predictions)
 
 	if λ≠0 
-		Φ = measure(F,branches,θ)
+		Φ = measure(F,branches,θ,targets)
 		return errors(predictions,targets) - λ*log(Φ)
 	else
 		return errors(predictions,targets)
@@ -17,17 +17,17 @@ function errors( predictions::AbstractVector{<:BorderedArray}, targets::StateSpa
 	return mean( p′-> mean( z->(z.p-p′)^2, predictions; type=:geometric ), targets.targets; type=:arithmetic )
 end
 
-function measure( F::Function, z::BorderedArray, θ::AbstractVector )
-	return 1 / ( 1 + abs( derivative( z->log(abs(det(∂Fu(F,z,θ)))), z, tangent_field(F,z,θ) ) )^(-1) ) # todo@(gszep) determinant calculation is computational bottleneck
+function measure( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace )
+	return window_function(targets.parameter,z) / ( 1 + abs( derivative( z->log(abs(det(∂Fu(F,z,θ)))), z, tangent_field(F,z,θ) ) )^(-1) ) # todo@(gszep) determinant calculation is computational bottleneck
 end
 
 ################################################################################
-function measure( F::Function, branch::Branch, θ::AbstractVector )
-	return sum( s -> measure(F,s.z,θ)*s.ds, branch )
+function measure( F::Function, branch::Branch, θ::AbstractVector, targets::StateSpace )
+	return sum( s -> measure(F,s.z,θ,targets)*s.ds, branch )
 end
 
-function measure( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector )
-	return sum( branch -> measure(F,branch,θ), branches )
+function measure( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace )
+	return sum( branch -> measure(F,branch,θ,targets), branches )
 end
 
 ########################################################## bifurcation distance and velocity
@@ -44,21 +44,8 @@ function tangent_field( F::Function, z::BorderedArray, θ::AbstractVector )
 end
 
[…]

### 2021-05-27 src/Plots.jl

@@ -26,11 +26,11 @@ function plot(steady_states::Vector{<:Branch},data::StateSpace; determinant=true
 
 			for idx ∈ 1:dim(branch)
 
-				plot!( parameter, map(s->s.z.u[idx],branch), subplot=1,
+				plot!( parameter, map(s->s.z.u[idx],branch), subplot=1, alpha=map(s->window_function(data.parameter,s.z),branch),
 					color=map( stable -> stable ? :darkblue : :lightblue, stability ) )
 			end
 
-			plot!( parameter, determinants, subplot=2,
+			plot!( parameter, determinants, subplot=2, alpha=map(s->window_function(data.parameter,s.z),branch),
 				color=map( stable -> stable ? :red : :pink, stability )
 			)
 		end

### 2021-05-27 src/Utils.jl

@@ -1,13 +1,12 @@
 ############################################################################ hyperparameter updates
-function getParameters(data::StateSpace{N,T}; maxIter::Int=800, tol=1e-6, padding=0.2, kwargs...) where {N,T<:Number}
+function getParameters(data::StateSpace{N,T}; maxIter::Int=800, tol=1e-6, kwargs...) where {N,T<:Number}
 	newtonOptions = NewtonPar(verbose=false,maxIter=maxIter,tol=T(tol))
 
 	# support for StaticArrays github.com/JuliaArrays/StaticArrays.jl/issues/73
 	newtonOptions = @set newtonOptions.linsolver.useFactorization = false
-	padding = padding*(maximum(data.parameter)-minimum(data.parameter))
 	return ContinuationPar(
 
-        pMin=minimum(data.parameter)-padding, pMax=maximum(data.parameter)+padding, maxSteps=10*length(data.parameter),
+        pMin=minimum(data.parameter), pMax=maximum(data.parameter), maxSteps=10*length(data.parameter),
         ds=step(data.parameter), dsmax=step(data.parameter), dsmin=step(data.parameter),
 
 		newtonOptions=newtonOptions, detectFold=false, detectBifurcation=true,

### 2021-05-28 docs/appendix.tex
Updates from Overleaf

@@
 
-Suppose there exists a one dimensional space curve $\mathcal{C(\theta)}$ embedded in $z\in\Reals^{N+1}$ whos geometry changes depending on input parameters $\theta\in\Reals^M$.
+Suppose there exists a one dimensional space curve $\mathcal{C(\theta)}$ embedded in $z\in\Reals^{N+1}$ whose geometry changes depending on input parameters $\theta\in\Reals^M$.
 This curve could be open or closed and changes in $\theta$ could change the curve topology as well.

### 2021-05-28 docs/article.tex
Updates from Overleaf

@@
 This helps systematically guide both model and experimental design.
-The collected observations, however, are often indirectly proportional to the variables defined in the model; in microscopy, for example, data are reported in arbitrary fluorescence units allowing the observer to shift and scale data arbitrarily.
-The model, on the other hand, describes the dynamics of concentrations of relevant proteins.
+The collected observations, however, are often indirectly proportional to the variables defined in the model.
+In fluorescence microscopy, a commonly used technique in biological studies, data are reported in arbitrary fluorescence units, allowing the observer to shift and scale data arbitrarily.
+Models of cellular networks, on the other hand, often seek to describe the dynamics of concentrations of relevant proteins \cite{Tyson2001NetworkPhysiology,Grant2020InterpretationCircuit}.
 Furthermore, the observed qualitative changes in response to changes in experimental conditions are more robust and reproducible across studies than the quantitative details.
 For example, several studies are likely to agree that the human immune system activates above a threshold concentration of a pathogen and deactivates at a lower threshold concentration, but may disagree on the exact quantities of the thresholds or the magnitudes of the immune response.
-Bifurcation theory provides us a framework for studying these qualitative changes in a manner that is independent of quantitative details \cite{Kuznetsov2004TopologicalSystems}.
+Bifurcation theory provides us a framework for studying these qualitative changes in a manner that is independent of quantitative details \cite{Kuznetsov2004TopologicalSystems,Tyson2001NetworkPhysiology}.
 The emerging picture suggests that identification of the qualitative behaviour -- the bifurcation diagram -- should precede any attempt at inferring other properties of a system \cite{Stumpf2019ParameterBifurcations}.
 
-Inferring the parameters of a model directly from a bifurcation diagram is difficult because it is not obvious how one should change the parameters to create a bifurcation.
+Inferring the parameters of a model directly from a bifurcation diagram is difficult because it is not obvious how multiple  parameters in concert control the existence and position of a bifurcation.
 It could even be impossible for the model to bifurcate in the manner desired.
@@
 Progress has been made is cases where model structure and stability conditions are used to refine the search space \cite{Otero-Muras2018Optimization-basedModels,Otero-Muras2014ACurves} yet the resulting objectives are still not explicit in the bifurcation targets and also not differentiable.
+In the emerging field of Scientific machine learning \cite{Rackauckas2017Differentialequations.jlJulia,Rackauckas2018ASolutions,Rackauckas2020UniversalLearning}, parameters of structured mechanistic models are favoured over flexible models in larger parameter spaces.
 A scalable method for navigating the space of bifurcation diagrams would enable design of differential equations with high-level qualitative constraints.
@@
 
-Back-propagation through differential equation solvers has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jl-AEquations} that enabled scalable parameter inference for differential equations from trajectory data.
+Back-propagation through differential equation solvers has been a breakthrough over the past couple of years \cite{Chen2018NeuralEquations,Rackauckas2019DiffEqFlux.jlEquations} that enabled scalable parameter inference for differential equations from trajectory data.
 Although one could use trajectory data to create the aforementioned qualitative constraints \cite{Ranciati2017BayesianParameters,Khadivar2021LearningBifurcations} this would entail over-constraining information originating from the kinetics and dynamical transients of the model.
@@
 
-In this work, we propose a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
+In this work, we propose a gradient-based semi-supervised approach for inferring the parameters of differential equation models that give rise to bifurcations at specified values of a control parameter.
 The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
 We apply the strategy of implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} to calculate gradients.
-To compute the diagram we use a predictor-corrector method called deflated pseudo-arclength continuation \cite{Farrell2016TheDiagrams,Veltz2019PseudoArcLengthContinuation.jl} developed for partial differential equations.
+To compute the diagram we use a predictor-corrector method called deflated continuation \cite{Farrell2016TheDiagrams,Veltz2020BifurcationKit.jl} developed for partial differential equations.
 In the case of partial differential equations, the computational complexity of calculating even a single bifurcation diagram can be unbounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
@@
 
-For a given set of parameters $\theta$ one could compute the set of predicted bifurcations $\predictions(\theta)$ using parameter continuation methods \cite{Veltz2019PseudoArcLengthContinuation.jl,Farrell2016TheDiagrams}.
+For a given set of parameters $\theta$ one could compute the set of predicted bifurcations $\predictions(\theta)$ using parameter continuation methods \cite{Veltz2020BifurcationKit.jl,Farrell2016TheDiagrams}.
 Our goal is to find optimal parameters $\theta^*$ that match predictions $\predictions(\theta^*)$ to specified targets $\targets$.
@@
 However, this can result in multiple predictions matching the same target.
-Therefore, we choose a geometric mean over the predictions and an arithmetic mean over targets to account for cases where $|\predictions|\neq|\targets|$ and discourage multiple predictions matching the same target (in cases where $|\predictions|\geq|\targets|$).
-Such an error term is only zero when each target is matched by at least one prediction
-\begin{align}
-    \frac{1}{|\targets|}\sum_{p'\in\targets}
+Therefore, we choose a geometric mean over the predictions and an arithmetic mean over targets to account for cases where $|\predictions|\neq|\targets|$ and discourage multiple predictions matching the same target (in cases where $|\predictions|\geq|\targets|$):
+\begin{equation}
+    \error(\theta, \targets) = \frac{1}{|\targets|}\sum_{p'\in\targets}
     \prod_{p(\theta)\in\predictions(\theta)}|p(\theta)-p'|
     ^{\frac{1}{|\predictions|}}
-    \label{eq:error}
-\end{align}
+    %\label{eq:error}
+\end{equation}
+The supervised term is only zero when each target is matched by at least one prediction.
 
 \subsubsection{Unsupervised term: encouraging bifurcations}
-\begin{wrapfigure}{r}{7cm}
+
+\begin{wrapfigure}{r}{0.5\textwidth}
     \centering
     \includegraphics[width=\linewidth]{bifurcation-measure}
-    \caption{Bifurcation measure $\measure(s)$ and determinant $\Det$ along the arclength $s$ of two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars}
+    \caption{Bifurcation measure $\measure(s)$ and determinant $\Det$ along the arclength $s$ of two different bifurcation curves demonstrating how maximising the measure along the curve maintains the existing bifurcation marked by a circle, while encouraging new bifurcations marked by stars.}
     \label{fig:measure}
@@
 \end{equation}
-This measure is maximal at bifurcations and has finite gradients in non-bifurcating regimes (Figure \ref{fig:measure}).
+
+The bifurcation measure $\measure(s)$ is maximal at bifurcations and has finite gradients in non-bifurcating regimes (Figure \ref{fig:measure}).
 More specifically, the measure $\measure(s)$ is one at bifurcation points and goes to zero an odd number of times between bifurcations.
-This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(s)$ going to zero for each turning point.
+This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(s)$ going to zero for each turning point (see Figure \ref{fig:measure}).
+As the determinant $\Det$ diverges, we approach regimes far away from any closest bifurcation and hence $\measure(s)\rightarrow0$.
+But we would still like to have non-zero gradients with respect to $\theta$ in this regime, and so the measure is designed to go to zero slowly.
+Finally, we note that $\measure(s)$ has the desirable property of being invariant to certain classes of transformations on $\rates(u,p)$ that do not change bifurcation locations (details provided in Appendix \ref{appendix:measure}).
 
-To quantify the overall propensity for bifurcations over a diagram, we define the total bifurcation measure $\Psi(\theta)$ as
+While the calculation of the determinant is straightforward, its directional derivative is not.
+Fortunately, a tangent field representation of the bifurcation curve allows us to compute this derivative implicitly; see Appendix \ref{appendix:tangent-fields} for details.
+
+Equipped with a measure that quantifies the appearance of bifurcations along a bifurcation arc, we now define the total measure for a bifurcation diagram as
 \begin{equation}
@@
         \mathrm{d}s
-    }
+    }.
 \end{equation}
 Here we denote $\int_{\rates(u,p)=0}\mathrm{d}s$ as the sum of the line integrals in $(u,p)\in\Reals^{N+1}$ defined by the level set $\rates(u,p)=0$ with $s$ being an arbitrary parametrisation of the curves.
+The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the controlled condition region $p$ is densely packed with bifurcations.
 
-While the calculation of the determinant is straightforward, its directional derivative is not.
-Fortunately, a tangent field representation of the bifurcation curve allows us to compute this derivative implicitly; see Appendix \ref{appendix:tangent-fields} for details.
-
-As the determinant $\Det$ diverges we approach regimes far away from any closest bifurcation and hence $\measure(s)\rightarrow0$.
-We would still like to have non-zero gradients with respect to $\theta$ in this regime, and therefore $\measure(s)$ was designed to go to zero slowly.
-The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the controlled condition region $p$ is densely packed with bifurcations.
-The measure $\measure(s)$ is also invariant under certain classes of transformations on $\rates(u,p)$ which do not change bifurcation locations, proofs of which are provided in Appendix \ref{appendix:measure}.
-The total measure $\Psi(\theta)$ is added to the supervised term as if it were a likelihood; this defines the semi-supervised cost function as
+The total measure $\Psi(\theta)$ is added to the supervised term as if it were a likelihood. This defines the semi-supervised cost function as
 \begin{align}
@@
     \big(|\predictions|-|\targets|\big)\,\,\lambda \log\Psi(\theta)+
-    \frac{1}{|\targets|}\sum_{p'}
-    \prod_{p(\theta)}|p(\theta)-p'|
-    ^{\frac{1}{|\predictions|}}
+    \error(\theta, \targets),
     \label{eq:loss}
 \end{align}
+where we introduce the hyperparameter $\lambda$ to enable control over the relative contribution of the unsupervised term.
+The pre-factor $|\targets|-|\predictions|$ in the unsupervised term ensures that the gradients are always pushing optimisers towards a state where $|\targets|=|\predictions|$.
+This can be seen as a step-wise annealing of the unsupervised term until the desired state is reached.
 
@@
 
-The pre-factor $|\targets|-|\predictions|$ in the unsupervised term ensures that the gradients are always pushing optimisers towards a state where $|\targets|=|\predictions|$.
-This can be seen as a step-wise annealing of the unsupervised term until the desired state is reached.
-Note that while individual bifurcations $p(\theta)$ depend smoothly on $\theta$, the total number of predictions $|\predictions|$ does not have gradient contributions with respect to $\theta$.
-We can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ and total measure $\Psi(\theta)$
+To make use of gradient-based optimisers to locate desired bifurcation diagrams, we here show how to differentiate the cost function.
+First, we note that while individual bifurcations $p(\theta)$ depend smoothly on $\theta$, the total number of predictions $|\predictions|$ does not have gradient contributions with respect to $\theta$.
+Therefore, we can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ (within the definition of $\error$) and the total measure $\Psi(\theta)$.
+Therefore,
 \begin{align}
@@
 \end{align}
+
 In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to calculate the gradient $\frac{\partial\loss}{\partial\theta}$ without having to differentiate through the operations of the solver that finds the bifurcation diagram $\rates(u,p)=0$ and the bifurcation locations $p(\theta)$.
@@
 
-In order to maintain biological interpretability, optimisation was restricted to the positive parameter regime by transforming the parameters to log-space $\theta\rightarrow10^\theta$.
-At the beginning of each optimisation run an initial $\theta$ was chosen in the log-space by sampling from a multivariate normal distribution with mean zero and standard deviation one.
-
 \begin{figure}[ht]
@@
 
-\clearpage
+In order to maintain biological interpretability, optimisation was restricted to the positive parameter regime by transforming the parameters to log-space $\theta\rightarrow10^\theta$.
+At the beginning of each optimisation run an initial $\theta$ was chosen in the log-space by sampling from a multivariate normal distribution with mean zero and standard deviation one.
+
+%\clearpage
 \subsection{Complexity}
-
-\begin{wrapfigure}{r}{7cm}
-\centering
-\includegraphics[width=\linewidth]{scaling}
-\caption{Complexity scaling of calculating the gradient of the cost function. Calculations were performed on an Intel Core i7-6700HQ CPU @ 2.60GHz x 8 without GPU acceleration}
-\label{fig:scaling}
-\end{wrapfigure}
 
@@
 This calculation scales like $N^3$ where $N$ is the number of state space variables (Figure \ref{fig:scaling}).
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+\centering
+\includegraphics[width=0.48\textwidth]{scaling}
+\caption{Complexity scaling of calculating the gradient of the cost function. Calculations were performed on an Intel Core i7-6700HQ CPU @ 2.60GHz x 8 without GPU acceleration}
+\label{fig:scaling}
+\end{wrapfigure}
 
@@
 
-\section{Acknowledgements}
-We would like to acknowledge Kieran Cooney for the fruitful conversations that helped guide the derivations and computational approach.
+% \section{Acknowledgements}
+% We would like to acknowledge Kieran Cooney for the fruitful conversations that helped guide the derivations and computational approach.
 

### 2021-05-28 docs/checklist.tex

@@
   \item If your work uses existing assets, did you cite the creators?
-    \answerYes{\texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} for continuation, \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia} for automatic differentiation and \texttt{Flux.jl} \cite{Innes2018FashionableFlux,Innes2018Flux:Julia} for optimisation}
+    \answerYes{\texttt{BifurcationKit.jl} \cite{Veltz2020BifurcationKit.jl} for continuation, \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia} for automatic differentiation and \texttt{Flux.jl} \cite{Innes2018FashionableFlux,Innes2018Flux:Julia} for optimisation}
   \item Did you mention the license of the assets?

### 2021-05-28 docs/article.tex
submitting to neurips

@@
 \begin{abstract}
-    Estimation of parameters in differential equation models can be achieved by applying learning algorithms to time-series data.
+    Estimation of parameters in differential equation models can be achieved by applying learning algorithms to quantitative time-series data.
 However, sometimes it is only possible to measure qualitative changes of a system in response to a controlled condition.
@@
 This helps systematically guide both model and experimental design.
-The collected observations, however, are often indirectly proportional to the variables defined in the model.
-In fluorescence microscopy, a commonly used technique in biological studies, data are reported in arbitrary fluorescence units, allowing the observer to shift and scale data arbitrarily.
-Models of cellular networks, on the other hand, often seek to describe the dynamics of concentrations of relevant proteins \cite{Tyson2001NetworkPhysiology,Grant2020InterpretationCircuit}.
-Furthermore, the observed qualitative changes in response to changes in experimental conditions are more robust and reproducible across studies than the quantitative details.
+While we would like to understand the quantitative details of a system, often only qualitative changes in response to varying experimental conditions can be robustly measured across independent studies \cite{Tyson2001NetworkPhysiology,Grant2020InterpretationCircuit}.
 For example, several studies are likely to agree that the human immune system activates above a threshold concentration of a pathogen and deactivates at a lower threshold concentration, but may disagree on the exact quantities of the thresholds or the magnitudes of the immune response.
-Bifurcation theory provides us a framework for studying these qualitative changes in a manner that is independent of quantitative details \cite{Kuznetsov2004TopologicalSystems,Tyson2001NetworkPhysiology}.
+Bifurcation theory provides us a framework for studying these transitions in a manner that is independent of quantitative details \cite{Kuznetsov2004TopologicalSystems}.
 The emerging picture suggests that identification of the qualitative behaviour -- the bifurcation diagram -- should precede any attempt at inferring other properties of a system \cite{Stumpf2019ParameterBifurcations}.
@@
 Several approaches exist to place bifurcations to desired locations once a manifold is present \cite{Iwasaki1997AnType,Lu2006InverseSystems,Dobson2004DistanceBifurcations} yet typically resort to sampling techniques to search for them in the first place \cite{Chickarmane2005BifurcationTool,Conrad2006BifurcationClock}.
-Progress has been made is cases where model structure and stability conditions are used to refine the search space \cite{Otero-Muras2018Optimization-basedModels,Otero-Muras2014ACurves} yet the resulting objectives are still not explicit in the bifurcation targets and also not differentiable.
-In the emerging field of Scientific machine learning \cite{Rackauckas2017Differentialequations.jlJulia,Rackauckas2018ASolutions,Rackauckas2020UniversalLearning}, parameters of structured mechanistic models are favoured over flexible models in larger parameter spaces.
+Progress has been made in cases where model structure and stability conditions are used to refine the search space \cite{Otero-Muras2018Optimization-basedModels,Otero-Muras2014ACurves} yet the resulting objectives are still not explicit in the bifurcation targets and also not differentiable.
+In the emerging field of scientific machine learning \cite{Rackauckas2017Differentialequations.jlJulia,Rackauckas2018ASolutions,Rackauckas2020UniversalLearning}, parameters of structured mechanistic models are favoured over flexible models in larger parameter spaces.
 A scalable method for navigating the space of bifurcation diagrams would enable design of differential equations with high-level qualitative constraints.
@@
 
-In this work, we propose a gradient-based semi-supervised approach for inferring the parameters of differential equation models that give rise to bifurcations at specified values of a control parameter.
+The problem of inferring differential equation parameters against a user-specified bifurcation diagram decomposes into two parts: searching for bifurcating regimes and matching the locations of bifurcation points to desired values.
+Matching bifurcation locations is a supervised problem where the data are expressed as bifurcations points \cite{Lu2006InverseSystems,Conrad2006BifurcationClock}.
+Searching for bifurcations is an unsupervised problem because when bifurcations are not present, there is no distance defined between data and prediction \cite{Chickarmane2005BifurcationTool}.
+Therefore only properties of the model can be used to start the search.
+We propose a semi-supervised approach for performing both tasks in an end-to-end fashion.
 The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
@@
 To compute the diagram we use a predictor-corrector method called deflated continuation \cite{Farrell2016TheDiagrams,Veltz2020BifurcationKit.jl} developed for partial differential equations.
-In the case of partial differential equations, the computational complexity of calculating even a single bifurcation diagram can be unbounded since superpositions of localised solutions may give rise to uncountably many branches \cite{Avitabile2010ToEquation}.
-The complexity of computing a single branch, however, is bounded by the complexities of the chosen eigenvalue solver and corrector, and further decreased by adaptive stepping procedures \cite{Aruliah2016AlgorithmContinuation}.
 
@@
     \item An end-to-end differentiable method for locating bifurcations in parameter space and then matching their dependency on a control condition to user-specified locations
-    \item Implementation of the method as a Julia package\\  \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
+    \item Implementation of the method as a Julia package\\  \href{https://github.com/anon/Code.jl}{\texttt{see attached zip file}}
+% \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
     \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
@@
 \end{align}
-
 In the context of the differential equations, and not considering limit cycles for now, a co-dimension one bifurcation can be defined by a set of conditions on the determinant of the Jacobian $\Det$.
@@
 The determinant approaching zero means that the dynamics of the system are slowing down, which is an important indicator for the onset of a transition between qualitative behaviours.
-Furthermore, the slowing down must necessarily be followed by breakdown of stability; for this to be true it is sufficient to require that the determinant cross zero with a finite slope, meaning that its directional derivative along the diagram $\frac{d}{ds}\Det$ is not zero.
+Furthermore, the slowing down must necessarily be followed by breakdown of stability; for this to be true it is sufficient \textit{but not necessary} to require that the determinant cross zero with a finite slope, meaning that its directional derivative along the diagram $\frac{d}{ds}\Det$ is not zero.
 The set of predicted values for the control condition $\predictions(\theta)\subset\Reals$ at which bifurcations occur are defined as
@@
 \end{align}
-
-To illustrate how the shapes of bifurcation diagrams compare with the determinant of the Jacobian, we show minimal models for saddle-node (Figure \ref{fig:saddle-node}) and pitchfork bifurcations (Figure \ref{fig:pitchfork}).
-Indeed the predictions $\predictions(\theta)$ are defined by zero crossings in the determinant with a finite slope.
+The most common bifurcations between steady states, not including limit cycles, are saddle-nodes and pitchforks \cite{Haragus2011LocalSystems}.
+The relationships between our conditions for bifurcation \eqref{eq:predictions} and the conditions for saddle-node are detailed in Appendix \ref{appendix:conditions}.
+Saddle-node bifurcations, which often appear in pairs (Figure \ref{fig:saddle-node}) are defined by stable and unstable fixed points meeting and disappearing.
+Pitchfork bifurcations are special cases of the saddle-node where a single steady state splits into two stable and one unstable steady state (Figure \ref{fig:pitchfork}).
+To illustrate these bifurcations, we define minimal models (Figure \ref{fig:minimal-models}) that span the space of saddle-node and pitchforks, where indeed zero crossings in the determinant with a finite slope define the set of prediction $\predictions(\theta)$.
 The location of these crossings in general may not match the targets $\targets$.
@@
 
-\clearpage
 \section{Proposed Method}
@@
 
-To identify parameter sets that give rise to bifurcation diagrams with specified turning points, we propose a semi-supervised cost function that comprises two terms.
+To identify parameter sets that give rise to bifurcation diagrams with specified bifurcation points, we propose a semi-supervised cost function that comprises two terms.
 The role of the supervised term is simply to reward predicted bifurcations to coincide with the specified target locations.
 This of course relies on such bifurcations existing.
-Therefore, the role of the unsupervised term is to encourage an optimiser to move towards parameter regimes that do exhibit bifurcations.
+The role of the unsupervised term is to encourage an optimiser to move towards parameter regimes that do exhibit bifurcations.
 
-\subsubsection{Supervised term: matched bifurcations to target locations}
+\subsubsection{Supervised term: matching bifurcations to target locations}
 
 In order for predicted bifurcations $p(\theta)\in\predictions(\theta)$ to match targets $p'\in\targets$ we need to evaluate an error term $|p(\theta)-p'|$.
-A naive approach might take an average over the norms for each prediction-target pair.
-However, this can result in multiple predictions matching the same target.
-Therefore, we choose a geometric mean over the predictions and an arithmetic mean over targets to account for cases where $|\predictions|\neq|\targets|$ and discourage multiple predictions matching the same target (in cases where $|\predictions|\geq|\targets|$):
+A naive approach might take an average over the norms for all prediction-target pairs.
+However this gives rise to unwanted cross-terms and the possibility of multiple predictions matching the same target without any penalty for unmatched targets.
+Therefore, we choose a geometric mean over the predictions and an arithmetic mean over targets:
 \begin{equation}
@@
     ^{\frac{1}{|\predictions|}}
-    %\label{eq:error}
+    \label{eq:error}
 \end{equation}
-The supervised term is only zero when each target is matched by at least one prediction.
+The supervised term is only zero when each target is matched by at least one prediction and allows for cases where the number of predictions is greater than or equal to the number of targets $|\predictions| \geq |\targets|$.
+An alternative approach, which undesirably introduces more hyper-parameters, would be to let each prediction $\predictions(\theta)$ represent the centroid of a mixture distribution and use expectation-maximisation to match the centroids to targets $\targets$.
 
 \subsubsection{Unsupervised term: encouraging bifurcations}
-
 \begin{wrapfigure}{r}{0.5\textwidth}
@@
 
-We can see from Figures \ref{fig:minimal-models} and definitions \eqref{eq:predictions} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant $\Det=0$.
-Meanwhile the directional derivative along the bifurcation curve is finite $\frac{d}{ds}\Det\neq 0$.
-Using these quantities we can define a positive semi-definite measure $\measure(s)$ of zero crossings in the determinant along the bifurcation curve, which we define as
+We can see from Figure \ref{fig:minimal-models} and definition \eqref{eq:predictions} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant crosses zero $\Det=0$ with a finite slope $\frac{d}{ds}\Det\neq 0$.
+Using these quantities we can define a positive semi-definite measure $\measure(s)$ of zero crossings in the determinant along a curve parameterised by $s$ which we define as
 \begin{equation}
@@
 More specifically, the measure $\measure(s)$ is one at bifurcation points and goes to zero an odd number of times between bifurcations.
-This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero and hence the measure $\measure(s)$ going to zero for each turning point (see Figure \ref{fig:measure}).
-As the determinant $\Det$ diverges, we approach regimes far away from any closest bifurcation and hence $\measure(s)\rightarrow0$.
-But we would still like to have non-zero gradients with respect to $\theta$ in this regime, and so the measure is designed to go to zero slowly.
-Finally, we note that $\measure(s)$ has the desirable property of being invariant to certain classes of transformations on $\rates(u,p)$ that do not change bifurcation locations (details provided in Appendix \ref{appendix:measure}).
+This is because $\Det$ must eventually turn around in order to return back to zero, resulting in the directional derivative $\frac{d}{ds}\Det$ going to zero.
+Hence the measure $\measure(s)$ goes to zero for each turning point (see Figure \ref{fig:measure}).
 
-While the calculation of the determinant is straightforward, its directional derivative is not.
-Fortunately, a tangent field representation of the bifurcation curve allows us to compute this derivative implicitly; see Appendix \ref{appendix:tangent-fields} for details.
+On the other hand, as the determinant $\Det$ diverges, we approach regimes far away from any bifurcations and hence $\measure(s)\rightarrow0$.
+Since we would still like to have non-zero gradients with respect to $\theta$ in these regimes we designed the measure to go to zero sufficiently slowly.
+\clearpage
 
-Equipped with a measure that quantifies the appearance of bifurcations along a bifurcation arc, we now define the total measure for a bifurcation diagram as
+% Finally, we note that $\measure(s)$ has the desirable property of being invariant to certain classes of transformations on $\rates(u,p)$ that do not change bifurcation locations (details provided in Appendix \ref{appendix:measure}).
+
+While the calculation of the determinant is straightforward, its directional derivative requires a tangent vector to the bifurcation curve.
+Fortunately the tangent vector $\tangent(s)$ at the solution $u(s),p(s)$ anywhere along the curve $s$ can be calculated as the nullspace of the rectangular $N\times(N+1)$ Jacobian
+\begin{equation}
+    \left.
+    \frac{\partial\rates}{\partial(u,p)}
+    \right|_{\rates(u(s),p(s))=0}
+    \cdot\tangent(s)=0
+    \label{eq:tangent}
+\end{equation}
+This equation guarantees that the tangent vector $\tangent(s)$ is orthogonal to all hyper-planes defined by the components of $\rates$.
+In this setting the dimension of the nullspace is always known, and therefore can reliably be calculated using QR factorisation methods \cite{Drmac2008OnStudy}.
+
+Equipped with a measure that quantifies the appearance of bifurcations along a bifurcation arc we can define the total measure for a bifurcation diagram as
 \begin{equation}
@@
     }.
+    \label{eq:total-measure}
 \end{equation}
@@
 The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the controlled condition region $p$ is densely packed with bifurcations.
-
-The total measure $\Psi(\theta)$ is added to the supervised term as if it were a likelihood. This defines the semi-supervised cost function as
+The total measure $\Psi(\theta)$ is added to the supervised term as if it were a likelihood.
+This defines the semi-supervised cost function as
 \begin{align}
     \loss(\theta|\targets):=
-    \big(|\predictions|-|\targets|\big)\,\,\lambda \log\Psi(\theta)+
+    \big(|\predictions|-|\targets|\big)\log\Psi(\theta)+
     \error(\theta, \targets),
@@
 \end{align}
-where we introduce the hyperparameter $\lambda$ to enable control over the relative contribution of the unsupervised term.
 The pre-factor $|\targets|-|\predictions|$ in the unsupervised term ensures that the gradients are always pushing optimisers towards a state where $|\targets|=|\predictions|$.
@@
 
-To make use of gradient-based optimisers to locate desired bifurcation diagrams, we here show how to differentiate the cost function.
+To make use of gradient-based optimisers to locate desired bifurcation diagrams, we show here how to differentiate the cost function.
 First, we note that while individual bifurcations $p(\theta)$ depend smoothly on $\theta$, the total number of predictions $|\predictions|$ does not have gradient contributions with respect to $\theta$.
-Therefore, we can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ (within the definition of $\error$) and the total measure $\Psi(\theta)$.
+Therefore, we can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ within the definition the supervised term \eqref{eq:error} and the total measure \eqref{eq:total-measure}.
 Therefore,
@@
 \end{align}
-
 In a similar vein to back-propagation through neural differential equations \cite{Chen2018NeuralEquations} we would like to be able to calculate the gradient $\frac{\partial\loss}{\partial\theta}$ without having to differentiate through the operations of the solver that finds the bifurcation diagram $\rates(u,p)=0$ and the bifurcation locations $p(\theta)$.
@@
 {\phantomsubcaption\label{fig:two-state-optima:models}}
-% \includegraphics[width=9cm]{two-state-optima}
-\includegraphics[width=9cm]{two-state-optima.png}
+\includegraphics[width=9cm]{two-state-optima}
 \begin{picture}(0,0) \put(-9,6.3){\subref{fig:two-state-optima:parameters}} \end{picture}
@@
 
-Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods, the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
-However instead of evaluating the inversions directly, we solve a system of linear equations, applying the same strategy as implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels}.
-This leaves us with the computational bottleneck of calculating the determinant of the state space Jacobian, required in the both the bifurcation measure \eqref{eq:measure} and gradient \eqref{eq:gradient-bifurcation}.
-This calculation scales like $N^3$ where $N$ is the number of state space variables (Figure \ref{fig:scaling}).
+Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods, which includes the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
+Instead of evaluating the inversions directly, we solve a system of linear equations, applying the same strategy as implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels}.
+This leaves us with the computational bottleneck of calculating the determinant of the state space Jacobian, required in both the bifurcation measure \eqref{eq:measure} and gradient \eqref{eq:gradient-bifurcation}.
+This calculation scales like $N^2$ where $N$ is the number of state space variables (Figure \ref{fig:scaling}).
 
-\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{figure}[ht]%{r}{0.5\textwidth}
 \centering
@@
 \label{fig:scaling}
-\end{wrapfigure}
+\end{figure}
 
 For this calculation a model \eqref{eq:scaling} was chosen that so that it is extensible both in the number of parameters $M$ and the number of states $N$. 
+\begin{equation}
+\begin{cases}
+    \,\,\partial_t u_1 = \sin^{2}\!p - (\theta_1\sin^{2}\!p+1)u_1\\
+    \,\,\partial_t u_n = u_{n-1} - (\mu_n^2+1) u_n & 2\le n\leq N
+\end{cases}
+\label{eq:scaling}
+\end{equation}
 In this model only the first state $u_1$ defines the shape of the bifurcation diagram, while the remaining states are merely linearly proportional to the first.
@@
 Note that any more efficient calculation must still permit back-propagation through it.
-\begin{equation}
-\begin{cases}
-    \,\,\partial_t u_1 = \sin^{2}\!p - (\theta_1\sin^{2}\!p+1)u_1\\
-    \,\,\partial_t u_n = u_{n-1} - (\mu_n^2+1) u_n & 2\le n\leq N
-\end{cases}
-\label{eq:scaling}
-\end{equation}
+
 \section{Conclusion \& Broader Impact}
@@
 By applying implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} and the generalised Leibniz rule \cite{Flanders1973DifferentiationSign} to the geometry of the implicitly defined steady states \cite{Goldman2005CurvatureSurfaces} it is possible to use automatic differentiation methods to efficiently calculate gradients.
-We defined a bifurcation measure that that uses the determinant of the state-space Jacobian as an indicator for bifurcating parameter regimes in the unsupervised term of the cost function.
+We defined a bifurcation measure that uses the determinant of the state-space Jacobian as an indicator for bifurcating parameter regimes in the unsupervised term of the cost function.
 The gradients of the cost can be efficiently computed using automatic differentiation methods.
@@
 
-Although we did not consider limit cycles, our approach can be modified using the conditions for Andronov-Hopf bifurcations instead of the conditions on the Jacobian determinant.
+Although we did not consider limit cycles, our approach can be modified using the conditions for Poincar\'e-Andronov-Hopf bifurcations instead of the conditions on the Jacobian determinant.
 Our approach generalises naturally to bifurcation manifolds such as limit point curves or surfaces.

### 2021-06-08 docs/appendix.tex
submitting to neurips
arxiv pre-print

@@
 
-\section{Properties of the Bifurcation Measure}
-\label{appendix:measure}
+\section{Conditions for Bifurcations}
+\label{appendix:conditions}
+
+A saddle-node bifurcation is the generic bifurcation of equilibria and additional conditions are required at a bifurcation point to obtain a transcritical or pitchfork bifurcations \cite{Kuznetsov2004TopologicalSystems}.
+If we are not specifically looking for transcritical or pitchfork bifurcations, it is sufficient to consider a non-zero Hessian determinant and non-zero vector slope
+\begin{equation}
+    \left|\frac{\partial^2\rates}{\partial u^2}\right| \neq 0\qquad \frac{\partial\rates}{\partial p}\neq 0
+    \label{eq:conditions}
+\end{equation}
+In the multidimensional case we would like to express the above conditions succinctly in terms of a scalar derivative.
+Let us consider the non-degeneracy condition along the equilibrium manifold in terms of the state-space determinant
+\begin{equation}
+    \frac{d}{ds}\left|\frac{\partial\rates}{\partial u}\right|\neq 0
+\end{equation}
+Applying Jacobi's formula and expanding the directional derivative in terms of basis vectors $\hat{u}(s)$ and $\hat{p}(s)$ that are tangent to the equilibrium manifold at a particular location $s$ we get
+\begin{equation}
+     \hat{p}(s)\,\mathrm{Tr}\left[
+        \frac{\partial^2\rates}{\partial u\partial p}
+        \,\mathrm{adj}\frac{\partial\rates}{\partial u}
+        \right]
+        +
+    \sum_{n=1}^N\hat{u}_{n}(s)\,\mathrm{Tr}\left[
+        \frac{\partial^2\rates}{\partial u \partial u_n}
+        \,\mathrm{adj}\frac{\partial F}{\partial u}
+        \right]
+  \neq 0
+\end{equation}
+We can see this expression contains the vector Hessian $\frac{\partial^2\rates}{\partial u^2}$ whos determinant is non-zero for the non-degenerate case.
+The slope $\frac{\partial\rates}{\partial p}$ is non-zero when the transversality condition holds.
 
@@
 The last integrand term can be seen as the divergence the vector field $\varphi_\theta(z)$ projected onto the one dimensional space curve.
-
-% \section{Simplification of two-state model}
-
-% Consider the general two-state model:
-% \begin{equation}
-%     \dfrac{dv_i}{dt} = \dfrac{a_i + b_i (K_i v_j)^2}{1 + (K_i v_j)^2} - \mu_i v_i
-% \end{equation}
-
-% Without loss of generality, we can rescale $v_i = U_i u_i$. Then,
-% \begin{equation}
-%     \dfrac{U_idu_i}{dt} = \dfrac{a_i + b_i (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i U_i u_i
-% \end{equation}
-
-% Dividing through by $U_i$, we obtain
-% \begin{equation}
-%     \dfrac{du_i}{dt} = \dfrac{\frac{a_i}{U_i} + \frac{b_i}{U_i} (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i u_i
-% \end{equation}
-
-% By choosing $U_i = b_i$, the model reduces to:
-% \begin{equation}
-%     \dfrac{du_1}{dt} = \dfrac{\hat{a}_1 + (p u_2)^2}{1 + (p u_2)^2} - \mu_1 u_1, \quad
-%     \dfrac{du_2}{dt} = \dfrac{\hat{a}_2 + (k u_1)^2}{1 + (k u_1)^2} - \mu_2 u_2
-% \end{equation}
-% where $p = \frac{K_1}{\sqrt{b_1}}$ and $k = \frac{K_2}{\sqrt{b_2}}$.

### 2021-06-08 docs/article.tex
arxiv pre-print

@@
     \item An end-to-end differentiable method for locating bifurcations in parameter space and then matching their dependency on a control condition to user-specified locations
-    \item Implementation of the method as a Julia package\\  \href{https://github.com/anon/Code.jl}{\texttt{see attached zip file}}
-% \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
+    \item Implementation of the method as a Julia package\\
+\href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
     \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
@@
 
-% \section{Acknowledgements}
-% We would like to acknowledge Kieran Cooney for the fruitful conversations that helped guide the derivations and computational approach.
+\section{Acknowledgements}
+We would like to acknowledge Kieran Cooney for the fruitful conversations that helped guide the derivations and computational approach.
+A special thanks go to Romain Veltz and the Julia community for helpful pointers on package development and discussions over Slack.
 
@@
 
-\clearpage\pagenumbering{gobble}
-\section*{Checklist}
-\input{checklist}
+% \clearpage\pagenumbering{gobble}
+% \section*{Checklist}
+% \input{checklist}
 

### 2021-06-11 src/BifurcationFit.jl
rename package

@@ -1,4 +1,4 @@
-module FluxContinuation
+module BifurcationFit
 
 	using BifurcationKit: ContIterable, newton, ContinuationPar, NewtonPar, DeflationOperator
 	using BifurcationKit: BorderedArray, AbstractLinearSolver, AbstractEigenSolver, BorderingBLS

### 2021-06-11 test/figures.jl

@@ -1,4 +1,4 @@
-using FluxContinuation
+using BifurcationFit
 using ForwardDiff
 
 using LaTeXStrings,Plots
@@ -66,29 +66,51 @@ include("scaling/scaling.jl")
 using StatsBase: mean,std
 using LaTeXStrings
 
-complexity_mean = Float64[]
-complexity_std = Float64[]
+continuation_mean = Float64[]
+continuation_std = Float64[]
 
-for n ∈	1:40
-	ts = [scaling(n,3) for _ ∈ 1:10]
-	push!(complexity_mean,mean(ts))
-	push!(complexity_std,std(ts))
+forward_mean = Float64[]
+forward_std = Float64[]
+
+backward_mean = Float64[]
+backward_std = Float64[]
+
+for n ∈	41:100
+
+	ts = [scaling_continuation(n,3) for _ ∈ 1:10]
+	push!(continuation_mean,mean(ts))
+	push!(continuation_std,std(ts))
+
+	ts = [scaling_forward(n,3) for _ ∈ 1:10]
+	push!(forward_mean,mean(ts))
+	push!(forward_std,std(ts))
+
+	# ts = [scaling_backward(n,3) for _ ∈ 1:10]
+	# push!(backward_mean,mean(ts))
+	# push!(backward_std,std(ts))	
 end
-scaling(40,3)
 
-default();default(grid=false)
-plot( 10:30, x->x^3/1800, label=L"N^3", color=:red, linewidth=2, legend=:topleft)
-scatter!( complexity_mean,yerror=complexity_std, label="", yscale=:log10, xscale=:log10, ms=3, color=:black,
-	xlabel=L"\mathrm{States}\,\,N", ylabel=L"\mathrm{Execution\,\,Time}\,\,/\,\,\mathrm{sec}", size=(400,400) )
+default();default(grid=false,label="", yscale=:log10, xscale=:log10,ms=3, ylim=(8e-2,1e1), msc=:auto, lc=:black,
+	xlabel=L"\mathrm{States}\,\,N",ylabel=L"\mathrm{Execution\,\,Time}\,\,/\,\,\mathrm{sec}", size=(400,400),
+	linewidth=2, legend=:topleft)
 
-xticks!([1,3,5,10,20,30],["1","3","5","10","20","30"])
+plot( 10:40, x->x^2/180, label=L"N^2", lc=:red )
[…]

### 2021-06-11 test/runtests.jl

@@ -1,7 +1,7 @@
 using FiniteDiff: finite_difference_gradient
-using FluxContinuation, Test
+using BifurcationFit, Test
 
-error_tolerance = 0.06
+error_tolerance = 0.05
 @testset "Gradients" begin
 
 	@testset "Saddle Node" begin

### 2021-06-11 docs/checklist.tex
submitting to neurips
rename package

@@
   \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerYes{Figures can be reproduced using unit tests in the GitHub repository for \texttt{BifurcationFit.jl}}
+    \answerYes{Figures can be reproduced using unit tests in the GitHub repository for \texttt{BifurcationInference.jl}}
   \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
@@
   \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerYes{Julia package implementing our method \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}}
+    \answerYes{Julia package implementing our method \href{https://github.com/anon/Code.jl}{\texttt{Julia package in attached zip file}}}
   \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?

### 2021-06-11 src/BifurcationInference.jl

@@ -1,4 +1,4 @@
-module BifurcationFit
+module BifurcationInference
 
 	using BifurcationKit: ContIterable, newton, ContinuationPar, NewtonPar, DeflationOperator
 	using BifurcationKit: BorderedArray, AbstractLinearSolver, AbstractEigenSolver, BorderingBLS

### 2021-06-11 test/figures.jl

@@ -1,4 +1,4 @@
-using BifurcationFit
+using BifurcationInference
 using ForwardDiff
 
 using LaTeXStrings,Plots
@@ -110,7 +110,7 @@ savefig("docs/figures/scaling.pdf")
 ####################################################################################################
 ####################################################################################################
 ################################################################################ two-state-optima
-using BifurcationFit,Plots,StaticArrays
+using BifurcationInference,Plots,StaticArrays
 using Clustering,JLD
 using StatsBase: median
 

### 2021-06-11 test/runtests.jl

@@ -1,5 +1,5 @@
 using FiniteDiff: finite_difference_gradient
-using BifurcationFit, Test
+using BifurcationInference, Test
 
 error_tolerance = 0.05
 @testset "Gradients" begin

### 2021-06-11 docs/article.tex
rename package
update appendix

@@
-\input{style}
+\documentclass{article}
+\PassOptionsToPackage{numbers,compress}{natbib}
+\usepackage{lipsum}
+
+% \usepackage{neurips_2021}
+\usepackage[preprint]{neurips_2021}
+% \usepackage[final]{neurips_2021}
+
+\usepackage[T1]{fontenc}
+\usepackage[latin1]{inputenc}
+\usepackage{url}
+
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks=true,
+    linkcolor=black,
+    filecolor=blue,
+    citecolor=blue,      
+    urlcolor=cyan,
+}
+
+\usepackage{graphicx,xcolor,float}
+\graphicspath{{./figures/}{docs/figures/}}
+
+\usepackage{booktabs,amsmath,amsfonts}
+\usepackage{nicefrac,microtype}
+
+\usepackage{caption,subcaption}
+\renewcommand\thesubfigure{\Alph{subfigure}}
+
+\usepackage[ruled,vlined]{algorithm2e}
+\usepackage{wrapfig}
 
@@
 \begin{document}
-\input{notation}
+
+\newcommand{\rates}{F_{\theta}}
+\newcommand{\tangent}{T_{\theta}}
+\newcommand{\steadystates}{\partial S_{\theta}}
+
+\newcommand{\Det}{\left| \frac{\partial\rates}{\partial u} \right|}
+\newcommand{\measure}{\varphi_{\theta}}
+
+\newcommand{\predictions}{\mathcal{P}}
+\newcommand{\targets}{\mathcal{D}}
+\newcommand{\loss}{L}
+\newcommand{\error}{E}
+\newcommand{\Reals}{\mathbb{R}}
+
 \maketitle
@@
     \item Implementation of the method as a Julia package\\
-\href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
+\href{https://github.com/gszep/BifurcationInference.jl}{\texttt{github.com/gszep/BifurcationInference.jl}}
     \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
@@
 The determinant of the Jacobian quantifies the rate at which trajectories in a local patch of state-space $u\in\Reals^N$ converge or diverge.
-The determinant approaching zero means that the dynamics of the system are slowing down, which is an important indicator for the onset of a transition between qualitative behaviours.
-Furthermore, the slowing down must necessarily be followed by breakdown of stability; for this to be true it is sufficient \textit{but not necessary} to require that the determinant cross zero with a finite slope, meaning that its directional derivative along the diagram $\frac{d}{ds}\Det$ is not zero.
+Let $s\in\Reals$ parametrise the curves that trace out the bifurcation diagram.
+Any location on the curve $u(s)$ and $p(s)$ must satisfy the steady-state of equations \eqref{eq:model}.
+The determinant approaching zero along the diagram means that the dynamics of the system are slowing down, which is an important indicator for the onset of a transition between qualitative behaviours.
+Furthermore, the slowing down must necessarily be followed by a breakdown of stability; for this to be true it is sufficient \textit{but not necessary} to require that the determinant cross zero with a finite slope, meaning that its directional derivative along the diagram $\frac{d}{ds}\Det$ is not zero.
 The set of predicted values for the control condition $\predictions(\theta)\subset\Reals$ at which bifurcations occur are defined as
@@
 \end{align}
+A proof of how the conditions \eqref{eq:predictions} are sufficient but not necessary for local bifurcations is detailed in Appendix \ref{appendix:conditions}.
 The most common bifurcations between steady states, not including limit cycles, are saddle-nodes and pitchforks \cite{Haragus2011LocalSystems}.
-The relationships between our conditions for bifurcation \eqref{eq:predictions} and the conditions for saddle-node are detailed in Appendix \ref{appendix:conditions}.
 Saddle-node bifurcations, which often appear in pairs (Figure \ref{fig:saddle-node}) are defined by stable and unstable fixed points meeting and disappearing.
@@
 {\phantomsubcaption\label{fig:pitchfork}}
-\includegraphics[width=6cm]{saddle-node}
-\begin{picture}(0,0) \put(-5.8,10){\subref{fig:saddle-node}} \end{picture}
-\includegraphics[width=6cm]{pitchfork}
-\begin{picture}(0,0) \put(-5.8,10){\subref{fig:pitchfork}} \end{picture}
+\includegraphics[width=5.5cm]{saddle-node}
+\begin{picture}(0,0)\put(-5.5,9){\subref{fig:saddle-node}} \end{picture}
+\includegraphics[width=5.5cm]{pitchfork}
+\begin{picture}(0,0)\put(-5.5,9){\subref{fig:pitchfork}} \end{picture}
 \caption{Illustration of bifurcation diagrams for minimal models of bifurcations.
@@
 
-% \clearpage\pagenumbering{gobble}
-% \section*{Checklist}
-% \input{checklist}
-
 \clearpage\pagenumbering{arabic}\setcounter{page}{1}

### 2021-06-11 docs/appendix.tex

--- 2021-05-28 docs/appendix.tex
@@
 
-\section{Properties of the Bifurcation Measure}
-\label{appendix:measure}
+\section{Conditions for Bifurcations}
+\label{appendix:conditions}
+
+The conditions for bifurcations \eqref{eq:predictions} are expressed at a parametrised location $s\in\Reals$ on the equilibrium manifold.
+A local bifurcation \textit{not involving limit cycles} can be defined as a change of sign of the real part of an eigenvalue of the state-space Jacobian.
+In this section we show that the conditions
+\begin{equation}
+    \Det_{s=s^*} = 0
+    \qquad
+    \frac{d}{ds}\Det_{s=s^*} \neq 0
+    \label{eq:conditions}
+\end{equation}
+are sufficient for a local bifurcation to occur at a specific location $s^*$ along the equilibrium manifold.
+Here we denote $\Det$ as the determinant of the state-space Jacobian.
+We begin by expressing the determinant is the product of eigenvalues
+\begin{equation}
+    \Det=\prod_{n=1}^N\lambda_n(s)
+    \label{eq:determinant}
+\end{equation}
+Applying the product rule when differentiating yields
+\begin{equation}
+    \frac{d}{ds}\Det=
+    \sum_{n=1}^N\frac{d\lambda_n}{ds}\prod_{n'\neq n}\lambda_{n'}(s)
+\end{equation}
+As $\Det_{s=s^*}=0$ then by \eqref{eq:determinant} one of the eigenvalues must vanish, say $\lambda^*(s)$ at  $s=s^*$ then
+\begin{equation}
+    \frac{d}{ds}\left|\frac{\partial\rates}{\partial u}\right|_{s=s^*}=
+    \frac{d\lambda^*}{ds}\prod_{\lambda\neq\lambda^*}\lambda(s)\,\bigg|_{s=s^*}
+\end{equation}
+which implies the following conditions on the eigenvalues
+\begin{align}
+    \left.\lambda^*(s)\right|_{s=s^*}=0
+    \qquad
+    \left.\frac{d\lambda^*}{ds}\right|_{s=s^*}\neq 0
+    \label{eq:sufficient-necessary}
+    \\
+    \lambda(s^*) \neq 0\quad\forall\lambda\neq\lambda*
+    \qquad\quad
+    \label{eq:sufficient}
+\end{align}
+Conditions \eqref{eq:sufficient-necessary} guarantee that eigenvalue $\lambda^*(s)$ crosses zero at $s^*$ with a finite slope and are therefore sufficient and necessary for a local non-degenerate bifurcation not involving limit cycles.
+Condition \eqref{eq:sufficient} restricts the bifurcation at $s^*$ to be such that only one eigenvalue may vanish with all others being finite.
+Therefore conditions \eqref{eq:conditions} are sufficient \textit{but not necessary} for local non-degenerate bifurcations at $s^*$ not involving limit cycles.
+
+In the case where more than one eigenvalue vanishes at the bifurcation the second condition \eqref{eq:conditions} is broken.
+This is because the determinant scales like $(s-s^*)^k$ where $k$ is the number of eigenvalues crossing zero.
+When the second condition \eqref{eq:conditions} is broken the measure \eqref{eq:measure} becomes zero.
+This means that the unsupervised term does not encourage bifurcations which have more than one eigenvalue crossing zero.
 

### 2021-10-22 src/Plots.jl
plotting imaginary parts of eigenvalues

@@ -3,62 +3,45 @@ using LaTeXStrings
 using Plots
 
 import Plots: plot,plot!
-function plot(steady_states::Vector{<:Branch},data::StateSpace; determinant=true, kwargs...)
+function plot(steady_states::Vector{<:Branch},data::StateSpace; kwargs...)
 
-	if determinant
+	layout = @layout [a;b{1.0w,0.5h}]
+	default(); default( ; grid=false,label="",margin=1mm,linewidth=2, lims=:round, kwargs...)
 
-		layout = @layout [a;b{1.0w,0.5h}]
-		default(); default( ; grid=false,label="",margin=1mm,linewidth=2,kwargs...)
-		figure = plot(layout = layout, link = :x, size=(300,500) )
+	figure = plot(layout = layout, link = :x, size=(300,500), ylabel=L"\mathrm{steady\,\,states}\quad |\,u\,\,|" )
+	hline!( [0],subplot=2,linewidth=1,color=:black, xlabel=L"\mathrm{control\,\,condition,}p",
+		xmirror=true, topmargin=-5mm, ylabel=L"\mathrm{spectrum}\,\quad \rho(\lambda)")
+	
+	vline!(data.targets,subplot=1,linewidth=1,color=:gold)
+	vline!(data.targets,subplot=2,linewidth=1,color=:gold)
 
-		hline!([0],subplot=1,linewidth=0,color=:black, ylabel=L"\mathrm{steady\,\,states}\quad F_{\theta}(u,p)=0")
-		hline!([0],subplot=2,linewidth=1,color=:black, xlabel=L"\mathrm{control\,\,condition,}p", xmirror=true, topmargin=-5mm,
-			ylabel=L"\mathrm{determinant}\,\quad\left|\!\!\!\!\frac{\partial F_{\theta}}{\partial u}\right|")
-		
-		vline!(data.targets,subplot=1,linewidth=1,color=:gold)
-		vline!(data.targets,subplot=2,linewidth=1,color=:gold)
+	for branch ∈ steady_states
 
-		for branch ∈ steady_states
+		alpha = map(s->window_function(data.parameter,s.z),branch)
+		parameter = map( s -> s.z.p, branch)
+		states = map(s-> norm(s.z.u), branch)
 
-			stability = map( s -> all(real(s.λ).<0), branch)
-			determinants = map( s -> prod(real(s.λ)), branch)
-			parameter = map( s -> s.z.p, branch)
[…]

### 2021-10-22 src/BifurcationInference.jl
compat with BifurcationKit 0.1.6

@@ -2,9 +2,9 @@ module BifurcationInference
 
 	using BifurcationKit: ContIterable, newton, ContinuationPar, NewtonPar, DeflationOperator
 	using BifurcationKit: BorderedArray, AbstractLinearSolver, AbstractEigenSolver, BorderingBLS
-	using BifurcationKit: ContState, solution, computeEigenvalues!, detectBifucation
+	using BifurcationKit: ContState, detectBifucation
 
-	using ForwardDiff: gradient,jacobian
+	using ForwardDiff: Dual,tagtype,derivative,gradient,jacobian
 	using Flux: Momentum,update!
 
 	using Setfield: @lens,@set,setproperties

### 2021-10-22 src/Plots.jl

@@ -6,9 +6,9 @@ import Plots: plot,plot!
 function plot(steady_states::Vector{<:Branch},data::StateSpace; kwargs...)
 
 	layout = @layout [a;b{1.0w,0.5h}]
-	default(); default( ; grid=false,label="",margin=1mm,linewidth=2, lims=:round, kwargs...)
+	default(); default( ; grid=false,label="",margin=1mm,linewidth=2,lims=:round, kwargs...)
 
-	figure = plot(layout = layout, link = :x, size=(300,500), ylabel=L"\mathrm{steady\,\,states}\quad |\,u\,\,|" )
+	figure = plot(layout = layout, link = :x, xlim=extrema(data.parameter), size=(300,500), ylabel=L"\mathrm{steady\,\,states}\quad |\,u\,\,|" )
 	hline!( [0],subplot=2,linewidth=1,color=:black, xlabel=L"\mathrm{control\,\,condition,}p",
 		xmirror=true, topmargin=-5mm, ylabel=L"\mathrm{spectrum}\,\quad \rho(\lambda)")
 	

### 2021-10-22 src/BifurcationInference.jl
compat

@@ -121,7 +121,7 @@ module BifurcationInference
 						end
 
 						midpoint = sum( s -> s.z.p, branch ) / length(branch)
-						if minimum(pRange) < midpoint && midpoint < maximum(pRange)
+						if minimum(pRange) < midpoint < maximum(pRange)
 							push!(branches,pMin < 0 ? reverse(branch) : branch) end
 
 					catch error

### 2021-10-22 src/Structures.jl

@@ -47,7 +47,7 @@ Branch{V,T} = Vector{BranchPoint{V,T}}
 dim(branch::Branch) = length(branch) > 0 ? length(first(branch).z)-1 : 0
 
 function push!(branch::Branch,state::ContState)
-    z,∂z = copy(solution(state)), [state.tau.u;state.tau.p]
+    z,∂z = copy(state.z_old), [state.tau.u;state.tau.p]
     push!(branch, ( z=z, λ=state.eigvals, ds=norm(∂z)*abs(state.ds), bif=detectBifucation(state) ))
 end
 

### 2021-10-22 src/Objectives.jl
using eigen gradient

@@ -17,8 +17,47 @@ function errors( predictions::AbstractVector{<:BorderedArray}, targets::StateSpa
 	return mean( p′-> mean( z->(z.p-p′)^2, predictions; type=:geometric ), targets.targets; type=:arithmetic )
 end
 
+realeigvals(J) = real(eigen(J).values)
 function measure( F::Function, z::BorderedArray, θ::AbstractVector, targets::StateSpace )
-	return window_function(targets.parameter,z) / ( 1 + abs( derivative( z->log(abs(det(∂Fu(F,z,θ)))), z, tangent_field(F,z,θ) ) )^(-1) ) # todo@(gszep) determinant calculation is computational bottleneck
+	λ,dλ = realeigvals(∂Fu(F,z,θ)), derivative( z -> realeigvals(∂Fu(F,z,θ)), z, tangent_field(F,z,θ) )
+	return window_function(targets.parameter,z) * mapreduce( (λ,dλ) -> 2 / ( 2 + abs(sinh(2λ)/dλ) ), +, λ, dλ )
+end
+
+# Giles, M. 2008. An extended collection of matrix derivative results for forward and reverse mode automatic differentiation
+# credit: @mateuszbaran https://github.com/JuliaManifolds/Manifolds.jl/pull/27#issuecomment-521693995
+import LinearAlgebra: eigen
+
+function make_eigen_dual(val::Real, partial)
+    Dual{tagtype(partial)}(val, partial.partials)
+end
+
+function make_eigen_dual(val::Complex, partial::Complex)
+    Complex(Dual{tagtype(real(partial))}(real(val), real(partial).partials),
+        Dual{tagtype(imag(partial))}(imag(val), imag(partial).partials))
+end
+
+function eigen(A::StridedMatrix{Dual{T,V,N}}) where {T,V,N}
+
+    Av = map(a -> a.value, A)
+    λ,U = eigen(Av)
+
+    dΛ = U \ A * U
+    dλ = diag(dΛ)
+
+    F = similar(Av, eltype(λ))
+    for i ∈ axes(A,1), j ∈ axes(A,2)
+        if i == j
+            F[i, j] = 0
+        else
+            F[i, j] = inv(λ[j] - λ[i])
+        end
+    end
+    dU = U * (F .* dΛ)
+
+    for i ∈ eachindex(dU)
+        dU[i] = make_eigen_dual(U[i], dU[i])
+    end
+    return Eigen(dλ, dU)
 end
 
 ################################################################################

### 2021-10-22 src/Utils.jl

@@ -29,7 +29,7 @@ end
 ############################################################################# training loop
 function train!( F::Function, parameters::NamedTuple, data::StateSpace;
 				 iter::Int=200, optimiser=Momentum(0.001), plot_solution = false, 
-				 convergence = 0.01, kwargs...)
+				 convergence = 0.02, kwargs...)
 
 	hyperparameters = getParameters(data;kwargs...)
 	Loss = steady_states = nothing

### 2021-10-22 test/runtests.jl

@@ -2,7 +2,7 @@ using FiniteDiff: finite_difference_gradient
 using BifurcationInference, Test
 using Flux: Optimise
 
-error_tolerance = 0.003
+error_tolerance = 0.004
 @testset "Gradients" begin
 
 	@testset "Saddle Node" begin

### 2021-10-22 src/Gradients.jl
more robust bifurcation count

@@ -1,7 +1,7 @@
 ################################################################################
 function ∇loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
 
-	predictions = [ s.z for branch ∈ branches for s ∈ branch if s.bif ]
+	predictions = unique([ s.z for branch ∈ branches for s ∈ branch if s.bif ]; atol=2*step(targets.parameter))
 	λ = length(targets.targets)-length(predictions)
 
 	if λ≠0 

### 2021-10-22 src/Objectives.jl

@@ -1,7 +1,7 @@
 ################################################################################
 function loss( F::Function, branches::AbstractVector{<:Branch}, θ::AbstractVector, targets::StateSpace; kwargs...)
 
-	predictions = [ s.z for branch ∈ branches for s ∈ branch if s.bif ]
+	predictions = unique([ s.z for branch ∈ branches for s ∈ branch if s.bif ]; atol=2*step(targets.parameter))
 	λ = length(targets.targets)-length(predictions)
 
 	if λ≠0 

### 2021-10-22 src/Structures.jl

@@ -91,6 +91,6 @@ end
 
 #################################################### display methods
 show(io::IO, branches::Vector{Branch{V,T}}) where {V,T} = print(io,
-    "Vector{Branch}[dim=$(dim(first(branches))) bifurcations=$(length([ s.z for branch ∈ branches for s ∈ branch if s.bif ])) branches=$(length(branches)), states=$(sum(branch->length(branch),branches))]")
+    "Vector{Branch}[dim=$(dim(first(branches))) bifurcations=$(length(unique([ s.z for branch ∈ branches for s ∈ branch if s.bif ]; atol=0.1))) branches=$(length(branches)), states=$(sum(branch->length(branch),branches))]")
 show(io::IO, M::MIME"text/plain", branches::Vector{Branch{V,T}}) where {V,T} = show(io,branches)
 show(io::IO, states::StateSpace{N,T}) where {N,T} = print(io,"StateSpace{$N,$T}(parameters=$(states.parameter),targets=$(states.targets))")

### 2021-10-22 src/Utils.jl

@@ -53,7 +53,7 @@ function train!( F::Function, parameters::NamedTuple, data::StateSpace;
 		if isinf(Loss) throw("infinite loss") end
 		push!(trajectory,copy(parameters.θ))
 
-		predictions = [ s.z for branch ∈ steady_states for s ∈ branch if s.bif ]
+		predictions = unique([ s.z for branch ∈ steady_states for s ∈ branch if s.bif ]; atol=2*step(data.parameter))
 		if (Loss < convergence) & (length(predictions) == length(data.targets))
 			break
 		end

### 2021-10-22 test/runtests.jl
add double exclusive unit test

@@ -56,4 +56,15 @@ include("applied/two-state.jl")
 
 		@test all( z->any(isapprox.(X.targets,z.p; atol=error_tolerance)), bifurcations)
 	end
+
+	include("applied/double-exclusive.jl")
+	@testset "Double Exclusive" begin
+		parameters = ( θ=SizedVector{21}(-0.5498107083383627, -0.30983502459648393, 0.631939862866151, -1.3407390261868093, 0.6331343612914606, -0.4049028383285403, -0.8524704190293824, -0.11432918583984555, -1.7299987854365662, -2.079032179004907, 0.4916870522978877, -0.010788745553567142, -0.3878513750140796, -0.08334187371052072, -0.2772816729274815, 0.8285363415116678, 0.054887187754647675, -0.49494957503215126, -0.8693294194842291, 0.9289938806228071, 0.22800165770588346), p=minimum(X.parameter) )
+		trajectory = train!( F, parameters, X;  iter=300, optimiser=Optimise.ADAM(0.01) )
+
+		steady_states = deflationContinuation(F,X.roots,(p=minimum(X.parameter),θ=trajectory[end]),getParameters(X))
+		bifurcations = [ s.z for branch ∈ steady_states for s ∈ branch if s.bif ]
+
+		@test all( z->any(isapprox.(X.targets,z.p; atol=error_tolerance)), bifurcations)
+	end
 end

### 2021-10-23 docs/appendix.tex
overleaf

--- 2021-05-28 docs/appendix.tex
@@
 
-\section{Properties of the Bifurcation Measure}
-\label{appendix:measure}
+\clearpage
+\section{Bifurcation Measure Properties}
+\label{appendix:conditions}
+Consider a vector $v(s)\in\Reals^N$ parameterised by $s\in\Reals$ that is tangent to an equilibrium manifold defined by $\rates(u)=0$.
+The conditions for a non-degenerate static bifurcation at $s^*$ along such a tangent can be expressed in terms of an eigenvalue $\lambda(s)$ of the state-space Jacobian crossing zero with a finite slope.
+A bifurcation exists at $s^*$ if
+\begin{align}
+    \frac{\partial\rates}{\partial u}v(s)=\lambda(s) \,v(s)
+    \quad\exists\lambda:\quad
+    \left.\lambda(s)\right|_{s=s^*}=0
+    \qquad
+    \left.\frac{d\lambda}{ds}\right|_{s=s^*}\neq 0
+    \label{eq:conditions}
+\end{align}
+These conditions are necessary and sufficient for a non-degenerate static local breakdown of stability.
+For now we do not consider dynamic bifurcations involving limit cycles or imaginary parts of eigenvalues and restrict $\lambda\in\Reals$.
+Cases where both $\left.\lambda(s)\right|_{s=s^*}=0$ and $\left.\frac{d\lambda}{ds}\right|_{s=s^*}=0$ require investigation into higher order derivatives $\frac{d^n\lambda}{ds^n}$.
+These are the cases we refer to as \emph{degenerate} and are not considered here.
+
+Instead of considering conditions on each eigenvalue individually it is possible to use the determinant of the state-space Jacobian to detect whether the conditions \eqref{eq:conditions} are satisfied.
+The determinant can be expressed as the product of eigenvalues
+\begin{equation}
+    \Det=\prod_{n=1}^N\lambda_n(s)
+    \label{eq:determinant}
+\end{equation}
+Applying the product rule when differentiating yields
+\begin{align}
+    \frac{d}{ds}\Det&=
+    \sum_{n=1}^N\frac{d\lambda_n}{ds}\prod_{n'\neq n}\lambda_{n'}(s)\\
+    &=\Det\sum_{n=1}^N\frac{d\lambda_n}{ds}\lambda_{n}(s)^{-1}
+\end{align}
+Substituting this expression into measure \eqref{eq:measure}
+\begin{equation}
+    \measure(s)=
+    \left(1+\left|\sum_{n=1}^N\frac{d\lambda_n}{ds}\lambda_{n}(s)^{-1}\right|^{-1}\right)^{-1}
+\end{equation}
+Which implies the following
+\begin{equation}
+    \exists\lambda:\quad
+    \begin{cases}
+        \,\lambda(s)=0 \quad\frac{d\lambda}{ds}\neq 0\\
+        \,\lambda(s)\neq0 \quad\frac{d\lambda}{ds}\rightarrow\pm\infty
+    \end{cases}
+    \implies
+    \measure(s)=1
+\end{equation}
+If there exists an eigenvalue that satisfies conditions \eqref{eq:conditions} then the measure is equal to one.
+The measure also approaches one in cases where the rate of change of an eigenvalue with respect to a manifold $s$ location  diverges while not crossing zero.
+This gives rise to finite gradients in the eigenvalue term in regimes far away from any bifurcation.
+
+\section{Extension for Hopf Bifurcations}
+\label{appendix:hopf}
+
+In order to detect bifurcations involving limit cycles, the measure must be extended to detect changes in the real part $\Re\mathrm{e}[\lambda(s)]$ for any eigenvalue of the Jacobian.
+These conditions can no longer be compactly written in terms of the determinant.
+Instead, the measure can be defined as the sum of eigenvalue terms
+\begin{align}
+    \measure(s):=\sum_{\lambda(s)\in\frac{\partial\rates}{\partial u}}
+    \left(\left|\frac{d}{ds}\log\Re\mathrm{e}[\lambda(s)]\right|^{-1}+1\right)^{-1}
+    \label{eq:hopf-measure}
+\end{align}
+
+The directional derivative of the logarithm diverges under two conditions: when eigenvalues vanish $\lambda(s)=0$ and when the directional derivative $\frac{d}{ds}\Re\mathrm{e}[\lambda(s)]$ diverges.
+These properties are sufficient for detecting the onset of damped oscillations and emergence of limit cycles via Hopf bifurcation as shown in Figure \ref{fig:hopf-measure}.
+Eigenvalues with negative real part which gain a finite imaginary part give rise to damped oscillations.
+At this onset we observe a discontinuity in the derivative $\frac{d}{ds}\Re\mathrm{e}[\lambda(s)]$ which is detected by equation \eqref{eq:hopf-measure}.
+Once damped oscillations exist, flipping the stability of the stable fixed point gives rise to a limit cycle, which can be detected by inspecting $\Re\mathrm{e}[\lambda(s)]$.
+
+\begin{figure}[H]
+\centering
+    \includegraphics[width=60mm]{hopf-measure}
+    \caption{Bifurcation measure $\measure(s)$ and eigenvalues $\lambda(s)$ along the arclength $s$ for two different bifurcation curves demonstrating how the measure detects non-zero imaginary parts $\Im\mathrm{m}[\lambda]$ (onset of damped oscillations marked by circle) and sign changes in real parts $\Re\mathrm{e}[\lambda]$ (Hopf bifurcations marked by stars)}
+    \label{fig:hopf-measure}
+\end{figure}
+
+In principle it is possible to construct measures to detect a variety of bifurcations as long as the conditions can be expressed in terms of derivatives with respect to fixed-point manifold direction $s$.
+Measures can be used sequentially or in parallel to encourage optimisers to run through a sequence of bifurcations or place specific bifurcation types next to each other.
 
@@
 
-% \section{Simplification of two-state model}
+\section{A More Complex Model}
+\label{appendix:double-exclusive}
+To demonstrate the wider reaching applicability of our method we optimise the \emph{double exclusive reporter} \cite{Grant2020InterpretationCircuit}, a synthetic \emph{E.
+Coli} designed to exhibit a cusp bifurcation with respect to two input signals.
+Figure \ref{fig:double-exclusive-trajectories} shows the cost function with respect to iteration for sample trajectories.
+We note that due to a
 
-% Consider the general two-state model:
-% \begin{equation}
-%     \dfrac{dv_i}{dt} = \dfrac{a_i + b_i (K_i v_j)^2}{1 + (K_i v_j)^2} - \mu_i v_i
-% \end{equation}
-
-% Without loss of generality, we can rescale $v_i = U_i u_i$. Then,
-% \begin{equation}
-%     \dfrac{U_idu_i}{dt} = \dfrac{a_i + b_i (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i U_i u_i
-% \end{equation}
-
-% Dividing through by $U_i$, we obtain
-% \begin{equation}
-%     \dfrac{du_i}{dt} = \dfrac{\frac{a_i}{U_i} + \frac{b_i}{U_i} (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i u_i
-% \end{equation}
-
-% By choosing $U_i = b_i$, the model reduces to:
-% \begin{equation}
-%     \dfrac{du_1}{dt} = \dfrac{\hat{a}_1 + (p u_2)^2}{1 + (p u_2)^2} - \mu_1 u_1, \quad
-%     \dfrac{du_2}{dt} = \dfrac{\hat{a}_2 + (k u_1)^2}{1 + (k u_1)^2} - \mu_2 u_2
-% \end{equation}
-% where $p = \frac{K_1}{\sqrt{b_1}}$ and $k = \frac{K_2}{\sqrt{b_2}}$.
+\section{Steady State Solvers}
+\label{appendix:stedy-state-solvers}

### 2021-10-23 docs/article.tex
overleaf

@@
 % \usepackage{neurips_2021}
-\usepackage[preprint]{neurips_2021}
-% \usepackage[final]{neurips_2021}
+% \usepackage[preprint]{neurips_2021}
+\usepackage[final]{neurips_2021}
 
@@
 \author{
-        Gregory Szep\\King's College London\\London, WC2R 2LS\\
+        Gregory Szep\\King's College London\\London, WC2R 2LS, UK\\
         \texttt{gregory.szep@kcl.ac.uk}
     \And
-        Attila Csik\'asz-Nagy\\King's College London\\London, WC2R 2LS\\
-        \texttt{attila.csikasz-nagy@kcl.ac.uk}
+        Attila Csik\'asz-Nagy\\P\'azm\'any P\'eter Catholic University\\Budapest, 1083, Hungary\\
+        \texttt{csikasznagy@gmail.com}
     \And
-        Neil Dalchau\\Microsoft Research Cambridge\\Cambridge, CB1 2FB\\
-        \texttt{ndalchau@microsoft.com}
+        Neil Dalchau\\Microsoft Research Cambridge\\Cambridge, CB1 2FB, UK\\
+        \texttt{ndalchau@gmail.com}
 }
@@
 In dynamical systems theory, such change points are known as \textit{bifurcations} and lie on a function of the controlled condition called the \textit{bifurcation diagram}.
-In this work, we propose a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
-The cost function contains a supervised error term that is minimal when the model bifurcations match the specified targets and an unsupervised bifurcation measure which has gradients that push optimisers towards bifurcating parameter regimes.
+In this work, we propose a gradient-based approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
+The cost function contains an error term that is minimal when the model bifurcations match the specified targets and a bifurcation measure which has gradients that push optimisers towards bifurcating parameter regimes.
 The gradients can be computed without the need to differentiate through the operations of the solver that was used to compute the diagram.
@@
 Progress has been made in cases where model structure and stability conditions are used to refine the search space \cite{Otero-Muras2018Optimization-basedModels,Otero-Muras2014ACurves} yet the resulting objectives are still not explicit in the bifurcation targets and also not differentiable.
+It is always possible to bias goodness of fit during optimisation with domain-specific knowledge \cite{}, however these approaches do not aim to generalise across a wider set of qualitative behaviours.
 In the emerging field of scientific machine learning \cite{Rackauckas2017Differentialequations.jlJulia,Rackauckas2018ASolutions,Rackauckas2020UniversalLearning}, parameters of structured mechanistic models are favoured over flexible models in larger parameter spaces.
@@
 Therefore only properties of the model can be used to start the search.
-We propose a semi-supervised approach for performing both tasks in an end-to-end fashion.
+We propose an approach for performing both tasks in an end-to-end fashion.
 The bifurcation diagram encodes high-level qualitative information defined by state space structures, rather than kinetics.
@@
 \end{align}
-In the context of the differential equations, and not considering limit cycles for now, a co-dimension one bifurcation can be defined by a set of conditions on the determinant of the Jacobian $\Det$.
+In the context of the differential equations, and not considering limit cycles for now, we show that a static non-degenerate bifurcation can be defined by a set of conditions on the determinant of the Jacobian $\Det$.
 The determinant of the Jacobian quantifies the rate at which trajectories in a local patch of state-space $u\in\Reals^N$ converge or diverge.
@@
 Furthermore, the slowing down must necessarily be followed by a breakdown of stability; for this to be true it is sufficient \textit{but not necessary} to require that the determinant cross zero with a finite slope, meaning that its directional derivative along the diagram $\frac{d}{ds}\Det$ is not zero.
+This is the non-degeneracy condition.
 The set of predicted values for the control condition $\predictions(\theta)\subset\Reals$ at which bifurcations occur are defined as
@@
 \end{align}
-A proof of how the conditions \eqref{eq:predictions} are sufficient but not necessary for local bifurcations is detailed in Appendix \ref{appendix:conditions}.
+A proof of how the conditions \eqref{eq:predictions} are necessary and sufficient for static non-degenerate bifurcations is detailed in Appendix \ref{appendix:conditions}.
 The most common bifurcations between steady states, not including limit cycles, are saddle-nodes and pitchforks \cite{Haragus2011LocalSystems}.
 Saddle-node bifurcations, which often appear in pairs (Figure \ref{fig:saddle-node}) are defined by stable and unstable fixed points meeting and disappearing.
-Pitchfork bifurcations are special cases of the saddle-node where a single steady state splits into two stable and one unstable steady state (Figure \ref{fig:pitchfork}).
+Pitchfork bifurcations occur where a single steady state splits into two stable and one unstable steady state (Figure \ref{fig:pitchfork} shows an \textit{imperfect} pitchfork; a \textit{perfect} pitchfork arises when $\theta_1=0$).
 To illustrate these bifurcations, we define minimal models (Figure \ref{fig:minimal-models}) that span the space of saddle-node and pitchforks, where indeed zero crossings in the determinant with a finite slope define the set of prediction $\predictions(\theta)$.
@@
 \label{section:method}
-\subsection{Semi-supervised Cost Function}
+\subsection{Cost Function}
 
-To identify parameter sets that give rise to bifurcation diagrams with specified bifurcation points, we propose a semi-supervised cost function that comprises two terms.
-The role of the supervised term is simply to reward predicted bifurcations to coincide with the specified target locations.
+To identify parameter sets that give rise to bifurcation diagrams with specified bifurcation points, we propose a cost function that comprises two terms.
+The role of the error term is simply to reward predicted bifurcations to coincide with the specified target locations.
 This of course relies on such bifurcations existing.
-The role of the unsupervised term is to encourage an optimiser to move towards parameter regimes that do exhibit bifurcations.
+The role of the eigenvalue term is to encourage an optimiser to move towards parameter regimes that do exhibit bifurcations.
 
-\subsubsection{Supervised term: matching bifurcations to target locations}
+\subsubsection{Error term: matching bifurcations to target locations}
 
@@
 \end{equation}
-The supervised term is only zero when each target is matched by at least one prediction and allows for cases where the number of predictions is greater than or equal to the number of targets $|\predictions| \geq |\targets|$.
+The error term is only zero when each target is matched by at least one prediction and allows for cases where the number of predictions is greater than or equal to the number of targets $|\predictions| \geq |\targets|$.
 An alternative approach, which undesirably introduces more hyper-parameters, would be to let each prediction $\predictions(\theta)$ represent the centroid of a mixture distribution and use expectation-maximisation to match the centroids to targets $\targets$.
 
-\subsubsection{Unsupervised term: encouraging bifurcations}
+\subsubsection{Eigenvalue term: encouraging bifurcations}
 \begin{wrapfigure}{r}{0.5\textwidth}
@@
 \clearpage
-
-% Finally, we note that $\measure(s)$ has the desirable property of being invariant to certain classes of transformations on $\rates(u,p)$ that do not change bifurcation locations (details provided in Appendix \ref{appendix:measure}).
 
@@
 The total measure $\Psi(\theta)$ is normalised such that $\Psi(\theta)\rightarrow1$ in the regimes where the controlled condition region $p$ is densely packed with bifurcations.
-The total measure $\Psi(\theta)$ is added to the supervised term as if it were a likelihood.
-This defines the semi-supervised cost function as
+The total measure $\Psi(\theta)$ is added to the error term as if it were a likelihood.
+This defines the cost function as
 \begin{align}
@@
 \end{align}
-The pre-factor $|\targets|-|\predictions|$ in the unsupervised term ensures that the gradients are always pushing optimisers towards a state where $|\targets|=|\predictions|$.
-This can be seen as a step-wise annealing of the unsupervised term until the desired state is reached.
+The pre-factor $|\targets|-|\predictions|$ in the eigenvalue term ensures that the gradients are always pushing optimisers towards a state where $|\targets|=|\predictions|$.
+This can be seen as a step-wise annealing of the eigenvalue term until the desired state is reached.
 
-\subsection{Differentiating the semi-supervised cost function}
+\subsection{Differentiating the cost function}
 
@@
 First, we note that while individual bifurcations $p(\theta)$ depend smoothly on $\theta$, the total number of predictions $|\predictions|$ does not have gradient contributions with respect to $\theta$.
-Therefore, we can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ within the definition the supervised term \eqref{eq:error} and the total measure \eqref{eq:total-measure}.
+Therefore, we can safely drop the dependency in the prediction counter and now proceed in taking gradients with respect to $\theta$ knowing that the only dependencies we need to track are for individual bifurcations $p(\theta)$ within the definition the error term \eqref{eq:error} and the total measure \eqref{eq:total-measure}.
 Therefore,
@@
 
+Our method can be used for more complex models than the minimal models shown in Figure \ref{fig:minimal-models:results}.
+To demonstrate this we applied the model to a 4 state 40 parameter model \cite{Grant2020InterpretationCircuit} (see Appendix \ref{appendix:double-exclusive}).
+
 Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods, which includes the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
@@
 
-We proposed a gradient-based semi-supervised approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
+We proposed a gradient-based approach for inferring the parameters of differential equations that produce a user-specified bifurcation diagram.
 By applying implicit layers \cite{Look2020DifferentiableLayers,Bai2019DeepModels} and the generalised Leibniz rule \cite{Flanders1973DifferentiationSign} to the geometry of the implicitly defined steady states \cite{Goldman2005CurvatureSurfaces} it is possible to use automatic differentiation methods to efficiently calculate gradients.
-We defined a bifurcation measure that uses the determinant of the state-space Jacobian as an indicator for bifurcating parameter regimes in the unsupervised term of the cost function.
+We defined a bifurcation measure that uses the determinant of the state-space Jacobian as an indicator for bifurcating parameter regimes in the eigenvalue term of the cost function.
 The gradients of the cost can be efficiently computed using automatic differentiation methods.
@@
 
-Although we did not consider limit cycles, our approach can be modified using the conditions for Poincar\'e-Andronov-Hopf bifurcations instead of the conditions on the Jacobian determinant.
+Although we did not consider limit cycles, the bifurcation measure can be extended to detect Poincar\'e-Andronov-Hopf bifurcations alongside changes in stability of fixed points (see Appendix \ref{appendix:hopf} for details).This measure enables detection of the onset of damped oscillations and/or the emergence of limit cycles (Figure \ref{fig:hopf-measure}).
+Used together with a steady state solver that detects periodic solutions and gradient-based optimisation, we can specify regions of damped oscillation and limit cycles.
 Our approach generalises naturally to bifurcation manifolds such as limit point curves or surfaces.
 This is because the normal components of implicit derivatives can still be calculated for under-determined systems of equations \cite{Jos2011OnSurface,Tao2016Near-IsometricTracking,Fujisawa2013CalculationInvariance}.
-In the case of manifolds it would be more appropriate to use isosurface extraction algorithms rather than continuation to obtain the steady-state manifold.
+In the case of manifolds it would be more appropriate to use isosurface extraction algorithms rather than continuation to obtain the steady-state manifold (further discussion in Appendix \ref{appendix:stedy-state-solvers}).
 Our approach does not depend on the details of the steady-state solver and therefore can still be applied.

### 2021-10-23 docs/appendix.tex

@@
 Coli} designed to exhibit a cusp bifurcation with respect to two input signals.
-Figure \ref{fig:double-exclusive-trajectories} shows the cost function with respect to iteration for sample trajectories.
+Let one of the input signals be the control condition $p$ while the other input is held constant at $1$.
+Figure \ref{fig:double-exclusive-optima} shows the cost function with respect to iteration for sample trajectories.
 We note that due to a
+
+\begin{figure}[ht]
+\centering
+\setlength\unitlength{1cm}
+{\phantomsubcaption\label{fig:double-exclusive-optima:parameters}}
+{\phantomsubcaption\label{fig:double-exclusive-optima:models}}
+\includegraphics[width=13cm]{double-exclusive-optima}
+\begin{picture}(0,0) \put(-9,6.3){\subref{fig:double-exclusive-optima:parameters}} \end{picture}
+\begin{picture}(0,0) \put(-3,6.3){\subref{fig:double-exclusive-optima:models}}
+\end{picture}
+\caption{Bifurcation inference for the \emph{double exclusive reporter}.
+\subref{fig:double-exclusive-optima:parameters}.
+Optimal parameter estimates $\theta^*$ for the targets $\targets=\{4,5\}$ reveal four clusters of qualitatively different regimes: mutual activation ($a_1 < 1$; cluster 1) and mutual inhibition ($a_1 > 1$; cluster 2).
+\subref{fig:double-exclusive-optima:models}.
+Example bifurcation diagrams indicate positively and negatively correlated dependencies between the two model states, as a function of the control condition.}
+\label{fig:double-exclusive-optima}
+\end{figure}
 

### 2021-10-25 docs/appendix.tex
Updates from Overleaf

@@
 \includegraphics[width=13cm]{double-exclusive-optima}
-\begin{picture}(0,0) \put(-9,6.3){\subref{fig:double-exclusive-optima:parameters}} \end{picture}
-\begin{picture}(0,0) \put(-3,6.3){\subref{fig:double-exclusive-optima:models}}
+\begin{picture}(0,0) \put(-13,7){\subref{fig:double-exclusive-optima:parameters}} \end{picture}
+\begin{picture}(0,0) \put(-7,7){\subref{fig:double-exclusive-optima:models}}
 \end{picture}

### 2022-03-20 src/Structures.jl
compat with BifurcationKit v0.1.11

@@ -47,7 +47,7 @@ Branch{V,T} = Vector{BranchPoint{V,T}}
 dim(branch::Branch) = length(branch) > 0 ? length(first(branch).z)-1 : 0
 
 function push!(branch::Branch,state::ContState)
-    z,∂z = copy(state.z_old), [state.tau.u;state.tau.p]
+    z,∂z = copy(state.z), [state.τ.u;state.τ.p]
     push!(branch, ( z=z, λ=state.eigvals, ds=norm(∂z)*abs(state.ds), bif=detectBifucation(state) ))
 end
 

### 2022-03-20 docs/appendix.tex

@@
 
-\clearpage
 \section{Bifurcation Measure Properties}
 \label{appendix:conditions}
-Consider a vector $v(s)\in\Reals^N$ parameterised by $s\in\Reals$ that is tangent to an equilibrium manifold defined by $\rates(u)=0$.
+Consider a vector $v(s)\in\Reals^N$ parametrised by $s\in\Reals$ that is tangent to an equilibrium manifold defined by $\rates(u)=0$.
 The conditions for a non-degenerate static bifurcation at $s^*$ along such a tangent can be expressed in terms of an eigenvalue $\lambda(s)$ of the state-space Jacobian crossing zero with a finite slope.
@@
     \measure(s)=1
+    \label{eq:measure-conditions}
 \end{equation}
@@
 This gives rise to finite gradients in the eigenvalue term in regimes far away from any bifurcation.
-
-\section{Extension for Hopf Bifurcations}
-\label{appendix:hopf}
-
-In order to detect bifurcations involving limit cycles, the measure must be extended to detect changes in the real part $\Re\mathrm{e}[\lambda(s)]$ for any eigenvalue of the Jacobian.
-These conditions can no longer be compactly written in terms of the determinant.
-Instead, the measure can be defined as the sum of eigenvalue terms
-\begin{align}
-    \measure(s):=\sum_{\lambda(s)\in\frac{\partial\rates}{\partial u}}
-    \left(\left|\frac{d}{ds}\log\Re\mathrm{e}[\lambda(s)]\right|^{-1}+1\right)^{-1}
-    \label{eq:hopf-measure}
-\end{align}
-
-The directional derivative of the logarithm diverges under two conditions: when eigenvalues vanish $\lambda(s)=0$ and when the directional derivative $\frac{d}{ds}\Re\mathrm{e}[\lambda(s)]$ diverges.
-These properties are sufficient for detecting the onset of damped oscillations and emergence of limit cycles via Hopf bifurcation as shown in Figure \ref{fig:hopf-measure}.
-Eigenvalues with negative real part which gain a finite imaginary part give rise to damped oscillations.
-At this onset we observe a discontinuity in the derivative $\frac{d}{ds}\Re\mathrm{e}[\lambda(s)]$ which is detected by equation \eqref{eq:hopf-measure}.
-Once damped oscillations exist, flipping the stability of the stable fixed point gives rise to a limit cycle, which can be detected by inspecting $\Re\mathrm{e}[\lambda(s)]$.
-
-\begin{figure}[H]
-\centering
-    \includegraphics[width=60mm]{hopf-measure}
-    \caption{Bifurcation measure $\measure(s)$ and eigenvalues $\lambda(s)$ along the arclength $s$ for two different bifurcation curves demonstrating how the measure detects non-zero imaginary parts $\Im\mathrm{m}[\lambda]$ (onset of damped oscillations marked by circle) and sign changes in real parts $\Re\mathrm{e}[\lambda]$ (Hopf bifurcations marked by stars)}
-    \label{fig:hopf-measure}
-\end{figure}
-
-In principle it is possible to construct measures to detect a variety of bifurcations as long as the conditions can be expressed in terms of derivatives with respect to fixed-point manifold direction $s$.
-Measures can be used sequentially or in parallel to encourage optimisers to run through a sequence of bifurcations or place specific bifurcation types next to each other.
 
@@
 This curve could be open or closed and changes in $\theta$ could change the curve topology as well.
-Let the function $\gamma_{\theta}:\Reals\rightarrow\Reals^{N+1}$ be a parameterisation of the position vector along the curve within a fixed domain $s\in\mathcal{S}$.
-Note that the choice of parameterisation is arbitrary and our results should not depend on this choice.
-Furthermore, if we parametrise the curve $\mathcal{C}(\theta)$ with respect to a fixed domain $\mathcal{S}$ the dependence on $\theta$ is picked up by the parameterisation $\gamma_{\theta}(s)$.
+Let the function $\gamma_{\theta}:\Reals\rightarrow\Reals^{N+1}$ be a parametrisation of the position vector along the curve within a fixed domain $s\in\mathcal{S}$.
+Note that the choice of parametrisation is arbitrary and our results should not depend on this choice.
+Furthermore, if we parametrise the curve $\mathcal{C}(\theta)$ with respect to a fixed domain $\mathcal{S}$ the dependence on $\theta$ is picked up by the parametrisation $\gamma_{\theta}(s)$.
 We can write a line integral of any scalar function $L_{\theta}:\Reals^{N+1}\rightarrow\Reals$ on the curve as
@@
 \end{align}
-The tangential component of the deformation field is not uniquely determined because there is no unique way of parameterising a surface.
+The tangential component of the deformation field is not uniquely determined because there is no unique way of parametrising a surface.
 This is the subject of many computer graphics papers \cite{Jos2011OnSurface,Tao2016Near-IsometricTracking,Fujisawa2013CalculationInvariance}.
 We are however not interested in the continuous propagation of a mesh - as is the subject of those papers.
-In fact we are looking for a deformation field that is orthogonal to the tangent vector $\hat{\tangent}(z) \cdot\frac{d z}{d\theta} =0$ for the space curve, and therefore letting the tangential component of the deformation equal zero is a valid choice and we can it instead of the parameterised deformation
+In fact we are looking for a deformation field that is orthogonal to the tangent vector $\hat{\tangent}(z) \cdot\frac{d z}{d\theta} =0$ for the space curve, and therefore letting the tangential component of the deformation equal zero is a valid choice and we can it instead of the parametrised deformation
 \begin{align}
@@
 
-\section{A More Complex Model}
+\clearpage
+
+\section{Application of Bifurcation Inference to a Complex Model}
 \label{appendix:double-exclusive}
-To demonstrate the wider reaching applicability of our method we optimise the \emph{double exclusive reporter} \cite{Grant2020InterpretationCircuit}, a synthetic \emph{E.
-Coli} designed to exhibit a cusp bifurcation with respect to two input signals.
-Let one of the input signals be the control condition $p$ while the other input is held constant at $1$.
-Figure \ref{fig:double-exclusive-optima} shows the cost function with respect to iteration for sample trajectories.
-We note that due to a
+
+To demonstrate the wider reaching applicability of our method we optimise the \emph{double exclusive reporter} \cite{Grant2020InterpretationCircuit}, a synthetic gene circuit in \emph{E.
+coli} that was designed to exhibit a cusp bifurcation.
+The circuit behaviour is observed by measuring a fluorescent protein whose expression is controlled by transcription factors (regulatory proteins) LacI $(L)$ and TetR $(T)$, whose expression is in turn controlled by externally controllable \emph{input} signals $c_{6}$ and $c_{12}$.
+To apply the method, we consider one of the input signals be the control condition $c_{6}=p$, with the other packed together with the remaining 20 parameters into vector $\theta$.
+Once the optima $\theta^*$ have been obtained, we perform dimensionality reduction using \texttt{GigaSOM.jl} \cite{Kratochvil2020GigaSOM.jl:Datasets} so that the results can be visualised in a two dimensional embedding (Figure \ref{fig:double-exclusive-optima:parameters}).
+
+The embedding reveals four optimal parameter regions.
+We find that, as with the two-state model in the main text \eqref{eq:two-state}, there are two qualitatively distinct regimes: mutual activation (region 1) and inhibition (regions 2-4).
+The mutual inhibition region can be further subdivided into three regions that are geometrically equivalent, but kinetically distinct: region 3 has swapped kinetic roles for regulatory proteins LacI and TetR compared to region 2, and region 4 has additional damped oscillations in the dynamics across the whole range of \emph{input} $c_{6}$ (Figure \ref{fig:double-exclusive-optima:models}).
+The two dimensional embedding of sampled optima $\theta^*$ enables navigation the space of qualitative behaviours of the \emph{double exclusive reporter} and organisation in terms of geometric and kinetic equivalence.
 
@@
 \subref{fig:double-exclusive-optima:parameters}.
-Optimal parameter estimates $\theta^*$ for the targets $\targets=\{4,5\}$ reveal four clusters of qualitatively different regimes: mutual activation ($a_1 < 1$; cluster 1) and mutual inhibition ($a_1 > 1$; cluster 2).
+Optimal parameter estimates $\theta^*$ for the targets $\targets=\{1,2\}$ (indicated by yellow lines in panel B) reveal four regions  with two geometrically different regimes: mutual activation (region 1) and mutual inhibition (regions 2-4).
 \subref{fig:double-exclusive-optima:models}.
-Example bifurcation diagrams indicate positively and negatively correlated dependencies between the two model states, as a function of the control condition.}
+Example bifurcation diagrams indicate that region 2 has swapped kinetics between $L$ and $T$ to region 3.
+Region 4 has models with non-zero imaginary parts to eigenvalues indicating damped oscillations (shown in light green).}
 \label{fig:double-exclusive-optima}
@@
 
-\section{Steady State Solvers}
-\label{appendix:stedy-state-solvers}
+These results were obtained with a modification of the bifurcation measure \eqref{eq:measure} to improve convergence rates.
+In parameter regimes where bifurcations are not present, according to conditions \eqref{eq:measure-conditions}, maximising the measure $\measure(s)$ can lead to a divergence in directional derivative $\frac{d\lambda}{ds}\rightarrow\pm\infty$ rather than a creation of a bifurcation.
+To discourage this from happening we can flatten out the gradients in that regime by applying the $\tanh$ non-linearity to the determinant.
+This leads to
+\begin{align}
+    \measure(s):=
+    \left(1+\left|\frac{\tanh\Det}{\frac{d}{ds}\tanh\Det}\right|\right)^{-1}
+    \label{eq:measure-tanh}
+\end{align}
+
+\clearpage
+\section{Extension for Hopf Bifurcations}
+\label{appendix:hopf}
+
+In order to detect bifurcations involving limit cycles, the measure must be extended to detect changes in the real part $\Re\mathrm{e}[\lambda(s)]$ for any eigenvalue of the Jacobian.
+These conditions can no longer be compactly written in terms of the determinant.
+Instead, the measure can be defined as the sum of eigenvalue terms
+\begin{align}
+    \measure(s):=\sum_{\lambda(s)\in\frac{\partial\rates}{\partial u}}
+    \left(\left|\frac{d}{ds}\log\Re\mathrm{e}[\lambda(s)]\right|^{-1}+1\right)^{-1}
+    \label{eq:hopf-measure}
+\end{align}
+
+The directional derivative of the logarithm diverges under two conditions: when eigenvalues vanish $\lambda(s)=0$ and when the directional derivative $\frac{d}{ds}\Re\mathrm{e}[\lambda(s)]$ diverges.
+These properties are sufficient for detecting the onset of damped oscillations and emergence of limit cycles via Hopf bifurcation as shown in Figure \ref{fig:hopf-measure}.
+Eigenvalues with negative real part which gain a finite imaginary part give rise to damped oscillations.
+At this onset we observe a discontinuity in the derivative $\frac{d}{ds}\Re\mathrm{e}[\lambda(s)]$ which is detected by equation \eqref{eq:hopf-measure}.
+Once damped oscillations exist, flipping the stability of the stable fixed point gives rise to a limit cycle, which can be detected by inspecting $\Re\mathrm{e}[\lambda(s)]$.
+
+\begin{figure}[H]
+\centering
+    \includegraphics[width=60mm]{hopf-measure}
+    \caption{Bifurcation measure $\measure(s)$ and eigenvalues $\lambda(s)$ along the arclength $s$ for two different bifurcation curves demonstrating how the measure detects non-zero imaginary parts $\Im\mathrm{m}[\lambda]$ (onset of damped oscillations marked by circle) and sign changes in real parts $\Re\mathrm{e}[\lambda]$ (Hopf bifurcations marked by stars)}
+    \label{fig:hopf-measure}
+\end{figure}
+
+In principle it is possible to construct measures to detect a variety of bifurcations as long as the conditions can be expressed in terms of derivatives with respect to fixed-point manifold direction $s$.
+Measures can be used sequentially or in parallel to encourage optimisers to run through a sequence of bifurcations or place specific bifurcation types next to each other.

### 2022-03-20 docs/article.tex

@@
 It could even be impossible for the model to bifurcate in the manner desired.
+For models with a sufficiently small number of parameters, finding specific bifurcation diagrams is typically done by hand \cite{Csikasz-Nagy2006AnalysisRegulation}.
 Several approaches exist to place bifurcations to desired locations once a manifold is present \cite{Iwasaki1997AnType,Lu2006InverseSystems,Dobson2004DistanceBifurcations} yet typically resort to sampling techniques to search for them in the first place \cite{Chickarmane2005BifurcationTool,Conrad2006BifurcationClock}.
+It is always possible to design bespoke goodness of fit measures to find specific model behaviours, for example using the period and phase of limit cycle oscillations \cite{Locke2005ModellingThaliana}.
+However, this approach does not generalise across a wider set of qualitative behaviours.
 Progress has been made in cases where model structure and stability conditions are used to refine the search space \cite{Otero-Muras2018Optimization-basedModels,Otero-Muras2014ACurves} yet the resulting objectives are still not explicit in the bifurcation targets and also not differentiable.
-It is always possible to bias goodness of fit during optimisation with domain-specific knowledge \cite{}, however these approaches do not aim to generalise across a wider set of qualitative behaviours.
 In the emerging field of scientific machine learning \cite{Rackauckas2017Differentialequations.jlJulia,Rackauckas2018ASolutions,Rackauckas2020UniversalLearning}, parameters of structured mechanistic models are favoured over flexible models in larger parameter spaces.
@@
     \item An end-to-end differentiable method for locating bifurcations in parameter space and then matching their dependency on a control condition to user-specified locations
-    \item Implementation of the method as a Julia package\\
-\href{https://github.com/gszep/BifurcationInference.jl}{\texttt{github.com/gszep/BifurcationInference.jl}}
+    \item Implementation of the method as a Julia package
+\href{https://github.com/gszep/BifurcationInference.jl}{\texttt{BifurcationInference.jl}}
     \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
@@
 Let $\targets$ be the set of those values and let us hypothesise that these transitions occur due to bifurcations in the dynamics that drive the underlying mechanism.
-Let us model the mechanism with a parameterised set of differential equations for states $u\in\Reals^N$ with a vector function $\rates$ in a parameter space $\theta\in\Reals^M$.
+Let us model the mechanism with a parametrised set of differential equations for states $u\in\Reals^N$ with a vector function $\rates$ in a parameter space $\theta\in\Reals^M$.
 
@@
 Any location on the curve $u(s)$ and $p(s)$ must satisfy the steady-state of equations \eqref{eq:model}.
+Directional derivatives $\frac{d}{ds}$ along the diagram require the calculation of a vector that is tangent to the diagram (see Supplementary \ref{appendix:tangent-fields}).
 The determinant approaching zero along the diagram means that the dynamics of the system are slowing down, which is an important indicator for the onset of a transition between qualitative behaviours.
@@
 	\predictions(\theta):=\left\{\,
-	p\,\,\exists\,\,u:\,\,\rates(u,p)=0,\,\,\Det=0,
+	p\,\,|\,\,\exists\,\,u:\,\,\rates(u,p)=0,\,\,\Det=0,
 	\,\, \frac{d}{ds}\Det\neq0
@@
 \end{align}
-A proof of how the conditions \eqref{eq:predictions} are necessary and sufficient for static non-degenerate bifurcations is detailed in Appendix \ref{appendix:conditions}.
+A proof of how the conditions \eqref{eq:predictions} are necessary and sufficient for static non-degenerate bifurcations is detailed in Supplementary \ref{appendix:conditions}.
 The most common bifurcations between steady states, not including limit cycles, are saddle-nodes and pitchforks \cite{Haragus2011LocalSystems}.
@@
 {\phantomsubcaption\label{fig:pitchfork}}
-\includegraphics[width=5.5cm]{saddle-node}
-\begin{picture}(0,0)\put(-5.5,9){\subref{fig:saddle-node}} \end{picture}
-\includegraphics[width=5.5cm]{pitchfork}
-\begin{picture}(0,0)\put(-5.5,9){\subref{fig:pitchfork}} \end{picture}
+\includegraphics[width=4.95cm]{saddle-node}
+\begin{picture}(0,0)\put(-5,8){\subref{fig:saddle-node}} \end{picture}
+\includegraphics[width=4.95cm]{pitchfork}
+\begin{picture}(0,0)\put(-5,8){\subref{fig:pitchfork}} \end{picture}
 \caption{Illustration of bifurcation diagrams for minimal models of bifurcations.
@@
 We can see from Figure \ref{fig:minimal-models} and definition \eqref{eq:predictions} that predictions $p(\theta)$ can be identified by looking for points along the curve where the determinant crosses zero $\Det=0$ with a finite slope $\frac{d}{ds}\Det\neq 0$.
-Using these quantities we can define a positive semi-definite measure $\measure(s)$ of zero crossings in the determinant along a curve parameterised by $s$ which we define as
+Using these quantities we can define a positive semi-definite measure $\measure(s)$ of zero crossings in the determinant along a curve parametrised by $s$ which we define as
 \begin{equation}
@@
 To calculate the gradient of the measure $\frac{\partial \Psi}{\partial\theta}$ we need to differentiate line integrals that depend on $\theta$.
-Fortunately this can be done by the application of the generalised Leibniz integral rule, details of which can be found in Appendix \ref{appendix:leibniz-rule}.
+Fortunately this can be done by the application of the generalised Leibniz integral rule, details of which can be found in Supplementary \ref{appendix:leibniz-rule}.
 
@@
 \label{section:results}
-In this section, we apply the method first to minimal examples that can produce saddle-node and pitchfork bifurcations, and then a more complex model that has multiple parametric regimes that produce saddle-node bifurcations.
-
-\subsection{Minimal Models}
-Optimisations of two parameters $(\theta_1,\theta_2)$ using simple gradient descent from \texttt{Flux.jl} with learning rate $\eta=0.01$ for the minimal saddle-node and pitchfork models yield trajectories approaching lines of global minima in the cost function (Figures \ref{fig:minimal-models:results}) which represent a set of geometrically equivalent models.
-Two bifurcation diagrams are geometrically equivalent if the number, type and locations of bifurcations match the specified targets $\targets$.
+In this section, we apply the method first to minimal examples that can produce saddle-node and pitchfork bifurcations (both $N=1$, $M=2$), and then a slightly more complex model ($N=2$, $M=5$) that has multiple parametric regimes producing saddle-node bifurcations.
+We also demonstrate our method on a model of greater complexity, to convince the reader that the method can be used on more realistic examples with practical significance.
+In Supplementary \ref{appendix:double-exclusive} we demonstrate the identification of saddle-node bifurcations and damped oscillations in a model ($N=4$, $M=21$) of a synthetic gene circuit in \emph{E.
+coli} \cite{Grant2020InterpretationCircuit}.
 
@@
 \centering
-\includegraphics[width=6cm]{saddle-landscape.png}
-\includegraphics[width=6cm]{saddle-optima.png}
-\includegraphics[width=6cm]{pitchfork-landscape.png}
-\includegraphics[width=6cm]{pitchfork-optima.png}
+\includegraphics[width=5.5cm]{saddle-landscape.png}
+\includegraphics[width=5.5cm]{saddle-optima.png}
+\includegraphics[width=5.5cm]{pitchfork-landscape.png}
+\includegraphics[width=5.5cm]{pitchfork-optima.png}
 \caption{Saddle-node $\rates(u,p) = p + \theta_{1}u+\theta_{2}u^3$ and pitchfork $\rates(u,p) = \theta_{1} + u p +\theta_{2}u^3$ optimised with respect to $\theta$ so that predicted bifurcations $\predictions(\theta)$ match targets $\targets$ in control condition $p$.
@@
 \end{figure}
+
+\subsection{Minimal Models}
+Optimisations of two parameters $(\theta_1,\theta_2)$ using simple gradient descent from \texttt{Flux.jl} with learning rate $\eta=0.01$ for the minimal saddle-node and pitchfork models (Figure \ref{fig:minimal-models}) yield trajectories approaching lines of global minima in the cost function (Figures \ref{fig:minimal-models:results}) which represent a set of geometrically equivalent models.
+Two bifurcation diagrams are geometrically equivalent if the number, type and locations of bifurcations match the specified targets $\targets$.
 
@@
 \end{equation}
-which reveals that only $a_1$, $a_2$ and the ratio between the sensitivity and degradation parameters, $\frac{k}{\mu_1}$, affect the solutions to this equation, and hence the locations of the bifurcations (Figure \ref{fig:two-state-optima:parameters}).
-In 98\% of 800 runs, optimisation using the ADAM optimiser \cite{Kingma2014} from \texttt{Flux.jl} with learning rate $\eta=0.1$ converged to one of two clusters: mutual activation ($a_1 < 1, a_2 < 1$; cluster 1) and mutual inhibition ($a_1 > 1, a_2 > 1$; cluster 2) regimes.
-Example bifurcation diagrams illustrate how the bifurcation curves of each species are positively correlated in mutual activation and negatively correlated for mutual inhibition (Figure \ref{fig:two-state-optima:models}).
 
@@
 
+which reveals that only $a_1$, $a_2$ and the ratio between the sensitivity and degradation parameters, $\frac{k}{\mu_1}$, affect the solutions to this equation, and hence the locations of the bifurcations (Figure \ref{fig:two-state-optima:parameters}).
+In 98\% of 800 runs, optimisation using the ADAM optimiser \cite{Kingma2014} from \texttt{Flux.jl} with learning rate $\eta=0.1$ converged to one of two clusters: mutual activation ($a_1 < 1, a_2 < 1$; cluster 1) and mutual inhibition ($a_1 > 1, a_2 > 1$; cluster 2) regimes.
+Example bifurcation diagrams illustrate how the bifurcation curves of each species are positively correlated in mutual activation and negatively correlated for mutual inhibition (Figure \ref{fig:two-state-optima:models}).
+
 In order to maintain biological interpretability, optimisation was restricted to the positive parameter regime by transforming the parameters to log-space $\theta\rightarrow10^\theta$.
@@
 
-Our method can be used for more complex models than the minimal models shown in Figure \ref{fig:minimal-models:results}.
-To demonstrate this we applied the model to a 4 state 40 parameter model \cite{Grant2020InterpretationCircuit} (see Appendix \ref{appendix:double-exclusive}).
-
 Performing one iteration of the optimisation requires the computation of the gradient of the cost \eqref{eq:gradient-loss}, requiring a computation of the bifurcation diagram with parameter continuation methods, which includes the evaluation of matrix inversions \eqref{eq:gradient-bifurcation}.
@@
 This leaves us with the computational bottleneck of calculating the determinant of the state space Jacobian, required in both the bifurcation measure \eqref{eq:measure} and gradient \eqref{eq:gradient-bifurcation}.
-This calculation scales like $N^2$ where $N$ is the number of state space variables (Figure \ref{fig:scaling}).
+This calculation scales like $N^2$ where $N$ is the number of state space variables (Figure \ref{fig:scaling-states}).
 
@@
 \centering
-\includegraphics[width=0.48\textwidth]{scaling}
-\caption{Complexity scaling of calculating the gradient of the cost function. Calculations were performed on an Intel Core i7-6700HQ CPU @ 2.60GHz x 8 without GPU acceleration}
+\setlength\unitlength{1cm}
+{\phantomsubcaption\label{fig:scaling-states}}
+{\phantomsubcaption\label{fig:scaling-parameters}}
+\includegraphics[width=0.48\textwidth]{scaling-states}
+\includegraphics[width=0.48\textwidth]{scaling-parameters}
+\begin{picture}(0,0)
+\put(-13.5,6.5){\subref{fig:scaling-states}} \end{picture}
+\begin{picture}(0,0) \put(-7,6.5){\subref{fig:scaling-parameters}}
+\end{picture}
+\caption{\subref{fig:scaling-states}.
+Execution time (time to calculate cost gradient) with respect to states $N$.
+\subref{fig:scaling-parameters}.
+Convergence times (the time it takes to find and match a bifurcation to within 1\% of a specified target) with respect to the number of parameters $M$, comparing against a gradient-free approach: Nelder-Mead.
+Calculations were performed on an Intel Core i7-6700HQ CPU @ 2.60GHz x 8 without GPU acceleration.}
 \label{fig:scaling}
@@
 
-For this calculation a model \eqref{eq:scaling} was chosen that so that it is extensible both in the number of parameters $M$ and the number of states $N$. 
+For the complexity study, a model was designed so that it is extensible both in the number of parameters $M$ and the number of states $N$.
+There are many choices for this; we opted for a model of the form
 \begin{equation}
@@
 \end{equation}
+
 In this model only the first state $u_1$ defines the shape of the bifurcation diagram, while the remaining states are merely linearly proportional to the first.
@@
 Note that any more efficient calculation must still permit back-propagation through it.
+
+To demonstrate the benefits of the gradient-based aspect of our method we compare convergence times of gradient descent against a gradient-free approach.
+We use the Nelder-Mead method from \texttt{Optim.jl} \cite{KMogensen2018Optim:Julia} and obtain convergence times as the number of parameters $M$ is increased (Figure \ref{fig:scaling-parameters}).
+We observe that for our method convergence times scale like $\sqrt{M}$ compared to $M^2$ for the gradient-free approach.
 
@@
 We demonstrated our approach on models with one bifurcation parameter that can give rise to pitchforks and saddle-nodes.
-The estimated parameters form distinct clusters, allowing us to organise models in terms of topological and geometric equivalence.
-In the case of the genetic toggle switch we recovered mutual activation and inhibition regimes.
+The estimated parameters form distinct clusters, allowing us to organise models in terms of topological and geometric equivalence (Figure \ref{fig:minimal-models:results}).
+In the case of the genetic toggle switch (Figure \ref{fig:two-state-optima}) and a more complex model \cite{Grant2020InterpretationCircuit} (Figure \ref{fig:double-exclusive-optima}) we recovered mutual activation and inhibition regimes.
+In the more complex model we found a damped oscillatory regime that was not known about in the original paper.
 
-Although we did not consider limit cycles, the bifurcation measure can be extended to detect Poincar\'e-Andronov-Hopf bifurcations alongside changes in stability of fixed points (see Appendix \ref{appendix:hopf} for details).This measure enables detection of the onset of damped oscillations and/or the emergence of limit cycles (Figure \ref{fig:hopf-measure}).
+Although we did not consider limit cycles, the bifurcation measure can be extended to detect Poincar\'e-Andronov-Hopf bifurcations alongside changes in stability of fixed points (see Supplementary \ref{appendix:hopf} for details).
+This measure enables detection of the onset of damped oscillations and/or the emergence of limit cycles (Figure \ref{fig:hopf-measure}).
 Used together with a steady state solver that detects periodic solutions and gradient-based optimisation, we can specify regions of damped oscillation and limit cycles.
@@
 This is because the normal components of implicit derivatives can still be calculated for under-determined systems of equations \cite{Jos2011OnSurface,Tao2016Near-IsometricTracking,Fujisawa2013CalculationInvariance}.
-In the case of manifolds it would be more appropriate to use isosurface extraction algorithms rather than continuation to obtain the steady-state manifold (further discussion in Appendix \ref{appendix:stedy-state-solvers}).
+In the case of manifolds it would be more appropriate to use isosurface extraction algorithms rather than continuation to obtain the steady-state manifold.
 Our approach does not depend on the details of the steady-state solver and therefore can still be applied.
@@
 A special thanks go to Romain Veltz and the Julia community for helpful pointers on package development and discussions over Slack.
-
+This work was supported by Microsoft Research through its PhD Scholarship Programme and the EPSRC Centre for Doctoral Training in Cross-Disciplinary Approaches to Non-Equilibrium Systems (CANES, EP/L015854/1).
+\clearpage
 \bibliography{refs}
@@
 \clearpage\pagenumbering{arabic}\setcounter{page}{1}
-\section*{Appendix}
+\section*{Supplementary Material}
 \appendix
+\counterwithin{figure}{section}
+\counterwithin{equation}{section}
+
 \input{appendix}
